### 2023-05-31 ited-template.tex
Initial commit

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Author style for INFORMS Transactions on Education (ited)
%% Mirko Janc, Ph.D., INFORMS, pubtech@informs.org
%% ver. 0.92, June 2009 -- default options: single-spaced, double-blinded
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\documentclass[ited]{informs3}                      % for a regular run
%\documentclass[ited,nonblindrev]{informs3}          % for review, not blinded
%\documentclass[ited,blindrev,copyedit]{informs3}    % spaced for copyediting
\documentclass[ited,blindrev]{informs3}              % for review, blinded

% If hyperref is used, dvi-to-ps driver of choice must be declared as
%   an additional option to the \documentstyle. For example
%\documentclass[dvips,ited]{informs1}      % if dvips is used 
%\documentclass[dvipsone,ited]{informs1}   % if dvipsone is used, etc.

% Private macros here (check that there is no clash with the style)

% Natbib setup for author-year style
\usepackage{natbib}
 \bibpunct[, ]{(}{)}{,}{a}{}{,}%
 \def\bibfont{\small}%
 \def\bibsep{\smallskipamount}%
 \def\bibhang{24pt}%
 \def\newblock{\ }%
 \def\BIBand{and}%

%% Setup of theorem styles. Outcomment only one. 
%% Preferred default is the first option.
\TheoremsNumberedThrough     % Preferred (Theorem 1, Lemma 1, Theorem 2)
%\TheoremsNumberedByChapter  % (Theorem 1.1, Lema 1.1, Theorem 1.2)

%% Setup of the equation numbering system. Outcomment only one.
%% Preferred default is the first option.
\EquationsNumberedThrough    % Default: (1), (2), ...
%\EquationsNumberedBySection % (1.1), (1.2), ...

% In the reviewing and copyediting stage enter the manuscript number.
%\MANUSCRIPTNO{} % When the article is logged in and DOI assigned to it,
                 %   this manuscript number is no longer necessary

%%%%%%%%%%%%%%%%
\begin{document}
%%%%%%%%%%%%%%%%

% Outcomment only when entries are known. Otherwise leave as is and 
%   default values will be used.
%\setcounter{page}{1}
%\VOLUME{00}%
%\NO{0}%
%\MONTH{Xxxxx}% (month or a similar seasonal id)
%\YEAR{0000}% e.g., 2005
%\FIRSTPAGE{000}%
%\LASTPAGE{000}%
%\SHORTYEAR{00}% shortened year (two-digit)
%\ISSUE{0000} %
%\LONGFIRSTPAGE{0001} %
%\DOI{10.1287/xxxx.0000.0000}%

% Author's names for the running heads
% Sample depending on the number of authors;
% \RUNAUTHOR{Jones}
% \RUNAUTHOR{Jones and Wilson}
% \RUNAUTHOR{Jones, Miller, and Wilson}
% \RUNAUTHOR{Jones et al.} % for four or more authors
% Enter authors following the given pattern:
\RUNAUTHOR{}

% Title or shortened title suitable for running heads. Sample:
% \RUNTITLE{Bundling Information Goods of Decreasing Value}
% Enter the (shortened) title:
\RUNTITLE{}

% Full title. Sample:
% \TITLE{Bundling Information Goods of Decreasing Value}
% Enter the full title:
\TITLE{}

% Block of authors and their affiliations starts here:
% NOTE: Authors with same affiliation, if the order of authors allows, 
%   should be entered in ONE field, separated by a comma. 
%   \EMAIL field can be repeated if more than one author
\ARTICLEAUTHORS{%
\AUTHOR{Author1}
\AFF{Author1 affiliation, \EMAIL{}, \URL{}}
\AUTHOR{Author2}
\AFF{Author2 affiliation, \EMAIL{}, \URL{}}
% Enter all authors
} % end of the block

\ABSTRACT{%
Text of your abstract % Enter your abstract
}%

% Sample
%\KEYWORDS{deterministic inventory theory; infinite linear programming duality; 
%  existence of optimal policies; semi-Markov decision process; cyclic schedule}

% Fill in data. If unknown, outcomment the field
\KEYWORDS{}
\HISTORY{}

\maketitle
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Samples of sectioning (and labeling) in ITED
% NOTE: (1) \section and \subsection do NOT end with a period
%       (2) \subsubsection and lower need end punctuation
%       (3) capitalization is as shown (title style).
%
%\section{Introduction.}\label{intro} %%1.
%\subsection{Duality and the Classical EOQ Problem.}\label{class-EOQ} %% 1.1.
%\subsection{Outline.}\label{outline1} %% 1.2.
%\subsubsection{Cyclic Schedules for the General Deterministic SMDP.}
%  \label{cyclic-schedules} %% 1.2.1
%\section{Problem Description.}\label{problemdescription} %% 2.

% Text of your paper here

Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod
tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim
veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea
commodo consequat. Duis aute irure dolor in reprehenderit in voluptate
velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint
occaecat cupidatat non proident, sunt in culpa qui officia deserunt
mollit anim id est laborum.

Sed ut perspiciatis unde omnis iste natus error sit voluptatem
accusantium doloremque laudantium, totam rem aperiam, eaque ipsa quae ab
illo inventore veritatis et quasi architecto beatae vitae dicta sunt
explicabo. Nemo enim ipsam voluptatem quia voluptas sit aspernatur aut
odit aut fugit, sed quia consequuntur magni dolores eos qui ratione
voluptatem sequi nesciunt. Neque porro quisquam est, qui dolorem ipsum
quia dolor sit amet, consectetur, adipisci velit, sed quia non numquam
eius modi tempora incidunt ut labore et dolore magnam aliquam quaerat
voluptatem. Ut enim ad minima veniam, quis nostrum exercitationem ullam
corporis suscipit laboriosam, nisi ut aliquid ex ea commodi consequatur?
Quis autem vel eum iure reprehenderit qui in ea voluptate velit esse
quam nihil molestiae consequatur, vel illum qui dolorem eum fugiat quo
voluptas nulla pariatur?

At vero eos et accusamus et iusto odio dignissimos ducimus qui
blanditiis praesentium voluptatum deleniti atque corrupti quos dolores
et quas molestias excepturi sint occaecati cupiditate non provident,
similique sunt in culpa qui officia deserunt mollitia animi, id est
laborum et dolorum fuga. Et harum quidem rerum facilis est et expedita
distinctio. Nam libero tempore, cum soluta nobis est eligendi optio
cumque nihil impedit quo minus id quod maxime placeat facere possimus,
omnis voluptas assumenda est, omnis dolor repellendus. Temporibus autem
quibusdam et aut officiis debitis aut rerum necessitatibus saepe eveniet
ut et voluptates repudiandae sint et molestiae non recusandae. Itaque
earum rerum hic tenetur a sapiente delectus, ut aut reiciendis
voluptatibus maiores alias consequatur aut perferendis doloribus
asperiores repellat.

% Acknowledgments here
\ACKNOWLEDGMENT{%
% Enter the text of acknowledgments here
}% Leave this (end of acknowledgment)

% Appendix here
% Options are (1) APPENDIX (with or without general title) or 
%             (2) APPENDICES (if it has more than one unrelated sections)
% Outcomment the appropriate case if necessary
%
% \begin{APPENDIX}{<Title of the Appendix>}
% \end{APPENDIX}
%
%   or 
%
% \begin{APPENDICES}
% \section{<Title of Section A>}
% \section{<Title of Section B>}
% etc
% \end{APPENDICES}

% References here (outcomment the appropriate case)

% CASE 1: BiBTeX used to constantly update the references 
%   (while the paper is being written).
%\bibliographystyle{informs2014} % outcomment this and next line in Case 1
%\bibliography{<your bib file(s)>} % if more than one, comma separated

% CASE 2: BiBTeX used to generate mypaper.bbl (to be further fine tuned)
%\input{mypaper.bbl} % outcomment this line in Case 2

\end{document}

### 2023-06-01 python_excel_ite.tex
rough structure

--- 2023-05-31 ited-template.tex
@@
 % Enter authors following the given pattern:
-\RUNAUTHOR{}
+\RUNAUTHOR{Isken}
 
@@
 % Enter the (shortened) title:
-\RUNTITLE{}
+\RUNTITLE{Python for spreadsheet type modeling}
 
@@
 % Enter the full title:
-\TITLE{}
+\TITLE{Using Python for spreadsheet type modeling}
 
@@
 \ARTICLEAUTHORS{%
-\AUTHOR{Author1}
-\AFF{Author1 affiliation, \EMAIL{}, \URL{}}
-\AUTHOR{Author2}
-\AFF{Author2 affiliation, \EMAIL{}, \URL{}}
+\AUTHOR{Mark W. Isken}
+\AFF{Oakland University, \EMAIL{isken@oakland.edu}, \URL{http://www.sba.oakland.edu/faculty/isken/}}
 % Enter all authors
@@
 \ABSTRACT{%
-Text of your abstract % Enter your abstract
+TODO: Text of your abstract % Enter your abstract
 }%
@@
 % Fill in data. If unknown, outcomment the field
-\KEYWORDS{}
+\KEYWORDS{python, spreadsheet modeling}
 \HISTORY{}
@@
 
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod
-tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim
-veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea
-commodo consequat. Duis aute irure dolor in reprehenderit in voluptate
-velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint
-occaecat cupidatat non proident, sunt in culpa qui officia deserunt
-mollit anim id est laborum.
+\section{Introduction}
 
-Sed ut perspiciatis unde omnis iste natus error sit voluptatem
-accusantium doloremque laudantium, totam rem aperiam, eaque ipsa quae ab
-illo inventore veritatis et quasi architecto beatae vitae dicta sunt
-explicabo. Nemo enim ipsam voluptatem quia voluptas sit aspernatur aut
-odit aut fugit, sed quia consequuntur magni dolores eos qui ratione
-voluptatem sequi nesciunt. Neque porro quisquam est, qui dolorem ipsum
-quia dolor sit amet, consectetur, adipisci velit, sed quia non numquam
-eius modi tempora incidunt ut labore et dolore magnam aliquam quaerat
-voluptatem. Ut enim ad minima veniam, quis nostrum exercitationem ullam
-corporis suscipit laboriosam, nisi ut aliquid ex ea commodi consequatur?
-Quis autem vel eum iure reprehenderit qui in ea voluptate velit esse
-quam nihil molestiae consequatur, vel illum qui dolorem eum fugiat quo
-voluptas nulla pariatur?
+% The opening part of the what-if notebook has some language for the motivation behind this module.
 
-At vero eos et accusamus et iusto odio dignissimos ducimus qui
-blanditiis praesentium voluptatum deleniti atque corrupti quos dolores
-et quas molestias excepturi sint occaecati cupiditate non provident,
-similique sunt in culpa qui officia deserunt mollitia animi, id est
-laborum et dolorum fuga. Et harum quidem rerum facilis est et expedita
-distinctio. Nam libero tempore, cum soluta nobis est eligendi optio
-cumque nihil impedit quo minus id quod maxime placeat facere possimus,
-omnis voluptas assumenda est, omnis dolor repellendus. Temporibus autem
-quibusdam et aut officiis debitis aut rerum necessitatibus saepe eveniet
-ut et voluptates repudiandae sint et molestiae non recusandae. Itaque
-earum rerum hic tenetur a sapiente delectus, ut aut reiciendis
-voluptatibus maiores alias consequatur aut perferendis doloribus
-asperiores repellat.
+Why I created
+
+Excel is widely used for building and using models of business problems to explore the impact of various model inputs on key outputs.
+Built in "what if?" tools such as Excel [Data Tables](https://support.microsoft.com/en-us/office/calculate-multiple-results-by-using-a-data-table-e95e2487-6ca6-4413-ad12-77542a5ea50b) and [Goal Seek](https://support.microsoft.com/en-us/office/use-goal-seek-to-find-the-result-you-want-by-adjusting-an-input-value-320cb99e-f4a4-417f-b1c3-4f369d6e66c7) are well known to power spreadsheet modelers.
+How might we do similar modeling and analysis using Python?
+
+While Python has been gaining momentum in the business analytics world, it is often used for data wrangling, analysis and visualization of tablular data using tools like pandas and matplotlib or Seaborn.
+You can find some great examples at [Chris Moffit's Practical Business Python blog](https://pbpython.com/).
+I use Python all the time for such tasks and teach a course called [Practical Computing for Data Analytics](http://www.sba.oakland.edu/faculty/isken/courses/mis5470/) that is Python (and R) based.
+But, it got me to thinking.
+What about those things for which Excel is well suited such as building formula based models and doing sensitivity analysis on these models?
+What would those look like in Python?
+
+\section{Module structure}
+
+\section{The modeling notebooks}
+\subsection{Notebook 1: Modeling and data tables}
+\subsection{Notebook 2: Goal seek}
+\subsection{Notebook 3: Monte-Carlo simulation}
+
+\section{The packaging notebooks}
+\subsection{Notebook 4: Project packaging}
+\subsection{Notebook 5: Documentation}
+
+\section{The wrangling notebooks}
+\subsection{Notebook 6: Manipulating Excel files with Python}
+
+\section{Student assignment}
+
+\section{Random notes}
+Learning good Python programming practices via learning to model in similar way that students learn good spreadsheet practices through building spreadsheet models using the W\&A textbook.
+
+How aap course fits in relation to pcda course. Could also be second part of a fully Python based course. Builds on basic intro Python concepts and skills.
+
+The Quant Econ website is a good example of another use of Python for typical Excel things.
 

### 2023-06-02 python_excel_ite.tex
describing modeling notebooks

@@
 
+The "Excel with Python" module is comprised of three submodules.
+
 \section{The modeling notebooks}
+
+The first focuses on building a simple what-if model and doing sensitivity analysis much like one would do using Excel's Data-Table functionality.
+The next submodule explores the creation and use of goal seeking capability so that we can do things like finding the break-even point in our model.
+
 \subsection{Notebook 1: Modeling and data tables}

### 2023-06-02 python_excel_ite.tex
module positioning

--- 2023-06-01 python_excel_ite.tex
@@
 
-\section{Module structure}
+\section{Module positioning and structure}
 
-\section{The modeling notebooks}
+What kind of Python background do students need have in order to be ready for this EwP module? This module does presume a basic familiarity with Python fundamentals for data analysis such as:
+
+\begin{itemize}
+	\item variables, artithmetic and boolean operators, and basic data types,
+	\item data structures such as lists, dictionaries, tuples, NumPy arrays and pandas dataframes,
+	\item flow control such as branching with \texttt{if \ldots else} and looping with \texttt{for} and \texttt{while}
+	\item module imports,
+	\item using built in functions and accessing methods and properties of objects,
+	\item creating functions,
+	\item basic use of NumPy and pandas for data wrangling and analysis,
+	\item basic plotting with matplotlib.
+\end{itemize}
+
+The EwP module could easily be used as part of a semester long Python based analytics course.
+At our institution, this module is part of a business course entitled "Advanced Analytics with Python" (\href{http://www.sba.oakland.edu/faculty/isken/courses/mis6900}{AAP}).
+The students taking this course have already taken my "Practical Computing for Business Analytics" (\href{http://www.sba.oakland.edu/faculty/isken/courses/mis5470}{PCDA}) course in which they spend about half the semester learning R and the other half learning Python, both in the context of business analytics.
+Some Linux basics are also part of the PCDA course.
+The students have also already taken "Business Analytics" (\href{http://www.sba.oakland.edu/faculty/isken/courses/mis5460}{BA}) which is a spreadsheet based introduction to business analytics.
+The BA and PCDA courses have historically been offered in face to face mode in a computer teaching lab.
+Since 2020, these two courses have been also offered as online, asynchronous courses.
+The newer AAP course has been offered since 2021 in an online asynchronous mode.
+All three courses have extensive course websites that are publicly accessible, including access to all of the video content and files needed.
+We will describe the AAP course web in more detail below in the context of the EwP module.
+
+The EwP module is made of three submodules. The first of these focuses on building and using typical spreadsheet model using Python
+
+\section{Submodule 1: The modeling notebooks}
+
 \subsection{Notebook 1: Modeling and data tables}
@@
 
-\section{The packaging notebooks}
+\section{Submodule 2: The packaging notebooks}
 \subsection{Notebook 4: Project packaging}
@@
 
-\section{The wrangling notebooks}
+\section{Submodule 3: The wrangling notebook}
 \subsection{Notebook 6: Manipulating Excel files with Python}

### 2023-06-02 python_excel_ite.tex
working on submodule 1

@@
 
-The EwP module is made of three submodules. The first of these focuses on building and using typical spreadsheet model using Python
+The EwP module is made of three submodules.
+The first of these focuses on building and using a typical spreadsheet model using Python.
+Then the functionality developed for doing this type of modeling in Python is deployed as a reusable package.
+We end with an introduction to using Python to manipulate Excel files, an extremely common use case.
+
+Combination of Jupyter notebooks, Python scripts, videos walking through the notebooks and scripts, all accessible from submodule specific course web pages.
 
@@
 
+There are two overarching learning objectives for this first set of Jupyter notebooks.
+Students learn how Python might be used to do something they are quite familiar with doing in a spreadsheet - model building and sensitivity analysis.
+At the same time, students encounter, as they are needed, more advanced Python programming techniques such as object oriented programming, ADD MORE FROM THE MODULES.
+In this way, the computational problem to be solved drives the introduction of more advanced Python concepts and techniques instead of such things being presented in a vacuum.
+I make it a point to try to convey the notion that there are often multiple possible approaches by embarking on one path and then starting over and doing things in a different way.
+MORE NEEDED
+
 \subsection{Notebook 1: Modeling and data tables}
+
+\begin{quote}
+	For example, here's a high level screenshot of a model that I assign for homework in my [MIS 4460/5460 Business Analytics class](http://www.sba.oakland.edu/faculty/isken/courses/mis5460/) (a spreadsheet based modeling class).
+	It's a really simple model in which we are selling a single product that we produce.
+There is a fixed cost to producing the product as well as a variable production cost per unit.
+We can sell the product for some price and we believe that demand for the product is related to the selling price through a power function.
+Let's assume for now that we have sufficient capacity to produce to demand and that all inputs are deterministic (we'll deal with simulating uncertainty later in this document).
+
+	The details aren't so important right now as is the overall structure of the model.
+There's a few key inputs and some pretty straightforward formulas for computing cost, revenue and profit.
+Notice the 1-way Data Table being used to explore how profit varies for different selling prices.
+There's a graph driven by the Data Table and some Text Boxes used for annotation and summary interpretative comments.
+There's a button that launches Goal Seek to find the break even selling price and a 2-way Data Table (not shown) to explore the joint effect of selling price and variable cost.
+Classic Excel modeling stuff.
+How might we go about building a similar model using Python?
+
+	What if we wanted to push it a little further and model some of the key inputs with probability distributions to reflect our uncertainty about their values?
+In the Excel world, we might use add-ins such as @Risk which allow uncertain quantities to be directly modeled with probability distributions.
+For example, we might have a key input such as the exponent in the power function that relates selling price to demand that is highly uncertain.
+By modeling it with a probability distribution and then sampling from that distribution many times (essentially by recalcing the spreadsheet) we can generate a bunch of possible values for key outputs (e.g.
+profit) and use statistics to summarize these outputs using things like histograms and summary stats.
+Often this type of simulation model is referred to as a *Monte-Carlo* model to suggest repeated sampling from one or more probability distributions within an otherwise pretty static model.
+If you want to see such models in action, check out my [Simulation Modeling with Excel page](http://www.sba.oakland.edu/faculty/isken/courses/mis5460/simulation.html) from my Business Analytics course.
+Again, how might we do this with Python?
+\end{quote}
+
+The first focuses on building a simple what-if model and doing sensitivity analysis much like one would do using Excel's Data-Table functionality.
+
 \subsection{Notebook 2: Goal seek}
+
+The next submodule explores the creation and use of goal seeking capability so that we can do things like finding the break-even point in our model.
+
 \subsection{Notebook 3: Monte-Carlo simulation}

### 2023-06-03 python_excel_ite.tex
working on nb 1

@@
 
+\usepackage{hyperref}
+\usepackage{listings}
+\usepackage{xcolor}
+
 % If hyperref is used, dvi-to-ps driver of choice must be declared as
@@
 % Private macros here (check that there is no clash with the style)
+\definecolor{LightGray}{gray}{0.9}
+\definecolor{codegreen}{rgb}{0,0.6,0}
+\definecolor{codegray}{rgb}{0.5,0.5,0.5}
+\definecolor{codepurple}{rgb}{0.58,0,0.82}
+\definecolor{backcolour}{rgb}{0.95,0.95,0.92}
+\lstdefinestyle{mystyle}{
+	backgroundcolor=\color{LightGray},   
+    commentstyle=\color{codegreen},
+	keywordstyle=\color{magenta},
+	numberstyle=\tiny\color{codegreen},
+	stringstyle=\color{codepurple},
+	basicstyle=\footnotesize\ttfamily,
+	breakatwhitespace=false,         
+	breaklines=true,                 
+	captionpos=b,                    
+	keepspaces=true,                 
+	numbers=left,                    
+	numbersep=5pt,                  
+	showspaces=false,                
+	showstringspaces=false,
+	showtabs=false,                  
+	tabsize=2
+}
+
+\lstset{style=mystyle}
 
@@
 
+- more topics I wanted to cover than were possible in PCDA course
+- weave in more SE type material which is often overlooked in BA programs
+
 Excel is widely used for building and using models of business problems to explore the impact of various model inputs on key outputs.
@@
 
-What kind of Python background do students need have in order to be ready for this EwP module? This module does presume a basic familiarity with Python fundamentals for data analysis such as:
+PCDA ends with intro to sklearn and AAP begins with review of sklearn and some more ML content - logistic regression with regularization
+
+What kind of Python background do students need to have in order to be ready for this EwP module? This module does presume a basic familiarity with Python fundamentals for data analysis such as:
 
@@
 
+This example is based on one in the [spreadsheet modeling textbook(s) I've used in my classes since 2001](https://host.kelley.iu.edu/albrightbooks/).
+I started out using Practical Management Science by Winston and Albright and switched to their Business Analytics: Data Analysis and Decision Making (Albright and Winston) around 2013ish.
+In both books, they introduce the "Walton Bookstore" problem in the chapter on Monte-Carlo simulation.
+Here's the basic problem (with a few modifications):
+
+\begin{itemize}
+	\item we have to place an order for a perishable product (e.g. a calendar),
+	\item there's a known unit cost for each one ordered,
+	\item we have a known selling price,
+	\item demand is uncertain but we can model it with some simple probability distribution,
+	\item for each unsold item, we can get a partial refund of our unit cost,
+	\item we need to select the order quantity for our one order for the year; orders can only be in multiples of 25.
+\end{itemize}
+
+The notebook begins with a suggestion to start with a very simple model that ignores the demand uncertainty and that is not object oriented.
+Proceeding much like we would in Excel, we store base input values in variables.
+(unit cost, unit price, selling prce, unit refund and demand)
+
+\begin{lstlisting}[language=Python]
+order_cost = unit_cost * order_quantity
+sales_revenue = ??? * selling_price
+refund_revenue = ??? * unit_refund
+profit = sales_revenue + refund_revenue - order_cost
+\end{lstlisting}
+
 \subsection{Notebook 2: Goal seek}

### 2023-06-06 python_excel_ite.tex
up to oo section

@@
 % Private macros here (check that there is no clash with the style)
+\newcommand{\code}[1]{\texttt{#1}}
+
 \definecolor{LightGray}{gray}{0.9}
@@
 
+Limitations of spreadsheets such as reproducibility and automation.
+
 While Python has been gaining momentum in the business analytics world, it is often used for data wrangling, analysis and visualization of tablular data using tools like pandas and matplotlib or Seaborn.
@@
 What would those look like in Python?
+
+Quant Econ example
+
+PBB blog
+
+Teach relevant Python to business students using familiar and relevant modeling examples.
+Give sense of software development process and design decisions
+Model a learning journey
+Teach students to think and act like a software developer, not just an analyst
 
@@
 Combination of Jupyter notebooks, Python scripts, videos walking through the notebooks and scripts, all accessible from submodule specific course web pages.
+Some background on Jupyter notebooks for analytics.
+Questions and coding challenges sprinkled throughout the notebooks - answers at bottom.
 
@@
 
-This example is based on one in the [spreadsheet modeling textbook(s) I've used in my classes since 2001](https://host.kelley.iu.edu/albrightbooks/).
-I started out using Practical Management Science by Winston and Albright and switched to their Business Analytics: Data Analysis and Decision Making (Albright and Winston) around 2013ish.
+This example is based on one in the [spreadsheet modeling textbooks I've used in my classes since 2001](https://host.kelley.iu.edu/albrightbooks/).
 In both books, they introduce the "Walton Bookstore" problem in the chapter on Monte-Carlo simulation.
-Here's the basic problem (with a few modifications):
+Here is the basic problem (with a few modifications):
 
@@
 	\item we have to place an order for a perishable product (e.g. a calendar),
-	\item there's a known unit cost for each one ordered,
+	\item there is a known unit cost for each one ordered,
 	\item we have a known selling price,
@@
 
-The notebook begins with a suggestion to start with a very simple model that ignores the demand uncertainty and that is not object oriented.
+\subsubsection{Build first model - procedural approach}
+
+The notebook begins with a suggestion to start with a very simple model that ignores the uncertainty in demand and that is not object oriented.
 Proceeding much like we would in Excel, we store base input values in variables.
-(unit cost, unit price, selling prce, unit refund and demand)
+
+\begin{lstlisting}[language=Python]
+# Base inputs
+unit_cost = 7.50
+selling_price = 10.00
+unit_refund = 2.50
+
+# Demand parameters
+demand_mean = 193
+demand_sd = 40
+
+# Deterministic model
+demand = demand_mean
+
+# Initial value for order quantity - this is the decision variable
+order_quantity = 200
+\end{lstlisting}
+
+Then students are given some skeleton code and must finish the expressions for computing key intermediate outputs (\code{order_cost, sales_revenue, refund_revenue} and the final output, \code{profit}.
+Answers are provided at the bottom of the notebook so that students can attempt to finish the code but have a resource to correctly complete the code and move on.
+The screencast associated with the notebook also shows me completing the code.
+Using skeleton code in this way has worked well both for face to face lab based versions of this type of course as well as for online asynchronous versions.
 
@@
 \end{lstlisting}
+
+At this point we discuss a fundamental difference between the computing model of Excel and that of Jupyter notebooks.
+While spreadsheets respond automatically to changes, Jupyter notebooks require rerunning all code cells that include or follow the changed code.
+
+\subsubsection{Sensitivity analysis}
+
+Since order quantity is the key decision variable, we might want to see how profit changes for different order quantities in this first simplified model.
+In Excel, the Data Table tool provides an easy way to do this and is a staple of spreadsheet based management science textbooks.
+In a Data Table, a range of order quantities can be used as row or column input and one or more output values can be computed.
+In Python, we can rely on the built in vectorized behavior of the language.
+A range of order quantities is created as a NumPy array and then can be used directly to compute vectors of all of the intermediate and final output variables.
+Much like Data Tables, vectorized computations avoid having to explicitly iterate (loop) through a collection of input values.
+
+A natural next step is to do the equivalent of an Excel 2-way Data Table.
+Can't do 2 vectors since we need all combinations of entries in each vector.
+List comprehensions provide a compact way to create nested loops over the two vectors.
+By creating a profit function that takes all of the base inputs as arguments, we can construct a list comprehension that does what an Excel 2-Way Data Table does.
+Even better, this actually allows us to do the equivalent of an n-Way Data Table.
+
+\begin{lstlisting}[language=Python]
+	def bookstore_profit(unit_cost, selling_price, unit_refund, order_quantity, demand):
+	'''
+	Compute profit in bookstore model
+	'''
+		order_cost = unit_cost * order_quantity
+		sales_revenue = np.minimum(order_quantity, demand) * selling_price
+		refund_revenue = np.maximum(0, order_quantity - demand)
+		profit = sales_revenue + refund_revenue - order_cost
+		return profit
+
+	# Set up input vector ranges	
+	demand_range = np.arange(50, 301, 5)
+	order_quantity_range = np.arange(50, 301, 25)
+
+	# Create data table (as a list of tuples)
+	data_table_1 = [(d, oq, bookstore_profit(unit_cost, selling_price, unit_refund, oq, d)) 
+	for d in demand_range for oq in order_quantity_range]
+
+	# Convert to dataframe
+	dtbl_1_df = pd.DataFrame(data_table_1, columns=['Demand', 'OrderQuantity', 'Profit'])
+
+\end{lstlisting}
+
+TODO - add output listing and plot
+
+\subsubsection{The object nature of Python}
+
+Introducing basic OO concepts is a primary learning objective.
+It's not clear whether an OO or a non-OO approach will end up making the most sense for doing spreadsheet type modeling.
+This uncertainty is made explicit in the notebook and students are led on a bit of a journey of discovery as we explore different software designs for our modeling problem.
+I purposefully designed the notebooks in this way to show students that the path to a solution is not necessarily linear and that giving yourself the freedom to explore alternatives, many of which will be dead ends, can lead to a rich learning experience.
+It is also more realistic in the sense that this is how most real problems are solved.
+It's analogous to the idea that mathematical proofs usually don't reflect the myriad of failed approaches that were tried but that ultimately contributed to the final product.
+
+In preparation for building an object oriented version of the model, basic object concepts are reviewed.
+Since most students have done some Excel VBA programming in a previous course, I refer back to a few properties and methods of the Excel \code{Worksheet} object.
+In Python, everything is an object.
+Using Python lists as an example, we explore some of its attributes such as methods for appending items to a list and reversing the order of its elements.
+The \code{dir} function is used to see all of an object's attributes and we see that even things like integers are objects in Python.
+With this basic object refresher, we are ready to do some actual object oriented programming and create our own objects - well, actually classes.
+
+For our initial design of a \code{BookstoreModel} class we make all of the base inputs attributes (properties) and then add method attributes to compute outputs such as costs, revenues and profits.
+Student can easily see the mapping between things in the initial procedural model and this new object oriented version of the model.
+I highlight several important concepts and syntactical details of doing OO programming Python.
+Again, the goal is to weave in these more advanced Python programming concepts within the context of a very familiar modeling problem.
+This section of the notebook ends with us using our \code{BookstoreModel} class to an actual model objects with all of its base inputs instantiated with values.
+Before moving on students are presented with a few challenges involving enhancements to the \code{BookstoreModel} class.
 
@@
 
-\section{Student assignment}
+\section{Classroom experience}
 

### 2023-06-07 python_excel_ite.tex
finished oo notebook

@@
 % Enter the (shortened) title:
-\RUNTITLE{Python for spreadsheet type modeling}
+\RUNTITLE{Python for spreadsheet modeling}
 
@@
 % Enter the full title:
-\TITLE{Using Python for spreadsheet type modeling}
+\TITLE{Python for spreadsheet modeling}
 
@@
 Questions and coding challenges sprinkled throughout the notebooks - answers at bottom.
+While AAP is online asynch, this same approach has been used very successfully in a computer teaching lab for the PCDA course.
+The screencasts are highly valued by the students as they can rewatch things as needed.
+Many are working professionals and a 2.5 hour evening class in a warm computer lab can be taxing.
 
-\section{Submodule 1: The modeling notebooks}
+\section{Submodule 1: What-if analysis with Python}
 
@@
 Students learn how Python might be used to do something they are quite familiar with doing in a spreadsheet - model building and sensitivity analysis.
-At the same time, students encounter, as they are needed, more advanced Python programming techniques such as object oriented programming, ADD MORE FROM THE MODULES.
+At the same time, students encounter, as they are needed, more advanced Python programming techniques such as object oriented programming and manipulation of tuples, lists and dictionaries.
 In this way, the computational problem to be solved drives the introduction of more advanced Python concepts and techniques instead of such things being presented in a vacuum.
 I make it a point to try to convey the notion that there are often multiple possible approaches by embarking on one path and then starting over and doing things in a different way.
-MORE NEEDED
+Each notebook is accompanied by a few to several videos that walk the students through the material - a guided tour with challenges to overcome along the way.
 
@@
 
-For our initial design of a \code{BookstoreModel} class we make all of the base inputs attributes (properties) and then add method attributes to compute outputs such as costs, revenues and profits.
-Student can easily see the mapping between things in the initial procedural model and this new object oriented version of the model.
+\subsubsection{Creating an object oriented model}
+
+For our initial design of a \code{BookstoreModel} class we make all of the base inputs class attributes (properties) and then add method attributes to compute outputs such as costs, revenues and profits.
+Students can easily see the mapping between things in the initial procedural model and this new object oriented version of the model.
 I highlight several important concepts and syntactical details of doing OO programming Python.
 Again, the goal is to weave in these more advanced Python programming concepts within the context of a very familiar modeling problem.
-This section of the notebook ends with us using our \code{BookstoreModel} class to an actual model objects with all of its base inputs instantiated with values.
+This section of the notebook ends with us using our \code{BookstoreModel} class to create a model object with all of its base inputs instantiated with values.
 Before moving on students are presented with a few challenges involving enhancements to the \code{BookstoreModel} class.
+Interspersing short coding challenges throughout the notebook gives students a chance to test themselves on their understanding and to get some coding practice.
+
+Then a few design dilemmas are posed and some non-working approaches illustrated as we try to decide on how to implement an n-way data table function for this new object oriented model.
+These meta reflections are intentionally included to better mimic actual software development and reinforce the idea that software development is much more than simply writing code to implement a perfectly thought out design.
+We forge ahead with some preliminary ideas and are confronted with an intermediate problem related to generating a list of dictionaries that represent combinations of inputs (i.e.
+scenarios) to evaluate by our n-way data table function.
+This provides a great opportunity to remind students of one of the great strengths of the open source software ecosystem - we can leverage work done in other packages and even look at actual source code to see how something was implemented.
+In this case, the well known scikit-learn package has a \code{ParameterGrid} function that solves our scenario generation problem quite nicely.
+This example also reminds the students of the value of being able to make sense of API documentation and of object oriented code written by others.
+Most data science related Python packages are object oriented in nature.
+
+Whew!
+This first notebook draws to a close with the creation of an n-way data table function that allows us to do sensitivity analysis with our objected oriented model.
+The payoff is the creation of the following faceted plot showing how profit varies for different order quantities and demand levels.
+We also recap the main things we learned in this notebook and reiterate that we will now move on to adding goal seeking capability to our model.
+
+INSERT PLOT HERE
 

### 2023-06-08 python_excel_ite.tex
finished goal seek and simulation notebooks

@@
 % The opening part of the what-if notebook has some language for the motivation behind this module.
+Spreadsheet based modeling wave of the early 2000s. Business schools with business analytics programs are the context. Grad and undergrad.
 
@@
 Each notebook is accompanied by a few to several videos that walk the students through the material - a guided tour with challenges to overcome along the way.
+The screenshot in \ref{fig:activities} shows the portion of the course web page that provides the high level roadmap through the notebooks and associated screencasts.
 
@@
 Whew!
-This first notebook draws to a close with the creation of an n-way data table function that allows us to do sensitivity analysis with our objected oriented model.
+This first notebook draws to a close with the creation of an \code{data_table} function that allows us to do sensitivity analysis with our objected oriented model.
+Unlike Excel's Data Table tool, our \code{data_table} function can handle an arbitrary number of inputs and outputs.
 The payoff is the creation of the following faceted plot showing how profit varies for different order quantities and demand levels.
@@
 
-The next submodule explores the creation and use of goal seeking capability so that we can do things like finding the break-even point in our model.
+No self respecting modeling tool would be complete without goal seeking capability to do things such as finding the break even point in the bookstore model.
+For many students, Excel's Goal Seek tool can seem almost magical when encountered for the first time.
+Even after they get a sense of what it is doing, understanding the importance of the initial guess and how it can lead to Goal Seek reporting different solutions, is not very transparent.
+With Python, students can gain a much better understanding of how tools like Goal Seek really work.
+
+Before trying to build our \code{goal_seek} function we take a brief detour into basic root finding algorithms through packages like SciPy as well as in standalone Python scripts that are widely available in the Python data science ecosystem (refs).
+We compare the output of Excel's Goal Seek with different starting values to the values obtained by different root finding algorithms implemented in Python.
+The importance of starting values and also the range of different root finding algorithms available is often surprising to business students who have spent most of their life in Excel.
+
+A \code{} function is created that implements a simple bisection search and we use it to find the break even demand point in our bookstore model.
+With \code{data_table} and \code{goal_seek} functions implemented, we are ready to move on to doing Monte-Carlo simulation.
 
 \subsection{Notebook 3: Monte-Carlo simulation}
+
+In the Excel modeling world, Monte-Carlo simulation can be done without add-ins, but packages like @Risk \cite{} can make the process much easier.
+In particular, random variable generation for a wide range of probability distributions is facilitated by @Risk.
+In Python, the NumPy package provide random variate generation for a large number of distributions (https://numpy.org/doc/stable/reference/random/legacy.html#distributions) and similarly, SciPy provides functions for computing distribution related quantities for many probability distributions \cite{}.
+Much like we did when creating the \code{data_table} function, we start with adding uncertainty to a single input variable in the bookstore model and rely on Python's vectorized computing capabilities to generate the profit associated with each realization of the random input.
+Then we can do standard statistical analysis of the simulation output using Python packages such as Pandas and SciPy.
+In spreadsheet based simulation textbooks, the notion of the ``Flaw of Averages'' popularized by Sam Savage \cite{} is illustrated.
+We do the same with our Python based model, showing how replacing random demand by its mean can result in a wildly optimistic estimate of the mean profit found through analyzing the simulation output.
+Much like doing simulation in Excel without add-ins, using Python makes transparent the quite simple and brute force nature of Monte-Carlo simulation.
+
+Moving on to multiple uncertain inputs raises challenges much like those we faced when moving from a 1-way to an n-way data table function.
+In addition to modeling multiple uncertain inputs, we want to be able to specify a range of scenarios to run, much like we did for the \code{data_table} function.
+The following set of design criteria are presented for a \code{simulate} function.
+
+\begin{itemize}
+	\tightlist
+	\item
+	The first argument will be a model object (i.e.~something like the
+	\texttt{BookstoreModel} model) that contains an \texttt{update}
+	method. Soon, we should add an abstract \texttt{Model} class from
+	which specific models such as the \texttt{BookstoreModel} class can be
+	created. The abstract class will contain the \texttt{update} method.
+	\item
+	The random inputs will be passed in as a dictionary whose keys are the
+	input variables being modeled as random and whose values are an
+	iterable representing the draws from some probability distribution.
+	Structurally, this is similar to how inputs are specified in the
+	\texttt{data\_table} function.
+	\item
+	We can optionally pass in a dictionary of scenario inputs. This is
+	exactly like the \texttt{data\_table} variable input.
+
+	\begin{itemize}
+		\tightlist
+		\item
+		If no scenario input dictionary is passed in, a single simulation
+		scenario is run using the current input values in the model object,
+		\item
+		If a scenario input dictionary is passed in, then a simulation
+		scenario is run for every combination of parameters in the
+		dictionary. Again, this is just like we do in the
+		\texttt{data\_table} function.
+	\end{itemize}
+	\item
+	The output will be a set of dictionaries containing dataframes of
+	simulation output as well standard summary stats and plots.
+\end{itemize}
+
+This is followed by a well commented first attempt at a \code{simulate} function.
+By this point, students are well equipped to understand the code as it builds on basic Python concepts and more advanced things that we have already covered in this EwP module.
+The rest of the notebook is spent trying out the \code{simulate} function and ends with creating a plot that many of them have likely created before with @Risk along with a faceted plot of histograms of profit created with the popular Seaborn package.
+
+INSERT GROUPED BOX PLOT HERE and FACETED HISTOGRAMS
 

### 2023-06-08 python_excel_ite.tex
starting packaging

@@
 \section{Submodule 2: The packaging notebooks}
+
+The first submodule and its three notebooks are very much about the type of modeling typically done in a spreadsheet based course.
+In a spreadsheet base course, students learn numerous Excel functions and techniques in the process of building and exercising such models.
+Similarly, numerous Python techniques are scattered throughout the three modeling focused notebooks.
+Now we shift gears and move beyond what might be considered prototyping and into the realm of deployment.
+
+\begin{quote}
+	content...
+
+In the first three notebooks, we've developed some Python approaches to
+typical Excel ``what if?'' analyses. Along the way we explored some
+slightly more advanced Python topics (for relative newcomers to Python)
+such as:
+
+\begin{itemize}
+	\tightlist
+	\item
+	List comprehensions,
+	\item
+	Basic OO programming, creating our own classes and using things like
+	\texttt{setattr} and \texttt{getattr},
+	\item
+	Leveraging scikit-learn's \texttt{ParameterGrid} class,
+	\item
+	Faceted plots using Seaborn and matplotlib,
+	\item
+	Tuple unpacking, zip, and itertools,
+	\item
+	Safely copying objects,
+	\item
+	Root finding with, and without, \texttt{scipy.optimize},
+	\item
+	Using \texttt{numpy.random} to generate random variates from various
+	probability distributions,
+	\item
+	Using \texttt{scipy.stats} to compute probabilities and percentiles,
+	\item
+	Importing our \texttt{data\_table}, \texttt{goal\_seek} and
+	\texttt{simulate} functions from a module.
+\end{itemize}
+
+Now that we've got a critical mass of "proof of concept" code, let's figure out how to structure our project and create a deployable package.
+In addition, let's rethink our OO design and add some much needed documentation to the code.
+\end{quote}
+
+Creating a deployable Python package is
+
 \subsection{Notebook 4: Project packaging}

### 2023-06-08 python_excel_ite.tex
rough draft of the three submodules

@@
 
-Combination of Jupyter notebooks, Python scripts, videos walking through the notebooks and scripts, all accessible from submodule specific course web pages.
+Combination of Jupyter notebooks, Python scripts, videos walking through the notebooks and scripts, all publicly accessible from submodule specific course web pages.
+GIVE URL.
 Some background on Jupyter notebooks for analytics.
@@
 
-\section{Submodule 2: The packaging notebooks}
-
 The first submodule and its three notebooks are very much about the type of modeling typically done in a spreadsheet based course.
-In a spreadsheet base course, students learn numerous Excel functions and techniques in the process of building and exercising such models.
+In a spreadsheet based course, students learn numerous Excel functions and techniques in the process of building and exercising such models.
 Similarly, numerous Python techniques are scattered throughout the three modeling focused notebooks.
 Now we shift gears and move beyond what might be considered prototyping and into the realm of deployment.
+
+\section{Submodule 2: The packaging notebooks}
 
@@
 
-Creating a deployable Python package is
+Creating a deployable Python package is a good exercise in designing and creating reusable software artifacts.
+Thinking about how others will interact with your software usually leads to better designed software.
+Basic software engineering practices are often overlooked in business analytics programs but are certainly desirable in the professional world.
+In this submodule we create a Python package that can be imported and used by other modelers.
+This will require some design changes to our software and will also involve moving code out of Jupyter notebooks and into Python script files.
+While Jupyter notebooks can be integrated into production environments \cite{netflix}, it is much more common for code to live in a Python script file.
+In addition to creating a deployable package, this submodule also discusses the importance of documentation in its many forms in a Python project.
 
 \subsection{Notebook 4: Project packaging}
+
+The first notebook in this submodule starts by explaining what Python package are and their role in sharing code.
+We then revisit a concept from earlier in the course - creating a good project folder structure.
+For this we use what is known as a \textit{cookiecutter} that can automatically generate a project folder structure and key files from the answers to a few prompts.
+Their are a few different folder layouts used in Python projects and there is a bit of discussion on the folder stucture we will use and resources given for exploring this issue further.
+Students are reminded of the importance of version control and we initialize a git repository for our budding package.
+The basics of version control were already covered in the first week of the course.
+My experience is that most students have little exposure to simple software project management concepts such as project folder structures and version control.
+This can lead to a jumble of files with incoherent names used as a proxy for version control - whatif_v1.py, whatif_v2.py, whatif_final.py, whatif_finalfinal.py, ...
+
+Key software design changes are then made and discussed which will make our code easier on our users.
+These changes include creating an abstract \code{Model} base class from which different model classes can be created and moving the \code{data_table}, \code{goal_seek}, and \code{simulate} functions into the \code{Model} base class.
+All of the code is moved into a code module named \code{whatif.py}.
+We are now ready to learn about the basics of turning our code into an installable package.
+This requires a basic understanding of two key Python files that are ubiquitous in Python pacakges - \code{__init__.py} and \code{setup.py}.
+Next comes a discussion of options for installing our package locally so that we can use it for different modeling projects.
+This includes an explanation of how and where Python searches for packages when \code{import} statements are encountered in code.
+We defer discussion of uploading our package to a public code sharing platform such as PyPI.
+All of this can be quite complex and I have made every effort to distill things down to a minimal level of complexity needed to convey the important points.
+I stress that there is much complexity lurking here and it is something they will need to revisit time and time again.
+
+This notebook ends by pointing the students to another notebook in which our newly deployed whatif package is used for a completely different model.
+The example used is the New Car Simulation that has been a staple of Winston and Albright's textbooks for many years \cite{bibid}.
+
 \subsection{Notebook 5: Documentation}
 
-\section{Submodule 3: The wrangling notebook}
-\subsection{Notebook 6: Manipulating Excel files with Python}
+This deployment focused submodule ends with a short notebook discussing the different types of documentation needed in a typical project.
+This includes things like code comments, docstrings, readme files and generating documentation by writing restrutured text files and using Sphinx.
+This is one of those tasks that few people like to do but can be very important for the long term success of a project.
+Even if you are just creating a tool for yourself, you will forget the details over time and documentation can be very beneficial in refreshing your memory.
+
+\section{Submodule 3: Excel data wrangling with Python}
+
+The EwP module ends with what has become a relatively common use of Python with Excel - automating various Excel data wrangling tasks. Examples include:
+
+\begin{itemize}
+	\tightlist
+	\item
+	You have a whole folder full of csv (or Excel) files with the same
+	file structure and you need to combine them into a single file. You
+	might also need to make some changes to the consolidated file.
+	\item
+	You have an Excel file with multiple sheets of similarly structured
+	data and you want to consolidate them into a single sheet.
+	\item
+	You have an Excel file with data in ``wide'' format and you need to
+	convert it to long format, and then perhaps export out individual
+	files (one per the key column(s) in the long formatted data).
+	\item
+	You have an Excel file acting as a simple flatfile database.
+	Periodically, you get new Excel files that need to get appended to the
+	``database'' file.
+\end{itemize}
+
+A single Jupyter notebook is used to give a taste of using Python to automate the process of working with Excel files by tackling each of the four examples above.
+Each of these examples are based on real problems I encountered either in research or industrial projects.
+TRANSITION NEEDED Several web based resources are shared with one particularly good one being the Practical Business Python blog written by Chris Moffitt.
+It includes numerous posts on different aspects of Python and Excel integration.
+There are a number of Python based tools that either include functionality for working with Excel files or are dedicated to specific Excel related operations.
+These include:
+
+\begin{itemize}
+	\tightlist
+	\item
+	pandas -
+	\href{https://pandas.pydata.org/docs/reference/api/pandas.read_excel.html}{read\_excel},
+	\href{https://pandas.pydata.org/docs/reference/api/pandas.ExcelWriter.html}{ExcelWriter},
+	\href{https://pandas.pydata.org/docs/reference/api/pandas.DataFrame.to_excel.html}{to\_excel}
+	\item
+	\href{https://openpyxl.readthedocs.io/en/stable/}{openpyxl} - can
+	read, write, and modify Excel files
+	\item
+	\href{https://xlsxwriter.readthedocs.io/}{XlsxWriter} - powerful but
+	can't edit existing Excel files
+	https://xlsxwriter.readthedocs.io/working\_with\_tables.html
+	\item
+	\href{https://www.xlwings.org/}{xlwings} - provides ``glue'' between
+	Python and Excel
+\end{itemize}
+
+\subsection{Example 1: Concatenating many csv files}
+
+While this first example is not necessarily Excel related, using Excel to consolidate csv files through a series of copy-paste operations is something that many Excel users will admit to doing.
+We also use this example to introduce the newish pathlib Python library for working with file paths.
+After getting familiar with pathlib, students are led through the development of a short procedure that creates a list of the csv files to concatenate and then uses the pandas library to read and combine the files in a \code{DataFrame} and then writes out the consolidated data as a csv file.
+
+\subsection{Example 2: Consolidating data from multiple sheets in a Excel file using pandas}
+
+Two different approaches are shown to consolidate data from multiple sheets into a single sheet.
+One approach uses the pandas library while the other relies on a library specifically created for working with Excel files from Python, openpyxl.
+One advantage to using Python rather than Visual Basic for Applications for this task is that Python is well suited for working with data in many different formats.
+The Excel file may be just one data source out of a collection of data sources that are used to create some sort of combined database.
+The generality of Python is a strength in such scenarios.
+
+\subsection{Example 3: Dealing with wide data and a multi-row header}
+
+Reshaping data from wide to long is a common data wrangling task especially in preparation for plotting.
+In this example, not only do students have to grapple with reshaping the data in an Excel workbook, they also have to deal with the all too common problem of multi-row header lines in a spreadsheet.
+Excel's flexibility makes it easy to create such multi-row headers and pity the poor analyst who then has to deal with it when trying to move the data into some sort of data frame or database table.
+This example is posed as a challenge to the students.
+Examples of the desired outputs are shown and a few hints given.
+The answer is provided at the end of the notebook.
+This example really shows the power of Python in creating a reproducible and automated approach to dealing with a poorly structured Excel workbook, creating not only restructured data frames but also non-trivial plots.
+Attempting to do all of this in Excel, even with VBA, is not an easy task.
+
+\subsection{Example 4: Appending new spreadsheet data in consolidated Excel workbook}
+
+Excel is not a database.
+But, that does not stop people from using it as a simple flat-file database.
+In such a scenario, it is not uncommon to have to periodically append new data to the end of a range of data in an Excel workbook.
+Again, a combination of openpyxl and pandas allows us to meet this challenge.
+
+\subsection{More on Python/Excel integration}
+
+The examples above barely scratch the surface of the use cases for Python integration with Excel.
+Our notebook ends with a shout out to the highly regarding Practical Business with Python blog and a list of specific examples that students can explore on their own.
+
+\begin{itemize}
+	\tightlist
+	\item
+	\href{https://pbpython.com/excel-file-combine.html}{Combining Data
+		From Multiple Excel Files} - file globbing, concatenating dataframes,
+	\item
+	\href{https://pbpython.com/pandas-excel-range.html}{Reading Poorly
+		Structured Excel Files with Pandas} - advanced use of
+	\texttt{read\_excel}, accessing ranges and Tables
+	\item
+	\href{https://pbpython.com/excel-pandas-comp.html}{Common Excel Tasks
+		Demonstrated in Pandas} - totals rows, fuzzy string matching
+	\item
+	\href{https://pbpython.com/excel-pandas-comp-2.html}{Common Excel
+		Tasks Demonstrated in Pandas - Part 2} - selection and filtering
+	\item
+	\href{https://pbpython.com/improve-pandas-excel-output.html}{Improving
+		pandas Excel output} - using XlsxWriter to format Excel workbooks from
+	Python
+	\item
+	\href{https://pbpython.com/advanced-excel-workbooks.html}{Creating
+		Advanced Excel Workbooks} - XlsxWriter, inserting VBA from Python(!),
+	using COM to merge sheets
+	\item
+	\href{https://pbpython.com/xlwings-pandas-excel.html}{Interactive Data
+		Analysis with Python and Excel} - using xlwings to ``glue'' Python and
+	Excel together, using sqlalchemy to interact with databases
+\end{itemize}
 
@@
 The Quant Econ website is a good example of another use of Python for typical Excel things.
+
+Everything discussed in this article is available from the publicly accessible course website.
 

### 2023-06-08 python_excel_ite.tex
very rough first draft

@@
 
-\section{Classroom experience}
+\section{Classroom experience and the future}
+
+This module has been included in my Advanced Analytics with Python course for the past three years.
+The course is offered in a seven-week accelerated format during May and June.
+It is an online and asynchronous course.
+The EwP module is covered over the course of two weeks and is the subject of one of the major homework assignments for the semester.
+The first part of the assignment involves creating a basic model, similar in complexity to the Bookstore Model, and doing sensitivity analysis, goal seeking and simulation with the model.
+It requires the students to import and use the whatif.py package we created as part of the EwP module.
+The second part of the assignment is based on the last submodule and requires the students to complete a task involving creation of a multi-sheet Excel workbook from a set of csv files using Python.
+Then, using the openpyxl library, they have to create and add formulas to the workbook and format them properly.
+The ranges used in the formulas vary by sheet and require the students to determine the number of rows in each sheet of data and use that information to construct the appropriate formula.
+An example assignment is available from \href{}{here}.
+The module has been well received and several of the working students have reported back to me that they have found use cases for these techniques in their day to day analytical activities.
+
+Rumours have swirled for a number of years that Microsoft was considering adding Python as a first class language alternative to VBA with their MS Office suite of packages.
+While this has not yet materialized, there continues to be significant interest in using Python to manipulate not only Excel workbooks but other office documents such as Powerpoint presentations.
+Given that business students are very familiar with both Excel and Powerpoint, inclusion of this material in a business school Python based analytics course seems quite appropriate.
+Students can learn more advanced and extremely useful Python concepts and techniques within a context with which they are intimately familiar.
+It is directly relevant to the job content of many of the students.
+This teaching module is a natural follow up to the commonly taught spreadsheet based analytics courses found in many business schools and provides an interesting way to teach and learn Python for analytics.
 

### 2023-06-09 python_excel_ite.tex
very rough first draft
working on intro

@@
 % Enter the (shortened) title:
-\RUNTITLE{Python for spreadsheet modeling}
+\RUNTITLE{Spreadsheet modeling with Python}
 
@@
 % Enter the full title:
-\TITLE{Python for spreadsheet modeling}
+\TITLE{Spreadsheet modeling with Python}
 
@@
 % The opening part of the what-if notebook has some language for the motivation behind this module.
-Spreadsheet based modeling wave of the early 2000s. Business schools with business analytics programs are the context. Grad and undergrad.
+% TODO - opening should really grab reader
+Spreadsheet based modeling wave of the early 2000s. TMS conferences. Ground breaking textbooks such as Ragsdale and W\&A.
+
+Advent of programmatic analytics using tools like R and Python have changed the teaching landscape.
+Motivated by reproducibility, open source, free, and ease of integration with web and cloud databases.
+
+Created PCDA course in response.
+Challenges with finding good textbooks.
+Developed much of my own curricular materials.
+Taught using combo of R Markdown documents and Jupyter notebooks.
+Literate programming.
+PCDA course focused mostly on data wrangling, EDA, statistical and machine learning based predictive models though I did include a Python based simulation of the Monte Hall 3-door problem.
+In teaching EDA using tools like ggplot2, I would challenge students to think about how they would go about trying to do some of the plots we were making in Excel.
+There is also some limited coverage of using Python to open Excel files and do some sheet manipulation.
+However, there was no real attempt to do the kind of modeling that was the focus of textbooks like PMS or DADM.
+
+% TODO - how should pandas appear in text
+Python has continued to make inroads to analytical territory that was formerly dominated by Excel.
+The widely used Python data analysis library, pandas, was created by Wes McKinney while working as a quantitative financial analyst.
+However, Python was mainly used for data wrangling, visualization and predictive modeling using various flavors of regression and tree based models.
+The blog \href{https://pbpython.com/}{Practical Business Python} written primarily by Chris Moffitt contributed greatly to making Python more accessible to business analytics practitioners through high quality posts that were extremely practical and useful.
+For some time I had been thinking about exploring the general problem of doing spreadsheet type modeling with Python.
+
+Business schools with business analytics programs are the context. Grad and undergrad.
 
@@
 
-Excel is widely used for building and using models of business problems to explore the impact of various model inputs on key outputs.
-Built in "what if?" tools such as Excel [Data Tables](https://support.microsoft.com/en-us/office/calculate-multiple-results-by-using-a-data-table-e95e2487-6ca6-4413-ad12-77542a5ea50b) and [Goal Seek](https://support.microsoft.com/en-us/office/use-goal-seek-to-find-the-result-you-want-by-adjusting-an-input-value-320cb99e-f4a4-417f-b1c3-4f369d6e66c7) are well known to power spreadsheet modelers.
-How might we do similar modeling and analysis using Python?
+Excel is widely used for building and using models of business problems
+to explore the impact of various model inputs on key outputs. Built in
+``what if analysis'' tools such as \href{https://support.microsoft.com/en-us/office/calculate-multiple-results-by-using-a-data-table-e95e2487-6ca6-4413-ad12-77542a5ea50b}{Data
+	Tables} and
+\href{https://support.microsoft.com/en-us/office/use-goal-seek-to-find-the-result-you-want-by-adjusting-an-input-value-320cb99e-f4a4-417f-b1c3-4f369d6e66c7}{Goal
+	Seek} are well known to power spreadsheet modelers. How might we do
+similar modeling and analysis using Python?
 
@@
 
-TODO - add output listing and plot
+% TODO - add output listing and plot
 
@@
 
+% todo Check the hyperlinks above
+
 \section{Classroom experience and the future}

### 2023-06-12 python_excel_ite.tex
working on intro
working on intro sections

@@
 Python has continued to make inroads to analytical territory that was formerly dominated by Excel.
-The widely used Python data analysis library, pandas, was created by Wes McKinney while working as a quantitative financial analyst.
+The widely used Python data analysis library, Pandas, was created by Wes McKinney while working as a quantitative financial analyst.
 However, Python was mainly used for data wrangling, visualization and predictive modeling using various flavors of regression and tree based models.
 The blog \href{https://pbpython.com/}{Practical Business Python} written primarily by Chris Moffitt contributed greatly to making Python more accessible to business analytics practitioners through high quality posts that were extremely practical and useful.
-For some time I had been thinking about exploring the general problem of doing spreadsheet type modeling with Python.
+But, it got me to thinking.
+What about those things for which Excel is well suited such as building formula based models and doing sensitivity analysis on these models?
+What would those look like in Python?
 
@@
 
+Created the ``Excel with Python'' (EwP) module as part
+
 \section{Module positioning and structure}
 
-PCDA ends with intro to sklearn and AAP begins with review of sklearn and some more ML content - logistic regression with regularization
+At our institution, this module is part of a business course entitled ``Advanced Analytics with Python'' (\href{http://www.sba.oakland.edu/faculty/isken/courses/mis6900}{AAP}).
+The students taking this course have already taken my ``Practical Computing for Business Analytics'' (\href{http://www.sba.oakland.edu/faculty/isken/courses/mis5470}{PCDA}) course in which they spend about half the semester learning R and the other half learning Python, both in the context of business analytics.
+Basic Linux skills are also part of the PCDA course.
+The PCDA course ends with intro to the sklearn library for doing predictive modeling in Python and the AAP course begins where the PCDA course ends.
+The EwP module could certainly be used as part of a semester long Python based analytics course.
+The students have also already taken "Business Analytics" (\href{http://www.sba.oakland.edu/faculty/isken/courses/mis5460}{BA}) which is a spreadsheet based introduction to business analytics.
+The BA and PCDA courses have historically been offered in face to face mode in a computer teaching lab.
+Since 2020, these two courses have been also offered as online, asynchronous courses.
+The newer AAP course has been offered since 2021 in an online asynchronous mode.
+All three courses have extensive course websites that are publicly accessible, including access to all of the video content and files needed.
+We will describe the AAP course web in more detail below in the context of the EwP module.
 
-What kind of Python background do students need to have in order to be ready for this EwP module? This module does presume a basic familiarity with Python fundamentals for data analysis such as:
+What kind of Python background do students need to have in order to be ready for the EwP module? This module presumes a basic familiarity with Python fundamentals for data analysis such as:
 
@@
 
-The EwP module could easily be used as part of a semester long Python based analytics course.
-At our institution, this module is part of a business course entitled "Advanced Analytics with Python" (\href{http://www.sba.oakland.edu/faculty/isken/courses/mis6900}{AAP}).
-The students taking this course have already taken my "Practical Computing for Business Analytics" (\href{http://www.sba.oakland.edu/faculty/isken/courses/mis5470}{PCDA}) course in which they spend about half the semester learning R and the other half learning Python, both in the context of business analytics.
-Some Linux basics are also part of the PCDA course.
-The students have also already taken "Business Analytics" (\href{http://www.sba.oakland.edu/faculty/isken/courses/mis5460}{BA}) which is a spreadsheet based introduction to business analytics.
-The BA and PCDA courses have historically been offered in face to face mode in a computer teaching lab.
-Since 2020, these two courses have been also offered as online, asynchronous courses.
-The newer AAP course has been offered since 2021 in an online asynchronous mode.
-All three courses have extensive course websites that are publicly accessible, including access to all of the video content and files needed.
-We will describe the AAP course web in more detail below in the context of the EwP module.
-
 The EwP module is made of three submodules.
@@
 
-Combination of Jupyter notebooks, Python scripts, videos walking through the notebooks and scripts, all publicly accessible from submodule specific course web pages.
+All of the Python modules in the PCDA course as well as the entire AAP course are taught using a combination of Jupyter notebooks, Python scripts, and videos walking through the notebooks, all publicly accessible from extensive course web pages.
 GIVE URL.
@@
 We then revisit a concept from earlier in the course - creating a good project folder structure.
-For this we use what is known as a \textit{cookiecutter} that can automatically generate a project folder structure and key files from the answers to a few prompts.
+For this we use what is known as a \textit{cookiecutter} that can automatically generate a project folder structure and key files from the answers to a few prompts \cite{cookiecutter}.
 Their are a few different folder layouts used in Python projects and there is a bit of discussion on the folder stucture we will use and resources given for exploring this issue further.
@@
 While this first example is not necessarily Excel related, using Excel to consolidate csv files through a series of copy-paste operations is something that many Excel users will admit to doing.
-We also use this example to introduce the newish pathlib Python library for working with file paths.
+We also use this example to introduce the relatively new pathlib library for working with file paths.
 After getting familiar with pathlib, students are led through the development of a short procedure that creates a list of the csv files to concatenate and then uses the pandas library to read and combine the files in a \code{DataFrame} and then writes out the consolidated data as a csv file.

### 2023-06-13 python_excel_ite.tex
working on first notebook

@@
 
+Integrated or separate lit review section?
+
 \section{Module positioning and structure}
 
-At our institution, this module is part of a business course entitled ``Advanced Analytics with Python'' (\href{http://www.sba.oakland.edu/faculty/isken/courses/mis6900}{AAP}).
+At our institution, The EwP module is part of a business course entitled ``Advanced Analytics with Python'' (\href{http://www.sba.oakland.edu/faculty/isken/courses/mis6900}{AAP}).
 The students taking this course have already taken my ``Practical Computing for Business Analytics'' (\href{http://www.sba.oakland.edu/faculty/isken/courses/mis5470}{PCDA}) course in which they spend about half the semester learning R and the other half learning Python, both in the context of business analytics.
@@
 
-What kind of Python background do students need to have in order to be ready for the EwP module? This module presumes a basic familiarity with Python fundamentals for data analysis such as:
+What kind of Python background do students need to have in order to be ready for the EwP module? A basic familiarity with Python fundamentals for data analysis is needed, including:
 
 \begin{itemize}
-	\item variables, artithmetic and boolean operators, and basic data types,
+	\item variables, arithmetic and boolean operators, and basic data types,
 	\item data structures such as lists, dictionaries, tuples, NumPy arrays and pandas dataframes,
@@
 GIVE URL.
-Some background on Jupyter notebooks for analytics.
+Some background on Jupyter notebooks for analytics - see downloaded papers.
 Questions and coding challenges sprinkled throughout the notebooks - answers at bottom.
@@
 In this way, the computational problem to be solved drives the introduction of more advanced Python concepts and techniques instead of such things being presented in a vacuum.
-I make it a point to try to convey the notion that there are often multiple possible approaches by embarking on one path and then starting over and doing things in a different way.
-Each notebook is accompanied by a few to several videos that walk the students through the material - a guided tour with challenges to overcome along the way.
+I make it a point to try to convey the notion that there are often multiple possible approaches by embarking on one path and there can be value to starting over and doing things in a different way.
+Each notebook is accompanied by videos that walk the students through the material and act as a guided tour with challenges to overcome along the way.
 The screenshot in \ref{fig:activities} shows the portion of the course web page that provides the high level roadmap through the notebooks and associated screencasts.
@@
 \subsection{Notebook 1: Modeling and data tables}
+
+The stage is set with an overview of a typical spreadsheet based model.
 
@@
 
-The first focuses on building a simple what-if model and doing sensitivity analysis much like one would do using Excel's Data-Table functionality.
+\begin{figure}[!ht]
+	\centering
+	\begin{minipage}{.80\textwidth}
+		\includegraphics[width=1.0\textwidth]{images/output_2_0.png}
+	\end{minipage}
+	\caption{Typical spreadsheet model}
+	\label{fig:ldr_pvsa}
+\end{figure}
 
-This example is based on one in the [spreadsheet modeling textbooks I've used in my classes since 2001](https://host.kelley.iu.edu/albrightbooks/).
+The example used throughout the rest of the notebook is based on one in the [spreadsheet modeling textbooks I've used in my classes since 2001](https://host.kelley.iu.edu/albrightbooks/).
 In both books, they introduce the "Walton Bookstore" problem in the chapter on Monte-Carlo simulation.
@@
 
-The notebook begins with a suggestion to start with a very simple model that ignores the uncertainty in demand and that is not object oriented.
-Proceeding much like we would in Excel, we store base input values in variables.
+We start with a very simple model that ignores the uncertainty in demand and that is not object oriented. Proceeding much like we would in Excel, we store base input values in variables.
 
@@
 In a Data Table, a range of order quantities can be used as row or column input and one or more output values can be computed.
-In Python, we can rely on the built in vectorized behavior of the language.
-A range of order quantities is created as a NumPy array and then can be used directly to compute vectors of all of the intermediate and final output variables.
-Much like Data Tables, vectorized computations avoid having to explicitly iterate (loop) through a collection of input values.
+In Python, we can rely on the vectorized behavior of the language.
+A range of order quantities is created as a NumPy array that can be used directly to compute vectors of all of the intermediate and final output variables.
+Much like Data Tables, vectorized computations avoid having to explicitly iterate, or loop, through a collection of input values.
 

### 2023-06-14 python_excel_ite.tex
working on first noteobok

@@
 	\caption{Typical spreadsheet model}
-	\label{fig:ldr_pvsa}
+	\label{fig:xl_model}
 \end{figure}
@@
 
-Then students are given some skeleton code and must finish the expressions for computing key intermediate outputs (\code{order_cost, sales_revenue, refund_revenue} and the final output, \code{profit}.
+Then students are given some skeleton code and must finish the expressions for computing key intermediate outputs (\code{order_cost, sales_revenue, refund_revenue}) and the final output, \code{profit}.
 Answers are provided at the bottom of the notebook so that students can attempt to finish the code but have a resource to correctly complete the code and move on.
@@
 
-At this point we discuss a fundamental difference between the computing model of Excel and that of Jupyter notebooks.
-While spreadsheets respond automatically to changes, Jupyter notebooks require rerunning all code cells that include or follow the changed code.
+At this point we discuss a fundamental difference between the computing model of Excel and that of Jupyter notebooks -- while spreadsheets respond automatically to changes, Jupyter notebooks require rerunning all code cells that include or follow the changed code.
 
@@
 A natural next step is to do the equivalent of an Excel 2-way Data Table.
-Can't do 2 vectors since we need all combinations of entries in each vector.
-List comprehensions provide a compact way to create nested loops over the two vectors.
 By creating a profit function that takes all of the base inputs as arguments, we can construct a list comprehension that does what an Excel 2-Way Data Table does.
-Even better, this actually allows us to do the equivalent of an n-Way Data Table.
+Even better, this actually allows us to do the equivalent of an n-Way Data Table with an arbitrary number of outputs - something Excel does not have.
+This example provides an effective way to show the power of creating reusable functions and Python's handy list comprehension tool.
 
@@
 
-Introducing basic OO concepts is a primary learning objective.
-It's not clear whether an OO or a non-OO approach will end up making the most sense for doing spreadsheet type modeling.
-This uncertainty is made explicit in the notebook and students are led on a bit of a journey of discovery as we explore different software designs for our modeling problem.
+It is not clear whether an OO or a non-OO approach will end up making the most sense for doing spreadsheet type modeling.
+This uncertainty is made explicit in the notebook and students are led on a bit of a journey of discovery as we explore different software designs for this modeling problem.
 I purposefully designed the notebooks in this way to show students that the path to a solution is not necessarily linear and that giving yourself the freedom to explore alternatives, many of which will be dead ends, can lead to a rich learning experience.
 It is also more realistic in the sense that this is how most real problems are solved.
-It's analogous to the idea that mathematical proofs usually don't reflect the myriad of failed approaches that were tried but that ultimately contributed to the final product.
+It is analogous to the idea that mathematical proofs usually do not reflect the myriad of failed approaches that were tried but that ultimately contributed to the final product.
 
@@
 The \code{dir} function is used to see all of an object's attributes and we see that even things like integers are objects in Python.
-With this basic object refresher, we are ready to do some actual object oriented programming and create our own objects - well, actually classes.
+With this basic object refresher, we are ready to do some actual object oriented programming and create our own classes.
 
@@
 In this case, the well known scikit-learn package has a \code{ParameterGrid} function that solves our scenario generation problem quite nicely.
-This example also reminds the students of the value of being able to make sense of API documentation and of object oriented code written by others.
-Most data science related Python packages are object oriented in nature.
+This example also reminds the students of the value of being able to make sense of API documentation and of object oriented code written by others -- most data science related Python packages are object oriented in nature.
 
-Whew!
-This first notebook draws to a close with the creation of an \code{data_table} function that allows us to do sensitivity analysis with our objected oriented model.
-Unlike Excel's Data Table tool, our \code{data_table} function can handle an arbitrary number of inputs and outputs.
+This first notebook draws to a close with the creation of an \code{data_table} function that allows us to do sensitivity analysis with the objected oriented model.
+Unlike Excel's Data Table tool, this \code{data_table} function can handle an arbitrary number of inputs and outputs.
 The payoff is the creation of the following faceted plot showing how profit varies for different order quantities and demand levels.
-We also recap the main things we learned in this notebook and reiterate that we will now move on to adding goal seeking capability to our model.
+We recap the main things learned in this notebook and prepare to add goal seeking capability to our model.
 
-INSERT PLOT HERE
+\begin{figure}[!ht]
+	\centering
+	\begin{minipage}{.80\textwidth}
+		\includegraphics[width=1.0\textwidth]{images/two_way_dt.png}
+	\end{minipage}
+	\caption{Plot based on 2-way data table}
+	\label{fig:2way_dt}
+\end{figure}
 

### 2023-06-15 python_excel_ite.tex
intro needs work

@@
 \usepackage{xcolor}
+
+% User added usepackage commands from Jupyter notebook preamble
+\usepackage{tcolorbox}
+%\usepackage{graphicx}
+%\usepackage{float}
+%\usepackage{xcolor} % Allow colors to be defined
+\usepackage{enumerate} % Needed for markdown enumerations to work
+\usepackage{geometry} % Used to adjust the document margins
+%\usepackage{amsmath} % Equations
+%\usepackage{amssymb} % Equations
+\usepackage{textcomp} % defines textquotesingle
+% Hack from http://tex.stackexchange.com/a/47451/13684:
+\AtBeginDocument{%
+	\def\PYZsq{\textquotesingle}% Upright quotes in Pygmentized code
+}
+\usepackage{upquote} % Upright quotes for verbatim code
+\usepackage{eurosym} % defines \euro
+
+\usepackage{fancyvrb} % verbatim replacement that allows latex
+
+% Private macros for Jupyter notebook styling
+
+% Colors for the hyperref package
+\definecolor{urlcolor}{rgb}{0,.145,.698}
+\definecolor{linkcolor}{rgb}{.71,0.21,0.01}
+\definecolor{citecolor}{rgb}{.12,.54,.11}
+
+% ANSI colors
+\definecolor{ansi-black}{HTML}{3E424D}
+\definecolor{ansi-black-intense}{HTML}{282C36}
+\definecolor{ansi-red}{HTML}{E75C58}
+\definecolor{ansi-red-intense}{HTML}{B22B31}
+\definecolor{ansi-green}{HTML}{00A250}
+\definecolor{ansi-green-intense}{HTML}{007427}
+\definecolor{ansi-yellow}{HTML}{DDB62B}
+\definecolor{ansi-yellow-intense}{HTML}{B27D12}
+\definecolor{ansi-blue}{HTML}{208FFB}
+\definecolor{ansi-blue-intense}{HTML}{0065CA}
+\definecolor{ansi-magenta}{HTML}{D160C4}
+\definecolor{ansi-magenta-intense}{HTML}{A03196}
+\definecolor{ansi-cyan}{HTML}{60C6C8}
+\definecolor{ansi-cyan-intense}{HTML}{258F8F}
+\definecolor{ansi-white}{HTML}{C5C1B4}
+\definecolor{ansi-white-intense}{HTML}{A1A6B2}
+\definecolor{ansi-default-inverse-fg}{HTML}{FFFFFF}
+\definecolor{ansi-default-inverse-bg}{HTML}{000000}
+
+% Pygments definitions
+\makeatletter
+\def\PY@reset{\let\PY@it=\relax \let\PY@bf=\relax%
+	\let\PY@ul=\relax \let\PY@tc=\relax%
+	\let\PY@bc=\relax \let\PY@ff=\relax}
+\def\PY@tok#1{\csname PY@tok@#1\endcsname}
+\def\PY@toks#1+{\ifx\relax#1\empty\else%
+	\PY@tok{#1}\expandafter\PY@toks\fi}
+\def\PY@do#1{\PY@bc{\PY@tc{\PY@ul{%
+				\PY@it{\PY@bf{\PY@ff{#1}}}}}}}
+\def\PY#1#2{\PY@reset\PY@toks#1+\relax+\PY@do{#2}}
+
+\@namedef{PY@tok@w}{\def\PY@tc##1{\textcolor[rgb]{0.73,0.73,0.73}{##1}}}
+\@namedef{PY@tok@c}{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.24,0.48,0.48}{##1}}}
+\@namedef{PY@tok@cp}{\def\PY@tc##1{\textcolor[rgb]{0.61,0.40,0.00}{##1}}}
+\@namedef{PY@tok@k}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\@namedef{PY@tok@kp}{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\@namedef{PY@tok@kt}{\def\PY@tc##1{\textcolor[rgb]{0.69,0.00,0.25}{##1}}}
+\@namedef{PY@tok@o}{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\@namedef{PY@tok@ow}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
+\@namedef{PY@tok@nb}{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\@namedef{PY@tok@nf}{\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
+\@namedef{PY@tok@nc}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
+\@namedef{PY@tok@nn}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
+\@namedef{PY@tok@ne}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.80,0.25,0.22}{##1}}}
+\@namedef{PY@tok@nv}{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
+\@namedef{PY@tok@no}{\def\PY@tc##1{\textcolor[rgb]{0.53,0.00,0.00}{##1}}}
+\@namedef{PY@tok@nl}{\def\PY@tc##1{\textcolor[rgb]{0.46,0.46,0.00}{##1}}}
+\@namedef{PY@tok@ni}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.44,0.44,0.44}{##1}}}
+\@namedef{PY@tok@na}{\def\PY@tc##1{\textcolor[rgb]{0.41,0.47,0.13}{##1}}}
+\@namedef{PY@tok@nt}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\@namedef{PY@tok@nd}{\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
+\@namedef{PY@tok@s}{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\@namedef{PY@tok@sd}{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\@namedef{PY@tok@si}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.64,0.35,0.47}{##1}}}
+\@namedef{PY@tok@se}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.67,0.36,0.12}{##1}}}
+\@namedef{PY@tok@sr}{\def\PY@tc##1{\textcolor[rgb]{0.64,0.35,0.47}{##1}}}
+\@namedef{PY@tok@ss}{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
+\@namedef{PY@tok@sx}{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\@namedef{PY@tok@m}{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\@namedef{PY@tok@gh}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
+\@namedef{PY@tok@gu}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.50,0.00,0.50}{##1}}}
+\@namedef{PY@tok@gd}{\def\PY@tc##1{\textcolor[rgb]{0.63,0.00,0.00}{##1}}}
+\@namedef{PY@tok@gi}{\def\PY@tc##1{\textcolor[rgb]{0.00,0.52,0.00}{##1}}}
+\@namedef{PY@tok@gr}{\def\PY@tc##1{\textcolor[rgb]{0.89,0.00,0.00}{##1}}}
+\@namedef{PY@tok@ge}{\let\PY@it=\textit}
+\@namedef{PY@tok@gs}{\let\PY@bf=\textbf}
+\@namedef{PY@tok@gp}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
+\@namedef{PY@tok@go}{\def\PY@tc##1{\textcolor[rgb]{0.44,0.44,0.44}{##1}}}
+\@namedef{PY@tok@gt}{\def\PY@tc##1{\textcolor[rgb]{0.00,0.27,0.87}{##1}}}
+\@namedef{PY@tok@err}{\def\PY@bc##1{{\setlength{\fboxsep}{\string -\fboxrule}\fcolorbox[rgb]{1.00,0.00,0.00}{1,1,1}{\strut ##1}}}}
+\@namedef{PY@tok@kc}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\@namedef{PY@tok@kd}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\@namedef{PY@tok@kn}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\@namedef{PY@tok@kr}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\@namedef{PY@tok@bp}{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\@namedef{PY@tok@fm}{\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
+\@namedef{PY@tok@vc}{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
+\@namedef{PY@tok@vg}{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
+\@namedef{PY@tok@vi}{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
+\@namedef{PY@tok@vm}{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
+\@namedef{PY@tok@sa}{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\@namedef{PY@tok@sb}{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\@namedef{PY@tok@sc}{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\@namedef{PY@tok@dl}{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\@namedef{PY@tok@s2}{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\@namedef{PY@tok@sh}{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\@namedef{PY@tok@s1}{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\@namedef{PY@tok@mb}{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\@namedef{PY@tok@mf}{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\@namedef{PY@tok@mh}{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\@namedef{PY@tok@mi}{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\@namedef{PY@tok@il}{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\@namedef{PY@tok@mo}{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\@namedef{PY@tok@ch}{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.24,0.48,0.48}{##1}}}
+\@namedef{PY@tok@cm}{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.24,0.48,0.48}{##1}}}
+\@namedef{PY@tok@cpf}{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.24,0.48,0.48}{##1}}}
+\@namedef{PY@tok@c1}{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.24,0.48,0.48}{##1}}}
+\@namedef{PY@tok@cs}{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.24,0.48,0.48}{##1}}}
+
+\def\PYZbs{\char`\\}
+\def\PYZus{\char`\_}
+\def\PYZob{\char`\{}
+\def\PYZcb{\char`\}}
+\def\PYZca{\char`\^}
+\def\PYZam{\char`\&}
+\def\PYZlt{\char`\<}
+\def\PYZgt{\char`\>}
+\def\PYZsh{\char`\#}
+\def\PYZpc{\char`\%}
+\def\PYZdl{\char`\$}
+\def\PYZhy{\char`\-}
+\def\PYZsq{\char`\'}
+\def\PYZdq{\char`\"}
+\def\PYZti{\char`\~}
+% for compatibility with earlier versions
+\def\PYZat{@}
+\def\PYZlb{[}
+\def\PYZrb{]}
+\makeatother
+
+    % Define a nice break command that doesn't care if a line doesn't already
+% exist.
+\def\br{\hspace*{\fill} \\* }
+
+% For linebreaks inside Verbatim environment from package fancyvrb.
+\makeatletter
+\newbox\Wrappedcontinuationbox
+\newbox\Wrappedvisiblespacebox
+\newcommand*\Wrappedvisiblespace {\textcolor{red}{\textvisiblespace}}
+\newcommand*\Wrappedcontinuationsymbol {\textcolor{red}{\llap{\tiny$\m@th\hookrightarrow$}}}
+\newcommand*\Wrappedcontinuationindent {3ex }
+\newcommand*\Wrappedafterbreak {\kern\Wrappedcontinuationindent\copy\Wrappedcontinuationbox}
+% Take advantage of the already applied Pygments mark-up to insert
+% potential linebreaks for TeX processing.
+%        {, <, #, %, $, ' and ": go to next line.
+%        _, }, ^, &, >, - and ~: stay at end of broken line.
+% Use of \textquotesingle for straight quote.
+\newcommand*\Wrappedbreaksatspecials {%
+	\def\PYGZus{\discretionary{\char`\_}{\Wrappedafterbreak}{\char`\_}}%
+	\def\PYGZob{\discretionary{}{\Wrappedafterbreak\char`\{}{\char`\{}}%
+	\def\PYGZcb{\discretionary{\char`\}}{\Wrappedafterbreak}{\char`\}}}%
+	\def\PYGZca{\discretionary{\char`\^}{\Wrappedafterbreak}{\char`\^}}%
+	\def\PYGZam{\discretionary{\char`\&}{\Wrappedafterbreak}{\char`\&}}%
+	\def\PYGZlt{\discretionary{}{\Wrappedafterbreak\char`\<}{\char`\<}}%
+	\def\PYGZgt{\discretionary{\char`\>}{\Wrappedafterbreak}{\char`\>}}%
+	\def\PYGZsh{\discretionary{}{\Wrappedafterbreak\char`\#}{\char`\#}}%
+	\def\PYGZpc{\discretionary{}{\Wrappedafterbreak\char`\%}{\char`\%}}%
+	\def\PYGZdl{\discretionary{}{\Wrappedafterbreak\char`\$}{\char`\$}}%
+	\def\PYGZhy{\discretionary{\char`\-}{\Wrappedafterbreak}{\char`\-}}%
+	\def\PYGZsq{\discretionary{}{\Wrappedafterbreak\textquotesingle}{\textquotesingle}}%
+	\def\PYGZdq{\discretionary{}{\Wrappedafterbreak\char`\"}{\char`\"}}%
+	\def\PYGZti{\discretionary{\char`\~}{\Wrappedafterbreak}{\char`\~}}%
+}
+% Some characters . , ; ? ! / are not pygmentized.
+% This macro makes them "active" and they will insert potential linebreaks
+\newcommand*\Wrappedbreaksatpunct {%
+	\lccode`\~`\.\lowercase{\def~}{\discretionary{\hbox{\char`\.}}{\Wrappedafterbreak}{\hbox{\char`\.}}}%
+	\lccode`\~`\,\lowercase{\def~}{\discretionary{\hbox{\char`\,}}{\Wrappedafterbreak}{\hbox{\char`\,}}}%
+	\lccode`\~`\;\lowercase{\def~}{\discretionary{\hbox{\char`\;}}{\Wrappedafterbreak}{\hbox{\char`\;}}}%
+	\lccode`\~`\:\lowercase{\def~}{\discretionary{\hbox{\char`\:}}{\Wrappedafterbreak}{\hbox{\char`\:}}}%
+	\lccode`\~`\?\lowercase{\def~}{\discretionary{\hbox{\char`\?}}{\Wrappedafterbreak}{\hbox{\char`\?}}}%
+	\lccode`\~`\!\lowercase{\def~}{\discretionary{\hbox{\char`\!}}{\Wrappedafterbreak}{\hbox{\char`\!}}}%
+	\lccode`\~`\/\lowercase{\def~}{\discretionary{\hbox{\char`\/}}{\Wrappedafterbreak}{\hbox{\char`\/}}}%
+	\catcode`\.\active
+	\catcode`\,\active
+	\catcode`\;\active
+	\catcode`\:\active
+	\catcode`\?\active
+	\catcode`\!\active
+	\catcode`\/\active
+	\lccode`\~`\~
+}
+\makeatother
+
+\let\OriginalVerbatim=\Verbatim
+\makeatletter
+\renewcommand{\Verbatim}[1][1]{%
+	%\parskip\z@skip
+	\sbox\Wrappedcontinuationbox {\Wrappedcontinuationsymbol}%
+	\sbox\Wrappedvisiblespacebox {\FV@SetupFont\Wrappedvisiblespace}%
+	\def\FancyVerbFormatLine ##1{\hsize\linewidth
+		\vtop{\raggedright\hyphenpenalty\z@\exhyphenpenalty\z@
+			\doublehyphendemerits\z@\finalhyphendemerits\z@
+			\strut ##1\strut}%
+	}%
+	% If the linebreak is at a space, the latter will be displayed as visible
+	% space at end of first line, and a continuation symbol starts next line.
+	% Stretch/shrink are however usually zero for typewriter font.
+	\def\FV@Space {%
+		\nobreak\hskip\z@ plus\fontdimen3\font minus\fontdimen4\font
+		\discretionary{\copy\Wrappedvisiblespacebox}{\Wrappedafterbreak}
+		{\kern\fontdimen2\font}%
+	}%
+
+	% Allow breaks at special characters using \PYG... macros.
+	\Wrappedbreaksatspecials
+	% Breaks at punctuation characters . , ; ? ! and / need catcode=\active
+	\OriginalVerbatim[#1,codes*=\Wrappedbreaksatpunct]%
+}
+\makeatother
+
+% Exact colors from NB
+\definecolor{incolor}{HTML}{303F9F}
+\definecolor{outcolor}{HTML}{D84315}
+\definecolor{cellborder}{HTML}{CFCFCF}
+\definecolor{cellbackground}{HTML}{F7F7F7}
+
+% prompt
+\makeatletter
+\newcommand{\boxspacing}{\kern\kvtcb@left@rule\kern\kvtcb@boxsep}
+\makeatother
+\newcommand{\prompt}[4]{
+	{\ttfamily\llap{{\color{#2}[#3]:\hspace{3pt}#4}}\vspace{-\baselineskip}}
+}
+
+% Prevent overflowing lines due to hard-to-break entities
+\sloppy
+% Setup hyperref package
+\hypersetup{
+	breaklinks=true,  % so long urls are correctly broken across lines
+	colorlinks=true,
+	urlcolor=urlcolor,
+	linkcolor=linkcolor,
+	citecolor=citecolor,
+}
+% Slightly bigger margins than the latex defaults
+
+\geometry{verbose,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}
 
@@
 Taught using combo of R Markdown documents and Jupyter notebooks.
-Literate programming.
+Literate computing \cite{Jupyter_proposal}.
+Educational narratives that support iterative exploration and development.
+The ability to weave standard language with formal programming languages is critical to creating the desired educational artifact.
+Jupyter used by many universities as a key component of their educational technology stack.
+
 PCDA course focused mostly on data wrangling, EDA, statistical and machine learning based predictive models though I did include a Python based simulation of the Monte Hall 3-door problem.
@@
 
+Jupyter notebooks for teaching data science are an established delivery mechanism.
+Computational notebook allowing mixing of explanatory text, written in Markdown with executable code (and its output) in Python or a number of other languages.
+Literate programming.
+Can also include equations, visualizations, other multi-media.
+It's free, open source, runs on all platforms, huge ecosystem, used in industry, used widely in education.
+
 \section{Submodule 1: What-if analysis with Python}
@@
 Students learn how Python might be used to do something they are quite familiar with doing in a spreadsheet - model building and sensitivity analysis.
-At the same time, students encounter, as they are needed, more advanced Python programming techniques such as object oriented programming and manipulation of tuples, lists and dictionaries.
+At the same time, students encounter, as they are needed, more advanced Python programming material.
 In this way, the computational problem to be solved drives the introduction of more advanced Python concepts and techniques instead of such things being presented in a vacuum.
@@
 
-The stage is set with an overview of a typical spreadsheet based model.
+The stage is set with an overview of a typical spreadsheet based model that most of the students have seen before.
 
@@
 
+With this as motivation, we then introduce a similar modeling problem.
 The example used throughout the rest of the notebook is based on one in the [spreadsheet modeling textbooks I've used in my classes since 2001](https://host.kelley.iu.edu/albrightbooks/).
@@
 	\item we have to place an order for a perishable product (e.g. a calendar),
-	\item there is a known unit cost for each one ordered,
-	\item we have a known selling price,
+	\item there is a known unit cost for each unit ordered,
+	\item we have a known and fixed selling price,
 	\item demand is uncertain but we can model it with some simple probability distribution,
@@
 
-Then students are given some skeleton code and must finish the expressions for computing key intermediate outputs (\code{order_cost, sales_revenue, refund_revenue}) and the final output, \code{profit}.
+\begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
+	%\prompt{In}{incolor}{3}{\boxspacing}
+	\begin{Verbatim}[commandchars=\\\{\}]
+	\PY{c+c1}{\PYZsh{} Base inputs}
+	\PY{n}{unit\PYZus{}cost} \PY{o}{=} \PY{l+m+mf}{7.50}
+	\PY{n}{selling\PYZus{}price} \PY{o}{=} \PY{l+m+mf}{10.00}
+	\PY{n}{unit\PYZus{}refund} \PY{o}{=} \PY{l+m+mf}{2.50}
+	\end{Verbatim}
+\end{tcolorbox}
+
+Students are given skeleton code as scaffolding and must finish the expressions for computing key intermediate outputs (\code{order\_cost, sales\_revenue, refund\_revenue}) and the final output, \code{profit}.
 Answers are provided at the bottom of the notebook so that students can attempt to finish the code but have a resource to correctly complete the code and move on.
-The screencast associated with the notebook also shows me completing the code.
-Using skeleton code in this way has worked well both for face to face lab based versions of this type of course as well as for online asynchronous versions.
+The screencast associated with the notebook also shows me completing the code in a subsequent code cell.
+Using skeleton code in this way has worked well both for lab based versions of this type of course as well as for online asynchronous versions.
 
@@
 At this point we discuss a fundamental difference between the computing model of Excel and that of Jupyter notebooks -- while spreadsheets respond automatically to changes, Jupyter notebooks require rerunning all code cells that include or follow the changed code.
+This difference in behavior between Jupyter notebooks and spreadsheets might be one of the most important concepts for students to internalize.
+A well known challenge associated with computational notebooks is the ease of running code cells out of order and introducing inadvertent errors \cite{bibid}.
+This point along with strategies to avoid it are discussed during the first week of the course which focuses on notebook computing best practices as well as an historical look at the use of computational notebooks \cite{bibid}.
 
@@
 
-Since order quantity is the key decision variable, we might want to see how profit changes for different order quantities in this first simplified model.
+Since order quantity is the key decision variable, we might want to see how profit changes for different order quantities.
 In Excel, the Data Table tool provides an easy way to do this and is a staple of spreadsheet based management science textbooks.
 In a Data Table, a range of order quantities can be used as row or column input and one or more output values can be computed.
-In Python, we can rely on the vectorized behavior of the language.
+In Python, we can rely on the vectorized behavior of the language \cite{bibid}.
 A range of order quantities is created as a NumPy array that can be used directly to compute vectors of all of the intermediate and final output variables.
@@
 Even better, this actually allows us to do the equivalent of an n-Way Data Table with an arbitrary number of outputs - something Excel does not have.
-This example provides an effective way to show the power of creating reusable functions and Python's handy list comprehension tool.
+This example provides an effective way to show the power of creating reusable functions and Python's handy list comprehension construct.
 
@@
 
-It is not clear whether an OO or a non-OO approach will end up making the most sense for doing spreadsheet type modeling.
+It is not clear whether an object oriented (OO) or a non-OO approach will end up making the most sense for doing spreadsheet type modeling.
 This uncertainty is made explicit in the notebook and students are led on a bit of a journey of discovery as we explore different software designs for this modeling problem.
@@
 Students can easily see the mapping between things in the initial procedural model and this new object oriented version of the model.
-I highlight several important concepts and syntactical details of doing OO programming Python.
+I highlight several important concepts and syntactical details of doing object oriented programming Python.
 Again, the goal is to weave in these more advanced Python programming concepts within the context of a very familiar modeling problem.
@@
 These meta reflections are intentionally included to better mimic actual software development and reinforce the idea that software development is much more than simply writing code to implement a perfectly thought out design.
-We forge ahead with some preliminary ideas and are confronted with an intermediate problem related to generating a list of dictionaries that represent combinations of inputs (i.e.
-scenarios) to evaluate by our n-way data table function.
+We forge ahead with some preliminary ideas and are confronted with an intermediate problem related to generating a list of dictionaries that represent combinations of inputs (i.e.\ scenarios) to evaluate by our n-way data table function.
 This provides a great opportunity to remind students of one of the great strengths of the open source software ecosystem - we can leverage work done in other packages and even look at actual source code to see how something was implemented.
-In this case, the well known scikit-learn package has a \code{ParameterGrid} function that solves our scenario generation problem quite nicely.
-This example also reminds the students of the value of being able to make sense of API documentation and of object oriented code written by others -- most data science related Python packages are object oriented in nature.
+In this case, the well known scikit-learn package \cite{bibid} has a \code{ParameterGrid} function that solves our scenario generation problem quite nicely.
+This example also reminds the students of the value of being able to make sense of API documentation and of object oriented code written by others as most data science related Python packages are object oriented in nature.
 
-This first notebook draws to a close with the creation of an \code{data_table} function that allows us to do sensitivity analysis with the objected oriented model.
-Unlike Excel's Data Table tool, this \code{data_table} function can handle an arbitrary number of inputs and outputs.
+This first notebook draws to a close with the creation of a \code{data\_table} function that allows us to do sensitivity analysis with the objected oriented model.
+Unlike Excel's Data Table tool, this \code{data\_table} function can handle an arbitrary number of inputs and outputs.
 The payoff is the creation of the following faceted plot showing how profit varies for different order quantities and demand levels.
@@
 For many students, Excel's Goal Seek tool can seem almost magical when encountered for the first time.
-Even after they get a sense of what it is doing, understanding the importance of the initial guess and how it can lead to Goal Seek reporting different solutions, is not very transparent.
+Even after they get a sense of what it is doing, understanding the importance of the initial solution guess and how it can lead to Goal Seek reporting different solutions, is not very transparent.
 With Python, students can gain a much better understanding of how tools like Goal Seek really work.
 
-Before trying to build our \code{goal_seek} function we take a brief detour into basic root finding algorithms through packages like SciPy as well as in standalone Python scripts that are widely available in the Python data science ecosystem (refs).
+Before trying to build our \code{goal\_seek} function we take a brief detour into basic root finding algorithms through packages like SciPy \cite{bibid} as well as in standalone Python scripts that are widely available in the Python data science ecosystem \cite{bibid}.
 We compare the output of Excel's Goal Seek with different starting values to the values obtained by different root finding algorithms implemented in Python.
-The importance of starting values and also the range of different root finding algorithms available is often surprising to business students who have spent most of their life in Excel.
+The importance of starting values and also the range of different root finding algorithms available is often surprising to business students who have spent most of their analytical life in Excel.
 
-A \code{} function is created that implements a simple bisection search and we use it to find the break even demand point in our bookstore model.
-With \code{data_table} and \code{goal_seek} functions implemented, we are ready to move on to doing Monte-Carlo simulation.
+A \code{goal\_seek} function is created that implements a simple bisection search and we use it to find the break even demand point in our bookstore model.
+With \code{data\_table} and \code{goal\_seek} functions implemented, we are ready to move on to doing Monte-Carlo simulation.
 
@@
 In the Excel modeling world, Monte-Carlo simulation can be done without add-ins, but packages like @Risk \cite{} can make the process much easier.
-In particular, random variable generation for a wide range of probability distributions is facilitated by @Risk.
-In Python, the NumPy package provide random variate generation for a large number of distributions (https://numpy.org/doc/stable/reference/random/legacy.html#distributions) and similarly, SciPy provides functions for computing distribution related quantities for many probability distributions \cite{}.
-Much like we did when creating the \code{data_table} function, we start with adding uncertainty to a single input variable in the bookstore model and rely on Python's vectorized computing capabilities to generate the profit associated with each realization of the random input.
-Then we can do standard statistical analysis of the simulation output using Python packages such as Pandas and SciPy.
+In particular, random variable generation for a wide range of probability distributions is facilitated by add-ins like @Risk.
+In Python, the NumPy package provide random variate generation for a large number of distributions and similarly, SciPy provides functions for computing distribution related quantities for many probability distributions \cite{}.
+Much like we did when creating the \code{data\_table} function, we start with adding uncertainty to a single input variable in the bookstore model and rely on Python's vectorized computing capabilities to generate the profit associated with each realization of the random input.
+Then we can do standard statistical analysis of the simulation output using Python packages such as pandas and SciPy.
 In spreadsheet based simulation textbooks, the notion of the ``Flaw of Averages'' popularized by Sam Savage \cite{} is illustrated.
@@
 Moving on to multiple uncertain inputs raises challenges much like those we faced when moving from a 1-way to an n-way data table function.
-In addition to modeling multiple uncertain inputs, we want to be able to specify a range of scenarios to run, much like we did for the \code{data_table} function.
+In addition to modeling multiple uncertain inputs, we want to be able to specify a range of scenarios to run, much like we did for the \code{data\_table} function.
 The following set of design criteria are presented for a \code{simulate} function.
@@
 \begin{itemize}
-	\tightlist
 	\item
 	The first argument will be a model object (i.e.~something like the
-	\texttt{BookstoreModel} model) that contains an \texttt{update}
-	method. Soon, we should add an abstract \texttt{Model} class from
+	\code{BookstoreModel} model) that contains an \code{update}
+	method. Soon, we should add an abstract \code{Model} class from
 	which specific models such as the \texttt{BookstoreModel} class can be
-	created. The abstract class will contain the \texttt{update} method.
+	created. The abstract class will contain the \code{update} method.
 	\item
@@
 	Structurally, this is similar to how inputs are specified in the
-	\texttt{data\_table} function.
+	\code{data\_table} function.
 	\item
 	We can optionally pass in a dictionary of scenario inputs. This is
-	exactly like the \texttt{data\_table} variable input.
+	exactly like the \code{data\_table} variable input.
 
 	\begin{itemize}
-		\tightlist
 		\item
@@
 		dictionary. Again, this is just like we do in the
-		\texttt{data\_table} function.
+		\code{data\_table} function.
 	\end{itemize}
@@
 By this point, students are well equipped to understand the code as it builds on basic Python concepts and more advanced things that we have already covered in this EwP module.
-The rest of the notebook is spent trying out the \code{simulate} function and ends with creating a plot that many of them have likely created before with @Risk along with a faceted plot of histograms of profit created with the popular Seaborn package.
+The rest of the notebook is spent trying out the \code{simulate} function and ends with creating a grouped boxplot that many of them have likely created before with @Risk along with a faceted plot of histograms of profit created with the popular Seaborn package.
 
-INSERT GROUPED BOX PLOT HERE and FACETED HISTOGRAMS
+\begin{figure}[!ht]
+	\centering
+	\begin{minipage}{.80\textwidth}
+		\includegraphics[width=1.0\textwidth]{images/boxplots.png}
+	\end{minipage}
+	\caption{Profit boxplots by order quantity}
+	\label{fig:2way_dt}
+\end{figure}
 
@@
 \begin{quote}
-	content...
 
@@
 \begin{itemize}
-	\tightlist
 	\item
@@
 
-Now that we've got a critical mass of "proof of concept" code, let's figure out how to structure our project and create a deployable package.
+Now that we've got a critical mass of ``proof of concept'' code, let's figure out how to structure our project and create a deployable package.
 In addition, let's rethink our OO design and add some much needed documentation to the code.
@@
 This will require some design changes to our software and will also involve moving code out of Jupyter notebooks and into Python script files.
-While Jupyter notebooks can be integrated into production environments \cite{netflix}, it is much more common for code to live in a Python script file.
+While Jupyter notebooks can be integrated into production environments \cite{netflix}, it is much more common for code to live in Python script files.
 In addition to creating a deployable package, this submodule also discusses the importance of documentation in its many forms in a Python project.
@@
 
-The first notebook in this submodule starts by explaining what Python package are and their role in sharing code.
+The first notebook in this submodule starts by explaining what Python packages are and their role in sharing code.
 We then revisit a concept from earlier in the course - creating a good project folder structure.
-For this we use what is known as a \textit{cookiecutter} that can automatically generate a project folder structure and key files from the answers to a few prompts \cite{cookiecutter}.
-Their are a few different folder layouts used in Python projects and there is a bit of discussion on the folder stucture we will use and resources given for exploring this issue further.
+For this we use what is known as a \code{cookiecutter} that can automatically generate a project folder structure and key files from the answers to a few prompts \cite{cookiecutter}.
+There are a few different folder layouts used in Python projects and there is a bit of discussion on the folder structure we will use and resources given for exploring this issue further.
 Students are reminded of the importance of version control and we initialize a git repository for our budding package.
@@
 My experience is that most students have little exposure to simple software project management concepts such as project folder structures and version control.
-This can lead to a jumble of files with incoherent names used as a proxy for version control - whatif_v1.py, whatif_v2.py, whatif_final.py, whatif_finalfinal.py, ...
+This can lead to a jumble of files with incoherent names used as a proxy for version control - \code{whatif\_v1.py}, \code{whatif\_v2.py}, \code{whatif\_final.py}, \code{whatif\_finalfinal.py}, and so on.
 
 Key software design changes are then made and discussed which will make our code easier on our users.
-These changes include creating an abstract \code{Model} base class from which different model classes can be created and moving the \code{data_table}, \code{goal_seek}, and \code{simulate} functions into the \code{Model} base class.
+These changes include creating an abstract \code{Model} base class from which different model classes can be created and moving the \code{data\_table}, \code{goal\_seek}, and \code{simulate} functions into the \code{Model} base class.
 All of the code is moved into a code module named \code{whatif.py}.
 We are now ready to learn about the basics of turning our code into an installable package.
-This requires a basic understanding of two key Python files that are ubiquitous in Python pacakges - \code{__init__.py} and \code{setup.py}.
-Next comes a discussion of options for installing our package locally so that we can use it for different modeling projects.
+This requires a basic understanding of two key Python files that are ubiquitous in Python pacakges - \code{\_\_init\_\_.py} and \code{setup.py}.
+Options are presented for installing our package locally so that we can use it for different modeling projects.
 This includes an explanation of how and where Python searches for packages when \code{import} statements are encountered in code.
@@
 
-While this first example is not necessarily Excel related, using Excel to consolidate csv files through a series of copy-paste operations is something that many Excel users will admit to doing.
-We also use this example to introduce the relatively new pathlib library for working with file paths.
-After getting familiar with pathlib, students are led through the development of a short procedure that creates a list of the csv files to concatenate and then uses the pandas library to read and combine the files in a \code{DataFrame} and then writes out the consolidated data as a csv file.
+While this first example is not necessarily Excel related, using Excel to consolidate csv files through a series of manual Copy-Paste operations is something that many Excel users will admit to having done.
+After getting familiar with pathlib, a library for working with file systems which has recently been added to Python's standard library, students are led through the development of a short procedure that creates a list of the csv files to concatenate and then uses the pandas library to read and combine the files in a \code{DataFrame} and then writes out the consolidated data as a csv file.
 
@@
 Two different approaches are shown to consolidate data from multiple sheets into a single sheet.
-One approach uses the pandas library while the other relies on a library specifically created for working with Excel files from Python, openpyxl.
-One advantage to using Python rather than Visual Basic for Applications for this task is that Python is well suited for working with data in many different formats.
+One approach uses the pandas library while the other relies on a library specifically created for working with Excel files from Python, openpyxl \cite{bibid}.
+One advantage to using Python rather than Visual Basic for Applications for this task is that Python is well suited for working with data in many different formats including csv, json, xml and a myriad of database formats.
 The Excel file may be just one data source out of a collection of data sources that are used to create some sort of combined database.
@@
 
-Reshaping data from wide to long is a common data wrangling task especially in preparation for plotting.
+Reshaping data \cite{bibid} from wide to long is a common data wrangling task especially in preparation for plotting.
 In this example, not only do students have to grapple with reshaping the data in an Excel workbook, they also have to deal with the all too common problem of multi-row header lines in a spreadsheet.
@@
 
-\subsection{Example 4: Appending new spreadsheet data in consolidated Excel workbook}
+\subsection{Example 4: Appending new spreadsheet data in a consolidated Excel workbook}
 
 Excel is not a database.
-But, that does not stop people from using it as a simple flat-file database.
+Of course, that does not stop people from using it as a simple flat-file database.
 In such a scenario, it is not uncommon to have to periodically append new data to the end of a range of data in an Excel workbook.
@@
 The first part of the assignment involves creating a basic model, similar in complexity to the Bookstore Model, and doing sensitivity analysis, goal seeking and simulation with the model.
-It requires the students to import and use the whatif.py package we created as part of the EwP module.
+It requires the students to import and use the \code{whatif.py} package we created as part of the EwP module.
 The second part of the assignment is based on the last submodule and requires the students to complete a task involving creation of a multi-sheet Excel workbook from a set of csv files using Python.
@@
 
-Rumours have swirled for a number of years that Microsoft was considering adding Python as a first class language alternative to VBA with their MS Office suite of packages.
-While this has not yet materialized, there continues to be significant interest in using Python to manipulate not only Excel workbooks but other office documents such as Powerpoint presentations.
+Rumours have swirled for a number of years that Microsoft was considering adding Python as a first class language alternative to VBA with their MS Office suite of packages \cite{bibid}.
+While this has not yet materialized, there continues to be significant interest in using Python to manipulate not only Excel workbooks but other office documents such as Powerpoint presentations \cite{bibid}.
 Given that business students are very familiar with both Excel and Powerpoint, inclusion of this material in a business school Python based analytics course seems quite appropriate.
 Students can learn more advanced and extremely useful Python concepts and techniques within a context with which they are intimately familiar.
-It is directly relevant to the job content of many of the students.
+Manipulation of MS Office directly relevant to the job content of many of the students.
 This teaching module is a natural follow up to the commonly taught spreadsheet based analytics courses found in many business schools and provides an interesting way to teach and learn Python for analytics.
@@
 Learning good Python programming practices via learning to model in similar way that students learn good spreadsheet practices through building spreadsheet models using the W\&A textbook.
-
-How aap course fits in relation to pcda course. Could also be second part of a fully Python based course. Builds on basic intro Python concepts and skills.
-
-The Quant Econ website is a good example of another use of Python for typical Excel things.
-
-Everything discussed in this article is available from the publicly accessible course website.
 

### 2023-06-20 python_excel_ite.tex
found strategy for code samples using tcolorbox and minted

@@
 
-\usepackage{hyperref}
-\usepackage{listings}
-\usepackage{xcolor}
-
-% User added usepackage commands from Jupyter notebook preamble
-\usepackage{tcolorbox}
-%\usepackage{graphicx}
-%\usepackage{float}
-%\usepackage{xcolor} % Allow colors to be defined
-\usepackage{enumerate} % Needed for markdown enumerations to work
-\usepackage{geometry} % Used to adjust the document margins
-%\usepackage{amsmath} % Equations
-%\usepackage{amssymb} % Equations
-\usepackage{textcomp} % defines textquotesingle
-% Hack from http://tex.stackexchange.com/a/47451/13684:
-\AtBeginDocument{%
-	\def\PYZsq{\textquotesingle}% Upright quotes in Pygmentized code
-}
-\usepackage{upquote} % Upright quotes for verbatim code
-\usepackage{eurosym} % defines \euro
-
-\usepackage{fancyvrb} % verbatim replacement that allows latex
-
-% Private macros for Jupyter notebook styling
-
-% Colors for the hyperref package
-\definecolor{urlcolor}{rgb}{0,.145,.698}
-\definecolor{linkcolor}{rgb}{.71,0.21,0.01}
-\definecolor{citecolor}{rgb}{.12,.54,.11}
-
-% ANSI colors
-\definecolor{ansi-black}{HTML}{3E424D}
-\definecolor{ansi-black-intense}{HTML}{282C36}
-\definecolor{ansi-red}{HTML}{E75C58}
-\definecolor{ansi-red-intense}{HTML}{B22B31}
-\definecolor{ansi-green}{HTML}{00A250}
-\definecolor{ansi-green-intense}{HTML}{007427}
-\definecolor{ansi-yellow}{HTML}{DDB62B}
-\definecolor{ansi-yellow-intense}{HTML}{B27D12}
-\definecolor{ansi-blue}{HTML}{208FFB}
-\definecolor{ansi-blue-intense}{HTML}{0065CA}
-\definecolor{ansi-magenta}{HTML}{D160C4}
-\definecolor{ansi-magenta-intense}{HTML}{A03196}
-\definecolor{ansi-cyan}{HTML}{60C6C8}
-\definecolor{ansi-cyan-intense}{HTML}{258F8F}
-\definecolor{ansi-white}{HTML}{C5C1B4}
-\definecolor{ansi-white-intense}{HTML}{A1A6B2}
-\definecolor{ansi-default-inverse-fg}{HTML}{FFFFFF}
-\definecolor{ansi-default-inverse-bg}{HTML}{000000}
-
-% Pygments definitions
-\makeatletter
-\def\PY@reset{\let\PY@it=\relax \let\PY@bf=\relax%
-	\let\PY@ul=\relax \let\PY@tc=\relax%
-	\let\PY@bc=\relax \let\PY@ff=\relax}
-\def\PY@tok#1{\csname PY@tok@#1\endcsname}
-\def\PY@toks#1+{\ifx\relax#1\empty\else%
-	\PY@tok{#1}\expandafter\PY@toks\fi}
-\def\PY@do#1{\PY@bc{\PY@tc{\PY@ul{%
-				\PY@it{\PY@bf{\PY@ff{#1}}}}}}}
-\def\PY#1#2{\PY@reset\PY@toks#1+\relax+\PY@do{#2}}
-
-\@namedef{PY@tok@w}{\def\PY@tc##1{\textcolor[rgb]{0.73,0.73,0.73}{##1}}}
-\@namedef{PY@tok@c}{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.24,0.48,0.48}{##1}}}
-\@namedef{PY@tok@cp}{\def\PY@tc##1{\textcolor[rgb]{0.61,0.40,0.00}{##1}}}
-\@namedef{PY@tok@k}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
-\@namedef{PY@tok@kp}{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
-\@namedef{PY@tok@kt}{\def\PY@tc##1{\textcolor[rgb]{0.69,0.00,0.25}{##1}}}
-\@namedef{PY@tok@o}{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
-\@namedef{PY@tok@ow}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
-\@namedef{PY@tok@nb}{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
-\@namedef{PY@tok@nf}{\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
-\@namedef{PY@tok@nc}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
-\@namedef{PY@tok@nn}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
-\@namedef{PY@tok@ne}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.80,0.25,0.22}{##1}}}
-\@namedef{PY@tok@nv}{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
-\@namedef{PY@tok@no}{\def\PY@tc##1{\textcolor[rgb]{0.53,0.00,0.00}{##1}}}
-\@namedef{PY@tok@nl}{\def\PY@tc##1{\textcolor[rgb]{0.46,0.46,0.00}{##1}}}
-\@namedef{PY@tok@ni}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.44,0.44,0.44}{##1}}}
-\@namedef{PY@tok@na}{\def\PY@tc##1{\textcolor[rgb]{0.41,0.47,0.13}{##1}}}
-\@namedef{PY@tok@nt}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
-\@namedef{PY@tok@nd}{\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
-\@namedef{PY@tok@s}{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
-\@namedef{PY@tok@sd}{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
-\@namedef{PY@tok@si}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.64,0.35,0.47}{##1}}}
-\@namedef{PY@tok@se}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.67,0.36,0.12}{##1}}}
-\@namedef{PY@tok@sr}{\def\PY@tc##1{\textcolor[rgb]{0.64,0.35,0.47}{##1}}}
-\@namedef{PY@tok@ss}{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
-\@namedef{PY@tok@sx}{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
-\@namedef{PY@tok@m}{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
-\@namedef{PY@tok@gh}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
-\@namedef{PY@tok@gu}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.50,0.00,0.50}{##1}}}
-\@namedef{PY@tok@gd}{\def\PY@tc##1{\textcolor[rgb]{0.63,0.00,0.00}{##1}}}
-\@namedef{PY@tok@gi}{\def\PY@tc##1{\textcolor[rgb]{0.00,0.52,0.00}{##1}}}
-\@namedef{PY@tok@gr}{\def\PY@tc##1{\textcolor[rgb]{0.89,0.00,0.00}{##1}}}
-\@namedef{PY@tok@ge}{\let\PY@it=\textit}
-\@namedef{PY@tok@gs}{\let\PY@bf=\textbf}
-\@namedef{PY@tok@gp}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
-\@namedef{PY@tok@go}{\def\PY@tc##1{\textcolor[rgb]{0.44,0.44,0.44}{##1}}}
-\@namedef{PY@tok@gt}{\def\PY@tc##1{\textcolor[rgb]{0.00,0.27,0.87}{##1}}}
-\@namedef{PY@tok@err}{\def\PY@bc##1{{\setlength{\fboxsep}{\string -\fboxrule}\fcolorbox[rgb]{1.00,0.00,0.00}{1,1,1}{\strut ##1}}}}
-\@namedef{PY@tok@kc}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
-\@namedef{PY@tok@kd}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
-\@namedef{PY@tok@kn}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
-\@namedef{PY@tok@kr}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
-\@namedef{PY@tok@bp}{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
-\@namedef{PY@tok@fm}{\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
-\@namedef{PY@tok@vc}{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
-\@namedef{PY@tok@vg}{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
-\@namedef{PY@tok@vi}{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
-\@namedef{PY@tok@vm}{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
-\@namedef{PY@tok@sa}{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
-\@namedef{PY@tok@sb}{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
-\@namedef{PY@tok@sc}{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
-\@namedef{PY@tok@dl}{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
-\@namedef{PY@tok@s2}{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
-\@namedef{PY@tok@sh}{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
-\@namedef{PY@tok@s1}{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
-\@namedef{PY@tok@mb}{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
-\@namedef{PY@tok@mf}{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
-\@namedef{PY@tok@mh}{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
-\@namedef{PY@tok@mi}{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
-\@namedef{PY@tok@il}{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
-\@namedef{PY@tok@mo}{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
-\@namedef{PY@tok@ch}{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.24,0.48,0.48}{##1}}}
-\@namedef{PY@tok@cm}{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.24,0.48,0.48}{##1}}}
-\@namedef{PY@tok@cpf}{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.24,0.48,0.48}{##1}}}
-\@namedef{PY@tok@c1}{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.24,0.48,0.48}{##1}}}
-\@namedef{PY@tok@cs}{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.24,0.48,0.48}{##1}}}
-
-\def\PYZbs{\char`\\}
-\def\PYZus{\char`\_}
-\def\PYZob{\char`\{}
-\def\PYZcb{\char`\}}
-\def\PYZca{\char`\^}
-\def\PYZam{\char`\&}
-\def\PYZlt{\char`\<}
-\def\PYZgt{\char`\>}
-\def\PYZsh{\char`\#}
-\def\PYZpc{\char`\%}
-\def\PYZdl{\char`\$}
-\def\PYZhy{\char`\-}
-\def\PYZsq{\char`\'}
-\def\PYZdq{\char`\"}
-\def\PYZti{\char`\~}
-% for compatibility with earlier versions
-\def\PYZat{@}
-\def\PYZlb{[}
-\def\PYZrb{]}
-\makeatother
-
-    % Define a nice break command that doesn't care if a line doesn't already
-% exist.
-\def\br{\hspace*{\fill} \\* }
-
-% For linebreaks inside Verbatim environment from package fancyvrb.
-\makeatletter
-\newbox\Wrappedcontinuationbox
-\newbox\Wrappedvisiblespacebox
-\newcommand*\Wrappedvisiblespace {\textcolor{red}{\textvisiblespace}}
-\newcommand*\Wrappedcontinuationsymbol {\textcolor{red}{\llap{\tiny$\m@th\hookrightarrow$}}}
-\newcommand*\Wrappedcontinuationindent {3ex }
-\newcommand*\Wrappedafterbreak {\kern\Wrappedcontinuationindent\copy\Wrappedcontinuationbox}
-% Take advantage of the already applied Pygments mark-up to insert
-% potential linebreaks for TeX processing.
-%        {, <, #, %, $, ' and ": go to next line.
-%        _, }, ^, &, >, - and ~: stay at end of broken line.
-% Use of \textquotesingle for straight quote.
-\newcommand*\Wrappedbreaksatspecials {%
-	\def\PYGZus{\discretionary{\char`\_}{\Wrappedafterbreak}{\char`\_}}%
-	\def\PYGZob{\discretionary{}{\Wrappedafterbreak\char`\{}{\char`\{}}%
-	\def\PYGZcb{\discretionary{\char`\}}{\Wrappedafterbreak}{\char`\}}}%
-	\def\PYGZca{\discretionary{\char`\^}{\Wrappedafterbreak}{\char`\^}}%
-	\def\PYGZam{\discretionary{\char`\&}{\Wrappedafterbreak}{\char`\&}}%
-	\def\PYGZlt{\discretionary{}{\Wrappedafterbreak\char`\<}{\char`\<}}%
-	\def\PYGZgt{\discretionary{\char`\>}{\Wrappedafterbreak}{\char`\>}}%
-	\def\PYGZsh{\discretionary{}{\Wrappedafterbreak\char`\#}{\char`\#}}%
-	\def\PYGZpc{\discretionary{}{\Wrappedafterbreak\char`\%}{\char`\%}}%
-	\def\PYGZdl{\discretionary{}{\Wrappedafterbreak\char`\$}{\char`\$}}%
-	\def\PYGZhy{\discretionary{\char`\-}{\Wrappedafterbreak}{\char`\-}}%
-	\def\PYGZsq{\discretionary{}{\Wrappedafterbreak\textquotesingle}{\textquotesingle}}%
-	\def\PYGZdq{\discretionary{}{\Wrappedafterbreak\char`\"}{\char`\"}}%
-	\def\PYGZti{\discretionary{\char`\~}{\Wrappedafterbreak}{\char`\~}}%
-}
-% Some characters . , ; ? ! / are not pygmentized.
-% This macro makes them "active" and they will insert potential linebreaks
-\newcommand*\Wrappedbreaksatpunct {%
-	\lccode`\~`\.\lowercase{\def~}{\discretionary{\hbox{\char`\.}}{\Wrappedafterbreak}{\hbox{\char`\.}}}%
-	\lccode`\~`\,\lowercase{\def~}{\discretionary{\hbox{\char`\,}}{\Wrappedafterbreak}{\hbox{\char`\,}}}%
-	\lccode`\~`\;\lowercase{\def~}{\discretionary{\hbox{\char`\;}}{\Wrappedafterbreak}{\hbox{\char`\;}}}%
-	\lccode`\~`\:\lowercase{\def~}{\discretionary{\hbox{\char`\:}}{\Wrappedafterbreak}{\hbox{\char`\:}}}%
-	\lccode`\~`\?\lowercase{\def~}{\discretionary{\hbox{\char`\?}}{\Wrappedafterbreak}{\hbox{\char`\?}}}%
-	\lccode`\~`\!\lowercase{\def~}{\discretionary{\hbox{\char`\!}}{\Wrappedafterbreak}{\hbox{\char`\!}}}%
-	\lccode`\~`\/\lowercase{\def~}{\discretionary{\hbox{\char`\/}}{\Wrappedafterbreak}{\hbox{\char`\/}}}%
-	\catcode`\.\active
-	\catcode`\,\active
-	\catcode`\;\active
-	\catcode`\:\active
-	\catcode`\?\active
-	\catcode`\!\active
-	\catcode`\/\active
-	\lccode`\~`\~
-}
-\makeatother
-
-\let\OriginalVerbatim=\Verbatim
-\makeatletter
-\renewcommand{\Verbatim}[1][1]{%
-	%\parskip\z@skip
-	\sbox\Wrappedcontinuationbox {\Wrappedcontinuationsymbol}%
-	\sbox\Wrappedvisiblespacebox {\FV@SetupFont\Wrappedvisiblespace}%
-	\def\FancyVerbFormatLine ##1{\hsize\linewidth
-		\vtop{\raggedright\hyphenpenalty\z@\exhyphenpenalty\z@
-			\doublehyphendemerits\z@\finalhyphendemerits\z@
-			\strut ##1\strut}%
-	}%
-	% If the linebreak is at a space, the latter will be displayed as visible
-	% space at end of first line, and a continuation symbol starts next line.
-	% Stretch/shrink are however usually zero for typewriter font.
-	\def\FV@Space {%
-		\nobreak\hskip\z@ plus\fontdimen3\font minus\fontdimen4\font
-		\discretionary{\copy\Wrappedvisiblespacebox}{\Wrappedafterbreak}
-		{\kern\fontdimen2\font}%
-	}%
-
-	% Allow breaks at special characters using \PYG... macros.
-	\Wrappedbreaksatspecials
-	% Breaks at punctuation characters . , ; ? ! and / need catcode=\active
-	\OriginalVerbatim[#1,codes*=\Wrappedbreaksatpunct]%
-}
-\makeatother
-
-% Exact colors from NB
-\definecolor{incolor}{HTML}{303F9F}
-\definecolor{outcolor}{HTML}{D84315}
-\definecolor{cellborder}{HTML}{CFCFCF}
-\definecolor{cellbackground}{HTML}{F7F7F7}
-
-% prompt
-\makeatletter
-\newcommand{\boxspacing}{\kern\kvtcb@left@rule\kern\kvtcb@boxsep}
-\makeatother
-\newcommand{\prompt}[4]{
-	{\ttfamily\llap{{\color{#2}[#3]:\hspace{3pt}#4}}\vspace{-\baselineskip}}
-}
-
-% Prevent overflowing lines due to hard-to-break entities
-\sloppy
-% Setup hyperref package
-\hypersetup{
-	breaklinks=true,  % so long urls are correctly broken across lines
-	colorlinks=true,
-	urlcolor=urlcolor,
-	linkcolor=linkcolor,
-	citecolor=citecolor,
-}
-% Slightly bigger margins than the latex defaults
-
-\geometry{verbose,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}
+\usepackage[most]{tcolorbox}
+\usepackage{minted}
+\tcbuselibrary{minted} % tcolorbox minted library, required to use the "minted" tcb listing engine (this library is not loaded by the option [most])
+%\usepackage{minted} % Allows input of raw code, such as Python code
+\usepackage[colorlinks]{hyperref} % ALWAYS load this package LAST
 
@@
 \newcommand{\code}[1]{\texttt{#1}}
-
-\definecolor{LightGray}{gray}{0.9}
-\definecolor{codegreen}{rgb}{0,0.6,0}
-\definecolor{codegray}{rgb}{0.5,0.5,0.5}
-\definecolor{codepurple}{rgb}{0.58,0,0.82}
-\definecolor{backcolour}{rgb}{0.95,0.95,0.92}
-\lstdefinestyle{mystyle}{
-	backgroundcolor=\color{LightGray},   
-    commentstyle=\color{codegreen},
-	keywordstyle=\color{magenta},
-	numberstyle=\tiny\color{codegreen},
-	stringstyle=\color{codepurple},
-	basicstyle=\footnotesize\ttfamily,
-	breakatwhitespace=false,         
-	breaklines=true,                 
-	captionpos=b,                    
-	keepspaces=true,                 
-	numbers=left,                    
-	numbersep=5pt,                  
-	showspaces=false,                
-	showstringspaces=false,
-	showtabs=false,                  
-	tabsize=2
-}
-
-\lstset{style=mystyle}
 
@@
 % TODO - opening should really grab reader
+Spreadsheet based modeling has transformed the teaching of management science in business schools.
+The late 1990's saw the first editions of game changing textbooks \cite{ } focused on the use of spreadsheets in teaching modeling and management science.
+A series of conferences focused on teaching management science (TMS) with spreadsheets started at Dartmouth in 1998 and several presentations from that conference made up the bulk of papers in the first two issues of the INFORMS Transactions on Education journal \cite{bibid}.
+I attended TMS III at Indiana University and left with numerous ideas that influenced my own spreadsheet based modeling course.
+The next few decades saw numerous textbooks and teaching papers that prominently featured spreadsheet based modeling.
+
+Despite numerous calls for and predictions of the eradication of spreadsheets from the business world,
+
 Spreadsheet based modeling wave of the early 2000s. TMS conferences. Ground breaking textbooks such as Ragsdale and W\&A.
+The first two issues of ITE are paper based on TMS 1 at Dartmouth in 1998. I attended TMS 3 in Bloomington in 2000.
+
+First edition of Ragsdale was 1995 PMS was 1997.
+
+While spreadsheets have certainly changed and expanded in their capabilities (PowerPivot, um...) since the 90's...
+
+Irony of Mgt Sci crisis of 1990's where course was dropped by AACSB and ignored by other business disciplines to an era where analytics has captured the day and now AI threatens even that.
+
+Powell's 2001 paper on the importance of modeling is no less relevant now than it was when published. I still many of his ideas in my spreadsheet based modeling course.
 
@@
 
-% TODO - how should pandas appear in text
 Python has continued to make inroads to analytical territory that was formerly dominated by Excel.
@@
 
-The stage is set with an overview of a typical spreadsheet based model that most of the students have seen before.
+The stage is set with an overview of a typical spreadsheet based model that most of the students have seen before. See Figure \ref{fig:xl_model} for a screenshot of the Excel model.
 
-\begin{quote}
-	For example, here's a high level screenshot of a model that I assign for homework in my [MIS 4460/5460 Business Analytics class](http://www.sba.oakland.edu/faculty/isken/courses/mis5460/) (a spreadsheet based modeling class).
+\begin{tcolorbox}
+\begin{quotation}
+\textsf{	
 	It's a really simple model in which we are selling a single product that we produce.
@@
 We can sell the product for some price and we believe that demand for the product is related to the selling price through a power function.
-Let's assume for now that we have sufficient capacity to produce to demand and that all inputs are deterministic (we'll deal with simulating uncertainty later in this document).
+Let's assume for now that we have sufficient capacity to produce to demand and that all inputs are deterministic (we'll deal with simulating uncertainty later in this document).}
 
-	The details aren't so important right now as is the overall structure of the model.
+	\textsf{The details aren't so important right now as is the overall structure of the model.
 There's a few key inputs and some pretty straightforward formulas for computing cost, revenue and profit.
@@
 How might we go about building a similar model using Python?
+}
 
-	What if we wanted to push it a little further and model some of the key inputs with probability distributions to reflect our uncertainty about their values?
+	\textsf{What if we wanted to push it a little further and model some of the key inputs with probability distributions to reflect our uncertainty about their values?
 In the Excel world, we might use add-ins such as @Risk which allow uncertain quantities to be directly modeled with probability distributions.
@@
 profit) and use statistics to summarize these outputs using things like histograms and summary stats.
-Often this type of simulation model is referred to as a *Monte-Carlo* model to suggest repeated sampling from one or more probability distributions within an otherwise pretty static model.
-If you want to see such models in action, check out my [Simulation Modeling with Excel page](http://www.sba.oakland.edu/faculty/isken/courses/mis5460/simulation.html) from my Business Analytics course.
-Again, how might we do this with Python?
-\end{quote}
+Often this type of simulation model is referred to as a Monte-Carlo model to suggest repeated sampling from one or more probability distributions within an otherwise pretty static model.
+Again, how might we do this with Python?}
+\end{quotation}
+\end{tcolorbox}
 
@@
 
-With this as motivation, we then introduce a similar modeling problem.
-The example used throughout the rest of the notebook is based on one in the [spreadsheet modeling textbooks I've used in my classes since 2001](https://host.kelley.iu.edu/albrightbooks/).
-In both books, they introduce the "Walton Bookstore" problem in the chapter on Monte-Carlo simulation.
-Here is the basic problem (with a few modifications):
+We then introduce a similar modeling problem to use throughout the rest of the notebook.
+The Walton Bookstore problem appears in the chapter on Monte-Carlo simulation in textbooks that I have used in my spreadsheet modeling course \cite{albrightBusinessAnalyticsData2016,winstonPracticalManagementScience2018}.
+The (slightly modified) problem is as follows:
 
@@
 
+The goal is to build a Python based version of this model that will allow us to do sensitivity analysis, find the break even point for demand and analyze the problem using Monte-Carlo simulation.
+
 \subsubsection{Build first model - procedural approach}
 
-We start with a very simple model that ignores the uncertainty in demand and that is not object oriented. Proceeding much like we would in Excel, we store base input values in variables.
+We start with a very simple model that ignores the uncertainty in demand and that is not object oriented. Proceeding much like we would in Excel, base input values are stored in variables.
 
-\begin{lstlisting}[language=Python]
+\begin{tcolorbox}
+\begin{minted}[gobble=0]{python}
 # Base inputs
@@
 order_quantity = 200
-\end{lstlisting}
-
-\begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
-	%\prompt{In}{incolor}{3}{\boxspacing}
-	\begin{Verbatim}[commandchars=\\\{\}]
-	\PY{c+c1}{\PYZsh{} Base inputs}
-	\PY{n}{unit\PYZus{}cost} \PY{o}{=} \PY{l+m+mf}{7.50}
-	\PY{n}{selling\PYZus{}price} \PY{o}{=} \PY{l+m+mf}{10.00}
-	\PY{n}{unit\PYZus{}refund} \PY{o}{=} \PY{l+m+mf}{2.50}
-	\end{Verbatim}
+\end{minted}
 \end{tcolorbox}
@@
 
-\begin{lstlisting}[language=Python]
+\begin{tcolorbox}
+\begin{minted}[gobble=0]{python}
 order_cost = unit_cost * order_quantity
@@
 profit = sales_revenue + refund_revenue - order_cost
-\end{lstlisting}
+\end{minted}
+\end{tcolorbox}
 
@@
 In a Data Table, a range of order quantities can be used as row or column input and one or more output values can be computed.
-In Python, we can rely on the vectorized behavior of the language \cite{bibid}.
+In Python, the vectorized behavior of the language \cite{bibid} provides a simple way to accomplish the same thing.
 A range of order quantities is created as a NumPy array that can be used directly to compute vectors of all of the intermediate and final output variables.
@@
 A natural next step is to do the equivalent of an Excel 2-way Data Table.
-By creating a profit function that takes all of the base inputs as arguments, we can construct a list comprehension that does what an Excel 2-Way Data Table does.
-Even better, this actually allows us to do the equivalent of an n-Way Data Table with an arbitrary number of outputs - something Excel does not have.
+By creating a profit function that takes all of the base inputs as arguments, a list comprehension can then be used to generate the equivalent of a 2-way Data Table.
+Even better, this actually allows us to do the equivalent of an n-way Data Table with an arbitrary number of outputs -- something Excel does not have.
 This example provides an effective way to show the power of creating reusable functions and Python's handy list comprehension construct.
 
-\begin{lstlisting}[language=Python]
-	def bookstore_profit(unit_cost, selling_price, unit_refund, order_quantity, demand):
+This code chunk,
+
+\begin{tcolorbox}[left=0mm]
+\begin{minted}[gobble=1]{python}
+	def bookstore_profit(unit_cost, selling_price, 
+	                     unit_refund, order_quantity, demand):
 	'''
@@
 	# Create data table (as a list of tuples)
-	data_table_1 = [(d, oq, bookstore_profit(unit_cost, selling_price, unit_refund, oq, d)) 
+	data_table_1 = [(d, oq, bookstore_profit(
+	                 unit_cost, selling_price, unit_refund, oq, d)) 
 	for d in demand_range for oq in order_quantity_range]
@@
 	# Convert to dataframe
-	dtbl_1_df = pd.DataFrame(data_table_1, columns=['Demand', 'OrderQuantity', 'Profit'])
+	dtbl_1_df = pd.DataFrame(data_table_1, 
+	                         columns=['Demand', 'OrderQuantity', 'Profit'])
 
-\end{lstlisting}
+\end{minted}
+\end{tcolorbox}
+
+results in the following output:
 
@@
 
+\begin{tcolorbox}
+	\begin{verbatim}
+	   Demand  OrderQuantity  Profit
+	      50             50   125.0
+	      50             75   -37.5
+	      50            100  -200.0
+	      50            125  -362.5
+	      50            150  -525.0
+	      ...
+	     300            275   687.5
+	     300            300   750.0
+\end{verbatim}	
+\end{tcolorbox}
+
 \subsubsection{The object nature of Python}
 
-It is not clear whether an object oriented (OO) or a non-OO approach will end up making the most sense for doing spreadsheet type modeling.
+It is not clear whether this non-object oriented approach will end up making the most sense for doing spreadsheet type modeling in Python.
 This uncertainty is made explicit in the notebook and students are led on a bit of a journey of discovery as we explore different software designs for this modeling problem.
@@
 For our initial design of a \code{BookstoreModel} class we make all of the base inputs class attributes (properties) and then add method attributes to compute outputs such as costs, revenues and profits.
-Students can easily see the mapping between things in the initial procedural model and this new object oriented version of the model.
-I highlight several important concepts and syntactical details of doing object oriented programming Python.
+Students can easily see the mapping between things in the initial procedural model and this new object oriented version.
+Several important concepts and syntactical details of doing object oriented programming Python are highlighted.
 Again, the goal is to weave in these more advanced Python programming concepts within the context of a very familiar modeling problem.
-This section of the notebook ends with us using our \code{BookstoreModel} class to create a model object with all of its base inputs instantiated with values.
-Before moving on students are presented with a few challenges involving enhancements to the \code{BookstoreModel} class.
+This section of the notebook ends with using the \code{BookstoreModel} class to create a new model object instance with all of its base inputs instantiated with values.
+Before moving, on students are presented with a few challenges involving enhancements to the \code{BookstoreModel} class.
 Interspersing short coding challenges throughout the notebook gives students a chance to test themselves on their understanding and to get some coding practice.
@@
 These meta reflections are intentionally included to better mimic actual software development and reinforce the idea that software development is much more than simply writing code to implement a perfectly thought out design.
-We forge ahead with some preliminary ideas and are confronted with an intermediate problem related to generating a list of dictionaries that represent combinations of inputs (i.e.\ scenarios) to evaluate by our n-way data table function.
-This provides a great opportunity to remind students of one of the great strengths of the open source software ecosystem - we can leverage work done in other packages and even look at actual source code to see how something was implemented.
+We forge ahead with some preliminary ideas and are confronted with an intermediate problem related to generating a list of dictionaries that represent combinations of inputs (i.e.\ scenarios) to evaluate.
+This provides a great opportunity to remind students of one of the great strengths of the open source software ecosystem -- we can leverage work done in other packages and even look at actual source code to see how something was implemented.
 In this case, the well known scikit-learn package \cite{bibid} has a \code{ParameterGrid} function that solves our scenario generation problem quite nicely.
-This example also reminds the students of the value of being able to make sense of API documentation and of object oriented code written by others as most data science related Python packages are object oriented in nature.
+This example also reminds the students of the value of being able to make sense of API documentation and of object oriented code written by others.
 
@@
 
-No self respecting modeling tool would be complete without goal seeking capability to do things such as finding the break even point in the bookstore model.
+No self respecting modeling tool would be complete without goal seeking capability to do things such as finding the break even demand point in the bookstore model.
 For many students, Excel's Goal Seek tool can seem almost magical when encountered for the first time.
@@
 Then we can do standard statistical analysis of the simulation output using Python packages such as pandas and SciPy.
-In spreadsheet based simulation textbooks, the notion of the ``Flaw of Averages'' popularized by Sam Savage \cite{} is illustrated.
+In spreadsheet based simulation textbooks, the notion of the ``Flaw of Averages'' popularized by Sam Savage \cite{} is often illustrated.
 We do the same with our Python based model, showing how replacing random demand by its mean can result in a wildly optimistic estimate of the mean profit found through analyzing the simulation output.
@@
 This is followed by a well commented first attempt at a \code{simulate} function.
-By this point, students are well equipped to understand the code as it builds on basic Python concepts and more advanced things that we have already covered in this EwP module.
+By this point, students are well equipped to understand the code as it builds on basic Python concepts and more advanced concepts that we have already covered in this EwP module.
 The rest of the notebook is spent trying out the \code{simulate} function and ends with creating a grouped boxplot that many of them have likely created before with @Risk along with a faceted plot of histograms of profit created with the popular Seaborn package.
@@
 Now we shift gears and move beyond what might be considered prototyping and into the realm of deployment.
+The notebook ends with these thoughts.
+
+\begin{tcolorbox}
+	\begin{quotation}
+
+\textsf{We have added a basic `simulate` function to our `data_table` and `goal_seek` functions. Python is proving to be quite nice for doing Excel-style "what if?" analysis.}
+
+\textsf{In Part 4 of this series, we'll make some improvements and do some clean-up on our classes and functions.
+We'll move everything into a single \code{whatif.py} module and learn how to create a Python package to make it easy to use and share our new functions.
+We'll try out our package on a new model and sketch out some ideas for future enhancements to the package.
+It's important we also start creating some basic documentation and a user guide.}
+\end{quotation}
+\end{tcolorbox}
 
@@
 
-\begin{quote}
-
-In the first three notebooks, we've developed some Python approaches to
-typical Excel ``what if?'' analyses. Along the way we explored some
-slightly more advanced Python topics (for relative newcomers to Python)
-such as:
-
-\begin{itemize}
-	\item
-	List comprehensions,
-	\item
-	Basic OO programming, creating our own classes and using things like
-	\texttt{setattr} and \texttt{getattr},
-	\item
-	Leveraging scikit-learn's \texttt{ParameterGrid} class,
-	\item
-	Faceted plots using Seaborn and matplotlib,
-	\item
-	Tuple unpacking, zip, and itertools,
-	\item
-	Safely copying objects,
-	\item
-	Root finding with, and without, \texttt{scipy.optimize},
-	\item
-	Using \texttt{numpy.random} to generate random variates from various
-	probability distributions,
-	\item
-	Using \texttt{scipy.stats} to compute probabilities and percentiles,
-	\item
-	Importing our \texttt{data\_table}, \texttt{goal\_seek} and
-	\texttt{simulate} functions from a module.
-\end{itemize}
-
-Now that we've got a critical mass of ``proof of concept'' code, let's figure out how to structure our project and create a deployable package.
-In addition, let's rethink our OO design and add some much needed documentation to the code.
-\end{quote}
-
-Creating a deployable Python package is a good exercise in designing and creating reusable software artifacts.
+Creating a deployable Python package is a good exercise in designing and creating a reusable software artifact.
 Thinking about how others will interact with your software usually leads to better designed software.
-Basic software engineering practices are often overlooked in business analytics programs but are certainly desirable in the professional world.
+Basic software engineering practices are often overlooked in business analytics classes but are certainly desirable in the professional world.
 In this submodule we create a Python package that can be imported and used by other modelers.
@@
 The basics of version control were already covered in the first week of the course.
-My experience is that most students have little exposure to simple software project management concepts such as project folder structures and version control.
+My experience is that most students have little exposure to software project management concepts such as project folder structures and version control.
 This can lead to a jumble of files with incoherent names used as a proxy for version control - \code{whatif\_v1.py}, \code{whatif\_v2.py}, \code{whatif\_final.py}, \code{whatif\_finalfinal.py}, and so on.
 
-Key software design changes are then made and discussed which will make our code easier on our users.
+Key software design changes are then made which will make our code easier on our users.
 These changes include creating an abstract \code{Model} base class from which different model classes can be created and moving the \code{data\_table}, \code{goal\_seek}, and \code{simulate} functions into the \code{Model} base class.
 All of the code is moved into a code module named \code{whatif.py}.
-We are now ready to learn about the basics of turning our code into an installable package.
+Students are now ready to learn about the basics of turning our code into an installable package.
 This requires a basic understanding of two key Python files that are ubiquitous in Python pacakges - \code{\_\_init\_\_.py} and \code{setup.py}.
@@
 The example used is the New Car Simulation that has been a staple of Winston and Albright's textbooks for many years \cite{bibid}.
+This new model differs structurally from the Bookstore Model, but is representative of many financial models, in that cash flows over multiple years need to be computed and summarized.
+It provides a good test for our new whatif package.
 
@@
 This includes things like code comments, docstrings, readme files and generating documentation by writing restrutured text files and using Sphinx.
-This is one of those tasks that few people like to do but can be very important for the long term success of a project.
+This is one of those tasks that few people like to do but can be very important for the long term success of a project \cite{bibid}.
 Even if you are just creating a tool for yourself, you will forget the details over time and documentation can be very beneficial in refreshing your memory.
@@
 
-The EwP module ends with what has become a relatively common use of Python with Excel - automating various Excel data wrangling tasks. Examples include:
+The EwP module ends with what has become a relatively common use of Python with Excel -- automating various Excel data wrangling tasks. Examples include:
 
@@
 A single Jupyter notebook is used to give a taste of using Python to automate the process of working with Excel files by tackling each of the four examples above.
-Each of these examples are based on real problems I encountered either in research or industrial projects.
-TRANSITION NEEDED Several web based resources are shared with one particularly good one being the Practical Business Python blog written by Chris Moffitt.
-It includes numerous posts on different aspects of Python and Excel integration.
-There are a number of Python based tools that either include functionality for working with Excel files or are dedicated to specific Excel related operations.
-These include:
+Each of these examples are based on a real problems I encountered either in research or industrial projects.
+Several web based resources are shared including the Practical Business Python blog \cite{bibid} written by Chris Moffitt, a data analytics professional.
+This blog, which focuses on helping Excel centric analytics professionals transition some of their work to Python, includes numerous posts on different aspects of Python and Excel integration.
+
+There are a number of Python based tools that either include functionality for working with Excel files or are dedicated to specific Excel related operations. These include:
 
@@
 The first part of the assignment involves creating a basic model, similar in complexity to the Bookstore Model, and doing sensitivity analysis, goal seeking and simulation with the model.
-It requires the students to import and use the \code{whatif.py} package we created as part of the EwP module.
-The second part of the assignment is based on the last submodule and requires the students to complete a task involving creation of a multi-sheet Excel workbook from a set of csv files using Python.
+It requires the students to import and use the \code{whatif.py} package we created as part of the module.
+The second part of the assignment is based on the last submodule and requires  students to complete a task involving creation of a multi-sheet Excel workbook from a set of csv files using Python.
 Then, using the openpyxl library, they have to create and add formulas to the workbook and format them properly.
-The ranges used in the formulas vary by sheet and require the students to determine the number of rows in each sheet of data and use that information to construct the appropriate formula.
+The ranges used in the formulas vary by sheet and require determination of the number of rows in each sheet of data and use that information to construct the appropriate formula.
 An example assignment is available from \href{}{here}.
@@
 
-Rumours have swirled for a number of years that Microsoft was considering adding Python as a first class language alternative to VBA with their MS Office suite of packages \cite{bibid}.
+Rumors have swirled for a number of years that Microsoft was considering adding Python as a first class language alternative to VBA with their MS Office suite of packages \cite{bibid}.
 While this has not yet materialized, there continues to be significant interest in using Python to manipulate not only Excel workbooks but other office documents such as Powerpoint presentations \cite{bibid}.

### 2023-06-20 test_minted.tex
found strategy for code samples using tcolorbox and minted

--- 2023-06-14 python_excel_ite.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Author style for INFORMS Transactions on Education (ited)
-%% Mirko Janc, Ph.D., INFORMS, pubtech@informs.org
-%% ver. 0.92, June 2009 -- default options: single-spaced, double-blinded
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\documentclass[ited]{informs3}                      % for a regular run
-%\documentclass[ited,nonblindrev]{informs3}          % for review, not blinded
-%\documentclass[ited,blindrev,copyedit]{informs3}    % spaced for copyediting
-\documentclass[ited,blindrev]{informs3}              % for review, blinded
+\documentclass{article}              % for review, blinded
 
-\usepackage{hyperref}
-\usepackage{listings}
-\usepackage{xcolor}
+\usepackage[most]{tcolorbox}
+\usepackage{minted}
+\tcbuselibrary{minted} % tcolorbox minted library, required to use the "minted" tcb listing engine (this library is not loaded by the option [most])
+%\usepackage{minted} % Allows input of raw code, such as Python code
 
-% If hyperref is used, dvi-to-ps driver of choice must be declared as
-%   an additional option to the \documentstyle. For example
-%\documentclass[dvips,ited]{informs1}      % if dvips is used 
-%\documentclass[dvipsone,ited]{informs1}   % if dvipsone is used, etc.
+\begin{document}
+	\begin{tcolorbox}
+	\begin{minted}[gobble=1]{python}
+	print("hello, world")
+	\end{minted}
+\end{tcolorbox}
 
-% Private macros here (check that there is no clash with the style)
-\newcommand{\code}[1]{\texttt{#1}}
-
-\definecolor{LightGray}{gray}{0.9}
-\definecolor{codegreen}{rgb}{0,0.6,0}
-\definecolor{codegray}{rgb}{0.5,0.5,0.5}
-\definecolor{codepurple}{rgb}{0.58,0,0.82}
-\definecolor{backcolour}{rgb}{0.95,0.95,0.92}
-\lstdefinestyle{mystyle}{
-	backgroundcolor=\color{LightGray},   
-    commentstyle=\color{codegreen},
-	keywordstyle=\color{magenta},
-	numberstyle=\tiny\color{codegreen},
-	stringstyle=\color{codepurple},
-	basicstyle=\footnotesize\ttfamily,
-	breakatwhitespace=false,         
-	breaklines=true,                 
-	captionpos=b,                    
-	keepspaces=true,                 
-	numbers=left,                    
-	numbersep=5pt,                  
-	showspaces=false,                
-	showstringspaces=false,
-	showtabs=false,                  
-	tabsize=2
-}
-
-\lstset{style=mystyle}
-
-% Natbib setup for author-year style
-\usepackage{natbib}
- \bibpunct[, ]{(}{)}{,}{a}{}{,}%
- \def\bibfont{\small}%
- \def\bibsep{\smallskipamount}%
- \def\bibhang{24pt}%
- \def\newblock{\ }%
- \def\BIBand{and}%
-
-%% Setup of theorem styles. Outcomment only one. 
-%% Preferred default is the first option.
-\TheoremsNumberedThrough     % Preferred (Theorem 1, Lemma 1, Theorem 2)
-%\TheoremsNumberedByChapter  % (Theorem 1.1, Lema 1.1, Theorem 1.2)
-
-%% Setup of the equation numbering system. Outcomment only one.
-%% Preferred default is the first option.
-\EquationsNumberedThrough    % Default: (1), (2), ...
-%\EquationsNumberedBySection % (1.1), (1.2), ...
-
-% In the reviewing and copyediting stage enter the manuscript number.
-%\MANUSCRIPTNO{} % When the article is logged in and DOI assigned to it,
-                 %   this manuscript number is no longer necessary
-
-%%%%%%%%%%%%%%%%
-\begin{document}
-%%%%%%%%%%%%%%%%
-
-% Outcomment only when entries are known. Otherwise leave as is and 
-%   default values will be used.
-%\setcounter{page}{1}
-%\VOLUME{00}%
-%\NO{0}%
-%\MONTH{Xxxxx}% (month or a similar seasonal id)
-%\YEAR{0000}% e.g., 2005
-%\FIRSTPAGE{000}%
-%\LASTPAGE{000}%
-%\SHORTYEAR{00}% shortened year (two-digit)
-%\ISSUE{0000} %
-%\LONGFIRSTPAGE{0001} %
-%\DOI{10.1287/xxxx.0000.0000}%
-
-% Author's names for the running heads
-% Sample depending on the number of authors;
-% \RUNAUTHOR{Jones}
-% \RUNAUTHOR{Jones and Wilson}
-% \RUNAUTHOR{Jones, Miller, and Wilson}
-% \RUNAUTHOR{Jones et al.} % for four or more authors
-% Enter authors following the given pattern:
-\RUNAUTHOR{Isken}
-
-% Title or shortened title suitable for running heads. Sample:
-% \RUNTITLE{Bundling Information Goods of Decreasing Value}
-% Enter the (shortened) title:
-\RUNTITLE{Spreadsheet modeling with Python}
-
-% Full title. Sample:
-% \TITLE{Bundling Information Goods of Decreasing Value}
-% Enter the full title:
-\TITLE{Spreadsheet modeling with Python}
-
-% Block of authors and their affiliations starts here:
-% NOTE: Authors with same affiliation, if the order of authors allows, 
-%   should be entered in ONE field, separated by a comma. 
-%   \EMAIL field can be repeated if more than one author
-\ARTICLEAUTHORS{%
-\AUTHOR{Mark W. Isken}
-\AFF{Oakland University, \EMAIL{isken@oakland.edu}, \URL{http://www.sba.oakland.edu/faculty/isken/}}
-% Enter all authors
-} % end of the block
-
-\ABSTRACT{%
-TODO: Text of your abstract % Enter your abstract
-}%
-
-% Sample
-%\KEYWORDS{deterministic inventory theory; infinite linear programming duality; 
-%  existence of optimal policies; semi-Markov decision process; cyclic schedule}
-
-% Fill in data. If unknown, outcomment the field
-\KEYWORDS{python, spreadsheet modeling}
-\HISTORY{}
-
-\maketitle
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% Samples of sectioning (and labeling) in ITED
-% NOTE: (1) \section and \subsection do NOT end with a period
-%       (2) \subsubsection and lower need end punctuation
-%       (3) capitalization is as shown (title style).
-%
-%\section{Introduction.}\label{intro} %%1.
-%\subsection{Duality and the Classical EOQ Problem.}\label{class-EOQ} %% 1.1.
-%\subsection{Outline.}\label{outline1} %% 1.2.
-%\subsubsection{Cyclic Schedules for the General Deterministic SMDP.}
-%  \label{cyclic-schedules} %% 1.2.1
-%\section{Problem Description.}\label{problemdescription} %% 2.
-
-% Text of your paper here
-
-\section{Introduction}
-
-% The opening part of the what-if notebook has some language for the motivation behind this module.
-% TODO - opening should really grab reader
-Spreadsheet based modeling wave of the early 2000s. TMS conferences. Ground breaking textbooks such as Ragsdale and W\&A.
-
-Advent of programmatic analytics using tools like R and Python have changed the teaching landscape.
-Motivated by reproducibility, open source, free, and ease of integration with web and cloud databases.
-
-Created PCDA course in response.
-Challenges with finding good textbooks.
-Developed much of my own curricular materials.
-Taught using combo of R Markdown documents and Jupyter notebooks.
-Literate programming.
-PCDA course focused mostly on data wrangling, EDA, statistical and machine learning based predictive models though I did include a Python based simulation of the Monte Hall 3-door problem.
-In teaching EDA using tools like ggplot2, I would challenge students to think about how they would go about trying to do some of the plots we were making in Excel.
-There is also some limited coverage of using Python to open Excel files and do some sheet manipulation.
-However, there was no real attempt to do the kind of modeling that was the focus of textbooks like PMS or DADM.
-
-% TODO - how should pandas appear in text
-Python has continued to make inroads to analytical territory that was formerly dominated by Excel.
-The widely used Python data analysis library, Pandas, was created by Wes McKinney while working as a quantitative financial analyst.
-However, Python was mainly used for data wrangling, visualization and predictive modeling using various flavors of regression and tree based models.
-The blog \href{https://pbpython.com/}{Practical Business Python} written primarily by Chris Moffitt contributed greatly to making Python more accessible to business analytics practitioners through high quality posts that were extremely practical and useful.
-But, it got me to thinking.
-What about those things for which Excel is well suited such as building formula based models and doing sensitivity analysis on these models?
-What would those look like in Python?
-
-Business schools with business analytics programs are the context. Grad and undergrad.
-
-Why I created
-
-- more topics I wanted to cover than were possible in PCDA course
-- weave in more SE type material which is often overlooked in BA programs
-
-Excel is widely used for building and using models of business problems
-to explore the impact of various model inputs on key outputs. Built in
-``what if analysis'' tools such as \href{https://support.microsoft.com/en-us/office/calculate-multiple-results-by-using-a-data-table-e95e2487-6ca6-4413-ad12-77542a5ea50b}{Data
-	Tables} and
-\href{https://support.microsoft.com/en-us/office/use-goal-seek-to-find-the-result-you-want-by-adjusting-an-input-value-320cb99e-f4a4-417f-b1c3-4f369d6e66c7}{Goal
-	Seek} are well known to power spreadsheet modelers. How might we do
-similar modeling and analysis using Python?
-
-Limitations of spreadsheets such as reproducibility and automation.
-
-While Python has been gaining momentum in the business analytics world, it is often used for data wrangling, analysis and visualization of tablular data using tools like pandas and matplotlib or Seaborn.
-You can find some great examples at [Chris Moffit's Practical Business Python blog](https://pbpython.com/).
-I use Python all the time for such tasks and teach a course called [Practical Computing for Data Analytics](http://www.sba.oakland.edu/faculty/isken/courses/mis5470/) that is Python (and R) based.
-But, it got me to thinking.
-What about those things for which Excel is well suited such as building formula based models and doing sensitivity analysis on these models?
-What would those look like in Python?
-
-Quant Econ example
-
-PBB blog
-
-Teach relevant Python to business students using familiar and relevant modeling examples.
-Give sense of software development process and design decisions
-Model a learning journey
-Teach students to think and act like a software developer, not just an analyst
-
-Created the ``Excel with Python'' (EwP) module as part
-
-Integrated or separate lit review section?
-
-\section{Module positioning and structure}
-
-At our institution, The EwP module is part of a business course entitled ``Advanced Analytics with Python'' (\href{http://www.sba.oakland.edu/faculty/isken/courses/mis6900}{AAP}).
-The students taking this course have already taken my ``Practical Computing for Business Analytics'' (\href{http://www.sba.oakland.edu/faculty/isken/courses/mis5470}{PCDA}) course in which they spend about half the semester learning R and the other half learning Python, both in the context of business analytics.
-Basic Linux skills are also part of the PCDA course.
-The PCDA course ends with intro to the sklearn library for doing predictive modeling in Python and the AAP course begins where the PCDA course ends.
-The EwP module could certainly be used as part of a semester long Python based analytics course.
-The students have also already taken "Business Analytics" (\href{http://www.sba.oakland.edu/faculty/isken/courses/mis5460}{BA}) which is a spreadsheet based introduction to business analytics.
-The BA and PCDA courses have historically been offered in face to face mode in a computer teaching lab.
-Since 2020, these two courses have been also offered as online, asynchronous courses.
-The newer AAP course has been offered since 2021 in an online asynchronous mode.
-All three courses have extensive course websites that are publicly accessible, including access to all of the video content and files needed.
-We will describe the AAP course web in more detail below in the context of the EwP module.
-
-What kind of Python background do students need to have in order to be ready for the EwP module? A basic familiarity with Python fundamentals for data analysis is needed, including:
-
-\begin{itemize}
-	\item variables, arithmetic and boolean operators, and basic data types,
-	\item data structures such as lists, dictionaries, tuples, NumPy arrays and pandas dataframes,
-	\item flow control such as branching with \texttt{if \ldots else} and looping with \texttt{for} and \texttt{while}
-	\item module imports,
-	\item using built in functions and accessing methods and properties of objects,
-	\item creating functions,
-	\item basic use of NumPy and pandas for data wrangling and analysis,
-	\item basic plotting with matplotlib.
-\end{itemize}
-
-The EwP module is made of three submodules.
-The first of these focuses on building and using a typical spreadsheet model using Python.
-Then the functionality developed for doing this type of modeling in Python is deployed as a reusable package.
-We end with an introduction to using Python to manipulate Excel files, an extremely common use case.
-
-All of the Python modules in the PCDA course as well as the entire AAP course are taught using a combination of Jupyter notebooks, Python scripts, and videos walking through the notebooks, all publicly accessible from extensive course web pages.
-GIVE URL.
-Some background on Jupyter notebooks for analytics - see downloaded papers.
-Questions and coding challenges sprinkled throughout the notebooks - answers at bottom.
-While AAP is online asynch, this same approach has been used very successfully in a computer teaching lab for the PCDA course.
-The screencasts are highly valued by the students as they can rewatch things as needed.
-Many are working professionals and a 2.5 hour evening class in a warm computer lab can be taxing.
-
-\section{Submodule 1: What-if analysis with Python}
-
-There are two overarching learning objectives for this first set of Jupyter notebooks.
-Students learn how Python might be used to do something they are quite familiar with doing in a spreadsheet - model building and sensitivity analysis.
-At the same time, students encounter, as they are needed, more advanced Python programming techniques such as object oriented programming and manipulation of tuples, lists and dictionaries.
-In this way, the computational problem to be solved drives the introduction of more advanced Python concepts and techniques instead of such things being presented in a vacuum.
-I make it a point to try to convey the notion that there are often multiple possible approaches by embarking on one path and there can be value to starting over and doing things in a different way.
-Each notebook is accompanied by videos that walk the students through the material and act as a guided tour with challenges to overcome along the way.
-The screenshot in \ref{fig:activities} shows the portion of the course web page that provides the high level roadmap through the notebooks and associated screencasts.
-
-\subsection{Notebook 1: Modeling and data tables}
-
-The stage is set with an overview of a typical spreadsheet based model.
-
-\begin{quote}
-	For example, here's a high level screenshot of a model that I assign for homework in my [MIS 4460/5460 Business Analytics class](http://www.sba.oakland.edu/faculty/isken/courses/mis5460/) (a spreadsheet based modeling class).
-	It's a really simple model in which we are selling a single product that we produce.
-There is a fixed cost to producing the product as well as a variable production cost per unit.
-We can sell the product for some price and we believe that demand for the product is related to the selling price through a power function.
-Let's assume for now that we have sufficient capacity to produce to demand and that all inputs are deterministic (we'll deal with simulating uncertainty later in this document).
-
-	The details aren't so important right now as is the overall structure of the model.
-There's a few key inputs and some pretty straightforward formulas for computing cost, revenue and profit.
-Notice the 1-way Data Table being used to explore how profit varies for different selling prices.
-There's a graph driven by the Data Table and some Text Boxes used for annotation and summary interpretative comments.
-There's a button that launches Goal Seek to find the break even selling price and a 2-way Data Table (not shown) to explore the joint effect of selling price and variable cost.
-Classic Excel modeling stuff.
-How might we go about building a similar model using Python?
-
-	What if we wanted to push it a little further and model some of the key inputs with probability distributions to reflect our uncertainty about their values?
-In the Excel world, we might use add-ins such as @Risk which allow uncertain quantities to be directly modeled with probability distributions.
-For example, we might have a key input such as the exponent in the power function that relates selling price to demand that is highly uncertain.
-By modeling it with a probability distribution and then sampling from that distribution many times (essentially by recalcing the spreadsheet) we can generate a bunch of possible values for key outputs (e.g.
-profit) and use statistics to summarize these outputs using things like histograms and summary stats.
-Often this type of simulation model is referred to as a *Monte-Carlo* model to suggest repeated sampling from one or more probability distributions within an otherwise pretty static model.
-If you want to see such models in action, check out my [Simulation Modeling with Excel page](http://www.sba.oakland.edu/faculty/isken/courses/mis5460/simulation.html) from my Business Analytics course.
-Again, how might we do this with Python?
-\end{quote}
-
-\begin{figure}[!ht]
-	\centering
-	\begin{minipage}{.80\textwidth}
-		\includegraphics[width=1.0\textwidth]{images/output_2_0.png}
-	\end{minipage}
-	\caption{Typical spreadsheet model}
-	\label{fig:xl_model}
-\end{figure}
-
-The example used throughout the rest of the notebook is based on one in the [spreadsheet modeling textbooks I've used in my classes since 2001](https://host.kelley.iu.edu/albrightbooks/).
-In both books, they introduce the "Walton Bookstore" problem in the chapter on Monte-Carlo simulation.
-Here is the basic problem (with a few modifications):
-
-\begin{itemize}
-	\item we have to place an order for a perishable product (e.g. a calendar),
-	\item there is a known unit cost for each one ordered,
-	\item we have a known selling price,
-	\item demand is uncertain but we can model it with some simple probability distribution,
-	\item for each unsold item, we can get a partial refund of our unit cost,
-	\item we need to select the order quantity for our one order for the year; orders can only be in multiples of 25.
-\end{itemize}
-
-\subsubsection{Build first model - procedural approach}
-
-We start with a very simple model that ignores the uncertainty in demand and that is not object oriented. Proceeding much like we would in Excel, we store base input values in variables.
-
-\begin{lstlisting}[language=Python]
-# Base inputs
-unit_cost = 7.50
-selling_price = 10.00
-unit_refund = 2.50
-
-# Demand parameters
-demand_mean = 193
-demand_sd = 40
-
-# Deterministic model
-demand = demand_mean
-
-# Initial value for order quantity - this is the decision variable
-order_quantity = 200
-\end{lstlisting}
-
-Then students are given some skeleton code and must finish the expressions for computing key intermediate outputs (\code{order_cost, sales_revenue, refund_revenue}) and the final output, \code{profit}.
-Answers are provided at the bottom of the notebook so that students can attempt to finish the code but have a resource to correctly complete the code and move on.
-The screencast associated with the notebook also shows me completing the code.
-Using skeleton code in this way has worked well both for face to face lab based versions of this type of course as well as for online asynchronous versions.
-
-\begin{lstlisting}[language=Python]
-order_cost = unit_cost * order_quantity
-sales_revenue = ??? * selling_price
-refund_revenue = ??? * unit_refund
-profit = sales_revenue + refund_revenue - order_cost
-\end{lstlisting}
-
-At this point we discuss a fundamental difference between the computing model of Excel and that of Jupyter notebooks -- while spreadsheets respond automatically to changes, Jupyter notebooks require rerunning all code cells that include or follow the changed code.
-
-\subsubsection{Sensitivity analysis}
-
-Since order quantity is the key decision variable, we might want to see how profit changes for different order quantities in this first simplified model.
-In Excel, the Data Table tool provides an easy way to do this and is a staple of spreadsheet based management science textbooks.
-In a Data Table, a range of order quantities can be used as row or column input and one or more output values can be computed.
-In Python, we can rely on the vectorized behavior of the language.
-A range of order quantities is created as a NumPy array that can be used directly to compute vectors of all of the intermediate and final output variables.
-Much like Data Tables, vectorized computations avoid having to explicitly iterate, or loop, through a collection of input values.
-
-A natural next step is to do the equivalent of an Excel 2-way Data Table.
-By creating a profit function that takes all of the base inputs as arguments, we can construct a list comprehension that does what an Excel 2-Way Data Table does.
-Even better, this actually allows us to do the equivalent of an n-Way Data Table with an arbitrary number of outputs - something Excel does not have.
-This example provides an effective way to show the power of creating reusable functions and Python's handy list comprehension tool.
-
-\begin{lstlisting}[language=Python]
-	def bookstore_profit(unit_cost, selling_price, unit_refund, order_quantity, demand):
-	'''
-	Compute profit in bookstore model
-	'''
-		order_cost = unit_cost * order_quantity
-		sales_revenue = np.minimum(order_quantity, demand) * selling_price
-		refund_revenue = np.maximum(0, order_quantity - demand)
-		profit = sales_revenue + refund_revenue - order_cost
-		return profit
-
-	# Set up input vector ranges	
-	demand_range = np.arange(50, 301, 5)
-	order_quantity_range = np.arange(50, 301, 25)
-
-	# Create data table (as a list of tuples)
-	data_table_1 = [(d, oq, bookstore_profit(unit_cost, selling_price, unit_refund, oq, d)) 
-	for d in demand_range for oq in order_quantity_range]
-
-	# Convert to dataframe
-	dtbl_1_df = pd.DataFrame(data_table_1, columns=['Demand', 'OrderQuantity', 'Profit'])
-
-\end{lstlisting}
-
-% TODO - add output listing and plot
-
-\subsubsection{The object nature of Python}
-
-It is not clear whether an OO or a non-OO approach will end up making the most sense for doing spreadsheet type modeling.
-This uncertainty is made explicit in the notebook and students are led on a bit of a journey of discovery as we explore different software designs for this modeling problem.
-I purposefully designed the notebooks in this way to show students that the path to a solution is not necessarily linear and that giving yourself the freedom to explore alternatives, many of which will be dead ends, can lead to a rich learning experience.
-It is also more realistic in the sense that this is how most real problems are solved.
-It is analogous to the idea that mathematical proofs usually do not reflect the myriad of failed approaches that were tried but that ultimately contributed to the final product.
-
-In preparation for building an object oriented version of the model, basic object concepts are reviewed.
-Since most students have done some Excel VBA programming in a previous course, I refer back to a few properties and methods of the Excel \code{Worksheet} object.
-In Python, everything is an object.
-Using Python lists as an example, we explore some of its attributes such as methods for appending items to a list and reversing the order of its elements.
-The \code{dir} function is used to see all of an object's attributes and we see that even things like integers are objects in Python.
-With this basic object refresher, we are ready to do some actual object oriented programming and create our own classes.
-
-\subsubsection{Creating an object oriented model}
-
-For our initial design of a \code{BookstoreModel} class we make all of the base inputs class attributes (properties) and then add method attributes to compute outputs such as costs, revenues and profits.
-Students can easily see the mapping between things in the initial procedural model and this new object oriented version of the model.
-I highlight several important concepts and syntactical details of doing OO programming Python.
-Again, the goal is to weave in these more advanced Python programming concepts within the context of a very familiar modeling problem.
-This section of the notebook ends with us using our \code{BookstoreModel} class to create a model object with all of its base inputs instantiated with values.
-Before moving on students are presented with a few challenges involving enhancements to the \code{BookstoreModel} class.
-Interspersing short coding challenges throughout the notebook gives students a chance to test themselves on their understanding and to get some coding practice.
-
-Then a few design dilemmas are posed and some non-working approaches illustrated as we try to decide on how to implement an n-way data table function for this new object oriented model.
-These meta reflections are intentionally included to better mimic actual software development and reinforce the idea that software development is much more than simply writing code to implement a perfectly thought out design.
-We forge ahead with some preliminary ideas and are confronted with an intermediate problem related to generating a list of dictionaries that represent combinations of inputs (i.e.
-scenarios) to evaluate by our n-way data table function.
-This provides a great opportunity to remind students of one of the great strengths of the open source software ecosystem - we can leverage work done in other packages and even look at actual source code to see how something was implemented.
-In this case, the well known scikit-learn package has a \code{ParameterGrid} function that solves our scenario generation problem quite nicely.
-This example also reminds the students of the value of being able to make sense of API documentation and of object oriented code written by others -- most data science related Python packages are object oriented in nature.
-
-This first notebook draws to a close with the creation of an \code{data_table} function that allows us to do sensitivity analysis with the objected oriented model.
-Unlike Excel's Data Table tool, this \code{data_table} function can handle an arbitrary number of inputs and outputs.
-The payoff is the creation of the following faceted plot showing how profit varies for different order quantities and demand levels.
-We recap the main things learned in this notebook and prepare to add goal seeking capability to our model.
-
-\begin{figure}[!ht]
-	\centering
-	\begin{minipage}{.80\textwidth}
-		\includegraphics[width=1.0\textwidth]{images/two_way_dt.png}
-	\end{minipage}
-	\caption{Plot based on 2-way data table}
-	\label{fig:2way_dt}
-\end{figure}
-
-\subsection{Notebook 2: Goal seek}
-
-No self respecting modeling tool would be complete without goal seeking capability to do things such as finding the break even point in the bookstore model.
-For many students, Excel's Goal Seek tool can seem almost magical when encountered for the first time.
-Even after they get a sense of what it is doing, understanding the importance of the initial guess and how it can lead to Goal Seek reporting different solutions, is not very transparent.
-With Python, students can gain a much better understanding of how tools like Goal Seek really work.
-
-Before trying to build our \code{goal_seek} function we take a brief detour into basic root finding algorithms through packages like SciPy as well as in standalone Python scripts that are widely available in the Python data science ecosystem (refs).
-We compare the output of Excel's Goal Seek with different starting values to the values obtained by different root finding algorithms implemented in Python.
-The importance of starting values and also the range of different root finding algorithms available is often surprising to business students who have spent most of their life in Excel.
-
-A \code{} function is created that implements a simple bisection search and we use it to find the break even demand point in our bookstore model.
-With \code{data_table} and \code{goal_seek} functions implemented, we are ready to move on to doing Monte-Carlo simulation.
-
-\subsection{Notebook 3: Monte-Carlo simulation}
-
-In the Excel modeling world, Monte-Carlo simulation can be done without add-ins, but packages like @Risk \cite{} can make the process much easier.
-In particular, random variable generation for a wide range of probability distributions is facilitated by @Risk.
-In Python, the NumPy package provide random variate generation for a large number of distributions (https://numpy.org/doc/stable/reference/random/legacy.html#distributions) and similarly, SciPy provides functions for computing distribution related quantities for many probability distributions \cite{}.
-Much like we did when creating the \code{data_table} function, we start with adding uncertainty to a single input variable in the bookstore model and rely on Python's vectorized computing capabilities to generate the profit associated with each realization of the random input.
-Then we can do standard statistical analysis of the simulation output using Python packages such as Pandas and SciPy.
-In spreadsheet based simulation textbooks, the notion of the ``Flaw of Averages'' popularized by Sam Savage \cite{} is illustrated.
-We do the same with our Python based model, showing how replacing random demand by its mean can result in a wildly optimistic estimate of the mean profit found through analyzing the simulation output.
-Much like doing simulation in Excel without add-ins, using Python makes transparent the quite simple and brute force nature of Monte-Carlo simulation.
-
-Moving on to multiple uncertain inputs raises challenges much like those we faced when moving from a 1-way to an n-way data table function.
-In addition to modeling multiple uncertain inputs, we want to be able to specify a range of scenarios to run, much like we did for the \code{data_table} function.
-The following set of design criteria are presented for a \code{simulate} function.
-
-\begin{itemize}
-	\tightlist
-	\item
-	The first argument will be a model object (i.e.~something like the
-	\texttt{BookstoreModel} model) that contains an \texttt{update}
-	method. Soon, we should add an abstract \texttt{Model} class from
-	which specific models such as the \texttt{BookstoreModel} class can be
-	created. The abstract class will contain the \texttt{update} method.
-	\item
-	The random inputs will be passed in as a dictionary whose keys are the
-	input variables being modeled as random and whose values are an
-	iterable representing the draws from some probability distribution.
-	Structurally, this is similar to how inputs are specified in the
-	\texttt{data\_table} function.
-	\item
-	We can optionally pass in a dictionary of scenario inputs. This is
-	exactly like the \texttt{data\_table} variable input.
-
-	\begin{itemize}
-		\tightlist
-		\item
-		If no scenario input dictionary is passed in, a single simulation
-		scenario is run using the current input values in the model object,
-		\item
-		If a scenario input dictionary is passed in, then a simulation
-		scenario is run for every combination of parameters in the
-		dictionary. Again, this is just like we do in the
-		\texttt{data\_table} function.
-	\end{itemize}
-	\item
-	The output will be a set of dictionaries containing dataframes of
-	simulation output as well standard summary stats and plots.
-\end{itemize}
-
-This is followed by a well commented first attempt at a \code{simulate} function.
-By this point, students are well equipped to understand the code as it builds on basic Python concepts and more advanced things that we have already covered in this EwP module.
-The rest of the notebook is spent trying out the \code{simulate} function and ends with creating a plot that many of them have likely created before with @Risk along with a faceted plot of histograms of profit created with the popular Seaborn package.
-
-INSERT GROUPED BOX PLOT HERE and FACETED HISTOGRAMS
-
-The first submodule and its three notebooks are very much about the type of modeling typically done in a spreadsheet based course.
-In a spreadsheet based course, students learn numerous Excel functions and techniques in the process of building and exercising such models.
-Similarly, numerous Python techniques are scattered throughout the three modeling focused notebooks.
-Now we shift gears and move beyond what might be considered prototyping and into the realm of deployment.
-
-\section{Submodule 2: The packaging notebooks}
-
-\begin{quote}
-	content...
-
-In the first three notebooks, we've developed some Python approaches to
-typical Excel ``what if?'' analyses. Along the way we explored some
-slightly more advanced Python topics (for relative newcomers to Python)
-such as:
-
-\begin{itemize}
-	\tightlist
-	\item
-	List comprehensions,
-	\item
-	Basic OO programming, creating our own classes and using things like
-	\texttt{setattr} and \texttt{getattr},
-	\item
-	Leveraging scikit-learn's \texttt{ParameterGrid} class,
-	\item
-	Faceted plots using Seaborn and matplotlib,
-	\item
-	Tuple unpacking, zip, and itertools,
-	\item
-	Safely copying objects,
-	\item
-	Root finding with, and without, \texttt{scipy.optimize},
-	\item
-	Using \texttt{numpy.random} to generate random variates from various
-	probability distributions,
-	\item
-	Using \texttt{scipy.stats} to compute probabilities and percentiles,
-	\item
-	Importing our \texttt{data\_table}, \texttt{goal\_seek} and
-	\texttt{simulate} functions from a module.
-\end{itemize}
-
-Now that we've got a critical mass of "proof of concept" code, let's figure out how to structure our project and create a deployable package.
-In addition, let's rethink our OO design and add some much needed documentation to the code.
-\end{quote}
-
-Creating a deployable Python package is a good exercise in designing and creating reusable software artifacts.
-Thinking about how others will interact with your software usually leads to better designed software.
-Basic software engineering practices are often overlooked in business analytics programs but are certainly desirable in the professional world.
-In this submodule we create a Python package that can be imported and used by other modelers.
-This will require some design changes to our software and will also involve moving code out of Jupyter notebooks and into Python script files.
-While Jupyter notebooks can be integrated into production environments \cite{netflix}, it is much more common for code to live in a Python script file.
-In addition to creating a deployable package, this submodule also discusses the importance of documentation in its many forms in a Python project.
-
-\subsection{Notebook 4: Project packaging}
-
-The first notebook in this submodule starts by explaining what Python package are and their role in sharing code.
-We then revisit a concept from earlier in the course - creating a good project folder structure.
-For this we use what is known as a \textit{cookiecutter} that can automatically generate a project folder structure and key files from the answers to a few prompts \cite{cookiecutter}.
-Their are a few different folder layouts used in Python projects and there is a bit of discussion on the folder stucture we will use and resources given for exploring this issue further.
-Students are reminded of the importance of version control and we initialize a git repository for our budding package.
-The basics of version control were already covered in the first week of the course.
-My experience is that most students have little exposure to simple software project management concepts such as project folder structures and version control.
-This can lead to a jumble of files with incoherent names used as a proxy for version control - whatif_v1.py, whatif_v2.py, whatif_final.py, whatif_finalfinal.py, ...
-
-Key software design changes are then made and discussed which will make our code easier on our users.
-These changes include creating an abstract \code{Model} base class from which different model classes can be created and moving the \code{data_table}, \code{goal_seek}, and \code{simulate} functions into the \code{Model} base class.
-All of the code is moved into a code module named \code{whatif.py}.
-We are now ready to learn about the basics of turning our code into an installable package.
-This requires a basic understanding of two key Python files that are ubiquitous in Python pacakges - \code{__init__.py} and \code{setup.py}.
-Next comes a discussion of options for installing our package locally so that we can use it for different modeling projects.
-This includes an explanation of how and where Python searches for packages when \code{import} statements are encountered in code.
-We defer discussion of uploading our package to a public code sharing platform such as PyPI.
-All of this can be quite complex and I have made every effort to distill things down to a minimal level of complexity needed to convey the important points.
-I stress that there is much complexity lurking here and it is something they will need to revisit time and time again.
-
-This notebook ends by pointing the students to another notebook in which our newly deployed whatif package is used for a completely different model.
-The example used is the New Car Simulation that has been a staple of Winston and Albright's textbooks for many years \cite{bibid}.
-
-\subsection{Notebook 5: Documentation}
-
-This deployment focused submodule ends with a short notebook discussing the different types of documentation needed in a typical project.
-This includes things like code comments, docstrings, readme files and generating documentation by writing restrutured text files and using Sphinx.
-This is one of those tasks that few people like to do but can be very important for the long term success of a project.
-Even if you are just creating a tool for yourself, you will forget the details over time and documentation can be very beneficial in refreshing your memory.
-
-\section{Submodule 3: Excel data wrangling with Python}
-
-The EwP module ends with what has become a relatively common use of Python with Excel - automating various Excel data wrangling tasks. Examples include:
-
-\begin{itemize}
-	\tightlist
-	\item
-	You have a whole folder full of csv (or Excel) files with the same
-	file structure and you need to combine them into a single file. You
-	might also need to make some changes to the consolidated file.
-	\item
-	You have an Excel file with multiple sheets of similarly structured
-	data and you want to consolidate them into a single sheet.
-	\item
-	You have an Excel file with data in ``wide'' format and you need to
-	convert it to long format, and then perhaps export out individual
-	files (one per the key column(s) in the long formatted data).
-	\item
-	You have an Excel file acting as a simple flatfile database.
-	Periodically, you get new Excel files that need to get appended to the
-	``database'' file.
-\end{itemize}
-
-A single Jupyter notebook is used to give a taste of using Python to automate the process of working with Excel files by tackling each of the four examples above.
-Each of these examples are based on real problems I encountered either in research or industrial projects.
-TRANSITION NEEDED Several web based resources are shared with one particularly good one being the Practical Business Python blog written by Chris Moffitt.
-It includes numerous posts on different aspects of Python and Excel integration.
-There are a number of Python based tools that either include functionality for working with Excel files or are dedicated to specific Excel related operations.
-These include:
-
-\begin{itemize}
-	\tightlist
-	\item
-	pandas -
-	\href{https://pandas.pydata.org/docs/reference/api/pandas.read_excel.html}{read\_excel},
-	\href{https://pandas.pydata.org/docs/reference/api/pandas.ExcelWriter.html}{ExcelWriter},
-	\href{https://pandas.pydata.org/docs/reference/api/pandas.DataFrame.to_excel.html}{to\_excel}
-	\item
-	\href{https://openpyxl.readthedocs.io/en/stable/}{openpyxl} - can
-	read, write, and modify Excel files
-	\item
-	\href{https://xlsxwriter.readthedocs.io/}{XlsxWriter} - powerful but
-	can't edit existing Excel files
-	https://xlsxwriter.readthedocs.io/working\_with\_tables.html
-	\item
-	\href{https://www.xlwings.org/}{xlwings} - provides ``glue'' between
-	Python and Excel
-\end{itemize}
-
-\subsection{Example 1: Concatenating many csv files}
-
-While this first example is not necessarily Excel related, using Excel to consolidate csv files through a series of copy-paste operations is something that many Excel users will admit to doing.
-We also use this example to introduce the relatively new pathlib library for working with file paths.
-After getting familiar with pathlib, students are led through the development of a short procedure that creates a list of the csv files to concatenate and then uses the pandas library to read and combine the files in a \code{DataFrame} and then writes out the consolidated data as a csv file.
-
-\subsection{Example 2: Consolidating data from multiple sheets in a Excel file using pandas}
-
-Two different approaches are shown to consolidate data from multiple sheets into a single sheet.
-One approach uses the pandas library while the other relies on a library specifically created for working with Excel files from Python, openpyxl.
-One advantage to using Python rather than Visual Basic for Applications for this task is that Python is well suited for working with data in many different formats.
-The Excel file may be just one data source out of a collection of data sources that are used to create some sort of combined database.
-The generality of Python is a strength in such scenarios.
-
-\subsection{Example 3: Dealing with wide data and a multi-row header}
-
-Reshaping data from wide to long is a common data wrangling task especially in preparation for plotting.
-In this example, not only do students have to grapple with reshaping the data in an Excel workbook, they also have to deal with the all too common problem of multi-row header lines in a spreadsheet.
-Excel's flexibility makes it easy to create such multi-row headers and pity the poor analyst who then has to deal with it when trying to move the data into some sort of data frame or database table.
-This example is posed as a challenge to the students.
-Examples of the desired outputs are shown and a few hints given.
-The answer is provided at the end of the notebook.
-This example really shows the power of Python in creating a reproducible and automated approach to dealing with a poorly structured Excel workbook, creating not only restructured data frames but also non-trivial plots.
-Attempting to do all of this in Excel, even with VBA, is not an easy task.
-
-\subsection{Example 4: Appending new spreadsheet data in consolidated Excel workbook}
-
-Excel is not a database.
-But, that does not stop people from using it as a simple flat-file database.
-In such a scenario, it is not uncommon to have to periodically append new data to the end of a range of data in an Excel workbook.
-Again, a combination of openpyxl and pandas allows us to meet this challenge.
-
-\subsection{More on Python/Excel integration}
-
-The examples above barely scratch the surface of the use cases for Python integration with Excel.
-Our notebook ends with a shout out to the highly regarding Practical Business with Python blog and a list of specific examples that students can explore on their own.
-
-\begin{itemize}
-	\tightlist
-	\item
-	\href{https://pbpython.com/excel-file-combine.html}{Combining Data
-		From Multiple Excel Files} - file globbing, concatenating dataframes,
-	\item
-	\href{https://pbpython.com/pandas-excel-range.html}{Reading Poorly
-		Structured Excel Files with Pandas} - advanced use of
-	\texttt{read\_excel}, accessing ranges and Tables
-	\item
-	\href{https://pbpython.com/excel-pandas-comp.html}{Common Excel Tasks
-		Demonstrated in Pandas} - totals rows, fuzzy string matching
-	\item
-	\href{https://pbpython.com/excel-pandas-comp-2.html}{Common Excel
-		Tasks Demonstrated in Pandas - Part 2} - selection and filtering
-	\item
-	\href{https://pbpython.com/improve-pandas-excel-output.html}{Improving
-		pandas Excel output} - using XlsxWriter to format Excel workbooks from
-	Python
-	\item
-	\href{https://pbpython.com/advanced-excel-workbooks.html}{Creating
-		Advanced Excel Workbooks} - XlsxWriter, inserting VBA from Python(!),
-	using COM to merge sheets
-	\item
-	\href{https://pbpython.com/xlwings-pandas-excel.html}{Interactive Data
-		Analysis with Python and Excel} - using xlwings to ``glue'' Python and
-	Excel together, using sqlalchemy to interact with databases
-\end{itemize}
-
-% todo Check the hyperlinks above
-
-\section{Classroom experience and the future}
-
-This module has been included in my Advanced Analytics with Python course for the past three years.
-The course is offered in a seven-week accelerated format during May and June.
-It is an online and asynchronous course.
-The EwP module is covered over the course of two weeks and is the subject of one of the major homework assignments for the semester.
-The first part of the assignment involves creating a basic model, similar in complexity to the Bookstore Model, and doing sensitivity analysis, goal seeking and simulation with the model.
-It requires the students to import and use the whatif.py package we created as part of the EwP module.
-The second part of the assignment is based on the last submodule and requires the students to complete a task involving creation of a multi-sheet Excel workbook from a set of csv files using Python.
-Then, using the openpyxl library, they have to create and add formulas to the workbook and format them properly.
-The ranges used in the formulas vary by sheet and require the students to determine the number of rows in each sheet of data and use that information to construct the appropriate formula.
-An example assignment is available from \href{}{here}.
-The module has been well received and several of the working students have reported back to me that they have found use cases for these techniques in their day to day analytical activities.
-
-Rumours have swirled for a number of years that Microsoft was considering adding Python as a first class language alternative to VBA with their MS Office suite of packages.
-While this has not yet materialized, there continues to be significant interest in using Python to manipulate not only Excel workbooks but other office documents such as Powerpoint presentations.
-Given that business students are very familiar with both Excel and Powerpoint, inclusion of this material in a business school Python based analytics course seems quite appropriate.
-Students can learn more advanced and extremely useful Python concepts and techniques within a context with which they are intimately familiar.
-It is directly relevant to the job content of many of the students.
-This teaching module is a natural follow up to the commonly taught spreadsheet based analytics courses found in many business schools and provides an interesting way to teach and learn Python for analytics.
-
-\section{Random notes}
-Learning good Python programming practices via learning to model in similar way that students learn good spreadsheet practices through building spreadsheet models using the W\&A textbook.
-
-How aap course fits in relation to pcda course. Could also be second part of a fully Python based course. Builds on basic intro Python concepts and skills.
-
-The Quant Econ website is a good example of another use of Python for typical Excel things.
-
-Everything discussed in this article is available from the publicly accessible course website.
-
-% Acknowledgments here
-\ACKNOWLEDGMENT{%
-% Enter the text of acknowledgments here
-}% Leave this (end of acknowledgment)
-
-% Appendix here
-% Options are (1) APPENDIX (with or without general title) or 
-%             (2) APPENDICES (if it has more than one unrelated sections)
-% Outcomment the appropriate case if necessary
-%
-% \begin{APPENDIX}{<Title of the Appendix>}
-% \end{APPENDIX}
-%
-%   or 
-%
-% \begin{APPENDICES}
-% \section{<Title of Section A>}
-% \section{<Title of Section B>}
-% etc
-% \end{APPENDICES}
-
-% References here (outcomment the appropriate case)
-
-% CASE 1: BiBTeX used to constantly update the references 
-%   (while the paper is being written).
-%\bibliographystyle{informs2014} % outcomment this and next line in Case 1
-%\bibliography{<your bib file(s)>} % if more than one, comma separated
-
-% CASE 2: BiBTeX used to generate mypaper.bbl (to be further fine tuned)
-%\input{mypaper.bbl} % outcomment this line in Case 2
-
+\begin{tcolorbox}[left=1mm]
+	this is code
+\end{tcolorbox}
 \end{document}

### 2023-06-20 python_excel_ite.tex
working on intro section

@@
 \tcbuselibrary{minted} % tcolorbox minted library, required to use the "minted" tcb listing engine (this library is not loaded by the option [most])
-%\usepackage{minted} % Allows input of raw code, such as Python code
 \usepackage[colorlinks]{hyperref} % ALWAYS load this package LAST
@@
 
-% The opening part of the what-if notebook has some language for the motivation behind this module.
-% TODO - opening should really grab reader
-Spreadsheet based modeling has transformed the teaching of management science in business schools.
-The late 1990's saw the first editions of game changing textbooks \cite{ } focused on the use of spreadsheets in teaching modeling and management science.
+Spreadsheet based modeling has transformed the teaching of management science in business schools over the past three decades.
+The late 1990's saw first editions of game changing textbooks \cite{winstonPracticalManagementScience2018,ragsdaleSpreadsheetModelingDecision2017} focused on the use of spreadsheets in teaching modeling and management science.
 A series of conferences focused on teaching management science (TMS) with spreadsheets started at Dartmouth in 1998 and several presentations from that conference made up the bulk of papers in the first two issues of the INFORMS Transactions on Education journal \cite{bibid}.
-I attended TMS III at Indiana University and left with numerous ideas that influenced my own spreadsheet based modeling course.
-The next few decades saw numerous textbooks and teaching papers that prominently featured spreadsheet based modeling.
+The next few decades saw numerous textbooks \cite{BusinessAnalyticsArt, BusinessAnalytics4th} and teaching papers \cite{INFORMSTransactionsEducation2014} that prominently featured spreadsheet based modeling.
 
-Despite numerous calls for and predictions of the eradication of spreadsheets from the business world,
+Despite numerous calls for and predictions of the eradication of spreadsheets from the business world, Google "spreadsheets are evil"
 
-Spreadsheet based modeling wave of the early 2000s. TMS conferences. Ground breaking textbooks such as Ragsdale and W\&A.
-The first two issues of ITE are paper based on TMS 1 at Dartmouth in 1998. I attended TMS 3 in Bloomington in 2000.
-
-First edition of Ragsdale was 1995 PMS was 1997.
+Excel not a database and shouldn't be used as one.
+There are better data visualizaiton and exploration tools.
+There are better dashboarding tools.
+Spreadsheet errors are a huge problem.
+Version control and documentation issues.
+Data size limitations.
+External link problems.
+But, spreadsheets are still very well suited to what if type modeling due to their flexibility, numerous built in functions, integrated visualization capabilities and ubiquity
 
@@
 Python has continued to make inroads to analytical territory that was formerly dominated by Excel.
-The widely used Python data analysis library, Pandas, was created by Wes McKinney while working as a quantitative financial analyst.
+The widely used Python data analysis library, Pandas, was created by Wes McKinney while working as a quantitative financial analyst \cite{bibid}.
 However, Python was mainly used for data wrangling, visualization and predictive modeling using various flavors of regression and tree based models.
@@
 
-PBB blog
-
 Teach relevant Python to business students using familiar and relevant modeling examples.
@@
 
-Integrated or separate lit review section?
+%TODO: Integrated or separate lit review section?
 
@@
 %   (while the paper is being written).
-%\bibliographystyle{informs2014} % outcomment this and next line in Case 1
-%\bibliography{<your bib file(s)>} % if more than one, comma separated
+\bibliographystyle{informs2014} % outcomment this and next line in Case 1
+\bibliography{python_excel.bib} % if more than one, comma separated
 

### 2023-06-22 python_excel_ite.tex
working on intro

@@
 
-Despite numerous calls for and predictions of the eradication of spreadsheets from the business world, Google "spreadsheets are evil"
+Despite numerous calls for and predictions of the eradication of spreadsheets from the business world, Excel and other spreadsheets are alive and well.
+Yes, spreadsheets have well known limitations.
+Spreadsheets are not databases and probably should not be used as one, especially for relational data models.
+There are better data visualization and exploration tools.
+There are better dashboarding tools that are more easily maintained in a production environment.
+Spreadsheet errors are a huge problem.
+Version control and documentation present challenges.
+There are data size limitations and external data links present their own set of challenges.
+But, spreadsheets are still very well suited to ``what if?'' type modeling due to their flexibility, numerous built in functions, and their integrated sensitivity analysis, goal seeking and visualization capabilities.
+And of course, they are ubiquitous and familiar to hordes of analysts and modelers.
 
-Excel not a database and shouldn't be used as one.
-There are better data visualizaiton and exploration tools.
-There are better dashboarding tools.
-Spreadsheet errors are a huge problem.
-Version control and documentation issues.
-Data size limitations.
-External link problems.
-But, spreadsheets are still very well suited to what if type modeling due to their flexibility, numerous built in functions, integrated visualization capabilities and ubiquity
+But, the analytics world has been changing.
+The advent of programmatic analytics using tools like R and Python have changed the practice of analytics as well as the teaching landscape \cite{bibid}.
+Limitations of spreadsheets with respect to reproducibility and automation have certainly played a role.
+Additionally, both R and Python are free, open source, and have a rich set of analytics related libraries that have contributed to their growing popularity.
+Python, being a general purpose programming language, also offers advantages when moving models and analytical analysis pipelines from the prototype stage into production.
+Enormous data science ecosystems have emerged around both R and Python providing a wealth of learning resources for both new and experienced analytics students and professionals.
 
-While spreadsheets have certainly changed and expanded in their capabilities (PowerPivot, um...) since the 90's...
+Numerous R and Python based data science courses have been developed and offered at universities and online learning platforms such as Coursera, EdX, Udemy and others.
+Some of these courses have made a large amount of their course content freely available for self-learners and for other educators.
+At our institution, we created a course called Practical Computing for Data Analytics (PCDA) in 2015 that introduced both R and Python (along with a little Linux) for use in data analytics work.
+The course is taught using combination of R Markdown documents \cite{bibid} and Jupyter notebooks \cite{bibid}.
+All of the course content is freely available from a publicly accessible course web site \url{URL}.
 
-Irony of Mgt Sci crisis of 1990's where course was dropped by AACSB and ignored by other business disciplines to an era where analytics has captured the day and now AI threatens even that.
+The PCDA course focuses mostly on data wrangling, exploratory data analysis (EDA), and statistical/machine learning based predictive models.
+It does include a Python based simulation of the Monte Hall 3-door problem.
+In teaching EDA using tools like ggplot2, I would challenge students to think about how they would go about trying to do some of the plots we were making in Excel.
+There is also some limited coverage of using Python to open Excel files and do some sheet manipulation.
+However, there was no real attempt to do the kind of modeling that was the focus of spreadsheet based modeling textbooks.
 
-Powell's 2001 paper on the importance of modeling is no less relevant now than it was when published. I still many of his ideas in my spreadsheet based modeling course.
-
-Advent of programmatic analytics using tools like R and Python have changed the teaching landscape.
-Motivated by reproducibility, open source, free, and ease of integration with web and cloud databases.
-
-Created PCDA course in response.
-Challenges with finding good textbooks.
-Developed much of my own curricular materials.
-Taught using combo of R Markdown documents and Jupyter notebooks.
 Literate computing \cite{Jupyter_proposal}.
@@
 Jupyter used by many universities as a key component of their educational technology stack.
-
-PCDA course focused mostly on data wrangling, EDA, statistical and machine learning based predictive models though I did include a Python based simulation of the Monte Hall 3-door problem.
-In teaching EDA using tools like ggplot2, I would challenge students to think about how they would go about trying to do some of the plots we were making in Excel.
-There is also some limited coverage of using Python to open Excel files and do some sheet manipulation.
-However, there was no real attempt to do the kind of modeling that was the focus of textbooks like PMS or DADM.
 
@@
 
-Limitations of spreadsheets such as reproducibility and automation.
-
 While Python has been gaining momentum in the business analytics world, it is often used for data wrangling, analysis and visualization of tablular data using tools like pandas and matplotlib or Seaborn.
@@
 Created the ``Excel with Python'' (EwP) module as part
+
+While spreadsheets have certainly changed and expanded in their capabilities (PowerPivot, um...) since the 90's...
+
+Irony of Mgt Sci crisis of 1990's where course was dropped by AACSB and ignored by other business disciplines to an era where analytics has captured the day and now AI threatens even that.
+
+Powell's 2001 paper on the importance of modeling is no less relevant now than it was when published. I still many of his ideas in my spreadsheet based modeling course.
 

### 2023-06-22 python_excel_ite.tex
working on intro

@@
 But, spreadsheets are still very well suited to ``what if?'' type modeling due to their flexibility, numerous built in functions, and their integrated sensitivity analysis, goal seeking and visualization capabilities.
-And of course, they are ubiquitous and familiar to hordes of analysts and modelers.
+Excel is widely used for building and using models of business problems
+to explore the impact of various model inputs on key outputs. Built in
+tools such as \href{https://support.microsoft.com/en-us/office/calculate-multiple-results-by-using-a-data-table-e95e2487-6ca6-4413-ad12-77542a5ea50b}{Data
+	Tables} and
+\href{https://support.microsoft.com/en-us/office/use-goal-seek-to-find-the-result-you-want-by-adjusting-an-input-value-320cb99e-f4a4-417f-b1c3-4f369d6e66c7}{Goal
+	Seek} are well known to power spreadsheet modelers.
 
@@
 At our institution, we created a course called Practical Computing for Data Analytics (PCDA) in 2015 that introduced both R and Python (along with a little Linux) for use in data analytics work.
-The course is taught using combination of R Markdown documents \cite{bibid} and Jupyter notebooks \cite{bibid}.
-All of the course content is freely available from a publicly accessible course web site \url{URL}.
+The course is taught using a combination of R Markdown documents \cite{bibid} and Jupyter notebooks \cite{bibid}.
+Jupyter notebooksLiterate computing \cite{Jupyter_proposal}.
+Educational narratives that support iterative exploration and development.
+The ability to weave standard language with formal programming languages is critical to creating the desired educational artifact.
+Jupyter used by many universities as a key component of their educational technology stack.All of the course content is freely available from a publicly accessible course web site \url{URL}.
 
@@
 It does include a Python based simulation of the Monte Hall 3-door problem.
-In teaching EDA using tools like ggplot2, I would challenge students to think about how they would go about trying to do some of the plots we were making in Excel.
+In teaching EDA using tools like ggplot2, students are challenged to think about how they would go about trying to use Excel to create some of the plots that are easy to do with R and Python based plotting libraries.
 There is also some limited coverage of using Python to open Excel files and do some sheet manipulation.
@@
 
-Literate computing \cite{Jupyter_proposal}.
-Educational narratives that support iterative exploration and development.
-The ability to weave standard language with formal programming languages is critical to creating the desired educational artifact.
-Jupyter used by many universities as a key component of their educational technology stack.
+Python has continued to make inroads to analytical territory that was formerly dominated by Excel.
+The widely used Python data analysis library, pandas, was created by Wes McKinney while working as a quantitative financial analyst \cite{bibid}.
 
-Python has continued to make inroads to analytical territory that was formerly dominated by Excel.
-The widely used Python data analysis library, Pandas, was created by Wes McKinney while working as a quantitative financial analyst \cite{bibid}.
-However, Python was mainly used for data wrangling, visualization and predictive modeling using various flavors of regression and tree based models.
+Python is widely used for data wrangling, visualization and predictive modeling using various flavors of regression and tree based models.
 The blog \href{https://pbpython.com/}{Practical Business Python} written primarily by Chris Moffitt contributed greatly to making Python more accessible to business analytics practitioners through high quality posts that were extremely practical and useful.
@@
 Powell's 2001 paper on the importance of modeling is no less relevant now than it was when published. I still many of his ideas in my spreadsheet based modeling course.
+
+Literate computing \cite{Jupyter_proposal}.
+Educational narratives that support iterative exploration and development.
+The ability to weave standard language with formal programming languages is critical to creating the desired educational artifact.
+Jupyter used by many universities as a key component of their educational technology stack.
 

### 2023-06-23 python_excel_ite.tex
working on intro

@@
 What would those look like in Python?
-
-Business schools with business analytics programs are the context. Grad and undergrad.
-
-Why I created
-
-- more topics I wanted to cover than were possible in PCDA course
-- weave in more SE type material which is often overlooked in BA programs
-
-Excel is widely used for building and using models of business problems
-to explore the impact of various model inputs on key outputs. Built in
-``what if analysis'' tools such as \href{https://support.microsoft.com/en-us/office/calculate-multiple-results-by-using-a-data-table-e95e2487-6ca6-4413-ad12-77542a5ea50b}{Data
-	Tables} and
-\href{https://support.microsoft.com/en-us/office/use-goal-seek-to-find-the-result-you-want-by-adjusting-an-input-value-320cb99e-f4a4-417f-b1c3-4f369d6e66c7}{Goal
-	Seek} are well known to power spreadsheet modelers. How might we do
-similar modeling and analysis using Python?
-
-While Python has been gaining momentum in the business analytics world, it is often used for data wrangling, analysis and visualization of tablular data using tools like pandas and matplotlib or Seaborn.
-You can find some great examples at [Chris Moffit's Practical Business Python blog](https://pbpython.com/).
-I use Python all the time for such tasks and teach a course called [Practical Computing for Data Analytics](http://www.sba.oakland.edu/faculty/isken/courses/mis5470/) that is Python (and R) based.
-But, it got me to thinking.
-What about those things for which Excel is well suited such as building formula based models and doing sensitivity analysis on these models?
-What would those look like in Python?
+Powell's 2001 paper on the importance of modeling is no less relevant now than it was when published.
+I still many of his ideas in my spreadsheet based modeling course.
 
@@
 
-Powell's 2001 paper on the importance of modeling is no less relevant now than it was when published. I still many of his ideas in my spreadsheet based modeling course.
-
-Literate computing \cite{Jupyter_proposal}.
-Educational narratives that support iterative exploration and development.
-The ability to weave standard language with formal programming languages is critical to creating the desired educational artifact.
-Jupyter used by many universities as a key component of their educational technology stack.
-
 %TODO: Integrated or separate lit review section?
@@
 At our institution, The EwP module is part of a business course entitled ``Advanced Analytics with Python'' (\href{http://www.sba.oakland.edu/faculty/isken/courses/mis6900}{AAP}).
-The students taking this course have already taken my ``Practical Computing for Business Analytics'' (\href{http://www.sba.oakland.edu/faculty/isken/courses/mis5470}{PCDA}) course in which they spend about half the semester learning R and the other half learning Python, both in the context of business analytics.
-Basic Linux skills are also part of the PCDA course.
+The students taking this course have already taken the PCDA course within which they learned fundamental Python programming within the context of data analytics.
 The PCDA course ends with intro to the sklearn library for doing predictive modeling in Python and the AAP course begins where the PCDA course ends.
+The AAP course was created
+
+Why I created
+
+- more topics I wanted to cover than were possible in PCDA course
+- weave in more SE type material which is often overlooked in BA programs
+
 The EwP module could certainly be used as part of a semester long Python based analytics course.
@@
 
-The EwP module is made of three submodules.
+The EwP module is made up of three submodules.
 The first of these focuses on building and using a typical spreadsheet model using Python.
@@
 It's free, open source, runs on all platforms, huge ecosystem, used in industry, used widely in education.
+Literate computing \cite{Jupyter_proposal}.
+Educational narratives that support iterative exploration and development.
+The ability to weave standard language with formal programming languages is critical to creating the desired educational artifact.
+Jupyter used by many universities as a key component of their educational technology stack.
 

### 2023-06-23 python_excel_ite.tex
working on intro

@@
 There are better dashboarding tools that are more easily maintained in a production environment.
-Spreadsheet errors are a huge problem.
+Spreadsheet errors can be a huge problem.
 Version control and documentation present challenges.
 There are data size limitations and external data links present their own set of challenges.
-But, spreadsheets are still very well suited to ``what if?'' type modeling due to their flexibility, numerous built in functions, and their integrated sensitivity analysis, goal seeking and visualization capabilities.
-Excel is widely used for building and using models of business problems
-to explore the impact of various model inputs on key outputs. Built in
-tools such as \href{https://support.microsoft.com/en-us/office/calculate-multiple-results-by-using-a-data-table-e95e2487-6ca6-4413-ad12-77542a5ea50b}{Data
-	Tables} and
-\href{https://support.microsoft.com/en-us/office/use-goal-seek-to-find-the-result-you-want-by-adjusting-an-input-value-320cb99e-f4a4-417f-b1c3-4f369d6e66c7}{Goal
-	Seek} are well known to power spreadsheet modelers.
+But, Excel is widely used for building and using models of business problems
+to explore the impact of various model inputs on key outputs.
+This ``what if?'' type modeling is facilitated by Excel's flexibility, numerous built in functions, integrated sensitivity analysis via \href{https://support.microsoft.com/en-us/office/calculate-multiple-results-by-using-a-data-table-e95e2487-6ca6-4413-ad12-77542a5ea50b}{Data Tables},  \href{https://support.microsoft.com/en-us/office/use-goal-seek-to-find-the-result-you-want-by-adjusting-an-input-value-320cb99e-f4a4-417f-b1c3-4f369d6e66c7}{Goal
+	Seek} or break-even analysis and visualization capabilities.
 
-But, the analytics world has been changing.
+Nevertheless, the analytics world has been changing.
 The advent of programmatic analytics using tools like R and Python have changed the practice of analytics as well as the teaching landscape \cite{bibid}.
@@
 
-Numerous R and Python based data science courses have been developed and offered at universities and online learning platforms such as Coursera, EdX, Udemy and others.
+Numerous R and Python based data science courses have been developed and offered at universities and through online learning platforms such as Coursera, EdX, Udemy and others.
 Some of these courses have made a large amount of their course content freely available for self-learners and for other educators.
-At our institution, we created a course called Practical Computing for Data Analytics (PCDA) in 2015 that introduced both R and Python (along with a little Linux) for use in data analytics work.
+At our institution, we have had a spreadsheet based modeling course in place since 2001.
+In 2015, we created a course called Practical Computing for Data Analytics (PCDA) that introduced both R and Python (along with a little Linux) for use in data analytics work.
 The course is taught using a combination of R Markdown documents \cite{bibid} and Jupyter notebooks \cite{bibid}.
-Jupyter notebooksLiterate computing \cite{Jupyter_proposal}.
-Educational narratives that support iterative exploration and development.
-The ability to weave standard language with formal programming languages is critical to creating the desired educational artifact.
-Jupyter used by many universities as a key component of their educational technology stack.All of the course content is freely available from a publicly accessible course web site \url{URL}.
+Both of these support \textit{literate computing} \cite{Jupyter_proposal} and allow us to create educational narratives that support iterative exploration and development.
+The ability to weave standard language with formal programming languages has made notebook computing a defacto standard in the data science education community.
+All of the course content, including many Jupyter notebooks and supporting screencasts, is freely available from a publicly accessible course web site \url{URL}.
 
-The PCDA course focuses mostly on data wrangling, exploratory data analysis (EDA), and statistical/machine learning based predictive models.
+The PCDA course, like many data science courses, focuses mostly on data wrangling, exploratory data analysis (EDA), and statistical/machine learning based predictive models.
 It does include a Python based simulation of the Monte Hall 3-door problem.
@@
 The widely used Python data analysis library, pandas, was created by Wes McKinney while working as a quantitative financial analyst \cite{bibid}.
-
-Python is widely used for data wrangling, visualization and predictive modeling using various flavors of regression and tree based models.
-The blog \href{https://pbpython.com/}{Practical Business Python} written primarily by Chris Moffitt contributed greatly to making Python more accessible to business analytics practitioners through high quality posts that were extremely practical and useful.
+Python is widely used for data wrangling, visualization and predictive modeling using various flavors of regression, tree based models and neural networks.
+The blog \href{https://pbpython.com/}{Practical Business Python} written primarily by Chris Moffitt contributed to making Python more accessible to business analytics practitioners through high quality posts that were extremely practical and useful.
 But, it got me to thinking.
@@
 What would those look like in Python?
-Powell's 2001 paper on the importance of modeling is no less relevant now than it was when published.
-I still many of his ideas in my spreadsheet based modeling course.
+The importance of modeling, articulated so well in the pages of this journal \cite{powellTeachingModelingManagement2001}, has not abated.
+What role does Python have to play in supporting the kind of modeling taught for years in our business schools?
 
-Quant Econ example
+In 2021, I published a series of blog posts that explored this question and formed the basis of the teaching module described in this paper.
+During the development of these posts, I came across the \href{https://quantecon.org/}{Quant Econ: open source code for economic modeling} project which \quote{is a nonprofit organization dedicated to development and documentation of open source computational tools for economics, econometrics, and decision making}.
+This is a mature project, run by renowned economists \cite{bibid}, with significant support.
+They have developed and released numerous high quality Jupyter notebooks exploring a range of computational economic and financial modeling topics.
+Here was exactly the kind Python for business usage I had been seeking.
+With this as inspiration, I decided to turn my blog post series into a teaching module called Excel with Python (EwP).
 
-Teach relevant Python to business students using familiar and relevant modeling examples.
-Give sense of software development process and design decisions
-Model a learning journey
-Teach students to think and act like a software developer, not just an analyst
+The EwP module is not as polished nor as advanced as the Quant Econ materials, but does add to the growing body of freely available teaching materials aimed at Python for business modeling.
+My goals for EwP are also a little different and include:
 
-Created the ``Excel with Python'' (EwP) module as part
+\begin{itemize}
+	\item Teach relevant Python to business students using familiar and relevant modeling examples,
+	\item Give sense of software development process and design decisions
+	\item Model a learning journey
+	\item Teach students to think and act like a software developer, not just an analyst
+\end{itemize}
 
-While spreadsheets have certainly changed and expanded in their capabilities (PowerPivot, um...) since the 90's...
-
-Irony of Mgt Sci crisis of 1990's where course was dropped by AACSB and ignored by other business disciplines to an era where analytics has captured the day and now AI threatens even that.
-
-%TODO: Integrated or separate lit review section?
+One criticism that has been levied against data science programs has been the lack of attention to basic software engineering principles.
+The image of newly minted grads developing spaghetti coded Jupyter notebooks not nearly ready for being put into production has reached meme like status.
+I hoped to address some of these issues in the EwP module.
+Sneakily, students would be exposed to some simple software engineering concepts during the process of building and analyzing familiar models.
+This is really no different than the integration of advanced Excel functions and techniques with spreadsheet modeling tutorials.
+Just as spreadsheet modeling students were able to easily wow their coworkers and managers with INDEX-MATCH and intermediate Pivot Table skills, I hoped that my students ability to properly document a Python function and use git and GitHub would pay off in the workplace.
 

### 2023-06-23 python_excel_ite.tex
intro section drafted

@@
 The ability to weave standard language with formal programming languages has made notebook computing a defacto standard in the data science education community.
-All of the course content, including many Jupyter notebooks and supporting screencasts, is freely available from a publicly accessible course web site \url{URL}.
+All of the PCDA course content, including many Jupyter notebooks and supporting screencasts, is freely available from a publicly accessible course web site \url{URL}.
 
@@
 \begin{itemize}
-	\item Teach relevant Python to business students using familiar and relevant modeling examples,
-	\item Give sense of software development process and design decisions
-	\item Model a learning journey
-	\item Teach students to think and act like a software developer, not just an analyst
+	\item teach relevant and more advanced Python to business students using familiar and simple modeling examples,
+	\item convey a sense of the software design and development process,
+	\item be a tour guide for a learning journey,
+	\item teach students to think and act like a software developer, not just an analyst.
 \end{itemize}
@@
 This is really no different than the integration of advanced Excel functions and techniques with spreadsheet modeling tutorials.
-Just as spreadsheet modeling students were able to easily wow their coworkers and managers with INDEX-MATCH and intermediate Pivot Table skills, I hoped that my students ability to properly document a Python function and use git and GitHub would pay off in the workplace.
+Just as spreadsheet modeling students were able to easily wow their coworkers and managers with INDEX-MATCH and intermediate Pivot Table skills, I hoped that my students ability to properly document a Python function and use git and GitHub would give them a little street cred.
 
@@
 Then the functionality developed for doing this type of modeling in Python is deployed as a reusable package.
-We end with an introduction to using Python to manipulate Excel files, an extremely common use case.
+We end with an introduction to using Python to manipulate Excel files.
 
+%TODO needs finishing
 All of the Python modules in the PCDA course as well as the entire AAP course are taught using a combination of Jupyter notebooks, Python scripts, and videos walking through the notebooks, all publicly accessible from extensive course web pages.

### 2023-06-27 python_excel_ite.tex
fixed bad quotation marks

@@
 
-\textsf{We have added a basic `simulate` function to our `data_table` and `goal_seek` functions. Python is proving to be quite nice for doing Excel-style "what if?" analysis.}
+\textsf{We have added a basic \code{simulate} function to our \code{data\_table} and \code{goal\_seek} functions. Python is proving to be quite nice for doing Excel-style "what if?" analysis.}
 
@@
 My experience is that most students have little exposure to software project management concepts such as project folder structures and version control.
-This can lead to a jumble of files with incoherent names used as a proxy for version control - \code{whatif\_v1.py}, \code{whatif\_v2.py}, \code{whatif\_final.py}, \code{whatif\_finalfinal.py}, and so on.
+This can lead to a jumble of files with incoherent names used as a proxy for version control -- \code{whatif\_v1.py}, \code{whatif\_v2.py}, \code{whatif\_final.py}, \code{whatif\_finalfinal.py}, and so on.
 

### 2023-06-29 python_excel_ite.tex
reworked intro

@@
 There are data size limitations and external data links present their own set of challenges.
-But, Excel is widely used for building and using models of business problems
+But, Excel continues to be used for a wide range of quantitative analysis related activities throughout the business world.
+One particular use is the building and exercising of models of business problems
 to explore the impact of various model inputs on key outputs.
@@
 
-Nevertheless, the analytics world has been changing.
+At the same time, the analytics world has been changing.
 The advent of programmatic analytics using tools like R and Python have changed the practice of analytics as well as the teaching landscape \cite{bibid}.
@@
 In 2015, we created a course called Practical Computing for Data Analytics (PCDA) that introduced both R and Python (along with a little Linux) for use in data analytics work.
-The course is taught using a combination of R Markdown documents \cite{bibid} and Jupyter notebooks \cite{bibid}.
-Both of these support \textit{literate computing} \cite{Jupyter_proposal} and allow us to create educational narratives that support iterative exploration and development.
-The ability to weave standard language with formal programming languages has made notebook computing a defacto standard in the data science education community.
-All of the PCDA course content, including many Jupyter notebooks and supporting screencasts, is freely available from a publicly accessible course web site \url{URL}.
 
@@
 
-In 2021, I published a series of blog posts that explored this question and formed the basis of the teaching module described in this paper.
+In 2021, I published a series of blog posts \cite{bibid} that explored this question and formed the basis of the teaching module described in this paper.
 During the development of these posts, I came across the \href{https://quantecon.org/}{Quant Econ: open source code for economic modeling} project which \quote{is a nonprofit organization dedicated to development and documentation of open source computational tools for economics, econometrics, and decision making}.
@@
 \begin{itemize}
-	\item teach relevant and more advanced Python to business students using familiar and simple modeling examples,
-	\item convey a sense of the software design and development process,
-	\item be a tour guide for a learning journey,
-	\item teach students to think and act like a software developer, not just an analyst.
+	\item teaching relevant and more advanced Python to business students using familiar and simple modeling examples,
+	\item conveying a sense of the software design and development process,
+	\item acting as a tour guide for a learning journey,
+	\item teaching students to think and act like a software developer, not just an analyst.
 \end{itemize}
 
+% TODO - Edit this paragraph
 One criticism that has been levied against data science programs has been the lack of attention to basic software engineering principles.
@@
 The PCDA course ends with intro to the sklearn library for doing predictive modeling in Python and the AAP course begins where the PCDA course ends.
-The AAP course was created
-
-Why I created
-
-- more topics I wanted to cover than were possible in PCDA course
-- weave in more SE type material which is often overlooked in BA programs
-
 The EwP module could certainly be used as part of a semester long Python based analytics course.
-The students have also already taken "Business Analytics" (\href{http://www.sba.oakland.edu/faculty/isken/courses/mis5460}{BA}) which is a spreadsheet based introduction to business analytics.
+Most students have also already taken "Business Analytics" (\href{http://www.sba.oakland.edu/faculty/isken/courses/mis5460}{BA}) which is a spreadsheet based introduction to business analytics.
+We do not strictly require the spreadsheet based course be taken before the PCDA course as that course draws students from outside of our business programs who are strictly interested in programmatic tools such as Python.
 The BA and PCDA courses have historically been offered in face to face mode in a computer teaching lab.
@@
 The newer AAP course has been offered since 2021 in an online asynchronous mode.
-All three courses have extensive course websites that are publicly accessible, including access to all of the video content and files needed.
+All three courses have \href{URL}{extensive course websites} that are publicly accessible, including access to all of the video content and files needed.
 We will describe the AAP course web in more detail below in the context of the EwP module.
@@
 
-%TODO needs finishing
+%TODO needs much editing and shaping
 All of the Python modules in the PCDA course as well as the entire AAP course are taught using a combination of Jupyter notebooks, Python scripts, and videos walking through the notebooks, all publicly accessible from extensive course web pages.
@@
 Many are working professionals and a 2.5 hour evening class in a warm computer lab can be taxing.
+
+The course is taught using a combination of R Markdown documents \cite{bibid} and Jupyter notebooks \cite{bibid}.
+Both are examples of \textit{literate computing} \cite{Jupyter_proposal} and allow us to create educational narratives that support iterative exploration and development.
+The ability to weave standard explanatory language with formal programming languages has made notebook computing a defacto standard in the data science education community.
+All of the PCDA course content, including many Jupyter notebooks and supporting screencasts, is freely available from a publicly accessible course web site \url{URL}.
 
@@
 Thinking about how others will interact with your software usually leads to better designed software.
-Basic software engineering practices are often overlooked in business analytics classes but are certainly desirable in the professional world.
 In this submodule we create a Python package that can be imported and used by other modelers.
@@
 In addition to creating a deployable package, this submodule also discusses the importance of documentation in its many forms in a Python project.
+
+% REMOVED
+% Basic software engineering practices are often overlooked in business analytics classes but are certainly desirable in the professional world.
 
@@
 Key software design changes are then made which will make our code easier on our users.
-These changes include creating an abstract \code{Model} base class from which different model classes can be created and moving the \code{data\_table}, \code{goal\_seek}, and \code{simulate} functions into the \code{Model} base class.
+These changes include creating an abstract \code{Model} base class from which different model classes can be created and moving the \code{data\_table}, \code{goal\_seek}, and \code{simulate} functions into the \code{Model} base class as methods.
 All of the code is moved into a code module named \code{whatif.py}.
@@
 This deployment focused submodule ends with a short notebook discussing the different types of documentation needed in a typical project.
-This includes things like code comments, docstrings, readme files and generating documentation by writing restrutured text files and using Sphinx.
+This includes code comments, docstrings, readme files and generating documentation by writing restrutured text files and using Sphinx.
 This is one of those tasks that few people like to do but can be very important for the long term success of a project \cite{bibid}.
@@
 	\href{https://xlsxwriter.readthedocs.io/}{XlsxWriter} - powerful but
-	can't edit existing Excel files
-	https://xlsxwriter.readthedocs.io/working\_with\_tables.html
+	cannot edit existing Excel files
 	\item
@@
 The ranges used in the formulas vary by sheet and require determination of the number of rows in each sheet of data and use that information to construct the appropriate formula.
-An example assignment is available from \href{}{here}.
+An example assignment is available from \href{some.url}{here}.
 The module has been well received and several of the working students have reported back to me that they have found use cases for these techniques in their day to day analytical activities.
 
-Rumors have swirled for a number of years that Microsoft was considering adding Python as a first class language alternative to VBA with their MS Office suite of packages \cite{bibid}.
+Rumors have swirled for a number of years that Microsoft was considering adding Python as a first class language alternative to VBA within their MS Office suite of packages \cite{bibid}.
 While this has not yet materialized, there continues to be significant interest in using Python to manipulate not only Excel workbooks but other office documents such as Powerpoint presentations \cite{bibid}.
@@
 Learning good Python programming practices via learning to model in similar way that students learn good spreadsheet practices through building spreadsheet models using the W\&A textbook.
+
+https://www.bleepingcomputer.com/news/microsoft/microsoft-considers-adding-python-as-an-official-scripting-language-to-excel/
+
+https://www.reddit.com/r/excel/comments/lrw21h/will_ms_excel_ever_use_python_as_scripting/
 

### 2023-06-29 python_excel_ite.tex
working on module positioning

@@
 
-% TODO - Edit this paragraph
-One criticism that has been levied against data science programs has been the lack of attention to basic software engineering principles.
-The image of newly minted grads developing spaghetti coded Jupyter notebooks not nearly ready for being put into production has reached meme like status.
+Recently there has been more recognition of the need for data scientists to learn basic software engineering principes \cite{SoftwareEngineeringData,SoftwareEngineeringDataa}.
+The image of newly minted data science graduates developing spaghetti coded Jupyter notebooks not nearly ready for being put into production has reached \href{https://www.reddit.com/r/ProgrammerHumor/comments/y2xe3n/like_every_time_ever_when_the_devops_engineer/}{meme like status}.
 I hoped to address some of these issues in the EwP module.
@@
 This is really no different than the integration of advanced Excel functions and techniques with spreadsheet modeling tutorials.
-Just as spreadsheet modeling students were able to easily wow their coworkers and managers with INDEX-MATCH and intermediate Pivot Table skills, I hoped that my students ability to properly document a Python function and use git and GitHub would give them a little street cred.
+Just as spreadsheet modeling students were able to easily wow their coworkers and managers with INDEX-MATCH and intermediate Pivot Table skills, I hoped that my students ability to properly document a Python function and use git and GitHub would give them a bit of street credibility.
+
+We will begin with a brief discussion of the overall positioning and structure of the EwP module within our business analytics programs.
+Then, each of the three main EwP submodules are described in some detail.
+Finally we conclude with a brief discussion of our experience in delivering this module and plans for the future.
 
@@
 The students taking this course have already taken the PCDA course within which they learned fundamental Python programming within the context of data analytics.
-The PCDA course ends with intro to the sklearn library for doing predictive modeling in Python and the AAP course begins where the PCDA course ends.
-The EwP module could certainly be used as part of a semester long Python based analytics course.
-Most students have also already taken "Business Analytics" (\href{http://www.sba.oakland.edu/faculty/isken/courses/mis5460}{BA}) which is a spreadsheet based introduction to business analytics.
-We do not strictly require the spreadsheet based course be taken before the PCDA course as that course draws students from outside of our business programs who are strictly interested in programmatic tools such as Python.
+The PCDA course ends with intro to the \href{https://scikit-learn.org/stable/index.html}{scikit-learn library} for doing predictive modeling in Python and the AAP course begins where the PCDA course ends.
+The EwP module could also be used as part of a semester long Python based analytics course.
+Most AAP students have also already taken "Business Analytics" (\href{http://www.sba.oakland.edu/faculty/isken/courses/mis5460}{BA}) which is a spreadsheet based introduction to business analytics.
+We do not strictly require the spreadsheet based course be taken before the PCDA or AAP courses as that course draws students from outside of our business programs who are strictly interested in programmatic tools such as Python.
 The BA and PCDA courses have historically been offered in face to face mode in a computer teaching lab.
@@
 The newer AAP course has been offered since 2021 in an online asynchronous mode.
-All three courses have \href{URL}{extensive course websites} that are publicly accessible, including access to all of the video content and files needed.
-We will describe the AAP course web in more detail below in the context of the EwP module.
+All three courses have \href{http://www.sba.oakland.edu/Faculty/isken/teaching.html}{extensive course websites} that are publicly accessible, including access to all of the video content and files needed.
 
@@
 %TODO needs much editing and shaping
-All of the Python modules in the PCDA course as well as the entire AAP course are taught using a combination of Jupyter notebooks, Python scripts, and videos walking through the notebooks, all publicly accessible from extensive course web pages.
-GIVE URL.
-Some background on Jupyter notebooks for analytics - see downloaded papers.
-Questions and coding challenges sprinkled throughout the notebooks - answers at bottom.
-While AAP is online asynch, this same approach has been used very successfully in a computer teaching lab for the PCDA course.
-The screencasts are highly valued by the students as they can rewatch things as needed.
-Many are working professionals and a 2.5 hour evening class in a warm computer lab can be taxing.
 
-The course is taught using a combination of R Markdown documents \cite{bibid} and Jupyter notebooks \cite{bibid}.
-Both are examples of \textit{literate computing} \cite{Jupyter_proposal} and allow us to create educational narratives that support iterative exploration and development.
-The ability to weave standard explanatory language with formal programming languages has made notebook computing a defacto standard in the data science education community.
-All of the PCDA course content, including many Jupyter notebooks and supporting screencasts, is freely available from a publicly accessible course web site \url{URL}.
+All of the Python modules in the PCDA course as well as the entire AAP course are taught using a combination of Jupyter notebooks, Python scripts, and videos walking through the notebooks.
+Jupyter notebooks, part of \href{https://jupyter.org/}{Project Jupyter}, are an example of \textit{literate computing} \cite{Jupyter_proposal} and allow one to create educational narratives that support iterative exploration and development.
+The ability to weave standard explanatory language with formal programming languages has made notebook computing a defacto standard in the data science education community \cite{perkel2018jupyter}.
+Jupyter notebooks can also include equations written in LaTex, data visualizations, and other multi-media.
+The technology is free, open source, runs on all platforms, and has a huge user community that has created a large number of resources for students, faculty and industry practitioners.
 
-Jupyter notebooks for teaching data science are an established delivery mechanism.
-Computational notebook allowing mixing of explanatory text, written in Markdown with executable code (and its output) in Python or a number of other languages.
-Literate programming.
-Can also include equations, visualizations, other multi-media.
-It's free, open source, runs on all platforms, huge ecosystem, used in industry, used widely in education.
-Literate computing \cite{Jupyter_proposal}.
-Educational narratives that support iterative exploration and development.
-The ability to weave standard language with formal programming languages is critical to creating the desired educational artifact.
-Jupyter used by many universities as a key component of their educational technology stack.
+As will be described below, Jupyter notebooks
+
+Jupyter as an authoring environment - blog posts were written as Jupyter notebooks.
+
+%TODO - more on structure of my Jupyter notebooks for teaching
+
+%TODO - more on Jupyter project
 
@@
 There are two overarching learning objectives for this first set of Jupyter notebooks.
-Students learn how Python might be used to do something they are quite familiar with doing in a spreadsheet - model building and sensitivity analysis.
+Students learn how Python might be used to do something they are quite familiar with doing in a spreadsheet -- model building and sensitivity analysis.
 At the same time, students encounter, as they are needed, more advanced Python programming material.
@@
 
-https://www.bleepingcomputer.com/news/microsoft/microsoft-considers-adding-python-as-an-official-scripting-language-to-excel/
+%https://www.bleepingcomputer.com/news/microsoft/microsoft-considers-adding-python-as-an-official-scripting-language-to-excel/
 
-https://www.reddit.com/r/excel/comments/lrw21h/will_ms_excel_ever_use_python_as_scripting/
+%https://www.reddit.com/r/excel/comments/lrw21h/will_ms_excel_ever_use_python_as_scripting/
 

### 2023-07-07 python_excel_ite.tex
working on jupyter notebook basics

@@
 
-The PCDA course, like many data science courses, focuses mostly on data wrangling, exploratory data analysis (EDA), and statistical/machine learning based predictive models.
-It does include a Python based simulation of the Monte Hall 3-door problem.
+The PCDA course, like many introductory data science courses, focuses mostly on data wrangling, exploratory data analysis (EDA), and statistical/machine learning based predictive models.
 In teaching EDA using tools like ggplot2, students are challenged to think about how they would go about trying to use Excel to create some of the plots that are easy to do with R and Python based plotting libraries.
 There is also some limited coverage of using Python to open Excel files and do some sheet manipulation.
-However, there was no real attempt to do the kind of modeling that was the focus of spreadsheet based modeling textbooks.
+However, other than a Python based simulation of the Monte Hall 3-door problem, there was no real attempt to do the kind of modeling that was the focus of spreadsheet based modeling textbooks.
 
@@
 Python is widely used for data wrangling, visualization and predictive modeling using various flavors of regression, tree based models and neural networks.
-The blog \href{https://pbpython.com/}{Practical Business Python} written primarily by Chris Moffitt contributed to making Python more accessible to business analytics practitioners through high quality posts that were extremely practical and useful.
+The blog \href{https://pbpython.com/}{Practical Business Python} written primarily by Chris Moffitt contributed to making Python more accessible to business analytics practitioners through high quality posts showing practical use of Python for business analysis problems.
 But, it got me to thinking.
@@
 The newer AAP course has been offered since 2021 in an online asynchronous mode.
-All three courses have \href{http://www.sba.oakland.edu/Faculty/isken/teaching.html}{extensive course websites} that are publicly accessible, including access to all of the video content and files needed.
+All three courses have \href{http://www.sba.oakland.edu/Faculty/isken/teaching.html}{extensive course websites} that are publicly accessible, including access to all of the video content and supporting files.
 
@@
 The technology is free, open source, runs on all platforms, and has a huge user community that has created a large number of resources for students, faculty and industry practitioners.
+
+Central to Jupyter notebooks is a mix of \textit{code cells} and \textit{markdown cells}.
+Code cells contain executable code while markdown cells contain plain text using \href{URL}{markdown} for styling text.
+For example, Figure \ref{cells_edit_model} shows two markdown cells separated by a code cell as they look when in editing mode.
+After running the cells, they appear as show in Figure \ref{cells_after_running}
 

### 2023-07-14 python_excel_ite.tex
intro pretty much done

@@
 The late 1990's saw first editions of game changing textbooks \cite{winstonPracticalManagementScience2018,ragsdaleSpreadsheetModelingDecision2017} focused on the use of spreadsheets in teaching modeling and management science.
-A series of conferences focused on teaching management science (TMS) with spreadsheets started at Dartmouth in 1998 and several presentations from that conference made up the bulk of papers in the first two issues of the INFORMS Transactions on Education journal \cite{bibid}.
+A series of conferences focused on teaching management science (TMS) with spreadsheets started at Dartmouth in 1998 and several presentations from that conference made up the bulk of papers in the first two issues of the INFORMS Transactions on Education journal.
 The next few decades saw numerous textbooks \cite{BusinessAnalyticsArt, BusinessAnalytics4th} and teaching papers \cite{INFORMSTransactionsEducation2014} that prominently featured spreadsheet based modeling.
@@
 At the same time, the analytics world has been changing.
-The advent of programmatic analytics using tools like R and Python have changed the practice of analytics as well as the teaching landscape \cite{bibid}.
+The advent of programmatic analytics using tools like R and Python have changed the practice of analytics as well as the teaching landscape.
 Limitations of spreadsheets with respect to reproducibility and automation have certainly played a role.
@@
 Some of these courses have made a large amount of their course content freely available for self-learners and for other educators.
-At our institution, we have had a spreadsheet based modeling course in place since 2001.
-In 2015, we created a course called Practical Computing for Data Analytics (PCDA) that introduced both R and Python (along with a little Linux) for use in data analytics work.
+In 2015, we created a course called \href{http://www.sba.oakland.edu/faculty/isken/courses/mis5470/}{Practical Computing for Data Analytics} (PCDA) that introduced both R and Python (along with a little Linux) for use in data analytics work.
+This new course was meant as a follow up to the \href{http://www.sba.oakland.edu/faculty/isken/courses/mis5460/}{spreadsheet based modeling course} we have had in place since 2001.
 
@@
 In teaching EDA using tools like ggplot2, students are challenged to think about how they would go about trying to use Excel to create some of the plots that are easy to do with R and Python based plotting libraries.
-There is also some limited coverage of using Python to open Excel files and do some sheet manipulation.
-However, other than a Python based simulation of the Monte Hall 3-door problem, there was no real attempt to do the kind of modeling that was the focus of spreadsheet based modeling textbooks.
+There is some limited coverage of using Python to open Excel files and do worksheet manipulation.
+However, other than a Python based simulation of the Monte Hall 3-door problem, there was no real attempt to do the kind of modeling that is the focus of most spreadsheet based modeling textbooks.
 
-Python has continued to make inroads to analytical territory that was formerly dominated by Excel.
-The widely used Python data analysis library, pandas, was created by Wes McKinney while working as a quantitative financial analyst \cite{bibid}.
+Python has continued to make inroads to analytical territory that was formerly dominated by spreadsheets.
+The widely used Python data analysis library, pandas, was created by Wes McKinney while working as a quantitative financial analyst \cite{PythonDataAnalysis}.
 Python is widely used for data wrangling, visualization and predictive modeling using various flavors of regression, tree based models and neural networks.
-The blog \href{https://pbpython.com/}{Practical Business Python} written primarily by Chris Moffitt contributed to making Python more accessible to business analytics practitioners through high quality posts showing practical use of Python for business analysis problems.
+The blog \href{https://pbpython.com/}{Practical Business Python} written primarily by Chris Moffitt has contributed to making Python more accessible to business analytics practitioners through high quality posts showing practical use of Python for business analysis problems.
 But, it got me to thinking.
-What about those things for which Excel is well suited such as building formula based models and doing sensitivity analysis on these models?
+What about those things for which Excel is particularly well suited such as building formula based models and doing sensitivity analysis on these models?
 What would those look like in Python?
@@
 
-In 2021, I published a series of blog posts \cite{bibid} that explored this question and formed the basis of the teaching module described in this paper.
+In 2021, I published a \href{https://bitsofanalytics.org/#category=excel}{series of blog posts} that explored this question and formed the basis of the teaching module described in this paper.
 During the development of these posts, I came across the \href{https://quantecon.org/}{Quant Econ: open source code for economic modeling} project which \quote{is a nonprofit organization dedicated to development and documentation of open source computational tools for economics, econometrics, and decision making}.
-This is a mature project, run by renowned economists \cite{bibid}, with significant support.
+This is a mature project, run by renowned economists, with significant support.
 They have developed and released numerous high quality Jupyter notebooks exploring a range of computational economic and financial modeling topics.
-Here was exactly the kind Python for business usage I had been seeking.
+Here was an example of the Python for business usage I had been seeking.
 With this as inspiration, I decided to turn my blog post series into a teaching module called Excel with Python (EwP).
@@
 The EwP module is not as polished nor as advanced as the Quant Econ materials, but does add to the growing body of freely available teaching materials aimed at Python for business modeling.
-My goals for EwP are also a little different and include:
+My goals for EwP are also a little different than those of Quant Econ and include:
 
@@
 
-Recently there has been more recognition of the need for data scientists to learn basic software engineering principes \cite{SoftwareEngineeringData,SoftwareEngineeringDataa}.
+Recently there has been more recognition of the need for data scientists to learn basic software engineering principes \cite{SoftwareEngineeringData,SoftwareEngineeringDataa,BuildingReproducibleAnalytical}.
 The image of newly minted data science graduates developing spaghetti coded Jupyter notebooks not nearly ready for being put into production has reached \href{https://www.reddit.com/r/ProgrammerHumor/comments/y2xe3n/like_every_time_ever_when_the_devops_engineer/}{meme like status}.
@@
 This is really no different than the integration of advanced Excel functions and techniques with spreadsheet modeling tutorials.
-Just as spreadsheet modeling students were able to easily wow their coworkers and managers with INDEX-MATCH and intermediate Pivot Table skills, I hoped that my students ability to properly document a Python function and use git and GitHub would give them a bit of street credibility.
+Just as spreadsheet modeling students were able to easily wow their coworkers and managers with INDEX-MATCH and intermediate Pivot Table skills, I hoped that my students ability to properly document a Python function and use git and GitHub would give them a bit more professional credibility.
 

### 2023-07-14 python_excel_ite.tex
module positioning and structure section done

@@
 At our institution, The EwP module is part of a business course entitled ``Advanced Analytics with Python'' (\href{http://www.sba.oakland.edu/faculty/isken/courses/mis6900}{AAP}).
-The students taking this course have already taken the PCDA course within which they learned fundamental Python programming within the context of data analytics.
+The students taking this course have already taken the PCDA course within which they learned fundamental Python programming within the context of data analytics over seven weeks.
 The PCDA course ends with intro to the \href{https://scikit-learn.org/stable/index.html}{scikit-learn library} for doing predictive modeling in Python and the AAP course begins where the PCDA course ends.
@@
 
-What kind of Python background do students need to have in order to be ready for the EwP module? A basic familiarity with Python fundamentals for data analysis is needed, including:
+What kind of Python background do students need to have in order to be ready for the EwP module? A basic familiarity with Python fundamentals for data analysis is required, including:
 
@@
 The first of these focuses on building and using a typical spreadsheet model using Python.
-Then the functionality developed for doing this type of modeling in Python is deployed as a reusable package.
-We end with an introduction to using Python to manipulate Excel files.
-
-%TODO needs much editing and shaping
+Then the functionality developed in the first submodule is deployed as a reusable package.
+The final submodule is an introduction to using Python to manipulate Excel files.
 
@@
 The ability to weave standard explanatory language with formal programming languages has made notebook computing a defacto standard in the data science education community \cite{perkel2018jupyter}.
-Jupyter notebooks can also include equations written in LaTex, data visualizations, and other multi-media.
-The technology is free, open source, runs on all platforms, and has a huge user community that has created a large number of resources for students, faculty and industry practitioners.
+Jupyter notebooks can also include equations written in \LaTeX, data visualizations, and other multi-media.
+The technology is free, open source, runs on most computing platforms, and has a huge user community that has created a large number of resources for students, faculty and industry practitioners.
 
@@
 For example, Figure \ref{cells_edit_model} shows two markdown cells separated by a code cell as they look when in editing mode.
-After running the cells, they appear as show in Figure \ref{cells_after_running}
-
-As will be described below, Jupyter notebooks
-
-Jupyter as an authoring environment - blog posts were written as Jupyter notebooks.
-
-%TODO - more on structure of my Jupyter notebooks for teaching
-
-%TODO - more on Jupyter project
+After running the cells, they appear as show in Figure \ref{cells_after_running}.
+Mixing code and markdown cells makes it quite easy to create interactive tutorials in which students can not only run pre-written code, but can also add their own code - all with the aid of supporting explanatory text.
+All of the blog posts upon which EwP is based were written in Jupyter notebooks.
+There are countless introductory tutorials for Jupyter notebooks, but the official Project Jupyter Documentations \cite{ProjectJupyterDocumentation} is a very good place to start if you are new to notebook computing.
 

### 2023-07-14 python_excel_ite.tex
subsection notebook 1 pretty much done, needs plot

@@
 After running the cells, they appear as show in Figure \ref{cells_after_running}.
-Mixing code and markdown cells makes it quite easy to create interactive tutorials in which students can not only run pre-written code, but can also add their own code - all with the aid of supporting explanatory text.
+Mixing code and markdown cells makes it quite easy to create interactive tutorials in which students can not only run pre-written code, but can also add their own code -- all with the aid of supporting explanatory text.
 All of the blog posts upon which EwP is based were written in Jupyter notebooks.
@@
 Students learn how Python might be used to do something they are quite familiar with doing in a spreadsheet -- model building and sensitivity analysis.
-At the same time, students encounter, as they are needed, more advanced Python programming material.
+At the same time, students encounter, as needed, more advanced Python programming concepts and techniques.
 In this way, the computational problem to be solved drives the introduction of more advanced Python concepts and techniques instead of such things being presented in a vacuum.
-I make it a point to try to convey the notion that there are often multiple possible approaches by embarking on one path and there can be value to starting over and doing things in a different way.
-Each notebook is accompanied by videos that walk the students through the material and act as a guided tour with challenges to overcome along the way.
+I make it a point to show that there are often multiple possible approaches to a computational problem by being willing to abandon one approach and trying another.
+Emphasis is given to the lessons learned from each path explored.
+Every notebook is accompanied by screencasts that walk the students through the material and act as a guided tour with challenges to overcome along the way.
 The screenshot in \ref{fig:activities} shows the portion of the course web page that provides the high level roadmap through the notebooks and associated screencasts.
@@
 	\textsf{The details aren't so important right now as is the overall structure of the model.
-There's a few key inputs and some pretty straightforward formulas for computing cost, revenue and profit.
+There are a few key inputs and some pretty straightforward formulas for computing cost, revenue and profit.
 Notice the 1-way Data Table being used to explore how profit varies for different selling prices.
@@
 
-We then introduce a similar modeling problem to use throughout the rest of the notebook.
+We then introduce a similar modeling problem that will be the focus of the rest of the notebook.
 The Walton Bookstore problem appears in the chapter on Monte-Carlo simulation in textbooks that I have used in my spreadsheet modeling course \cite{albrightBusinessAnalyticsData2016,winstonPracticalManagementScience2018}.
@@
 
-The goal is to build a Python based version of this model that will allow us to do sensitivity analysis, find the break even point for demand and analyze the problem using Monte-Carlo simulation.
+The overall goal is to build a Python based version of this model that will allow us to do sensitivity analysis, find the break even point for demand and analyze the problem using Monte-Carlo simulation.
 
@@
 This difference in behavior between Jupyter notebooks and spreadsheets might be one of the most important concepts for students to internalize.
-A well known challenge associated with computational notebooks is the ease of running code cells out of order and introducing inadvertent errors \cite{bibid}.
-This point along with strategies to avoid it are discussed during the first week of the course which focuses on notebook computing best practices as well as an historical look at the use of computational notebooks \cite{bibid}.
-
+A well known challenge associated with computational notebooks is the ease of running code cells out of order and introducing inadvertent errors.
+This point along with strategies to avoid it are discussed during the first week of the course which focuses on \href{http://www.sba.oakland.edu/faculty/isken/courses/mis6900/jupyter_notebooks.html}{notebook computing best practices as well as a historical look} at computational notebooks.
 \subsubsection{Sensitivity analysis}
@@
 In a Data Table, a range of order quantities can be used as row or column input and one or more output values can be computed.
-In Python, the vectorized behavior of the language \cite{bibid} provides a simple way to accomplish the same thing.
+In Python, the vectorized nature of the NumPy library provides a simple way to accomplish the same thing.
 A range of order quantities is created as a NumPy array that can be used directly to compute vectors of all of the intermediate and final output variables.
@@
 For our initial design of a \code{BookstoreModel} class we make all of the base inputs class attributes (properties) and then add method attributes to compute outputs such as costs, revenues and profits.
-Students can easily see the mapping between things in the initial procedural model and this new object oriented version.
+Students can easily see the mapping between elements in the initial procedural model and this new object oriented version.
 Several important concepts and syntactical details of doing object oriented programming Python are highlighted.
@@
 
-Then a few design dilemmas are posed and some non-working approaches illustrated as we try to decide on how to implement an n-way data table function for this new object oriented model.
+A few design dilemmas are posed and some non-working approaches illustrated as we try to decide on how to implement an n-way data table function for this new object oriented model.
 These meta reflections are intentionally included to better mimic actual software development and reinforce the idea that software development is much more than simply writing code to implement a perfectly thought out design.
-We forge ahead with some preliminary ideas and are confronted with an intermediate problem related to generating a list of dictionaries that represent combinations of inputs (i.e.\ scenarios) to evaluate.
+We forge ahead with some preliminary ideas and are confronted with an intermediate problem related to generating a list of dictionaries that represent combinations of inputs, or \textit{scenarios}, to evaluate.
 This provides a great opportunity to remind students of one of the great strengths of the open source software ecosystem -- we can leverage work done in other packages and even look at actual source code to see how something was implemented.
-In this case, the well known scikit-learn package \cite{bibid} has a \code{ParameterGrid} function that solves our scenario generation problem quite nicely.
+In this case, the well known scikit-learn package \cite{ScikitlearnMachineLearning} has a \code{ParameterGrid} function that solves our scenario generation problem quite nicely.
 This example also reminds the students of the value of being able to make sense of API documentation and of object oriented code written by others.
@@
 This first notebook draws to a close with the creation of a \code{data\_table} function that allows us to do sensitivity analysis with the objected oriented model.
-Unlike Excel's Data Table tool, this \code{data\_table} function can handle an arbitrary number of inputs and outputs.
 The payoff is the creation of the following faceted plot showing how profit varies for different order quantities and demand levels.

### 2023-07-14 python_excel_ite.tex
subsection notebook 2 pretty much done

@@
 This provides a great opportunity to remind students of one of the great strengths of the open source software ecosystem -- we can leverage work done in other packages and even look at actual source code to see how something was implemented.
-In this case, the well known scikit-learn package \cite{ScikitlearnMachineLearning} has a \code{ParameterGrid} function that solves our scenario generation problem quite nicely.
+In this case, the well known \href{https://scikit-learn.org/stable/}{scikit-learn} package \cite{ScikitlearnMachineLearning} has a \code{ParameterGrid} function that solves our scenario generation problem quite nicely.
 This example also reminds the students of the value of being able to make sense of API documentation and of object oriented code written by others.
@@
 
-Before trying to build our \code{goal\_seek} function we take a brief detour into basic root finding algorithms through packages like SciPy \cite{bibid} as well as in standalone Python scripts that are widely available in the Python data science ecosystem \cite{bibid}.
+Before trying to build our \code{goal\_seek} function we take a brief detour into basic root finding algorithms through packages like \href{https://docs.scipy.org/doc/scipy/tutorial/optimize.html#root-finding}{SciPy} \cite{SciPy} as well as in \href{https://github.com/patrickwalls/mathematical-python/}{open source notebooks and scripts} that are widely available in the Python data science ecosystem \cite{wallsMathematicalPython2023}.
 We compare the output of Excel's Goal Seek with different starting values to the values obtained by different root finding algorithms implemented in Python.

### 2023-07-14 python_excel_ite.tex
working on last submodule

@@
 Python is widely used for data wrangling, visualization and predictive modeling using various flavors of regression, tree based models and neural networks.
-The blog \href{https://pbpython.com/}{Practical Business Python} written primarily by Chris Moffitt has contributed to making Python more accessible to business analytics practitioners through high quality posts showing practical use of Python for business analysis problems.
+You can get a very good sense of this by visiting the very popular \href{https://www.reddit.com/r/datascience/}{r/datascience subreddit on Reddit} -- a well known online forum.
 But, it got me to thinking.
@@
 In a Data Table, a range of order quantities can be used as row or column input and one or more output values can be computed.
-In Python, the vectorized nature of the NumPy library provides a simple way to accomplish the same thing.
+In Python, the vectorized nature of the \href{https://numpy.org/}{NumPy} library \cite{NumPy}provides a simple way to accomplish the same thing.
 A range of order quantities is created as a NumPy array that can be used directly to compute vectors of all of the intermediate and final output variables.
@@
 
-In the Excel modeling world, Monte-Carlo simulation can be done without add-ins, but packages like @Risk \cite{} can make the process much easier.
+In the Excel modeling world, Monte-Carlo simulation can be done without add-ins, but packages like @Risk \cite{RISKRiskAnalysis} can make the process much easier.
 In particular, random variable generation for a wide range of probability distributions is facilitated by add-ins like @Risk.
-In Python, the NumPy package provide random variate generation for a large number of distributions and similarly, SciPy provides functions for computing distribution related quantities for many probability distributions \cite{}.
-Much like we did when creating the \code{data\_table} function, we start with adding uncertainty to a single input variable in the bookstore model and rely on Python's vectorized computing capabilities to generate the profit associated with each realization of the random input.
+In Python, the NumPy package provide random variate generation for a large number of distributions and similarly, SciPy provides functions for computing distribution related quantities for many probability distributions.
+Much like we did when creating the \code{data\_table} function, we start with adding uncertainty to a single input variable in the bookstore model and rely on NumPy's vectorized computing capabilities to generate the profit associated with each realization of the random input.
 Then we can do standard statistical analysis of the simulation output using Python packages such as pandas and SciPy.
-In spreadsheet based simulation textbooks, the notion of the ``Flaw of Averages'' popularized by Sam Savage \cite{} is often illustrated.
+In spreadsheet based simulation textbooks, the notion of the ``Flaw of Averages'' popularized by Sam Savage \cite{savageFlawAveragesWhy2012} is often illustrated.
 We do the same with our Python based model, showing how replacing random demand by its mean can result in a wildly optimistic estimate of the mean profit found through analyzing the simulation output.
@@
 This is followed by a well commented first attempt at a \code{simulate} function.
-By this point, students are well equipped to understand the code as it builds on basic Python concepts and more advanced concepts that we have already covered in this EwP module.
-The rest of the notebook is spent trying out the \code{simulate} function and ends with creating a grouped boxplot that many of them have likely created before with @Risk along with a faceted plot of histograms of profit created with the popular Seaborn package.
+By this point, students are well equipped to understand the code as it builds on basic Python concepts and more advanced concepts that we have already covered in the EwP module.
+The rest of the notebook is spent trying out the \code{simulate} function and ends with creating a grouped boxplot that many of them have likely created before with @Risk along with a faceted plot of histograms of profit created with the popular \href{https://seaborn.pydata.org/}{Seaborn} package.
 
@@
 
-\section{Submodule 2: The packaging notebooks}
+\section{Submodule 2: The packaging and documentation notebooks}
+
+%TODO - Need transition opening from modeling to packaging for reusability
+
+Don't want to be copying and pasting code for the next model. We have spent some effort in creating something that is reusable.
+VBA for Modelers got into the deployment process in the Excel world.
+
+Classroom based spreadsheet modeling problems often end here. The model is built and used to analyze a given problem.
 
@@
 The first notebook in this submodule starts by explaining what Python packages are and their role in sharing code.
-We then revisit a concept from earlier in the course - creating a good project folder structure.
-For this we use what is known as a \code{cookiecutter} that can automatically generate a project folder structure and key files from the answers to a few prompts \cite{cookiecutter}.
+We then revisit a concept from earlier in the course -- creating a good project folder structure.
+For this we use what is known as a \href{https://cookiecutter.readthedocs.io/en/stable/}{cookiecutter} that can automatically generate a project folder structure and key files from the answers to a few prompts \cite{CookiecutterBetterProject}.
 There are a few different folder layouts used in Python projects and there is a bit of discussion on the folder structure we will use and resources given for exploring this issue further.
@@
 
-This notebook ends by pointing the students to another notebook in which our newly deployed whatif package is used for a completely different model.
-The example used is the New Car Simulation that has been a staple of Winston and Albright's textbooks for many years \cite{bibid}.
-This new model differs structurally from the Bookstore Model, but is representative of many financial models, in that cash flows over multiple years need to be computed and summarized.
-It provides a good test for our new whatif package.
+This notebook ends by pointing the students to another notebook in which our newly deployed whatif package is used for a completely different model -- the New Car Simulation that has been a staple of Winston and Albright's textbooks for many years \cite{winstonPracticalManagementScience2018,albrightBusinessAnalyticsData2016}.
+This new model differs structurally from the Bookstore Model, but is representative of many financial models in that cash flows over multiple years need to be computed and summarized.
 
@@
 This deployment focused submodule ends with a short notebook discussing the different types of documentation needed in a typical project.
-This includes code comments, docstrings, readme files and generating documentation by writing restrutured text files and using Sphinx.
-This is one of those tasks that few people like to do but can be very important for the long term success of a project \cite{bibid}.
+This includes code comments, docstrings, \code{readme} files and generating documentation by writing \href{https://www.sphinx-doc.org/en/master/usage/restructuredtext/basics.html}{reStructuredText} files (similar but more powerful than markdown) and using \href{https://www.sphinx-doc.org/en/master/index.html}{Sphinx}, the Python documentation generation tool.
+This is one of those tasks that few people like to do but can be very important for the long term success of a project.
 Even if you are just creating a tool for yourself, you will forget the details over time and documentation can be very beneficial in refreshing your memory.
@@
 Each of these examples are based on a real problems I encountered either in research or industrial projects.
-Several web based resources are shared including the Practical Business Python blog \cite{bibid} written by Chris Moffitt, a data analytics professional.
+Several web based resources are shared including the \href{https://pbpython.com/}{Practical Business Python blog} written by Chris Moffitt, a data analytics professional \cite{moffittPracticalBusinessPython}.
 This blog, which focuses on helping Excel centric analytics professionals transition some of their work to Python, includes numerous posts on different aspects of Python and Excel integration.
@@
 While this first example is not necessarily Excel related, using Excel to consolidate csv files through a series of manual Copy-Paste operations is something that many Excel users will admit to having done.
-After getting familiar with pathlib, a library for working with file systems which has recently been added to Python's standard library, students are led through the development of a short procedure that creates a list of the csv files to concatenate and then uses the pandas library to read and combine the files in a \code{DataFrame} and then writes out the consolidated data as a csv file.
+After getting familiar with pathlib, a library for working with file systems which has recently been added to Python's standard library, students are led through the development of a short procedure that creates a list of the csv files to concatenate and then uses pandas to read and combine the files in a \code{DataFrame} and writes out the consolidated data as a csv file.
 
@@
 Two different approaches are shown to consolidate data from multiple sheets into a single sheet.
-One approach uses the pandas library while the other relies on a library specifically created for working with Excel files from Python, openpyxl \cite{bibid}.
+One approach uses pandas while the other relies on a library specifically created for working with Excel files from Python, openpyxl \cite{bibid}.
 One advantage to using Python rather than Visual Basic for Applications for this task is that Python is well suited for working with data in many different formats including csv, json, xml and a myriad of database formats.

### 2023-07-14 python_excel_ite.tex
first draft almost done - see TODOs

@@
 Two different approaches are shown to consolidate data from multiple sheets into a single sheet.
-One approach uses pandas while the other relies on a library specifically created for working with Excel files from Python, openpyxl \cite{bibid}.
+One approach uses pandas while the other relies on a library specifically created for working with Excel files from Python, \href{https://openpyxl.readthedocs.io/en/stable/}{openpyxl}.
 One advantage to using Python rather than Visual Basic for Applications for this task is that Python is well suited for working with data in many different formats including csv, json, xml and a myriad of database formats.
@@
 
-Reshaping data \cite{bibid} from wide to long is a common data wrangling task especially in preparation for plotting.
+Reshaping, or tidying, data \cite{wickhamTidyData2014} from wide to long format is a common data wrangling task especially in preparation for plotting.
 In this example, not only do students have to grapple with reshaping the data in an Excel workbook, they also have to deal with the all too common problem of multi-row header lines in a spreadsheet.
@@
 
-Rumors have swirled for a number of years that Microsoft was considering adding Python as a first class language alternative to VBA within their MS Office suite of packages \cite{bibid}.
-While this has not yet materialized, there continues to be significant interest in using Python to manipulate not only Excel workbooks but other office documents such as Powerpoint presentations \cite{bibid}.
-Given that business students are very familiar with both Excel and Powerpoint, inclusion of this material in a business school Python based analytics course seems quite appropriate.
+Rumors have swirled for a number of years that Microsoft was considering adding Python as a first class language alternative to VBA within their MS Office suite of packages \cite{MicrosoftConsidersAdding}.
+While this has not yet materialized, there continues to be significant interest in using Python to manipulate not only Excel workbooks but other office documents such as Powerpoint presentations \cite{PythonpptxPythonpptx21}.
+Given that the ubiquity of both Excel and Powerpoint in business education and practice, inclusion of this material in a business school Python based analytics course seems quite appropriate.
 Students can learn more advanced and extremely useful Python concepts and techniques within a context with which they are intimately familiar.
-Manipulation of MS Office directly relevant to the job content of many of the students.
-This teaching module is a natural follow up to the commonly taught spreadsheet based analytics courses found in many business schools and provides an interesting way to teach and learn Python for analytics.
-
-\section{Random notes}
-Learning good Python programming practices via learning to model in similar way that students learn good spreadsheet practices through building spreadsheet models using the W\&A textbook.
-
-%https://www.bleepingcomputer.com/news/microsoft/microsoft-considers-adding-python-as-an-official-scripting-language-to-excel/
-
-%https://www.reddit.com/r/excel/comments/lrw21h/will_ms_excel_ever_use_python_as_scripting/
+Manipulation of MS Office documents is directly relevant to the job content of many of the students.
+This teaching module is a natural complement to the commonly taught spreadsheet based analytics and introductory data science courses found in many business schools and provides an interesting way to teach and learn Python for analytics.
 

### 2023-07-18 python_excel_ite.tex
need software avail section and to fix refs with no authors

@@
 
-Spreadsheet based modeling has transformed the teaching of management science in business schools over the past three decades.
+Spreadsheet based modeling has transformed the teaching of management science in business schools over the past twenty plus years.
 The late 1990's saw first editions of game changing textbooks \cite{winstonPracticalManagementScience2018,ragsdaleSpreadsheetModelingDecision2017} focused on the use of spreadsheets in teaching modeling and management science.
@@
 
-% TODO - add output listing and plot
-
 \begin{tcolorbox}
@@
 \end{tcolorbox}
+
+While laid out differently, this is the same output produced by a 2-way Data Table in Excel. At this point, we pause to consider a higher level design question.
 
@@
 
-%TODO - Need transition opening from modeling to packaging for reusability
-
-Don't want to be copying and pasting code for the next model. We have spent some effort in creating something that is reusable.
-VBA for Modelers got into the deployment process in the Excel world.
-
-Classroom based spreadsheet modeling problems often end here. The model is built and used to analyze a given problem.
+Now that we have created some useful modeling functionality, our goal is to make it easier for a modeler to actually use it.
+Certainly we do not want to rely on copying, pasting and modifying code when building a new model.
+In the Python world, the natural next step is to deploy our code as a Python \textit{package} \cite{buezentomasWelcomePythonPackages} that can then be imported like any other Python library.
 
@@
 This will require some design changes to our software and will also involve moving code out of Jupyter notebooks and into Python script files.
-While Jupyter notebooks can be integrated into production environments \cite{netflix}, it is much more common for code to live in Python script files.
+While Jupyter notebooks can be integrated into production environments \cite{uffordInteractiveNotebookInnovation2019}, it is much more common for code to live in Python script files.
 In addition to creating a deployable package, this submodule also discusses the importance of documentation in its many forms in a Python project.
-
-% REMOVED
-% Basic software engineering practices are often overlooked in business analytics classes but are certainly desirable in the professional world.
 
@@
 
+\section{Software availability}
+
+%TODO - write this section and decide if any links need to also be added in body of paper above
+
 \section{Classroom experience and the future}
@@
 This teaching module is a natural complement to the commonly taught spreadsheet based analytics and introductory data science courses found in many business schools and provides an interesting way to teach and learn Python for analytics.
+
+%TODO - Fix all of the bib entries with no authors. It's mostly websites.
 

### 2023-07-19 python_excel_ite.tex
drafted software section

@@
 
-There are two overarching learning objectives for this first set of Jupyter notebooks.
+There are two overarching learning objectives for this \href{http://www.sba.oakland.edu/faculty/isken/courses/mis6900/mod3a_whatif.html}{first EwP submodule}.
 Students learn how Python might be used to do something they are quite familiar with doing in a spreadsheet -- model building and sensitivity analysis.
@@
 Thinking about how others will interact with your software usually leads to better designed software.
-In this submodule we create a Python package that can be imported and used by other modelers.
+In \href{http://www.sba.oakland.edu/faculty/isken/courses/mis6900/mod3b_whatif_packaging.html}{this submodule} we create a Python package that can be imported and used by other modelers.
 This will require some design changes to our software and will also involve moving code out of Jupyter notebooks and into Python script files.
@@
 
-The EwP module ends with what has become a relatively common use of Python with Excel -- automating various Excel data wrangling tasks. Examples include:
+The \href{http://www.sba.oakland.edu/faculty/isken/courses/mis6900/mod3c_python_excel.html}{EwP module ends} with what has become a relatively common use of Python with Excel -- automating various Excel data wrangling tasks.
+Examples include:
 
@@
 
-%TODO - write this section and decide if any links need to also be added in body of paper above
+Each of the submodules described above has its own dedicated course web page.
+Links to each submodule page can be found on \href{http://www.sba.oakland.edu/faculty/isken/courses/mis6900/mod3_excel_with_python.html}{this EwP landing page} within the AAP course website.
+Each page contains learning objectives, a link to a compressed file containing the Jupyter notebooks and other supporting files, a set of activities with accompanying screen casts, and links to additional web based materials to explore.
+
+There is also a [whatif Github repository](https://github.com/misken/whatif) for the \code{whatif} package. It also contains the original notebooks used in the blog posts that motivated this project.
 
@@
 This module has been included in my Advanced Analytics with Python course for the past three years.
-The course is offered in a seven-week accelerated format during May and June.
-It is an online and asynchronous course.
+The course is offered in a seven-week accelerated format during the summer in an online and asynchronous mode.
 The EwP module is covered over the course of two weeks and is the subject of one of the major homework assignments for the semester.
@@
 The module has been well received and several of the working students have reported back to me that they have found use cases for these techniques in their day to day analytical activities.
+
+%TODO - how to link to assignment
 

### 2023-07-19 python_excel_ite.tex
fixed missing bib items, now editing

@@
 A series of conferences focused on teaching management science (TMS) with spreadsheets started at Dartmouth in 1998 and several presentations from that conference made up the bulk of papers in the first two issues of the INFORMS Transactions on Education journal.
-The next few decades saw numerous textbooks \cite{BusinessAnalyticsArt, BusinessAnalytics4th} and teaching papers \cite{INFORMSTransactionsEducation2014} that prominently featured spreadsheet based modeling.
+The next few decades saw numerous textbooks \cite{powellstepheng.BusinessAnalyticsArt, camm2020business} and teaching papers \cite{INFORMSTransactionsEducation2014} that prominently featured spreadsheet based modeling.
 
@@
 Python has continued to make inroads to analytical territory that was formerly dominated by spreadsheets.
-The widely used Python data analysis library, pandas, was created by Wes McKinney while working as a quantitative financial analyst \cite{PythonDataAnalysis}.
+The widely used Python data analysis library, pandas, was created by Wes McKinney while working as a quantitative financial analyst \cite{mckinney2022python}.
 Python is widely used for data wrangling, visualization and predictive modeling using various flavors of regression, tree based models and neural networks.
@@
 In 2021, I published a \href{https://bitsofanalytics.org/#category=excel}{series of blog posts} that explored this question and formed the basis of the teaching module described in this paper.
-During the development of these posts, I came across the \href{https://quantecon.org/}{Quant Econ: open source code for economic modeling} project which \quote{is a nonprofit organization dedicated to development and documentation of open source computational tools for economics, econometrics, and decision making}.
+During the development of these posts, I came across the \href{https://quantecon.org/}{Quant Econ: open source code for economic modeling} project which ``is a nonprofit organization dedicated to development and documentation of open source computational tools for economics, econometrics, and decision making''.
 This is a mature project, run by renowned economists, with significant support.
@@
 
-Recently there has been more recognition of the need for data scientists to learn basic software engineering principes \cite{SoftwareEngineeringData,SoftwareEngineeringDataa,BuildingReproducibleAnalytical}.
+Recently there has been more recognition of the need for data scientists to learn basic software engineering principes \cite{nelsoncatherineSoftwareEngineeringData2024,treadwayandrewSoftwareEngineeringData2023,rodriguesBuildingReproducibleAnalytical2023}.
 The image of newly minted data science graduates developing spaghetti coded Jupyter notebooks not nearly ready for being put into production has reached \href{https://www.reddit.com/r/ProgrammerHumor/comments/y2xe3n/like_every_time_ever_when_the_devops_engineer/}{meme like status}.
@@
 
-At our institution, The EwP module is part of a business course entitled ``Advanced Analytics with Python'' (\href{http://www.sba.oakland.edu/faculty/isken/courses/mis6900}{AAP}).
+At our institution, The EwP module is part of a business course entitled  \href{http://www.sba.oakland.edu/faculty/isken/courses/mis6900}{Advanced Analytics with Python} (AAP).
 The students taking this course have already taken the PCDA course within which they learned fundamental Python programming within the context of data analytics over seven weeks.
@@
 The EwP module could also be used as part of a semester long Python based analytics course.
-Most AAP students have also already taken "Business Analytics" (\href{http://www.sba.oakland.edu/faculty/isken/courses/mis5460}{BA}) which is a spreadsheet based introduction to business analytics.
+Most AAP students have also already taken   \href{http://www.sba.oakland.edu/faculty/isken/courses/mis5460}{Business Analytics} (BA) which is a spreadsheet based introduction to business analytics.
 We do not strictly require the spreadsheet based course be taken before the PCDA or AAP courses as that course draws students from outside of our business programs who are strictly interested in programmatic tools such as Python.
@@
 All of the Python modules in the PCDA course as well as the entire AAP course are taught using a combination of Jupyter notebooks, Python scripts, and videos walking through the notebooks.
-Jupyter notebooks, part of \href{https://jupyter.org/}{Project Jupyter}, are an example of \textit{literate computing} \cite{Jupyter_proposal} and allow one to create educational narratives that support iterative exploration and development.
+Jupyter notebooks, part of \href{https://jupyter.org/}{Project Jupyter}, are an example of \textit{literate computing} \cite{perez2015project} and allow one to create educational narratives that support iterative exploration and development.
 The ability to weave standard explanatory language with formal programming languages has made notebook computing a defacto standard in the data science education community \cite{perkel2018jupyter}.
@@
 Code cells contain executable code while markdown cells contain plain text using \href{URL}{markdown} for styling text.
-For example, Figure \ref{cells_edit_model} shows two markdown cells separated by a code cell as they look when in editing mode.
-After running the cells, they appear as show in Figure \ref{cells_after_running}.
+For example, Figure \ref{fig:cells_edit_mode} shows two markdown cells separated by a code cell as they look when in editing mode.
+After running the cells, they appear as show in Figure \ref{fig:cells_after_running}.
 Mixing code and markdown cells makes it quite easy to create interactive tutorials in which students can not only run pre-written code, but can also add their own code -- all with the aid of supporting explanatory text.
 All of the blog posts upon which EwP is based were written in Jupyter notebooks.
-There are countless introductory tutorials for Jupyter notebooks, but the official Project Jupyter Documentations \cite{ProjectJupyterDocumentation} is a very good place to start if you are new to notebook computing.
+There are countless introductory tutorials for Jupyter notebooks, but the official Project Jupyter Documentations \cite{jupyterteamProjectJupyterDocumentation2015} is a very good place to start if you are new to notebook computing.
+
+\begin{figure*}[!htbp]
+	\centering
+	\caption{Notebook cells in edit model}
+	\includegraphics[scale=1.0]{images/cells_edit_mode.png}
+	\label{fig:cells_edit_mode}
+\end{figure*}
+
+\begin{figure*}[!htbp]
+	\centering
+	\caption{Notebook cells after execution}
+	\includegraphics[scale=1.0]{images/cells_after_running.png}
+	\label{fig:cells_after_running}
+\end{figure*}
 
@@
 Every notebook is accompanied by screencasts that walk the students through the material and act as a guided tour with challenges to overcome along the way.
-The screenshot in \ref{fig:activities} shows the portion of the course web page that provides the high level roadmap through the notebooks and associated screencasts.
 
@@
 This point along with strategies to avoid it are discussed during the first week of the course which focuses on \href{http://www.sba.oakland.edu/faculty/isken/courses/mis6900/jupyter_notebooks.html}{notebook computing best practices as well as a historical look} at computational notebooks.
+
 \subsubsection{Sensitivity analysis}
@@
 In a Data Table, a range of order quantities can be used as row or column input and one or more output values can be computed.
-In Python, the vectorized nature of the \href{https://numpy.org/}{NumPy} library \cite{NumPy}provides a simple way to accomplish the same thing.
+In Python, the vectorized nature of the \href{https://numpy.org/}{NumPy} library \cite{harris2020array} provides a simple way to accomplish the same thing.
 A range of order quantities is created as a NumPy array that can be used directly to compute vectors of all of the intermediate and final output variables.
@@
 This provides a great opportunity to remind students of one of the great strengths of the open source software ecosystem -- we can leverage work done in other packages and even look at actual source code to see how something was implemented.
-In this case, the well known \href{https://scikit-learn.org/stable/}{scikit-learn} package \cite{ScikitlearnMachineLearning} has a \code{ParameterGrid} function that solves our scenario generation problem quite nicely.
+In this case, the well known \href{https://scikit-learn.org/stable/}{scikit-learn} package \cite{scikit-learn} has a \code{ParameterGrid} function that solves our scenario generation problem quite nicely.
 This example also reminds the students of the value of being able to make sense of API documentation and of object oriented code written by others.
@@
 
-Before trying to build our \code{goal\_seek} function we take a brief detour into basic root finding algorithms through packages like \href{https://docs.scipy.org/doc/scipy/tutorial/optimize.html#root-finding}{SciPy} \cite{SciPy} as well as in \href{https://github.com/patrickwalls/mathematical-python/}{open source notebooks and scripts} that are widely available in the Python data science ecosystem \cite{wallsMathematicalPython2023}.
+Before trying to build our \code{goal\_seek} function we take a brief detour into basic root finding algorithms through packages like \href{https://docs.scipy.org/doc/scipy/tutorial/optimize.html#root-finding}{SciPy} \cite{2020SciPy-NMeth} as well as in \href{https://github.com/patrickwalls/mathematical-python/}{open source notebooks and scripts} that are widely available in the Python data science ecosystem \cite{wallsMathematicalPython2023}.
 We compare the output of Excel's Goal Seek with different starting values to the values obtained by different root finding algorithms implemented in Python.
@@
 
-In the Excel modeling world, Monte-Carlo simulation can be done without add-ins, but packages like @Risk \cite{RISKRiskAnalysis} can make the process much easier.
+In the Excel modeling world, Monte-Carlo simulation can be done without add-ins, but packages like @Risk \cite{palisadeRISKRiskAnalysis} can make the process much easier.
 In particular, random variable generation for a wide range of probability distributions is facilitated by add-ins like @Risk.
@@
 By this point, students are well equipped to understand the code as it builds on basic Python concepts and more advanced concepts that we have already covered in the EwP module.
-The rest of the notebook is spent trying out the \code{simulate} function and ends with creating a grouped boxplot that many of them have likely created before with @Risk along with a faceted plot of histograms of profit created with the popular \href{https://seaborn.pydata.org/}{Seaborn} package.
+The rest of the notebook is spent trying out the \code{simulate} function and ends with creating a grouped boxplot, see Figure \ref{fig:boxplot} that many of them have likely created before with @Risk along with a faceted plot of histograms of profit created with the popular \href{https://seaborn.pydata.org/}{Seaborn} package.
 
@@
 	\caption{Profit boxplots by order quantity}
-	\label{fig:2way_dt}
+	\label{fig:boxplot}
 \end{figure}
@@
 We then revisit a concept from earlier in the course -- creating a good project folder structure.
-For this we use what is known as a \href{https://cookiecutter.readthedocs.io/en/stable/}{cookiecutter} that can automatically generate a project folder structure and key files from the answers to a few prompts \cite{CookiecutterBetterProject}.
+For this we use what is known as a \href{https://cookiecutter.readthedocs.io/en/stable/}{cookiecutter} that can automatically generate a project folder structure and key files from the answers to a few prompts \cite{audreyroyandcookiecuttercommunityCookiecutterBetterProject2012}.
 There are a few different folder layouts used in Python projects and there is a bit of discussion on the folder structure we will use and resources given for exploring this issue further.
@@
 The ranges used in the formulas vary by sheet and require determination of the number of rows in each sheet of data and use that information to construct the appropriate formula.
-An example assignment is available from \href{some.url}{here}.
+An example assignment is available from \href{https://drive.google.com/file/d/1prf8acZXZ4RoYBgQ8e9-eijA65uV7T-V/view?usp=sharing}{here}.
 The module has been well received and several of the working students have reported back to me that they have found use cases for these techniques in their day to day analytical activities.
 
-%TODO - how to link to assignment
-
-Rumors have swirled for a number of years that Microsoft was considering adding Python as a first class language alternative to VBA within their MS Office suite of packages \cite{MicrosoftConsidersAdding}.
-While this has not yet materialized, there continues to be significant interest in using Python to manipulate not only Excel workbooks but other office documents such as Powerpoint presentations \cite{PythonpptxPythonpptx21}.
+Rumors have swirled for a number of years that Microsoft was considering adding Python as a first class language alternative to VBA within their MS Office suite of packages \cite{cimpanucatalinMicrosoftConsidersAdding}.
+While this has not yet materialized, there continues to be significant interest in using Python to manipulate not only Excel workbooks but other office documents such as Powerpoint presentations \cite{cannystevePythonpptx2013}.
 Given that the ubiquity of both Excel and Powerpoint in business education and practice, inclusion of this material in a business school Python based analytics course seems quite appropriate.
@@
 This teaching module is a natural complement to the commonly taught spreadsheet based analytics and introductory data science courses found in many business schools and provides an interesting way to teach and learn Python for analytics.
-
-%TODO - Fix all of the bib entries with no authors. It's mostly websites.
 

### 2023-07-19 python_excel_ite.tex
updated readme

@@
 	\caption{Notebook cells in edit model}
-	\includegraphics[scale=1.0]{images/cells_edit_mode.png}
+	\includegraphics[scale=0.5]{images/cells_edit_mode.png}
 	\label{fig:cells_edit_mode}
@@
 	\caption{Notebook cells after execution}
-	\includegraphics[scale=1.0]{images/cells_after_running.png}
+	\includegraphics[scale=0.5]{images/cells_after_running.png}
 	\label{fig:cells_after_running}
@@
 
-There is also a [whatif Github repository](https://github.com/misken/whatif) for the \code{whatif} package. It also contains the original notebooks used in the blog posts that motivated this project.
+There is also a \href{https://github.com/misken/whatif}{whatif Github repository} for the \code{whatif} package.
+It also contains the original notebooks used in the blog posts that motivated this project.
 

### 2023-07-19 python_excel_ite.tex
more citation cleaning

@@
 A series of conferences focused on teaching management science (TMS) with spreadsheets started at Dartmouth in 1998 and several presentations from that conference made up the bulk of papers in the first two issues of the INFORMS Transactions on Education journal.
-The next few decades saw numerous textbooks \cite{powellstepheng.BusinessAnalyticsArt, camm2020business} and teaching papers \cite{INFORMSTransactionsEducation2014} that prominently featured spreadsheet based modeling.
+The next few decades saw numerous textbooks \cite{powellBusinessAnalyticsArt, camm2020business} and teaching papers \cite{INFORMSTransactionsEducation2014} that prominently featured spreadsheet based modeling.
 
@@
 
-Recently there has been more recognition of the need for data scientists to learn basic software engineering principes \cite{nelsoncatherineSoftwareEngineeringData2024,treadwayandrewSoftwareEngineeringData2023,rodriguesBuildingReproducibleAnalytical2023}.
+Recently there has been more recognition of the need for data scientists to learn basic software engineering principes \cite{nelsonSoftwareEngineeringData2024,treadwaySoftwareEngineeringData2023,rodriguesBuildingReproducibleAnalytical2023}.
 The image of newly minted data science graduates developing spaghetti coded Jupyter notebooks not nearly ready for being put into production has reached \href{https://www.reddit.com/r/ProgrammerHumor/comments/y2xe3n/like_every_time_ever_when_the_devops_engineer/}{meme like status}.
@@
 Certainly we do not want to rely on copying, pasting and modifying code when building a new model.
-In the Python world, the natural next step is to deploy our code as a Python \textit{package} \cite{buezentomasWelcomePythonPackages} that can then be imported like any other Python library.
+In the Python world, the natural next step is to deploy our code as a Python \textit{package} \cite{buezenWelcomePythonPackages} that can then be imported like any other Python library.
 
@@
 Each of these examples are based on a real problems I encountered either in research or industrial projects.
-Several web based resources are shared including the \href{https://pbpython.com/}{Practical Business Python blog} written by Chris Moffitt, a data analytics professional \cite{moffittPracticalBusinessPython}.
+Several web based resources are shared including the \href{https://pbpython.com/}{Practical Business Python blog} written by Chris Moffitt, a data analytics professional \cite{moffittPracticalBusinessPython2022}.
 This blog, which focuses on helping Excel centric analytics professionals transition some of their work to Python, includes numerous posts on different aspects of Python and Excel integration.
@@
 The examples above barely scratch the surface of the use cases for Python integration with Excel.
-Our notebook ends with a shout out to the highly regarding Practical Business with Python blog and a list of specific examples that students can explore on their own.
+Our notebook ends with a shout out to the highly regarding \href{https://pbpython.com/}{Practical Business with Python} blog and a list of specific examples that students can explore on their own.
 
@@
 
-Rumors have swirled for a number of years that Microsoft was considering adding Python as a first class language alternative to VBA within their MS Office suite of packages \cite{cimpanucatalinMicrosoftConsidersAdding}.
-While this has not yet materialized, there continues to be significant interest in using Python to manipulate not only Excel workbooks but other office documents such as Powerpoint presentations \cite{cannystevePythonpptx2013}.
+Rumors have swirled for a number of years that Microsoft was considering adding Python as a first class language alternative to VBA within their MS Office suite of packages \cite{cimpanuMicrosoftConsidersAdding2017}.
+While this has not yet materialized, there continues to be significant interest in using Python to manipulate not only Excel workbooks but other office documents such as Powerpoint presentations \cite{cannyPythonpptx2013}.
 Given that the ubiquity of both Excel and Powerpoint in business education and practice, inclusion of this material in a business school Python based analytics course seems quite appropriate.

### 2023-07-27 python_excel_ite.tex
converting tcolorboxes to screenshot pngs

@@
 The late 1990's saw first editions of game changing textbooks \cite{winstonPracticalManagementScience2018,ragsdaleSpreadsheetModelingDecision2017} focused on the use of spreadsheets in teaching modeling and management science.
-A series of conferences focused on teaching management science (TMS) with spreadsheets started at Dartmouth in 1998 and several presentations from that conference made up the bulk of papers in the first two issues of the INFORMS Transactions on Education journal.
-The next few decades saw numerous textbooks \cite{powellBusinessAnalyticsArt, camm2020business} and teaching papers \cite{INFORMSTransactionsEducation2014} that prominently featured spreadsheet based modeling.
+A series of conferences focused on teaching management science with spreadsheets started at Dartmouth in 1998 and several presentations from that conference made up the bulk of papers in the first two issues of the INFORMS Transactions on Education (ITE) journal.
+The next few decades saw numerous textbooks \cite{powellBusinessAnalyticsArt, camm2020business} and teaching papers in journals such as \href{https://pubsonline.informs.org/journal/ited}{ITE}that prominently featured spreadsheet based modeling.
 
-Despite numerous calls for and predictions of the eradication of spreadsheets from the business world, Excel and other spreadsheets are alive and well.
+Despite numerous calls for, and predictions of, the eradication of spreadsheets from the business world, Excel and other spreadsheets are \href{https://benn.substack.com/p/the-next-billion-programmers}{alive and well}.
 Yes, spreadsheets have well known limitations.
@@
 There are better dashboarding tools that are more easily maintained in a production environment.
-Spreadsheet errors can be a huge problem.
-Version control and documentation present challenges.
-There are data size limitations and external data links present their own set of challenges.
+Spreadsheet errors can be a \href{https://eusprig.org/research-info/horror-stories/}{huge problem}.
+Version control is difficult as both data and logic live in the same spreadsheet document.
+Documentation of multi-step manual data manipulation processes present challenges, it it is done at all -- hindering reproducibility and automation.
+There are data size limitations and external data links present their \href{https://www.thespreadsheetguru.com/find-remove-external-links/}{own set of challenges}.
 But, Excel continues to be used for a wide range of quantitative analysis related activities throughout the business world.
 One particular use is the building and exercising of models of business problems
-to explore the impact of various model inputs on key outputs.
+to explore the impact of important model inputs on different output metrics.
 This ``what if?'' type modeling is facilitated by Excel's flexibility, numerous built in functions, integrated sensitivity analysis via \href{https://support.microsoft.com/en-us/office/calculate-multiple-results-by-using-a-data-table-e95e2487-6ca6-4413-ad12-77542a5ea50b}{Data Tables},  \href{https://support.microsoft.com/en-us/office/use-goal-seek-to-find-the-result-you-want-by-adjusting-an-input-value-320cb99e-f4a4-417f-b1c3-4f369d6e66c7}{Goal
@@
 At the same time, the analytics world has been changing.
-The advent of programmatic analytics using tools like R and Python have changed the practice of analytics as well as the teaching landscape.
+The advent of programmatic analytics using tools like \href{https://www.r-project.org/}{R} and \href{https://www.python.org/}{Python} have changed the practice of analytics as well as the teaching landscape.
 Limitations of spreadsheets with respect to reproducibility and automation have certainly played a role.
 Additionally, both R and Python are free, open source, and have a rich set of analytics related libraries that have contributed to their growing popularity.
+There is a widely used Python distribution, \href{https://www.anaconda.com/download}{Anaconda}, that specifically targets analytics and data science work.
 Python, being a general purpose programming language, also offers advantages when moving models and analytical analysis pipelines from the prototype stage into production.
@@
 
-Numerous R and Python based data science courses have been developed and offered at universities and through online learning platforms such as Coursera, EdX, Udemy and others.
-Some of these courses have made a large amount of their course content freely available for self-learners and for other educators.
-In 2015, we created a course called \href{http://www.sba.oakland.edu/faculty/isken/courses/mis5470/}{Practical Computing for Data Analytics} (PCDA) that introduced both R and Python (along with a little Linux) for use in data analytics work.
+Numerous R and Python based data science courses have been developed and offered at universities and through online learning platforms such as \href{https://www.coursera.org/browse/data-science}{Coursera}, \href{https://www.edx.org/learn/data-science}{EdX}, \href{https://www.datacamp.com/}{DataCamp}, \href{https://www.udemy.com/topic/data-analysis/}{Udemy} and many others.
+In 2015, at Oakland University, we created a course called \href{http://www.sba.oakland.edu/faculty/isken/courses/mis5470/}{Practical Computing for Data Analytics} (PCDA) that introduced both R and Python (along with a little Linux) for use in data analytics work.
 This new course was meant as a follow up to the \href{http://www.sba.oakland.edu/faculty/isken/courses/mis5460/}{spreadsheet based modeling course} we have had in place since 2001.
@@
 Recently there has been more recognition of the need for data scientists to learn basic software engineering principes \cite{nelsonSoftwareEngineeringData2024,treadwaySoftwareEngineeringData2023,rodriguesBuildingReproducibleAnalytical2023}.
-The image of newly minted data science graduates developing spaghetti coded Jupyter notebooks not nearly ready for being put into production has reached \href{https://www.reddit.com/r/ProgrammerHumor/comments/y2xe3n/like_every_time_ever_when_the_devops_engineer/}{meme like status}.
+The image of newly minted data science graduates developing spaghetti coded Jupyter notebooks not nearly ready for being put into production has reached \href{https://www.reddit.com/r/ProgrammerHumor/comments/y2xe3n/like_every_time_ever_when_the_devops_engineer/}{meme worthy status}.
 I hoped to address some of these issues in the EwP module.
-Sneakily, students would be exposed to some simple software engineering concepts during the process of building and analyzing familiar models.
+Students would be exposed to some simple software engineering concepts during the process of building and analyzing familiar models.
 This is really no different than the integration of advanced Excel functions and techniques with spreadsheet modeling tutorials.
@@
 
-We will begin with a brief discussion of the overall positioning and structure of the EwP module within our business analytics programs.
+We will begin with a brief discussion of the overall positioning and structure of the EwP module within our business analytics programs at Oakland University.
 Then, each of the three main EwP submodules are described in some detail.
-Finally we conclude with a brief discussion of our experience in delivering this module and plans for the future.
+Finally we conclude with a brief discussion of our experience in delivering this module.
 
@@
 
-The stage is set with an overview of a typical spreadsheet based model that most of the students have seen before. See Figure \ref{fig:xl_model} for a screenshot of the Excel model.
+The stage is set with an overview of a typical spreadsheet based model that most of the students have seen before -- see Figure \ref{fig:stage}.
+The notebook also includes a screenshot (Figure \ref{fig:xl_model}) of a typical Excel model.
 
-\begin{tcolorbox}
+\begin{tcolorbox}[boxrule=1pt,sharp corners]
 \begin{quotation}
@@
 Students are given skeleton code as scaffolding and must finish the expressions for computing key intermediate outputs (\code{order\_cost, sales\_revenue, refund\_revenue}) and the final output, \code{profit}.
-Answers are provided at the bottom of the notebook so that students can attempt to finish the code but have a resource to correctly complete the code and move on.
-The screencast associated with the notebook also shows me completing the code in a subsequent code cell.
-Using skeleton code in this way has worked well both for lab based versions of this type of course as well as for online asynchronous versions.
 
@@
 
+\begin{figure*}[!htbp]
+	\centering
+	\caption{Instantiating model inputs}
+	\includegraphics[scale=0.5]{images/model_inputs.png}
+	\label{fig:model_inputs}
+\end{figure*}
+
+Answers are provided at the bottom of the notebook so that students can attempt to finish the code but have a resource to correctly complete the code and move on.
+The screencast associated with the notebook also shows the code being completed in a subsequent code cell.
+Using skeleton code in this way has worked well both for lab based versions of this type of course as well as for online asynchronous versions.
+For a lab based course, students actively work on completing the code during class and the instructor can move around the lab acting as a guide and consultant.
+
 At this point we discuss a fundamental difference between the computing model of Excel and that of Jupyter notebooks -- while spreadsheets respond automatically to changes, Jupyter notebooks require rerunning all code cells that include or follow the changed code.
 This difference in behavior between Jupyter notebooks and spreadsheets might be one of the most important concepts for students to internalize.
-A well known challenge associated with computational notebooks is the ease of running code cells out of order and introducing inadvertent errors.
+A well known challenge associated with computational notebooks is the ease of running code cells out of order and introducing inadvertent errors \cite{grusDonNotebooks2018}.
 This point along with strategies to avoid it are discussed during the first week of the course which focuses on \href{http://www.sba.oakland.edu/faculty/isken/courses/mis6900/jupyter_notebooks.html}{notebook computing best practices as well as a historical look} at computational notebooks.

### 2023-07-27 python_excel_ite.tex
added refs for good jupyter notebook practices and reworked those paragraphs

@@
 All of the Python modules in the PCDA course as well as the entire AAP course are taught using a combination of Jupyter notebooks, Python scripts, and videos walking through the notebooks.
-Jupyter notebooks, part of \href{https://jupyter.org/}{Project Jupyter}, are an example of \textit{literate computing} \cite{perez2015project} and allow one to create educational narratives that support iterative exploration and development.
-The ability to weave standard explanatory language with formal programming languages has made notebook computing a defacto standard in the data science education community \cite{perkel2018jupyter}.
-Jupyter notebooks can also include equations written in \LaTeX, data visualizations, and other multi-media.
-The technology is free, open source, runs on most computing platforms, and has a huge user community that has created a large number of resources for students, faculty and industry practitioners.
+Jupyter notebooks, part of \href{https://jupyter.org/}{Project Jupyter}, are an example of \textit{literate computing} \cite{perez2015project} and allow one to create rich interactive documents that support iterative exploration and development.
 
-Central to Jupyter notebooks is a mix of \textit{code cells} and \textit{markdown cells}.
+Central to Jupyter notebooks is a mix of \textit{code cells}, \textit{markdown cells} and output cells.
 Code cells contain executable code while markdown cells contain plain text using \href{URL}{markdown} for styling text.
@@
 After running the cells, they appear as show in Figure \ref{fig:cells_after_running}.
+After executing a code cell, the output appears directly below it in an output cell.
 Mixing code and markdown cells makes it quite easy to create interactive tutorials in which students can not only run pre-written code, but can also add their own code -- all with the aid of supporting explanatory text.
 All of the blog posts upon which EwP is based were written in Jupyter notebooks.
+
+The ability to weave standard explanatory language with formal programming languages has made notebook computing a popular tool in the data science education community \cite{perkel2018jupyter}.
+Jupyter notebooks can also include equations written in \LaTeX, data visualizations, and other multi-media.
+The technology is free, open source, runs on most computing platforms, and has a huge user community that has created a large number of resources for students, faculty and industry practitioners.
 There are countless introductory tutorials for Jupyter notebooks, but the official Project Jupyter Documentations \cite{jupyterteamProjectJupyterDocumentation2015} is a very good place to start if you are new to notebook computing.
+As with all tools, there are pros and cons to their use.
+See \cite{rule2018ten, barba2019teaching} and \cite{johnson2020benefits} for guidance on Jupyter notebook best practices for both research and teaching.
 
@@
 This difference in behavior between Jupyter notebooks and spreadsheets might be one of the most important concepts for students to internalize.
-A well known challenge associated with computational notebooks is the ease of running code cells out of order and introducing inadvertent errors \cite{grusDonNotebooks2018}.
+A well known challenge, often referred to as the \textit{hidden state problem}, associated with computational notebooks is the ease of running code cells out of order and introducing inadvertent errors \cite{\cite{johnson2020benefits},grusDonNotebooks2018}.
 This point along with strategies to avoid it are discussed during the first week of the course which focuses on \href{http://www.sba.oakland.edu/faculty/isken/courses/mis6900/jupyter_notebooks.html}{notebook computing best practices as well as a historical look} at computational notebooks.

### 2023-07-27 python_excel_ite.tex
code chunks converted to screenshots

@@
 
-We start with a very simple model that ignores the uncertainty in demand and that is not object oriented. Proceeding much like we would in Excel, base input values are stored in variables.
-
-\begin{tcolorbox}
-\begin{minted}[gobble=0]{python}
-# Base inputs
-unit_cost = 7.50
-selling_price = 10.00
-unit_refund = 2.50
-
-# Demand parameters
-demand_mean = 193
-demand_sd = 40
-
-# Deterministic model
-demand = demand_mean
-
-# Initial value for order quantity - this is the decision variable
-order_quantity = 200
-\end{minted}
-\end{tcolorbox}
-
-Students are given skeleton code as scaffolding and must finish the expressions for computing key intermediate outputs (\code{order\_cost, sales\_revenue, refund\_revenue}) and the final output, \code{profit}.
-
-\begin{tcolorbox}
-\begin{minted}[gobble=0]{python}
-order_cost = unit_cost * order_quantity
-sales_revenue = ??? * selling_price
-refund_revenue = ??? * unit_refund
-profit = sales_revenue + refund_revenue - order_cost
-\end{minted}
-\end{tcolorbox}
+We start with a very simple model that ignores the uncertainty in demand and that is not object oriented.
+Proceeding much like we would in Excel, base input values are stored in variables.
+Students are given skeleton code as scaffolding and must finish the expressions for computing key intermediate outputs (\code{order\_cost, sales\_revenue, refund\_revenue}) and the final output, \code{profit} -- see Figure \ref{fig:model_inputs}
 
@@
 This difference in behavior between Jupyter notebooks and spreadsheets might be one of the most important concepts for students to internalize.
-A well known challenge, often referred to as the \textit{hidden state problem}, associated with computational notebooks is the ease of running code cells out of order and introducing inadvertent errors \cite{\cite{johnson2020benefits},grusDonNotebooks2018}.
+A well known challenge, often referred to as the \textit{hidden state problem}, associated with computational notebooks is the ease of running code cells out of order and introducing inadvertent errors \cite{johnson2020benefits,grusDonNotebooks2018}.
 This point along with strategies to avoid it are discussed during the first week of the course which focuses on \href{http://www.sba.oakland.edu/faculty/isken/courses/mis6900/jupyter_notebooks.html}{notebook computing best practices as well as a historical look} at computational notebooks.
@@
 This example provides an effective way to show the power of creating reusable functions and Python's handy list comprehension construct.
+As shown in Figure \ref{fig:function_datatable}, a few code and markdown cells are used to create the code, provide some explanation and show the output.
+This is the same output one would get with an Excel 2-way Data Table, just laid out differently.
+Then in Figure \ref{fig:datatable_scatter} we show how the output can be turned into a scatter plot with profit mapped to a color hue.
 
-This code chunk,
+\begin{figure*}[!htbp]
+	\centering
+	\caption{A 2-way data table}
+	\includegraphics[scale=0.5]{images/function_datatable.png}
+	\label{fig:function_datatable}
+\end{figure*}
 
-\begin{tcolorbox}[left=0mm]
-\begin{minted}[gobble=1]{python}
-	def bookstore_profit(unit_cost, selling_price, 
-	                     unit_refund, order_quantity, demand):
-	'''
-	Compute profit in bookstore model
-	'''
-		order_cost = unit_cost * order_quantity
-		sales_revenue = np.minimum(order_quantity, demand) * selling_price
-		refund_revenue = np.maximum(0, order_quantity - demand)
-		profit = sales_revenue + refund_revenue - order_cost
-		return profit
-
-	# Set up input vector ranges	
-	demand_range = np.arange(50, 301, 5)
-	order_quantity_range = np.arange(50, 301, 25)
-
-	# Create data table (as a list of tuples)
-	data_table_1 = [(d, oq, bookstore_profit(
-	                 unit_cost, selling_price, unit_refund, oq, d)) 
-	for d in demand_range for oq in order_quantity_range]
-
-	# Convert to dataframe
-	dtbl_1_df = pd.DataFrame(data_table_1, 
-	                         columns=['Demand', 'OrderQuantity', 'Profit'])
-
-\end{minted}
-\end{tcolorbox}
-
-results in the following output:
-
-\begin{tcolorbox}
-	\begin{verbatim}
-	   Demand  OrderQuantity  Profit
-	      50             50   125.0
-	      50             75   -37.5
-	      50            100  -200.0
-	      50            125  -362.5
-	      50            150  -525.0
-	      ...
-	     300            275   687.5
-	     300            300   750.0
-\end{verbatim}
-\end{tcolorbox}
-
-While laid out differently, this is the same output produced by a 2-way Data Table in Excel. At this point, we pause to consider a higher level design question.
+\begin{figure*}[!htbp]
+	\centering
+	\caption{Scatter plot based on 2-way data table}
+	\includegraphics[scale=0.5]{images/datatable_scatter.png}
+	\label{fig:datatable_scatter}
+\end{figure*}
 
@@
 
-It is not clear whether this non-object oriented approach will end up making the most sense for doing spreadsheet type modeling in Python.
+It is not clear whether this non-object oriented procedural approach will end up making the most sense for doing spreadsheet type modeling in Python.
 This uncertainty is made explicit in the notebook and students are led on a bit of a journey of discovery as we explore different software designs for this modeling problem.
@@
 This section of the notebook ends with using the \code{BookstoreModel} class to create a new model object instance with all of its base inputs instantiated with values.
-Before moving, on students are presented with a few challenges involving enhancements to the \code{BookstoreModel} class.
+Before moving on, students are presented with a challenges involving enhancements to the \code{BookstoreModel} class.
 Interspersing short coding challenges throughout the notebook gives students a chance to test themselves on their understanding and to get some coding practice.
@@
 In this case, the well known \href{https://scikit-learn.org/stable/}{scikit-learn} package \cite{scikit-learn} has a \code{ParameterGrid} function that solves our scenario generation problem quite nicely.
-This example also reminds the students of the value of being able to make sense of API documentation and of object oriented code written by others.
+This example also reminds the students of the value of being able to make sense of API documentation and to understand object oriented code written by others.
 
@@
 	\end{minipage}
-	\caption{Plot based on 2-way data table}
+	\caption{Faceted plot based on 2-way data table}
 	\label{fig:2way_dt}
@@
 
-In the Excel modeling world, Monte-Carlo simulation can be done without add-ins, but packages like @Risk \cite{palisadeRISKRiskAnalysis} can make the process much easier.
+In the Excel modeling world, Monte-Carlo simulation can be done without add-ins, but packages like \href{https://lumivero.com/products/at-risk/}{@Risk} can make the process much easier.
 In particular, random variable generation for a wide range of probability distributions is facilitated by add-ins like @Risk.
-In Python, the NumPy package provide random variate generation for a large number of distributions and similarly, SciPy provides functions for computing distribution related quantities for many probability distributions.
+In Python, the \href{https://numpy.org/doc/stable/reference/random/generator.html}{NumPy package provide random variate generation} for a large number of distributions and similarly, \href{https://docs.scipy.org/doc/scipy/reference/stats.html}{SciPy provides functions for computing distribution related quantities} for many probability distributions.
 Much like we did when creating the \code{data\_table} function, we start with adding uncertainty to a single input variable in the bookstore model and rely on NumPy's vectorized computing capabilities to generate the profit associated with each realization of the random input.

### 2023-07-27 python_excel_ite.tex
still need html versions of notebooks and associated links

@@
 
-\begin{itemize}
-	\item
-	The first argument will be a model object (i.e.~something like the
-	\code{BookstoreModel} model) that contains an \code{update}
-	method. Soon, we should add an abstract \code{Model} class from
-	which specific models such as the \texttt{BookstoreModel} class can be
-	created. The abstract class will contain the \code{update} method.
-	\item
-	The random inputs will be passed in as a dictionary whose keys are the
-	input variables being modeled as random and whose values are an
-	iterable representing the draws from some probability distribution.
-	Structurally, this is similar to how inputs are specified in the
-	\code{data\_table} function.
-	\item
-	We can optionally pass in a dictionary of scenario inputs. This is
-	exactly like the \code{data\_table} variable input.
-
-	\begin{itemize}
-		\item
-		If no scenario input dictionary is passed in, a single simulation
-		scenario is run using the current input values in the model object,
-		\item
-		If a scenario input dictionary is passed in, then a simulation
-		scenario is run for every combination of parameters in the
-		dictionary. Again, this is just like we do in the
-		\code{data\_table} function.
-	\end{itemize}
-	\item
-	The output will be a set of dictionaries containing dataframes of
-	simulation output as well standard summary stats and plots.
-\end{itemize}
+\begin{figure*}[!htbp]
+	\centering
+	\caption{Design specification for \code{simulate()} function}
+	\includegraphics[scale=0.5]{images/oo_sim_design_specs.png}
+	\label{fig:oo_sim_design_specs}
+\end{figure*}
 
@@
 By this point, students are well equipped to understand the code as it builds on basic Python concepts and more advanced concepts that we have already covered in the EwP module.
-The rest of the notebook is spent trying out the \code{simulate} function and ends with creating a grouped boxplot, see Figure \ref{fig:boxplot} that many of them have likely created before with @Risk along with a faceted plot of histograms of profit created with the popular \href{https://seaborn.pydata.org/}{Seaborn} package.
+The rest of the notebook is spent trying out the \code{simulate()} function and ends with creating a grouped boxplot, see Figure \ref{fig:boxplot} that many of them have likely created before with @Risk along with a faceted plot of histograms of profit created with the popular \href{https://seaborn.pydata.org/}{Seaborn} package.
 
@@
 Now we shift gears and move beyond what might be considered prototyping and into the realm of deployment.
-The notebook ends with these thoughts.
+Figure \ref{fig:sim_wrapup} shows how the notebook ends and gives students a preview of the next submodule.
 
-\begin{tcolorbox}
-	\begin{quotation}
-
-\textsf{We have added a basic \code{simulate} function to our \code{data\_table} and \code{goal\_seek} functions. Python is proving to be quite nice for doing Excel-style "what if?" analysis.}
-
-\textsf{In Part 4 of this series, we'll make some improvements and do some clean-up on our classes and functions.
-We'll move everything into a single \code{whatif.py} module and learn how to create a Python package to make it easy to use and share our new functions.
-We'll try out our package on a new model and sketch out some ideas for future enhancements to the package.
-It's important we also start creating some basic documentation and a user guide.}
-\end{quotation}
-\end{tcolorbox}
+\begin{figure*}[!htbp]
+	\centering
+	\caption{Closing thoughts in simulation notebook}
+	\includegraphics[scale=0.5]{images/sim_wrapup.png}
+	\label{fig:sim_wrapup}
+\end{figure*}
 
@@
 This includes an explanation of how and where Python searches for packages when \code{import} statements are encountered in code.
-We defer discussion of uploading our package to a public code sharing platform such as PyPI.
-All of this can be quite complex and I have made every effort to distill things down to a minimal level of complexity needed to convey the important points.
-I stress that there is much complexity lurking here and it is something they will need to revisit time and time again.
+There is much complexity lurking here and students are reminded that they will need to revisit this topic frequently.
+I have made every effort to distill things down to a minimal level of complexity needed to convey the important points.
 
@@
 This new model differs structurally from the Bookstore Model, but is representative of many financial models in that cash flows over multiple years need to be computed and summarized.
+This helps cement the idea that there is value in creating reusable code.
 
@@
 This deployment focused submodule ends with a short notebook discussing the different types of documentation needed in a typical project.
-This includes code comments, docstrings, \code{readme} files and generating documentation by writing \href{https://www.sphinx-doc.org/en/master/usage/restructuredtext/basics.html}{reStructuredText} files (similar but more powerful than markdown) and using \href{https://www.sphinx-doc.org/en/master/index.html}{Sphinx}, the Python documentation generation tool.
+This includes code comments, \href{https://peps.python.org/pep-0257/}{docstrings}, \code{readme} files and generating documentation by creating \href{https://www.sphinx-doc.org/en/master/usage/restructuredtext/basics.html}{reStructuredText} files (similar to but more powerful than markdown) and using \href{https://www.sphinx-doc.org/en/master/index.html}{Sphinx}, the Python documentation generation tool.
 This is one of those tasks that few people like to do but can be very important for the long term success of a project.
@@
 	\item
-	You have an Excel file with data in ``wide'' format and you need to
+	You have an Excel file with data in wide format and you need to
 	convert it to long format, and then perhaps export out individual
@@
 	\item
-	You have an Excel file acting as a simple flatfile database.
+	You have an Excel file acting as a simple flat file database.
 	Periodically, you get new Excel files that need to get appended to the
@@
 	\href{https://pandas.pydata.org/docs/reference/api/pandas.ExcelWriter.html}{ExcelWriter},
-	\href{https://pandas.pydata.org/docs/reference/api/pandas.DataFrame.to_excel.html}{to\_excel}
+	\href{https://pandas.pydata.org/docs/reference/api/pandas.DataFrame.to_excel.html}{to\_excel} -- reading and writing Excel files,
 	\item
 	\href{https://openpyxl.readthedocs.io/en/stable/}{openpyxl} - can
-	read, write, and modify Excel files
+	read, write, and modify Excel files,
 	\item
-	\href{https://xlsxwriter.readthedocs.io/}{XlsxWriter} - powerful but
-	cannot edit existing Excel files
+	\href{https://xlsxwriter.readthedocs.io/}{XlsxWriter} - create new Excel files but
+	cannot edit existing Excel files,
 	\item
-	\href{https://www.xlwings.org/}{xlwings} - provides ``glue'' between
-	Python and Excel
+	\href{https://www.xlwings.org/}{xlwings} - manipulate the Excel application from Python.
 \end{itemize}
@@
 One approach uses pandas while the other relies on a library specifically created for working with Excel files from Python, \href{https://openpyxl.readthedocs.io/en/stable/}{openpyxl}.
-One advantage to using Python rather than Visual Basic for Applications for this task is that Python is well suited for working with data in many different formats including csv, json, xml and a myriad of database formats.
+One advantage to using Python rather than Visual Basic for Applications for this task is that Python is well suited for working with data in many different formats including csv, JSON, XML and a myriad of database formats.
 The Excel file may be just one data source out of a collection of data sources that are used to create some sort of combined database.

### 2023-07-27 python_excel_ite.tex
draft ready for detailed final proofreading

@@
 % Enter the (shortened) title:
-\RUNTITLE{Spreadsheet modeling with Python}
+\RUNTITLE{Spreadsheet-ish modeling with Python}
 
@@
 % Enter the full title:
-\TITLE{Spreadsheet modeling with Python}
+\TITLE{Spreadsheet-ish modeling with Python}
 
@@
 The newer AAP course has been offered since 2021 in an online asynchronous mode.
-All three courses have \href{http://www.sba.oakland.edu/Faculty/isken/teaching.html}{extensive course websites} that are publicly accessible, including access to all of the video content and supporting files.
+All three courses have extensive course websites that are publicly accessible, including access to all of the video content and supporting files.
 
@@
 
-\subsection{Notebook 1: Modeling and data tables}
+\subsection{Notebook 1: Modeling and data tables (\href{http://www.sba.oakland.edu/faculty/isken/excel_with_python/what_if_1_model_datatable.html}{html version})}
 
-The stage is set with an overview of a typical spreadsheet based model that most of the students have seen before -- see Figure \ref{fig:stage}.
-The notebook also includes a screenshot (Figure \ref{fig:xl_model}) of a typical Excel model.
+The stage is set with an overview of a typical spreadsheet based model that most of the students have seen before.
 
@@
 
-\begin{figure}[!ht]
-	\centering
-	\begin{minipage}{.80\textwidth}
+The notebook also includes a screenshot (Figure \ref{fig:xl_model}) of a typical Excel model.
+
+\begin{figure}[!htbp]
 	\includegraphics[width=1.0\textwidth]{images/output_2_0.png}
-	\end{minipage}
 	\caption{Typical spreadsheet model}
@@
 
-\begin{figure*}[!htbp]
+\begin{figure}[!htbp]
 	\centering
+	\includegraphics[scale=0.5]{images/model_inputs.png}
 	\caption{Instantiating model inputs}
-	\includegraphics[scale=0.5]{images/model_inputs.png}
 	\label{fig:model_inputs}
-\end{figure*}
+\end{figure}
 
@@
 
-\begin{figure*}[!htbp]
+\begin{figure}[!htbp]
 	\centering
+	\includegraphics[scale=0.5]{images/function_datatable.png}
 	\caption{A 2-way data table}
-	\includegraphics[scale=0.5]{images/function_datatable.png}
 	\label{fig:function_datatable}
-\end{figure*}
+\end{figure}
 
-\begin{figure*}[!htbp]
+\begin{figure}[!htbp]
 	\centering
+	\includegraphics[scale=0.5]{images/datatable_scatter.png}
 	\caption{Scatter plot based on 2-way data table}
-	\includegraphics[scale=0.5]{images/datatable_scatter.png}
 	\label{fig:datatable_scatter}
-\end{figure*}
+\end{figure}
 
@@
 
-\begin{figure}[!ht]
+\begin{figure}[!htbp]
 	\centering
-	\begin{minipage}{.80\textwidth}
 	\includegraphics[width=1.0\textwidth]{images/two_way_dt.png}
-	\end{minipage}
 	\caption{Faceted plot based on 2-way data table}
@@
 
-\subsection{Notebook 2: Goal seek}
+\subsection{Notebook 2: Goal seek (\href{http://www.sba.oakland.edu/faculty/isken/excel_with_python/what_if_2_goalseek.html}{html version})}
 
@@
 
-\subsection{Notebook 3: Monte-Carlo simulation}
+\subsection{Notebook 3: Monte-Carlo simulation (\href{http://www.sba.oakland.edu/faculty/isken/excel_with_python/what_if_3_simulation.html}{html version})}
 
@@
 
-\begin{figure*}[!htbp]
+\begin{figure}[!htbp]
 	\centering
+	\includegraphics[scale=0.5]{images/oo_sim_design_specs.png}
 	\caption{Design specification for \code{simulate()} function}
-	\includegraphics[scale=0.5]{images/oo_sim_design_specs.png}
 	\label{fig:oo_sim_design_specs}
-\end{figure*}
+\end{figure}
 
@@
 
-\begin{figure}[!ht]
+\begin{figure}[!htbp]
 	\centering
-	\begin{minipage}{.80\textwidth}
 	\includegraphics[width=1.0\textwidth]{images/boxplots.png}
-	\end{minipage}
 	\caption{Profit boxplots by order quantity}
@@
 
-\begin{figure*}[!htbp]
+\begin{figure}[!htbp]
 	\centering
+	\includegraphics[scale=0.5]{images/sim_wrapup.png}
 	\caption{Closing thoughts in simulation notebook}
-	\includegraphics[scale=0.5]{images/sim_wrapup.png}
 	\label{fig:sim_wrapup}
-\end{figure*}
+\end{figure}
 
@@
 
-\subsection{Notebook 4: Project packaging}
+\subsection{Notebook 4: Project packaging (\href{http://www.sba.oakland.edu/faculty/isken/excel_with_python/what_if_4_project_packaging.html}{html version})}
 
@@
 
-\subsection{Notebook 5: Documentation}
+\subsection{Notebook 5: Documentation (\href{http://www.sba.oakland.edu/faculty/isken/excel_with_python/what_if_5_documentation.html}{html version})}
 

### 2023-07-27 python_excel_ite.tex
draft ready for detailed final proofreading
fixed missing items in references

@@
 A series of conferences focused on teaching management science with spreadsheets started at Dartmouth in 1998 and several presentations from that conference made up the bulk of papers in the first two issues of the INFORMS Transactions on Education (ITE) journal.
-The next few decades saw numerous textbooks \cite{powellBusinessAnalyticsArt, camm2020business} and teaching papers in journals such as \href{https://pubsonline.informs.org/journal/ited}{ITE}that prominently featured spreadsheet based modeling.
+The next few decades saw numerous textbooks \cite{powellBusinessAnalyticsArt2019, camm2020business} and teaching papers in journals such as \href{https://pubsonline.informs.org/journal/ited}{ITE}that prominently featured spreadsheet based modeling.
 
@@
 	\centering
+	\includegraphics[scale=0.5]{images/cells_edit_mode.png}
 	\caption{Notebook cells in edit model}
-	\includegraphics[scale=0.5]{images/cells_edit_mode.png}
 	\label{fig:cells_edit_mode}
@@
 	\centering
+	\includegraphics[scale=0.5]{images/cells_after_running.png}
 	\caption{Notebook cells after execution}
-	\includegraphics[scale=0.5]{images/cells_after_running.png}
 	\label{fig:cells_after_running}
@@
 Every notebook is accompanied by screencasts that walk the students through the material and act as a guided tour with challenges to overcome along the way.
+While all of the notebooks are publicly available (see Section \ref{sec:software}), links to an html version of each of the notebooks are provided throughout the document.
 
@@
 Certainly we do not want to rely on copying, pasting and modifying code when building a new model.
-In the Python world, the natural next step is to deploy our code as a Python \textit{package} \cite{buezenWelcomePythonPackages} that can then be imported like any other Python library.
+In the Python world, the natural next step is to deploy our code as a Python \textit{package} \cite{buezenWelcomePythonPackages2022} that can then be imported like any other Python library.
 
@@
 \section{Software availability}
+\label{sec:software}
 
@@
 \section{Classroom experience and the future}
+\label{sec:classroom}
 

### 2023-07-28 python_excel_ite.tex
draft done, needs final proofreading and editing

@@
 \ABSTRACT{%
-TODO: Text of your abstract % Enter your abstract
+A staple of many spreadsheet based management science textbooks and courses are the use of Excel for activities such as model building, sensitivity analysis, goal seeking and Monte-Carlo simulation.
+What might those things look like if done with Python?
+We describe a teaching module in which Python is used to do typical Excel based modeling and data wrangling tasks.
+In addition, students are exposed to basic software engineering principles including project folder structures, version control, object oriented programming and other more advanced Python skills, creating deployable packages and documentation.
+The module is supported with Jupyter notebooks, Python scripts, course web pages which include numerous screencasts, and a few GitHub repositories.
+All of the supporting materials are permissively licensed and freely accessible.
 }%

### 2023-07-28 python_excel_ite.tex
fixed course links to reflect new folder structure

@@
 Numerous R and Python based data science courses have been developed and offered at universities and through online learning platforms such as \href{https://www.coursera.org/browse/data-science}{Coursera}, \href{https://www.edx.org/learn/data-science}{EdX}, \href{https://www.datacamp.com/}{DataCamp}, \href{https://www.udemy.com/topic/data-analysis/}{Udemy} and many others.
-In 2015, at Oakland University, we created a course called \href{http://www.sba.oakland.edu/faculty/isken/courses/mis5470/}{Practical Computing for Data Analytics} (PCDA) that introduced both R and Python (along with a little Linux) for use in data analytics work.
-This new course was meant as a follow up to the \href{http://www.sba.oakland.edu/faculty/isken/courses/mis5460/}{spreadsheet based modeling course} we have had in place since 2001.
+In 2015, at Oakland University, we created a course called \href{http://www.sba.oakland.edu/faculty/isken/courses/pcda/}{Practical Computing for Data Analytics} (PCDA) that introduced both R and Python (along with a little Linux) for use in data analytics work.
+This new course was meant as a follow up to the \href{http://www.sba.oakland.edu/faculty/isken/courses/ba/}{spreadsheet based modeling course} we have had in place since 2001.
 
@@
 
-At our institution, The EwP module is part of a business course entitled  \href{http://www.sba.oakland.edu/faculty/isken/courses/mis6900}{Advanced Analytics with Python} (AAP).
+At our institution, The EwP module is part of a business course entitled  \href{http://www.sba.oakland.edu/faculty/isken/courses/aap}{Advanced Analytics with Python} (AAP).
 The students taking this course have already taken the PCDA course within which they learned fundamental Python programming within the context of data analytics over seven weeks.
@@
 The EwP module could also be used as part of a semester long Python based analytics course.
-Most AAP students have also already taken   \href{http://www.sba.oakland.edu/faculty/isken/courses/mis5460}{Business Analytics} (BA) which is a spreadsheet based introduction to business analytics.
+Most AAP students have also already taken   \href{http://www.sba.oakland.edu/faculty/isken/courses/ba}{Business Analytics} (BA) which is a spreadsheet based introduction to business analytics.
 We do not strictly require the spreadsheet based course be taken before the PCDA or AAP courses as that course draws students from outside of our business programs who are strictly interested in programmatic tools such as Python.
@@
 
-There are two overarching learning objectives for this \href{http://www.sba.oakland.edu/faculty/isken/courses/mis6900/mod3a_whatif.html}{first EwP submodule}.
+There are two overarching learning objectives for this \href{http://www.sba.oakland.edu/faculty/isken/courses/aap/mod3a_whatif.html}{first EwP submodule}.
 Students learn how Python might be used to do something they are quite familiar with doing in a spreadsheet -- model building and sensitivity analysis.
@@
 A well known challenge, often referred to as the \textit{hidden state problem}, associated with computational notebooks is the ease of running code cells out of order and introducing inadvertent errors \cite{johnson2020benefits,grusDonNotebooks2018}.
-This point along with strategies to avoid it are discussed during the first week of the course which focuses on \href{http://www.sba.oakland.edu/faculty/isken/courses/mis6900/jupyter_notebooks.html}{notebook computing best practices as well as a historical look} at computational notebooks.
+This point along with strategies to avoid it are discussed during the first week of the course which focuses on \href{http://www.sba.oakland.edu/faculty/isken/courses/aap/jupyter_notebooks.html}{notebook computing best practices as well as a historical look} at computational notebooks.
 
@@
 Thinking about how others will interact with your software usually leads to better designed software.
-In \href{http://www.sba.oakland.edu/faculty/isken/courses/mis6900/mod3b_whatif_packaging.html}{this submodule} we create a Python package that can be imported and used by other modelers.
+In \href{http://www.sba.oakland.edu/faculty/isken/courses/aap/mod3b_whatif_packaging.html}{this submodule} we create a Python package that can be imported and used by other modelers.
 This will require some design changes to our software and will also involve moving code out of Jupyter notebooks and into Python script files.
@@
 
-The \href{http://www.sba.oakland.edu/faculty/isken/courses/mis6900/mod3c_python_excel.html}{EwP module ends} with what has become a relatively common use of Python with Excel -- automating various Excel data wrangling tasks.
+The \href{http://www.sba.oakland.edu/faculty/isken/courses/aap/mod3c_python_excel.html}{EwP module ends} with what has become a relatively common use of Python with Excel -- automating various Excel data wrangling tasks.
 Examples include:
@@
 Each of the submodules described above has its own dedicated course web page.
-Links to each submodule page can be found on \href{http://www.sba.oakland.edu/faculty/isken/courses/mis6900/mod3_excel_with_python.html}{this EwP landing page} within the AAP course website.
+Links to each submodule page can be found on \href{http://www.sba.oakland.edu/faculty/isken/courses/aap/mod3_excel_with_python.html}{this EwP landing page} within the AAP course website.
 Each page contains learning objectives, a link to a compressed file containing the Jupyter notebooks and other supporting files, a set of activities with accompanying screen casts, and links to additional web based materials to explore.
-
+The source reStructuredText files for the course web pages can be found at \href{https://github.com/misken/aap}{this app GitHub repository}.
 There is also a \href{https://github.com/misken/whatif}{whatif Github repository} for the \code{whatif} package.
-It also contains the original notebooks used in the blog posts that motivated this project.
+It contains the original notebooks used in the blog posts that motivated this project.
 

### 2023-07-28 python_excel_ite.tex
final proofing through submodule1

@@
 \ABSTRACT{%
-A staple of many spreadsheet based management science textbooks and courses are the use of Excel for activities such as model building, sensitivity analysis, goal seeking and Monte-Carlo simulation.
-What might those things look like if done with Python?
+A staple of many spreadsheet based management science courses is the use of Excel for activities such as model building, sensitivity analysis, goal seeking and Monte-Carlo simulation.
+What might those things look like if carried out using Python?
 We describe a teaching module in which Python is used to do typical Excel based modeling and data wrangling tasks.
@@
 A series of conferences focused on teaching management science with spreadsheets started at Dartmouth in 1998 and several presentations from that conference made up the bulk of papers in the first two issues of the INFORMS Transactions on Education (ITE) journal.
-The next few decades saw numerous textbooks \cite{powellBusinessAnalyticsArt2019, camm2020business} and teaching papers in journals such as \href{https://pubsonline.informs.org/journal/ited}{ITE}that prominently featured spreadsheet based modeling.
+The next few decades saw numerous textbooks \cite{powellBusinessAnalyticsArt2019, camm2020business} and teaching papers in journals such as \href{https://pubsonline.informs.org/journal/ited}{ITE} that prominently featured spreadsheet based modeling.
 
@@
 Version control is difficult as both data and logic live in the same spreadsheet document.
-Documentation of multi-step manual data manipulation processes present challenges, it it is done at all -- hindering reproducibility and automation.
+Documentation of multi-step manual data manipulation processes is often lacking -- hindering reproducibility and automation.
 There are data size limitations and external data links present their \href{https://www.thespreadsheetguru.com/find-remove-external-links/}{own set of challenges}.
@@
 This ``what if?'' type modeling is facilitated by Excel's flexibility, numerous built in functions, integrated sensitivity analysis via \href{https://support.microsoft.com/en-us/office/calculate-multiple-results-by-using-a-data-table-e95e2487-6ca6-4413-ad12-77542a5ea50b}{Data Tables},  \href{https://support.microsoft.com/en-us/office/use-goal-seek-to-find-the-result-you-want-by-adjusting-an-input-value-320cb99e-f4a4-417f-b1c3-4f369d6e66c7}{Goal
-	Seek} or break-even analysis and visualization capabilities.
+	Seek} for break-even analysis and exploratory data analysis via plots and Pivot Tables.
 
@@
 Most AAP students have also already taken   \href{http://www.sba.oakland.edu/faculty/isken/courses/ba}{Business Analytics} (BA) which is a spreadsheet based introduction to business analytics.
-We do not strictly require the spreadsheet based course be taken before the PCDA or AAP courses as that course draws students from outside of our business programs who are strictly interested in programmatic tools such as Python.
 The BA and PCDA courses have historically been offered in face to face mode in a computer teaching lab.
@@
 All three courses have extensive course websites that are publicly accessible, including access to all of the video content and supporting files.
+These courses serve upper level undergraduate and graduate students.
+Many are from the School of Business Administration but they draw students from many programs across our campus.
 
@@
 \begin{itemize}
-	\item variables, arithmetic and boolean operators, and basic data types,
+	\item variables, arithmetic and boolean operators, basic data types,
 	\item data structures such as lists, dictionaries, tuples, NumPy arrays and pandas dataframes,
-	\item flow control such as branching with \texttt{if \ldots else} and looping with \texttt{for} and \texttt{while}
+	\item flow control such as branching with \texttt{if \ldots else} and iterating with \texttt{for} and \texttt{while}
 	\item module imports,
@@
 
-All of the Python modules in the PCDA course as well as the entire AAP course are taught using a combination of Jupyter notebooks, Python scripts, and videos walking through the notebooks.
+All of the Python modules in the PCDA course as well as the entire AAP course are taught using a combination of Jupyter notebooks, Python scripts, and videos that walk readers through the notebooks.
 Jupyter notebooks, part of \href{https://jupyter.org/}{Project Jupyter}, are an example of \textit{literate computing} \cite{perez2015project} and allow one to create rich interactive documents that support iterative exploration and development.
 
-Central to Jupyter notebooks is a mix of \textit{code cells}, \textit{markdown cells} and output cells.
+Central to Jupyter notebooks is a mix of \textit{code cells}, \textit{markdown cells} and \textit{output cells}.
 Code cells contain executable code while markdown cells contain plain text using \href{URL}{markdown} for styling text.
@@
 The stage is set with an overview of a typical spreadsheet based model that most of the students have seen before.
+The notebook also includes a screenshot (Figure \ref{fig:xl_model}) of this Excel model.
 
@@
 
-The notebook also includes a screenshot (Figure \ref{fig:xl_model}) of a typical Excel model.
-
 \begin{figure}[!htbp]
@@
 
-\subsubsection{Build first model - procedural approach}
+\subsubsection{Model building using a procedural programming approach}
 
-We start with a very simple model that ignores the uncertainty in demand and that is not object oriented.
+We start with a very simple model that ignores the uncertainty in demand and does not use object oriented programming concepts.
 Proceeding much like we would in Excel, base input values are stored in variables.
-Students are given skeleton code as scaffolding and must finish the expressions for computing key intermediate outputs (\code{order\_cost, sales\_revenue, refund\_revenue}) and the final output, \code{profit} -- see Figure \ref{fig:model_inputs}
+Students are given skeleton code as scaffolding and must finish the expressions for computing key intermediate outputs (\code{order\_cost, sales\_revenue, refund\_revenue}) and the final output, \code{profit} -- see Figure \ref{fig:model_inputs}.
 
@@
 	\includegraphics[scale=0.5]{images/model_inputs.png}
-	\caption{Instantiating model inputs}
+	\caption{Instantiating model inputs and creating output formulas}
 	\label{fig:model_inputs}
@@
 
-At this point we discuss a fundamental difference between the computing model of Excel and that of Jupyter notebooks -- while spreadsheets respond automatically to changes, Jupyter notebooks require rerunning all code cells that include or follow the changed code.
+At this point we discuss a fundamental difference between the computing model of Excel and that of Jupyter notebooks.
+While spreadsheets respond automatically to changes, Jupyter notebooks require rerunning all code cells that include or follow the changed code.
 This difference in behavior between Jupyter notebooks and spreadsheets might be one of the most important concepts for students to internalize.
@@
 Since order quantity is the key decision variable, we might want to see how profit changes for different order quantities.
-In Excel, the Data Table tool provides an easy way to do this and is a staple of spreadsheet based management science textbooks.
+In Excel, the Data Table tool provides a familiar and easy way to do this.
 In a Data Table, a range of order quantities can be used as row or column input and one or more output values can be computed.
@@
 A natural next step is to do the equivalent of an Excel 2-way Data Table.
-By creating a profit function that takes all of the base inputs as arguments, a list comprehension can then be used to generate the equivalent of a 2-way Data Table.
+By creating a profit function that takes all of the base inputs as arguments, a \href{https://realpython.com/list-comprehension-python/}{list comprehension} can then be used to generate the equivalent of a 2-way Data Table.
 Even better, this actually allows us to do the equivalent of an n-way Data Table with an arbitrary number of outputs -- something Excel does not have.
@@
 
-Moving on to multiple uncertain inputs raises challenges much like those we faced when moving from a 1-way to an n-way data table function.
-In addition to modeling multiple uncertain inputs, we want to be able to specify a range of scenarios to run, much like we did for the \code{data\_table} function.
-The following set of design criteria are presented for a \code{simulate} function.
+Moving on to multiple uncertain inputs raises challenges similar to those we faced when moving from a 1-way to an n-way data table function.
+In addition to modeling multiple uncertain inputs, we want to be able to specify a range of scenarios to run, just as we did for the \code{data\_table} function.
+Figure \ref{fig:oo_sim_design_specs} shows the design specifications for our \code{simulate} function.
 
@@
 By this point, students are well equipped to understand the code as it builds on basic Python concepts and more advanced concepts that we have already covered in the EwP module.
-The rest of the notebook is spent trying out the \code{simulate()} function and ends with creating a grouped boxplot, see Figure \ref{fig:boxplot} that many of them have likely created before with @Risk along with a faceted plot of histograms of profit created with the popular \href{https://seaborn.pydata.org/}{Seaborn} package.
+The rest of the notebook is spent trying out the \code{simulate()} function and ends with creating a grouped boxplot, see Figure \ref{fig:boxplot}, along with a faceted plot of profit histograms created with the popular \href{https://seaborn.pydata.org/}{Seaborn} package.
 
@@
 
-Now that we have created some useful modeling functionality, our goal is to make it easier for a modeler to actually use it.
+Now that we have created some useful modeling functionality, our goal is to make it easier for a modeler to actually use the code.
 Certainly we do not want to rely on copying, pasting and modifying code when building a new model.
@@
 
-Key software design changes are then made which will make our code easier on our users.
+Key software design changes are then made which will make using our code easier on our users.
 These changes include creating an abstract \code{Model} base class from which different model classes can be created and moving the \code{data\_table}, \code{goal\_seek}, and \code{simulate} functions into the \code{Model} base class as methods.

### 2023-07-28 python_excel_ite.tex
paper ready for submission

@@
 
-Key software design changes are then made which will make using our code easier on our users.
+Key software design changes are then made which will make the code easier to use.
 These changes include creating an abstract \code{Model} base class from which different model classes can be created and moving the \code{data\_table}, \code{goal\_seek}, and \code{simulate} functions into the \code{Model} base class as methods.
@@
 
-This notebook ends by pointing the students to another notebook in which our newly deployed whatif package is used for a completely different model -- the New Car Simulation that has been a staple of Winston and Albright's textbooks for many years \cite{winstonPracticalManagementScience2018,albrightBusinessAnalyticsData2016}.
+This notebook ends by pointing the students to another notebook in which our newly deployed whatif package is used for a completely different model -- the New Car Simulation that has been a part of Winston and Albright's textbooks for many years \cite{winstonPracticalManagementScience2018,albrightBusinessAnalyticsData2016}.
 This new model differs structurally from the Bookstore Model, but is representative of many financial models in that cash flows over multiple years need to be computed and summarized.
@@
 Several web based resources are shared including the \href{https://pbpython.com/}{Practical Business Python blog} written by Chris Moffitt, a data analytics professional \cite{moffittPracticalBusinessPython2022}.
-This blog, which focuses on helping Excel centric analytics professionals transition some of their work to Python, includes numerous posts on different aspects of Python and Excel integration.
+This blog, which focuses on helping Excel-centric analytics professionals transition some of their work to Python, includes numerous posts on different aspects of Python and Excel integration.
 
@@
 Each page contains learning objectives, a link to a compressed file containing the Jupyter notebooks and other supporting files, a set of activities with accompanying screen casts, and links to additional web based materials to explore.
-The source reStructuredText files for the course web pages can be found at \href{https://github.com/misken/aap}{this app GitHub repository}.
+The source reStructuredText files for the course web pages can be found at \href{https://github.com/misken/aap}{this AAP GitHub repository}.
 There is also a \href{https://github.com/misken/whatif}{whatif Github repository} for the \code{whatif} package.
@@
 The ranges used in the formulas vary by sheet and require determination of the number of rows in each sheet of data and use that information to construct the appropriate formula.
+Students must create a proper project folder structure and put their project under version control.
 An example assignment is available from \href{https://drive.google.com/file/d/1prf8acZXZ4RoYBgQ8e9-eijA65uV7T-V/view?usp=sharing}{here}.

### 2023-11-20 python_excel_ite.tex
edits for revision

@@
 \usepackage[most]{tcolorbox}
-\usepackage{minted}
-\tcbuselibrary{minted} % tcolorbox minted library, required to use the "minted" tcb listing engine (this library is not loaded by the option [most])
+\usepackage{graphicx}
+%\usepackage{minted}
+%\tcbuselibrary{minted} % tcolorbox minted library, required to use the "minted" tcb listing engine (this library is not loaded by the option [most])
 \usepackage[colorlinks]{hyperref} % ALWAYS load this package LAST
@@
 % In the reviewing and copyediting stage enter the manuscript number.
-%\MANUSCRIPTNO{} % When the article is logged in and DOI assigned to it,
+\MANUSCRIPTNO{} % When the article is logged in and DOI assigned to it,
                  %   this manuscript number is no longer necessary
@@
 % Enter the (shortened) title:
-\RUNTITLE{Spreadsheet-ish modeling with Python}
+\RUNTITLE{Spreadsheet modeling and wrangling with Python}
 
@@
 % Enter the full title:
-\TITLE{Spreadsheet-ish modeling with Python}
+\TITLE{Spreadsheet modeling and wrangling with Python}
 
@@
 \centering
-	\includegraphics[scale=0.5]{images/cells_edit_mode.png}
+\includegraphics[scale=0.5]{cells_edit_mode}
 \caption{Notebook cells in edit model}
@@
 \centering
-	\includegraphics[scale=0.5]{images/cells_after_running.png}
+\includegraphics[scale=0.5]{cells_after_running}
 \caption{Notebook cells after execution}
@@
 \begin{figure}[!htbp]
-	\includegraphics[width=1.0\textwidth]{images/output_2_0.png}
+\includegraphics[width=1.0\textwidth]{output_2_0}
 \caption{Typical spreadsheet model}
@@
 \centering
-	\includegraphics[scale=0.5]{images/model_inputs.png}
+\includegraphics[scale=0.5]{model_inputs}
 \caption{Instantiating model inputs and creating output formulas}
@@
 \centering
-	\includegraphics[scale=0.5]{images/function_datatable.png}
+\includegraphics[scale=0.5]{function_datatable}
 \caption{A 2-way data table}
@@
 \centering
-	\includegraphics[scale=0.5]{images/datatable_scatter.png}
+\includegraphics[scale=0.5]{datatable_scatter}
 \caption{Scatter plot based on 2-way data table}
@@
 \centering
-	\includegraphics[width=1.0\textwidth]{images/two_way_dt.png}
+\includegraphics[width=1.0\textwidth]{two_way_dt}
 \caption{Faceted plot based on 2-way data table}
@@
 \centering
-	\includegraphics[scale=0.5]{images/oo_sim_design_specs.png}
+\includegraphics[scale=0.5]{oo_sim_design_specs}
 \caption{Design specification for \code{simulate()} function}
@@
 \centering
-	\includegraphics[width=1.0\textwidth]{images/boxplots.png}
+\includegraphics[width=1.0\textwidth]{boxplots}
 \caption{Profit boxplots by order quantity}
@@
 \centering
-	\includegraphics[scale=0.5]{images/sim_wrapup.png}
+\includegraphics[scale=0.5]{sim_wrapup}
 \caption{Closing thoughts in simulation notebook}
@@
 %   (while the paper is being written).
-\bibliographystyle{informs2014} % outcomment this and next line in Case 1
-\bibliography{python_excel.bib} % if more than one, comma separated
+%\bibliographystyle{informs2014} % outcomment this and next line in Case 1
+%\bibliography{python_excel.bib} % if more than one, comma separated
 
@@
 %\input{mypaper.bbl} % outcomment this line in Case 2
+\begin{thebibliography}{29}
+	\providecommand{\natexlab}[1]{#1}
+	\providecommand{\url}[1]{\texttt{#1}}
+	\providecommand{\urlprefix}{URL }
+
+	\bibitem[{Albright \protect\BIBand{}
+		Winston(2016)}]{albrightBusinessAnalyticsData2016}
+	Albright SC, Winston WL (2016) \emph{Business {{Analytics}}: {{Data Analysis}}
+		\& {{Decision Making}} - {{Standalone}} Book} ({Boston, MA}: {Cengage
+		Learning}), 6th edition edition, ISBN 978-1-305-94754-2.
+
+	\bibitem[{{Audrey Roy and Cookiecutter
+			community}(2012)}]{audreyroyandcookiecuttercommunityCookiecutterBetterProject2012}
+	{Audrey Roy and Cookiecutter community} (2012) \emph{Cookiecutter: {{Better}}
+		Project Templates}.
+
+	\bibitem[{Barba et~al.(2019)Barba, Barker, Blank, Brown, Downey, George, Heagy,
+		Mandli, Moore, Lippert et~al.}]{barba2019teaching}
+	Barba LA, Barker LJ, Blank DS, Brown J, Downey AB, George T, Heagy LJ, Mandli
+	KT, Moore JK, Lippert D, et~al. (2019) Teaching and learning with
+	{{Jupyter}}. \emph{Recuperado: https://jupyter4edu. github.
+		io/jupyter-edu-book} .
+
+	\bibitem[{Buezen \protect\BIBand{}
+		Timbers(2022)}]{buezenWelcomePythonPackages2022}
+	Buezen T, Timbers T (2022) Welcome to {{Python Packages}}! \textemdash{}
+	{{Python Packages}}. https://py-pkgs.org/.
+
+	\bibitem[{Camm et~al.(2020)Camm, Cochran, Fry, \protect\BIBand{}
+		Ohlmann}]{camm2020business}
+	Camm JD, Cochran JJ, Fry MJ, Ohlmann JW (2020) \emph{Business Analytics}
+	({Cengage Learning}).
+
+	\bibitem[{Canny(2013)}]{cannyPythonpptx2013}
+	Canny S (2013) Python-pptx. https://python-pptx.readthedocs.io/en/latest/.
+
+	\bibitem[{Cimpanu(2017)}]{cimpanuMicrosoftConsidersAdding2017}
+	Cimpanu C (2017) Microsoft {{Considers Adding Python}} as an {{Official
+			Scripting Language}} to {{Excel}}.
+	https://www.bleepingcomputer.com/news/microsoft/microsoft-considers-adding-python-as-an-official-scripting-language-to-excel/.
+
+	\bibitem[{Grus(2018)}]{grusDonNotebooks2018}
+	Grus J (2018) I don't like notebooks ({JupyterCon 2018 (New York)}).
+
+	\bibitem[{Harris et~al.(2020)Harris, Millman, {van der Walt}, Gommers,
+		Virtanen, Cournapeau, Wieser, Taylor, Berg, Smith, Kern, Picus, Hoyer, {van
+			Kerkwijk}, Brett, Haldane, {del R{\'i}o}, Wiebe, Peterson,
+		{G{\'e}rard-Marchant}, Sheppard, Reddy, Weckesser, Abbasi, Gohlke,
+		\protect\BIBand{} Oliphant}]{harris2020array}
+	Harris CR, Millman KJ, {van der Walt} SJ, Gommers R, Virtanen P, Cournapeau D,
+	Wieser E, Taylor J, Berg S, Smith NJ, Kern R, Picus M, Hoyer S, {van
+		Kerkwijk} MH, Brett M, Haldane A, {del R{\'i}o} JF, Wiebe M, Peterson P,
+	{G{\'e}rard-Marchant} P, Sheppard K, Reddy T, Weckesser W, Abbasi H, Gohlke
+	C, Oliphant TE (2020) Array programming with {{NumPy}}. \emph{Nature}
+	585(7825):357--362,
+	\urlprefix\url{http://dx.doi.org/10.1038/s41586-020-2649-2}.
+
+	\bibitem[{Johnson(2020)}]{johnson2020benefits}
+	Johnson JW (2020) Benefits and pitfalls of jupyter notebooks in the classroom.
+	\emph{Proceedings of the 21st Annual Conference on Information Technology
+		Education}, 32--37.
+
+	\bibitem[{{Jupyter Team}(2015)}]{jupyterteamProjectJupyterDocumentation2015}
+	{Jupyter Team} (2015) Project jupyter documentation.
+	https://docs.jupyter.org/en/latest/.
+
+	\bibitem[{McKinney(2022)}]{mckinney2022python}
+	McKinney W (2022) \emph{Python for Data Analysis} ({O'Reilly}).
+
+	\bibitem[{Moffitt(2022)}]{moffittPracticalBusinessPython2022}
+	Moffitt C (2022) Practical {{Business Python}}. https://pbpython.com/.
+
+	\bibitem[{Nelson(2024)}]{nelsonSoftwareEngineeringData2024}
+	Nelson C (2024) \emph{Software {{Engineering}} for {{Data Scientists}}}
+	({O'Reilly}), ISBN 978-1-09-813619-2.
+
+	\bibitem[{Pedregosa et~al.(2011)Pedregosa, Varoquaux, Gramfort, Michel,
+		Thirion, Grisel, Blondel, Prettenhofer, Weiss, Dubourg, Vanderplas, Passos,
+		Cournapeau, Brucher, Perrot, \protect\BIBand{} Duchesnay}]{scikit-learn}
+	Pedregosa F, Varoquaux G, Gramfort A, Michel V, Thirion B, Grisel O, Blondel M,
+	Prettenhofer P, Weiss R, Dubourg V, Vanderplas J, Passos A, Cournapeau D,
+	Brucher M, Perrot M, Duchesnay E (2011) Scikit-learn: {{Machine}} learning in
+	{{Python}}. \emph{Journal of Machine Learning Research} 12:2825--2830.
+
+	\bibitem[{Perez \protect\BIBand{} Granger(2015)}]{perez2015project}
+	Perez F, Granger BE (2015) Project {{Jupyter}}: {{Computational}} narratives as
+	the engine of collaborative data science. \emph{Retrieved September}
+	11(207):108.
+
+	\bibitem[{Perkel(2018)}]{perkel2018jupyter}
+	Perkel JM (2018) Why {{Jupyter}} is data scientists' computational notebook of
+	choice. \emph{Nature} 563(7732):145--147.
+
+	\bibitem[{Powell(2001)}]{powellTeachingModelingManagement2001}
+	Powell SG (2001) Teaching {{Modeling}} in {{Management Science}}. \emph{INFORMS
+		Transactions on Education} 1(2):62--67, ISSN 1532-0545,
+	\urlprefix\url{http://dx.doi.org/10.1287/ited.1.2.62}.
+
+	\bibitem[{Powell \protect\BIBand{}
+		Baker(2019)}]{powellBusinessAnalyticsArt2019}
+	Powell SG, Baker KR (2019) \emph{Business {{Analytics}}: {{The Art}} of
+		{{Modeling With Spreadsheets}}, 5th {{Edition}} | {{Wiley}}}.
+
+	\bibitem[{Ragsdale(2017)}]{ragsdaleSpreadsheetModelingDecision2017}
+	Ragsdale C (2017) \emph{Spreadsheet {{Modeling}} \& {{Decision Analysis}}: {{A
+				Practical Introduction}} to {{Business Analytics}}} ({Australia}: {Cengage
+		Learning}), 8th edition edition, ISBN 978-1-305-94741-2.
+
+	\bibitem[{Rodrigues(2023)}]{rodriguesBuildingReproducibleAnalytical2023}
+	Rodrigues B (2023) Building reproducible analytical pipelines with {{R}}.
+	https://raps-with-r.dev/.
+
+	\bibitem[{Rule et~al.(2018)Rule, Birmingham, Zuniga, Altintas, Huang, Knight,
+		Moshiri, Nguyen, Rosenthal, P{\'e}rez et~al.}]{rule2018ten}
+	Rule A, Birmingham A, Zuniga C, Altintas I, Huang SC, Knight R, Moshiri N,
+	Nguyen MH, Rosenthal SB, P{\'e}rez F, et~al. (2018) Ten simple rules for
+	reproducible research in {{Jupyter}} notebooks. \emph{arXiv preprint
+		arXiv:1810.08055} .
+
+	\bibitem[{Savage(2012)}]{savageFlawAveragesWhy2012}
+	Savage SL (2012) \emph{The {{Flaw}} of {{Averages}}: {{Why We Underestimate
+				Risk}} in the {{Face}} of {{Uncertainty}}} ({Hoboken, New Jersey}: {Wiley}),
+	1st edition edition, ISBN 978-1-118-07375-9.
+
+	\bibitem[{Treadway(2023)}]{treadwaySoftwareEngineeringData2023}
+	Treadway A (2023) \emph{Software {{Engineering}} for {{Data Scientists}}}
+	({Manning}).
+
+	\bibitem[{Ufford et~al.(2019)Ufford, {Pacer, M}, {Seal, Matthew},
+		\protect\BIBand{} {Kelley, Kyle}}]{uffordInteractiveNotebookInnovation2019}
+	Ufford M, {Pacer, M}, {Seal, Matthew}, {Kelley, Kyle} (2019) Beyond
+	{{Interactive}}: {{Notebook Innovation}} at {{Netflix}}.
+	https://netflixtechblog.com/notebook-innovation-591ee3221233.
+
+	\bibitem[{Virtanen et~al.(2020)Virtanen, Gommers, Oliphant, Haberland, Reddy,
+		Cournapeau, Burovski, Peterson, Weckesser, Bright, {van der Walt}, Brett,
+		Wilson, Millman, Mayorov, Nelson, Jones, Kern, Larson, Carey, Polat, Feng,
+		Moore, VanderPlas, Laxalde, Perktold, Cimrman, Henriksen, Quintero, Harris,
+		Archibald, Ribeiro, Pedregosa, {van Mulbregt}, \protect\BIBand{} {SciPy 1.0
+			Contributors}}]{2020SciPy-NMeth}
+	Virtanen P, Gommers R, Oliphant TE, Haberland M, Reddy T, Cournapeau D,
+	Burovski E, Peterson P, Weckesser W, Bright J, {van der Walt} SJ, Brett M,
+	Wilson J, Millman KJ, Mayorov N, Nelson ARJ, Jones E, Kern R, Larson E, Carey
+	CJ, Polat {\.I}, Feng Y, Moore EW, VanderPlas J, Laxalde D, Perktold J,
+	Cimrman R, Henriksen I, Quintero EA, Harris CR, Archibald AM, Ribeiro AH,
+	Pedregosa F, {van Mulbregt} P, {SciPy 10 Contributors} (2020) {{SciPy}} 1.0:
+	{{Fundamental}} algorithms for scientific computing in python. \emph{Nature
+		Methods} 17:261--272,
+	\urlprefix\url{http://dx.doi.org/10.1038/s41592-019-0686-2}.
+
+	\bibitem[{Walls(2023)}]{wallsMathematicalPython2023}
+	Walls P (2023) Mathematical {{Python}}.
+	https://github.com/patrickwalls/mathematicalpython.
+
+	\bibitem[{Wickham(2014)}]{wickhamTidyData2014}
+	Wickham H (2014) Tidy {{Data}}. \emph{Journal of Statistical Software}
+	59:1--23, ISSN 1548-7660,
+	\urlprefix\url{http://dx.doi.org/10.18637/jss.v059.i10}.
+
+	\bibitem[{Winston \protect\BIBand{}
+		Albright(2018)}]{winstonPracticalManagementScience2018}
+	Winston WL, Albright SC (2018) \emph{Practical {{Management Science}}}
+	({Australia ; Boston, MA}: {Cengage Learning}), 6th edition edition, ISBN
+	978-1-337-40665-9.
+
+\end{thebibliography}
 

### 2023-11-20 python_excel_ite_r1.tex
created r1
edits for revision

--- 2023-07-28 python_excel_ite.tex
@@
 Spreadsheet based modeling has transformed the teaching of management science in business schools over the past twenty plus years.
-The late 1990's saw first editions of game changing textbooks \cite{winstonPracticalManagementScience2018,ragsdaleSpreadsheetModelingDecision2017} focused on the use of spreadsheets in teaching modeling and management science.
-A series of conferences focused on teaching management science with spreadsheets started at Dartmouth in 1998 and several presentations from that conference made up the bulk of papers in the first two issues of the INFORMS Transactions on Education (ITE) journal.
-The next few decades saw numerous textbooks \cite{powellBusinessAnalyticsArt2019, camm2020business} and teaching papers in journals such as \href{https://pubsonline.informs.org/journal/ited}{ITE} that prominently featured spreadsheet based modeling.
+The late 1990's saw first editions of game changing textbooks \citep{winstonPracticalManagementScience2018,ragsdaleSpreadsheetModelingDecision2017} focused on the use of spreadsheets in teaching modeling and management science.
+A series of conferences focused on teaching management science with spreadsheets started at Dartmouth in 1998 and several presentations from that conference made up the bulk of papers in the first two issues of the INFORMS Transactions on Education (ITE) journal \citep{baker2000:lp,bell2000:statistics,evans2000:simulation,carraway2000:mba,ragsdale2001:teaching,savage2001:blitzograms}.
+The next few decades saw numerous textbooks, including {\citet{powellBusinessAnalyticsArt2019, camm2020business}}, and teaching papers in journals such as \href{https://pubsonline.informs.org/journal/ited}{ITE} that prominently featured spreadsheet based modeling.
 
@@
 Numerous R and Python based data science courses have been developed and offered at universities and through online learning platforms such as \href{https://www.coursera.org/browse/data-science}{Coursera}, \href{https://www.edx.org/learn/data-science}{EdX}, \href{https://www.datacamp.com/}{DataCamp}, \href{https://www.udemy.com/topic/data-analysis/}{Udemy} and many others.
+One particularly relevant example are the courses in computation offered by the Naval Postgraduate School and described in \citet{alderson2022:interactive_computing}.
+In their courses, Jupyter notebooks (discussed in more detail in the next section) play a prominent role in providing an interactive development and learning environment for Python within an operations research context.
 In 2015, at Oakland University, we created a course called \href{http://www.sba.oakland.edu/faculty/isken/courses/pcda/}{Practical Computing for Data Analytics} (PCDA) that introduced both R and Python (along with a little Linux) for use in data analytics work.
@@
 Python has continued to make inroads to analytical territory that was formerly dominated by spreadsheets.
-The widely used Python data analysis library, pandas, was created by Wes McKinney while working as a quantitative financial analyst \cite{mckinney2022python}.
+The widely used Python data analysis library, pandas, was created by Wes McKinney while working as a quantitative financial analyst \citep{mckinney2022python}.
 Python is widely used for data wrangling, visualization and predictive modeling using various flavors of regression, tree based models and neural networks.
@@
 What would those look like in Python?
-The importance of modeling, articulated so well in the pages of this journal \cite{powellTeachingModelingManagement2001}, has not abated.
+The importance of modeling, articulated so well by \citet{powellTeachingModelingManagement2001} in the pages of this journal, has not abated.
 What role does Python have to play in supporting the kind of modeling taught for years in our business schools?
@@
 
-Recently there has been more recognition of the need for data scientists to learn basic software engineering principes \cite{nelsonSoftwareEngineeringData2024,treadwaySoftwareEngineeringData2023,rodriguesBuildingReproducibleAnalytical2023}.
+Recently there has been more recognition of the need for data scientists to learn basic software engineering principes \citep{nelsonSoftwareEngineeringData2024,treadwaySoftwareEngineeringData2023,rodriguesBuildingReproducibleAnalytical2023}.
 The image of newly minted data science graduates developing spaghetti coded Jupyter notebooks not nearly ready for being put into production has reached \href{https://www.reddit.com/r/ProgrammerHumor/comments/y2xe3n/like_every_time_ever_when_the_devops_engineer/}{meme worthy status}.
@@
 The PCDA course ends with intro to the \href{https://scikit-learn.org/stable/index.html}{scikit-learn library} for doing predictive modeling in Python and the AAP course begins where the PCDA course ends.
-The EwP module could also be used as part of a semester long Python based analytics course.
-Most AAP students have also already taken   \href{http://www.sba.oakland.edu/faculty/isken/courses/ba}{Business Analytics} (BA) which is a spreadsheet based introduction to business analytics.
+The EwP module follows immediately after the modules on machine learning and builds on the Python fundamentals which were covered in the PCDA course.
+The EwP module could also be used as part of a semester long Python based analytics course in which the Python modules covered in the PCDA course are followed by more advanced topics such as EwP or other topics in our AAP course.
+The only reason the EwP module is not included in the PCDA course is that Python is only covered in half of that course - the other half being an introduction to R and Linux.
+If you visit the open access \href{http://www.sba.oakland.edu/faculty/isken/courses/pcda/}{PCDA course website}  and \href{http://www.sba.oakland.edu/faculty/isken/courses/aap/}{AAP course website}, you can see the topic outline for each course and how the EwP module is positioned.
+
+Most AAP students have also already taken   \href{http://www.sba.oakland.edu/faculty/isken/courses/ba}{Business Analytics} (BA) which is a spreadsheet based introduction to business analytics (\cite{isken2003:muddy,isken2014:lab}).
 The BA and PCDA courses have historically been offered in face to face mode in a computer teaching lab.
@@
 \begin{itemize}
+	\item familiarity with Jupyter notebooks and an integrated development environment such as \href{https://www.spyder-ide.org/}{Spyder}, \href{https://code.visualstudio.com/}{VSCode} or \href{https://www.jetbrains.com/pycharm/}{PyCharm},
 	\item variables, arithmetic and boolean operators, basic data types,
@@
 All of the Python modules in the PCDA course as well as the entire AAP course are taught using a combination of Jupyter notebooks, Python scripts, and videos that walk readers through the notebooks.
-Jupyter notebooks, part of \href{https://jupyter.org/}{Project Jupyter}, are an example of \textit{literate computing} \cite{perez2015project} and allow one to create rich interactive documents that support iterative exploration and development.
+Jupyter notebooks, part of \href{https://jupyter.org/}{Project Jupyter}, are an example of \textit{literate computing} (\cite{perez2015project}) and allow one to create rich interactive documents that support iterative exploration and development.
 
@@
 
-The ability to weave standard explanatory language with formal programming languages has made notebook computing a popular tool in the data science education community \cite{perkel2018jupyter}.
+The ability to weave standard explanatory language with formal programming languages has made notebook computing a popular tool in the data science education community (\cite{perkel2018jupyter}).
 Jupyter notebooks can also include equations written in \LaTeX, data visualizations, and other multi-media.
 The technology is free, open source, runs on most computing platforms, and has a huge user community that has created a large number of resources for students, faculty and industry practitioners.
-There are countless introductory tutorials for Jupyter notebooks, but the official Project Jupyter Documentations \cite{jupyterteamProjectJupyterDocumentation2015} is a very good place to start if you are new to notebook computing.
+There are countless introductory tutorials for Jupyter notebooks, but the official Project Jupyter Documentation (\cite{jupyterteamProjectJupyterDocumentation2015}) is a very good place to start if you are new to notebook computing.
 As with all tools, there are pros and cons to their use.
@@
 
-There are two overarching learning objectives for this \href{http://www.sba.oakland.edu/faculty/isken/courses/aap/mod3a_whatif.html}{first EwP submodule}.
+There are two overarching learning objectives for this \href{http://www.sba.oakland.edu/faculty/isken/courses/aap/mod3a_whatif.html}{first EwP submodule} - see Figure \ref{fig:objectives_whatif} for how the objectives are presented to students via the course website.
 Students learn how Python might be used to do something they are quite familiar with doing in a spreadsheet -- model building and sensitivity analysis.
@@
 
+\begin{figure}[!htbp]
+	\includegraphics[width=1.0\textwidth]{objectives_whatif}
+	\caption{Learning objectives for Submodule 1}
+	\label{fig:objectives_whatif}
+\end{figure}
+
 \subsection{Notebook 1: Modeling and data tables (\href{http://www.sba.oakland.edu/faculty/isken/excel_with_python/what_if_1_model_datatable.html}{html version})}
@@
 The notebook also includes a screenshot (Figure \ref{fig:xl_model}) of this Excel model.
+Readers are encouraged to click the \href{http://www.sba.oakland.edu/faculty/isken/excel_with_python/what_if_1_model_datatable.html}{html version} link so that you can see all of the notebook details as you are reading through this section.
 
@@
 This difference in behavior between Jupyter notebooks and spreadsheets might be one of the most important concepts for students to internalize.
-A well known challenge, often referred to as the \textit{hidden state problem}, associated with computational notebooks is the ease of running code cells out of order and introducing inadvertent errors \cite{johnson2020benefits,grusDonNotebooks2018}.
+A well known challenge, often referred to as the \textit{hidden state problem}, associated with computational notebooks is the ease of running code cells out of order and introducing inadvertent errors -- see \cite{johnson2020benefits,grusDonNotebooks2018}.
 This point along with strategies to avoid it are discussed during the first week of the course which focuses on \href{http://www.sba.oakland.edu/faculty/isken/courses/aap/jupyter_notebooks.html}{notebook computing best practices as well as a historical look} at computational notebooks.
@@
 In a Data Table, a range of order quantities can be used as row or column input and one or more output values can be computed.
-In Python, the vectorized nature of the \href{https://numpy.org/}{NumPy} library \cite{harris2020array} provides a simple way to accomplish the same thing.
+In Python, the vectorized nature of the \href{https://numpy.org/}{NumPy} library (\cite{harris2020array}) provides a simple way to accomplish the same thing.
 A range of order quantities is created as a NumPy array that can be used directly to compute vectors of all of the intermediate and final output variables.
@@
 This provides a great opportunity to remind students of one of the great strengths of the open source software ecosystem -- we can leverage work done in other packages and even look at actual source code to see how something was implemented.
-In this case, the well known \href{https://scikit-learn.org/stable/}{scikit-learn} package \cite{scikit-learn} has a \code{ParameterGrid} function that solves our scenario generation problem quite nicely.
+In this case, the well known \href{https://scikit-learn.org/stable/}{scikit-learn} package (\cite{scikit-learn}) has a \code{ParameterGrid} function that solves our scenario generation problem quite nicely.
 This example also reminds the students of the value of being able to make sense of API documentation and to understand object oriented code written by others.
@@
 
-Before trying to build our \code{goal\_seek} function we take a brief detour into basic root finding algorithms through packages like \href{https://docs.scipy.org/doc/scipy/tutorial/optimize.html#root-finding}{SciPy} \cite{2020SciPy-NMeth} as well as in \href{https://github.com/patrickwalls/mathematical-python/}{open source notebooks and scripts} that are widely available in the Python data science ecosystem \cite{wallsMathematicalPython2023}.
+Before trying to build our \code{goal\_seek} function we take a brief detour into basic root finding algorithms through packages like \href{https://docs.scipy.org/doc/scipy/tutorial/optimize.html#root-finding}{SciPy} (\cite{2020SciPy-NMeth}) as well as in \href{https://github.com/patrickwalls/mathematical-python/}{open source notebooks and scripts} that are widely available in the Python data science ecosystem (\cite{wallsMathematicalPython2023}).
 We compare the output of Excel's Goal Seek with different starting values to the values obtained by different root finding algorithms implemented in Python.
@@
 Then we can do standard statistical analysis of the simulation output using Python packages such as pandas and SciPy.
-In spreadsheet based simulation textbooks, the notion of the ``Flaw of Averages'' popularized by Sam Savage \cite{savageFlawAveragesWhy2012} is often illustrated.
+In spreadsheet based simulation textbooks, the notion of the ``Flaw of Averages'' popularized by Sam Savage (\cite{savageFlawAveragesWhy2012}) is often illustrated.
 We do the same with our Python based model, showing how replacing random demand by its mean can result in a wildly optimistic estimate of the mean profit found through analyzing the simulation output.
@@
 Certainly we do not want to rely on copying, pasting and modifying code when building a new model.
-In the Python world, the natural next step is to deploy our code as a Python \textit{package} \cite{buezenWelcomePythonPackages2022} that can then be imported like any other Python library.
+In the Python world, the natural next step is to deploy our code as a Python \textit{package} (\cite{buezenWelcomePythonPackages2022}) that can then be imported like any other Python library.
+Figure \ref{fig:objectives_packaging} shows the learning objectives from this \href{http://www.sba.oakland.edu/faculty/isken/courses/aap/mod3b_whatif_packaging.html}{submodule page on the course website}.
+
+\begin{figure}[!htbp]
+	\centering
+	\includegraphics[scale=0.5]{objectives_packaging}
+	\caption{Learning objectives for Submodule 2}
+	\label{fig:objectives_packaging}
+\end{figure}
 
@@
 Thinking about how others will interact with your software usually leads to better designed software.
-In \href{http://www.sba.oakland.edu/faculty/isken/courses/aap/mod3b_whatif_packaging.html}{this submodule} we create a Python package that can be imported and used by other modelers.
+In this submodule we create a Python package that can be imported and used by other modelers.
 This will require some design changes to our software and will also involve moving code out of Jupyter notebooks and into Python script files.
@@
 We then revisit a concept from earlier in the course -- creating a good project folder structure.
-For this we use what is known as a \href{https://cookiecutter.readthedocs.io/en/stable/}{cookiecutter} that can automatically generate a project folder structure and key files from the answers to a few prompts \cite{audreyroyandcookiecuttercommunityCookiecutterBetterProject2012}.
+For this we use what is known as a \href{https://cookiecutter.readthedocs.io/en/stable/}{cookiecutter} that can automatically generate a project folder structure and key files from the answers to a few prompts (\cite{audreyroyandcookiecuttercommunityCookiecutterBetterProject2012}).
 There are a few different folder layouts used in Python projects and there is a bit of discussion on the folder structure we will use and resources given for exploring this issue further.
@@
 
-This notebook ends by pointing the students to another notebook in which our newly deployed whatif package is used for a completely different model -- the New Car Simulation that has been a part of Winston and Albright's textbooks for many years \cite{winstonPracticalManagementScience2018,albrightBusinessAnalyticsData2016}.
+This notebook ends by pointing the students to another notebook in which our newly deployed whatif package is used for a completely different model -- the New Car Simulation that has been a part of Winston and Albright's textbooks for many years (\cite{winstonPracticalManagementScience2018,albrightBusinessAnalyticsData2016}).
 This new model differs structurally from the Bookstore Model, but is representative of many financial models in that cash flows over multiple years need to be computed and summarized.
@@
 Each of these examples are based on a real problems I encountered either in research or industrial projects.
-Several web based resources are shared including the \href{https://pbpython.com/}{Practical Business Python blog} written by Chris Moffitt, a data analytics professional \cite{moffittPracticalBusinessPython2022}.
+Several web based resources are shared including the \href{https://pbpython.com/}{Practical Business Python blog} written by Chris Moffitt, a data analytics professional (\cite{moffittPracticalBusinessPython2022}).
 This blog, which focuses on helping Excel-centric analytics professionals transition some of their work to Python, includes numerous posts on different aspects of Python and Excel integration.

### 2023-11-21 response_to_r1.md
finished response to reviewers

# Response to reviewers of "Spreadsheet modeling and wrangling with Python"

I'd like to thank the reviewer and the editor for their thoughtful comments on this paper. I have
done by best to address them.

## Editor comments

**Comment 1**

I would like to apologize to the author for the delay in the review.  Although two reviewers accepted the assignment, one of them, at this point, is 65 days late with the review – therefore, I decided to proceed with a single review.

My first request to the author is to include other related ITE publications either as part of the literature review or to the other relevant sections.  Here is a short list of links to start:
•       https://pubsonline.informs.org/doi/pdf/10.1287/ited.2021.0261
•       https://pubsonline.informs.org/doi/full/10.1287/ited.2023.0285
•       https://pubsonline.informs.org/doi/pdf/10.1287/ited.2022.0275

**Response 1**

In the opening paragraph I refer to the seminal issues of ITE but didn't reference specific papers. I've rectified that and have added the following references.

- Alderson DL (2022) Interactive computing for accelerated learning in computation and data science. INFORMS Transactions on Education 22(2):130–145, URL http://dx.doi.org/10.1287/ited.
2021.0261.
- Baker K (2000) Gaining insight in linear programming from patterns in optimal solutions. INFORMS Transactions on Education 1(1):4–17, URL http://dx.doi.org/10.1287/ited.1.1.4.
- Bell PC (2000) Teaching business statistics with microsoft excel. INFORMS Transactions on Education 1(1):18–26, URL http://dx.doi.org/10.1287/ited.1.1.18.
- Carraway RL, Clyman DR (2000) Integrating spreadsheets into a case-based mba quantitative methods course: Real managers make real decisions. INFORMS Transactions on Education 1(1):38–46, URL http://dx.doi.org/10.1287/ited.1.1.38.
- Evans JR (2000) Spreadsheets as a tool for teaching simulation. INFORMS Transactions on Education 1(1):27–37, URL http://dx.doi.org/10.1287/ited.1.1.27.
- Isken MW (2003) Data cleansing and analysis as a prelude to model based decision support. INFORMS Transactions on Education 3(3):23–75, URL http://dx.doi.org/10.1287/ited.3.3.23.
- Isken MW (2014) Translating a lab based spreadsheet modeling course to an online format: Experience from a natural experiment. INFORMS Transactions on Education 14(3):120–128, URL http://dx.doi.org/10.1287/ited.2013.0123.
- Powell SG (2001) Teaching Modeling in Management Science. INFORMS Transactions on Education 1(2):62–67, ISSN 1532-0545, URL http://dx.doi.org/10.1287/ited.1.2.62.
- Ragsdale CT (2001) Teaching management science with spreadsheets: From decision models to decision support. INFORMS Transactions on Education 1(2):68–74, URL http://dx.doi.org/10.1287/ited.1.2.68.
- Savage S (2001) Blitzograms—interactive histograms. INFORMS Transactions on Education 1(2):77–87, URL http://dx.doi.org/10.1287/ited.1.2.77.

The publication by Alderson (2022) is very relevant and I'm not sure how I missed seeing that. I've integrated it into the paper and cited it. Thank you for bringing it to my attention.

"One particularly relevant example are the courses in computation offered by the Naval Postgraduate School and described in Alderson (2022). In their courses, Jupyter notebooks (discussed in more detail in the next section) play a prominent role in providing an interactive development and learning environment for Python within an operations research context."

In the first version I avoid citing my own work in ITE even though it is quite relevant. It now seems appropriate now to add these and I have done
that in the section describing my speadsheet based analytics course.

**Comment 2**

Second, I found it difficult to follow the structure of the environment within which the course it taught.  Could the author please include a graphic or a table to add a visualization to the explanation of where in the program’s curriculum the AAP course fits in, and where in the AAP course the EwP module fits in.

**Response 2**

I understand. I've added more explanation to the relevant paragraph. As opposed to adding a table or graphic that abstracts the details, I'm encouraging the readers to visit the relevant course landing pages as they include nicely structured topic outlines that make it very clear how EwP fits in to the Python material covered by our PCDA and AAP courses.

Here's the reworked paragraph:

"At our institution, The EwP module is part of a business course entitled  [Advanced Analytics with
Python](http://www.sba.oakland.edu/faculty/isken/courses/aap) (AAP). The students taking this course have already taken the PCDA course within which they learned fundamental Python programming within the context of data analytics over seven weeks. The PCDA course ends with intro to the [scikit-learn library](https://scikit-learn.org/stable/index.html) for doing predictive modeling in Python and the AAP course begins where the PCDA course ends. The EwP module follows immediately after the modules on machine learning and builds on the Python fundamentals which were covered in the PCDA course. The EwP module could also be used as part of a semester long Python based analytics course in which the Python modules covered in the PCDA course are followed by more advanced topics such as EwP or other topics in our AAP course. The only reason the EwP module is not included in the PCDA course is that Python is only covered in half of that course - the other half being an introduction to R and Linux. If you visit the open access [PCDA course
website](http://www.sba.oakland.edu/faculty/isken/courses/pcda/)  and [AAP course
website](http://www.sba.oakland.edu/faculty/isken/courses/aap/), you can see the topic outline for each course and how the EwP module is positioned."

I'd like to stress that a big part of my paper is that it isn't just a paper. It is backed by a huge amount of public, freely available and Creative Commons licensed course websites and course content. Anyone who is interested in adapting my materials for their course is free to do so. Everything other than all of the homework assignments are publicly accessible from the course websites. That's why throughout this paper I provide hyperlinks to relevant parts of the EwP course website or course content. This allows the reader to have a richer reading experience - instead of just a paper which necessarily is limited by what it can contain, the reader can view the entirety of the Jupyter notebooks, in html format, which make up the modules. ITE is an online journal and I'm trying to take advantage of the capabilities offered by such a platform.

**Comment 3**

Are the students already familiar with the Jupyter notebook environment from their PCDA course, and therefore the instructor continues using the same environment, or are the students using a different development environment in their preceding course?

**Response 3**

Yes, we use Jupyter notebooks as well as a integrated development environment (either Spyder, PyCharm, or VSCode) in the PCDA course. The same tools are used in the AAP course. All of the technical and logistical details of both courses are described in excruciating detail in the public course websites linked to in the paper.

**Comment 4**

The same with source control – it is not clear to me why the program chose to include software engineering related discussion about source control in the EwP module of the AAP course, could the author please add a bit of discussion around that?

**Response 4**

I didn't choose that. Version control is covered in the very first module of the AAP course and students use git and GitHub for the ML module that precedes the EwP module. The following sentence is in the original version of the paper:

"Students are reminded of the importance of version control and we initialize a git repository for our budding package. The basics of version control were already covered in the first week of the course."

Again, the AAP course website main page at [http://www.sba.oakland.edu/faculty/isken/courses/aap/](http://www.sba.oakland.edu/faculty/isken/courses/aap/) contains a detailed topic outline and the reader can see what is covered when.

**Comment 5**

Third, would it be possible to add the learning goals of the EwP module and suggested methods for assessments?

**Response 5**

Each submodule of EwP has a number of learning objectives that appear at the top of each
submodule course webpage. I've added screenshots of these objectives so that readers can see
how the objectives are presented to the students.

As for assessment, in the final section of the paper, *Classroom experience and the future*, we described and made available for download an assignment that we have used for assessment of this material. It covers all aspects of the EwP module.

**Comment 6**

Fourth, the author should also address the comments and suggestions provided by Reviewer 1.

Thank you for the submission to ITE and sharing your ideas and experiences with our academic community.  My apology again for the delay in the review process.

**Response 6**

See below for comments and responses to Reviewer 1.

## Reviewer #1 comments

Thank you taking the time to review my paper and to provide thoughtful feedback.

**Comment 1**

In the Introduction section, there appears to be too much emphasis on offering
the EwP approach as an alternative to Excel due to its limitations. I don’t think
many folks will buy the argument. The discussion seems to shortchange Excel’s
ability to set up connections with externally managed data while maintaining the
flexibility and ease of summarizing data in Excel. An alternative is to simply
suggest previous work in Excel by students provides a good launch point for
developing the more advanced programming skills in Python.

**Response 1**

I'm definitely not trying to offer EwP as an alternative to Excel. I've taught an Excel based spreadsheet modeling course for over 20 years and have used Excel extensively in both industry as well as in academic settings. I love spreadsheets and have for a long time. But, they do
have their limitations and the better our students are at assessing when a spreadsheet is and is not appropriate, the better off they will be. This is not some novel argument. Excel bashing is rampant in corporate IT and analytics settings. Quite honestly, even Jupyter notebooks have major shortcomings when it comes to productionizing analytical models implemented in Python and I mention this in the paper - and that's why I also make students use an actual IDE to write code in addition to the things we do in notebooks. There's a lot of nuance in knowing how to effectively wield tools like Excel and Python in professional analytics environments. This is a big part of the motivation behind the development of my courses and in writing this paper.

My intent is to provide an interesting way for business students to learn more advanced Python concepts through a very familiar lens of spreadsheet modeling. In addition, I want to expose them to the abilities of Python in automating manipulation of spreadsheets. Now with MS announcing the inclusion of Python within Excel, the complementary nature of these two powerhouse tools will become even more important. To that end, I've modified the beginning of that paragraph as follows:

"Despite numerous calls for, and predictions of, the eradication of spreadsheets from the business world, Excel and other spreadsheets are \href{https://benn.substack.com/p/the-next-billion-programmers}{alive and well}. Their flexibility is unparalleled and it is unrealistic to expect them to disappear anytime soon. If anything, the [recent announcement by Microsoft](https://techcommunity.microsoft.com/t5/excel-blog/announcing-python-in-excel-combining-the-power-of-python-and-the/ba-p/3893439)
 that users will be able combine Python scripts with Excel formulas within the same workbook, will likely solidify Excel's place in the analytics space."

**Comment 2**

In sections 3.2 and 3.3, it would have been nice to see more figures illustrating
the coding associated with Goal Seek and Monte Carlo simulation exercises.

**Response 2**

I wrestled with this when writing the paper. Instead of including more figures, I opted, as I described above, to include prominent hyperlinks to HTML versions of the notebooks in the relevant section headers.

For example:

3.3. Notebook 3: Monte-Carlo simulation (html version)

I also added explicit language early in the paper that highlighted these hyperlinks and encouraged readers to use them.

"While all of the notebooks are publicly available (see Section 6), links to html versions of
each of the notebooks are provided throughout the document and readers are encouraged to use
them while reading to see exactly how each topic is presented."

My intent is for the reader to open this html version in a browser tab and then they can see the entire notebook. Code snippets taken out of context are not nearly as useful as following the narrative flow of the notebook - this is the real strength of Jupyter notebooks for teaching.

If the reviewers really want me to add more code screenshots, I can, but I'd really prefer readers look at the (html versions of) the actual notebooks. Of course, the actual Jupyter notebooks are also available from the course website.

**Comment 3**

In section 7, the author(s) need to more fully address the roll out of Python in
Excel. Does this alter their pedagogy? It may not if the focus is develop the skills
outlined in the Introduction and listed below
o  teaching relevant and more advanced Python to business students using familiar and
simple modeling examples,  conveying a sense of the software design and development
process,  acting as a tour guide for a learning journey,  teaching students to think and
act like a software developer, not just an analyst
In any case, I think the paper needs a more comprehensive discussion regarding
this development.

**Response 3**

I've added reference to this development in the opening section as I mentioned in Response 1 above. I've also added a bit in the last section:

"Rumors have swirled for a number of years that Microsoft was considering adding Python as a first class language alternative to VBA within their MS Office suite of packages \cite{cimpanuMicrosoftConsidersAdding2017}. During the review process for this article, this development [finally
materialized](https://techcommunity.microsoft.com/t5/excel-blog/announcing-python-in-excel-combining-the-power-of-python-and-the/ba-p/3893439)
, albeit in a limited roll out to beta testers. This has potentially enormous implications for the future of both Excel and Python. At this point, it is probably safe to say that this development further underscores the importance of business analytics students adding Python to their set of analytical toolbox."

Other than that, I'm not in a position to say a whole lot more right now. The roll out was restricted to those with Beta Channel access (which I do not have). It appears right now that this feature is designed to run in the Azure cloud environment, which is a little concerning. It's also not clear if we'll be able to write full blown code modules in Python instead of Excel VBA. It's just too early to know how this is going to play out.
If and when this development is widely released, then certainly I will develop a teaching module to add to my AAP class and will likely add a very introductory module to my spreadsheet based analytics course. This will be a rich area for collaboration between the Excel and Python communities. Pretty exciting development but pales a bit in comparison to the chaos around Microsoft and OpenAI right now. :)

**Editing Comments**

On page 5, 2 nd paragraph – The sentence “ After running the cells, they appear as show
in Figure 2.” – The word show should be changed to shown

**Response**

Oops. It's fixed now.

### 2023-11-21 python_excel_ite_r1.tex
finished response to reviewers

@@
 Despite numerous calls for, and predictions of, the eradication of spreadsheets from the business world, Excel and other spreadsheets are \href{https://benn.substack.com/p/the-next-billion-programmers}{alive and well}.
+Their flexibility is unparalleled and it is unrealistic to expect them to disappear anytime soon.
+If anything, the \href{https://techcommunity.microsoft.com/t5/excel-blog/announcing-python-in-excel-combining-the-power-of-python-and-the/ba-p/3893439}{recent announcement by Microsoft} that users will be able combine Python scripts with Excel formulas within the same workbook, will likely solidify Excel's place in the analytics space.
 Yes, spreadsheets have well known limitations.
@@
 
-Most AAP students have also already taken   \href{http://www.sba.oakland.edu/faculty/isken/courses/ba}{Business Analytics} (BA) which is a spreadsheet based introduction to business analytics (\cite{isken2003:muddy,isken2014:lab}).
+Most AAP students have also already taken   \href{http://www.sba.oakland.edu/faculty/isken/courses/ba}{Business Analytics} (BA) which is a spreadsheet based introduction to business analytics \citep{isken2003:muddy,isken2014:lab}.
 The BA and PCDA courses have historically been offered in face to face mode in a computer teaching lab.
@@
 All of the Python modules in the PCDA course as well as the entire AAP course are taught using a combination of Jupyter notebooks, Python scripts, and videos that walk readers through the notebooks.
-Jupyter notebooks, part of \href{https://jupyter.org/}{Project Jupyter}, are an example of \textit{literate computing} (\cite{perez2015project}) and allow one to create rich interactive documents that support iterative exploration and development.
+Jupyter notebooks, part of \href{https://jupyter.org/}{Project Jupyter}, are an example of \textit{literate computing} \citep{perez2015project} and allow one to create rich interactive documents that support iterative exploration and development.
 
@@
 For example, Figure \ref{fig:cells_edit_mode} shows two markdown cells separated by a code cell as they look when in editing mode.
-After running the cells, they appear as show in Figure \ref{fig:cells_after_running}.
+After running the cells, they appear as shown in Figure \ref{fig:cells_after_running}.
 After executing a code cell, the output appears directly below it in an output cell.
@@
 
-The ability to weave standard explanatory language with formal programming languages has made notebook computing a popular tool in the data science education community (\cite{perkel2018jupyter}).
+The ability to weave standard explanatory language with formal programming languages has made notebook computing a popular tool in the data science education community \citep{perkel2018jupyter}.
 Jupyter notebooks can also include equations written in \LaTeX, data visualizations, and other multi-media.
 The technology is free, open source, runs on most computing platforms, and has a huge user community that has created a large number of resources for students, faculty and industry practitioners.
-There are countless introductory tutorials for Jupyter notebooks, but the official Project Jupyter Documentation (\cite{jupyterteamProjectJupyterDocumentation2015}) is a very good place to start if you are new to notebook computing.
+There are countless introductory tutorials for Jupyter notebooks, but the official Project Jupyter Documentation \citep{jupyterteamProjectJupyterDocumentation2015} is a very good place to start if you are new to notebook computing.
 As with all tools, there are pros and cons to their use.
-See \cite{rule2018ten, barba2019teaching} and \cite{johnson2020benefits} for guidance on Jupyter notebook best practices for both research and teaching.
+See \citet{rule2018ten, barba2019teaching} and \citet{johnson2020benefits} for guidance on Jupyter notebook best practices for both research and teaching.
 
@@
 Students learn how Python might be used to do something they are quite familiar with doing in a spreadsheet -- model building and sensitivity analysis.
-At the same time, students encounter, as needed, more advanced Python programming concepts and techniques.
+At the same time, they encounter more advanced Python programming concepts and techniques.
 In this way, the computational problem to be solved drives the introduction of more advanced Python concepts and techniques instead of such things being presented in a vacuum.
@@
 Every notebook is accompanied by screencasts that walk the students through the material and act as a guided tour with challenges to overcome along the way.
-While all of the notebooks are publicly available (see Section \ref{sec:software}), links to an html version of each of the notebooks are provided throughout the document.
+While all of the notebooks are publicly available (see Section \ref{sec:software}), links to html versions of each of the notebooks are provided throughout the document and readers are encouraged to use them while reading to see exactly how each topic is presented.
 
@@
 The notebook also includes a screenshot (Figure \ref{fig:xl_model}) of this Excel model.
-Readers are encouraged to click the \href{http://www.sba.oakland.edu/faculty/isken/excel_with_python/what_if_1_model_datatable.html}{html version} link so that you can see all of the notebook details as you are reading through this section.
 
@@
 We then introduce a similar modeling problem that will be the focus of the rest of the notebook.
-The Walton Bookstore problem appears in the chapter on Monte-Carlo simulation in textbooks that I have used in my spreadsheet modeling course \cite{albrightBusinessAnalyticsData2016,winstonPracticalManagementScience2018}.
+The Walton Bookstore problem appears in the chapter on Monte-Carlo simulation in textbooks that I have used in my spreadsheet modeling course \citep{albrightBusinessAnalyticsData2016,winstonPracticalManagementScience2018}.
 The (slightly modified) problem is as follows:
@@
 This difference in behavior between Jupyter notebooks and spreadsheets might be one of the most important concepts for students to internalize.
-A well known challenge, often referred to as the \textit{hidden state problem}, associated with computational notebooks is the ease of running code cells out of order and introducing inadvertent errors -- see \cite{johnson2020benefits,grusDonNotebooks2018}.
+A well known challenge, often referred to as the \textit{hidden state problem}, associated with computational notebooks is the ease of running code cells out of order and introducing inadvertent errors -- see \citet{johnson2020benefits,grusDonNotebooks2018}.
 This point along with strategies to avoid it are discussed during the first week of the course which focuses on \href{http://www.sba.oakland.edu/faculty/isken/courses/aap/jupyter_notebooks.html}{notebook computing best practices as well as a historical look} at computational notebooks.
@@
 In a Data Table, a range of order quantities can be used as row or column input and one or more output values can be computed.
-In Python, the vectorized nature of the \href{https://numpy.org/}{NumPy} library (\cite{harris2020array}) provides a simple way to accomplish the same thing.
+In Python, the vectorized nature of the \href{https://numpy.org/}{NumPy} library \citep{harris2020array} provides a simple way to accomplish the same thing.
 A range of order quantities is created as a NumPy array that can be used directly to compute vectors of all of the intermediate and final output variables.
@@
 
-Before trying to build our \code{goal\_seek} function we take a brief detour into basic root finding algorithms through packages like \href{https://docs.scipy.org/doc/scipy/tutorial/optimize.html#root-finding}{SciPy} (\cite{2020SciPy-NMeth}) as well as in \href{https://github.com/patrickwalls/mathematical-python/}{open source notebooks and scripts} that are widely available in the Python data science ecosystem (\cite{wallsMathematicalPython2023}).
+Before trying to build our \code{goal\_seek} function we take a brief detour into basic root finding algorithms through packages like \href{https://docs.scipy.org/doc/scipy/tutorial/optimize.html#root-finding}{SciPy} \citep{2020SciPy-NMeth} as well as in \href{https://github.com/patrickwalls/mathematical-python/}{open source notebooks and scripts} that are widely available in the Python data science ecosystem \citep{wallsMathematicalPython2023}.
 We compare the output of Excel's Goal Seek with different starting values to the values obtained by different root finding algorithms implemented in Python.
@@
 Then we can do standard statistical analysis of the simulation output using Python packages such as pandas and SciPy.
-In spreadsheet based simulation textbooks, the notion of the ``Flaw of Averages'' popularized by Sam Savage (\cite{savageFlawAveragesWhy2012}) is often illustrated.
+In spreadsheet based simulation textbooks, the notion of the ``Flaw of Averages'' \citep{savageFlawAveragesWhy2012} is often illustrated.
 We do the same with our Python based model, showing how replacing random demand by its mean can result in a wildly optimistic estimate of the mean profit found through analyzing the simulation output.
@@
 Certainly we do not want to rely on copying, pasting and modifying code when building a new model.
-In the Python world, the natural next step is to deploy our code as a Python \textit{package} (\cite{buezenWelcomePythonPackages2022}) that can then be imported like any other Python library.
+In the Python world, the natural next step is to deploy our code as a Python \textit{package} \citep{buezenWelcomePythonPackages2022} that can then be imported like any other Python library.
 Figure \ref{fig:objectives_packaging} shows the learning objectives from this \href{http://www.sba.oakland.edu/faculty/isken/courses/aap/mod3b_whatif_packaging.html}{submodule page on the course website}.
@@
 This will require some design changes to our software and will also involve moving code out of Jupyter notebooks and into Python script files.
-While Jupyter notebooks can be integrated into production environments \cite{uffordInteractiveNotebookInnovation2019}, it is much more common for code to live in Python script files.
+While Jupyter notebooks can be integrated into production environments \citep{uffordInteractiveNotebookInnovation2019}, it is much more common for code to live in Python script files.
 In addition to creating a deployable package, this submodule also discusses the importance of documentation in its many forms in a Python project.
@@
 We then revisit a concept from earlier in the course -- creating a good project folder structure.
-For this we use what is known as a \href{https://cookiecutter.readthedocs.io/en/stable/}{cookiecutter} that can automatically generate a project folder structure and key files from the answers to a few prompts (\cite{audreyroyandcookiecuttercommunityCookiecutterBetterProject2012}).
+For this we use what is known as a \href{https://cookiecutter.readthedocs.io/en/stable/}{cookiecutter} that can automatically generate a project folder structure and key files from the answers to a few prompts \citep{audreyroyandcookiecuttercommunityCookiecutterBetterProject2012}.
 There are a few different folder layouts used in Python projects and there is a bit of discussion on the folder structure we will use and resources given for exploring this issue further.
@@
 
-This notebook ends by pointing the students to another notebook in which our newly deployed whatif package is used for a completely different model -- the New Car Simulation that has been a part of Winston and Albright's textbooks for many years (\cite{winstonPracticalManagementScience2018,albrightBusinessAnalyticsData2016}).
+This notebook ends by pointing the students to another notebook in which our newly deployed whatif package is used for a completely different model -- the New Car Simulation that has been a part of Winston and Albright's textbooks for many years \citep{winstonPracticalManagementScience2018,albrightBusinessAnalyticsData2016}.
 This new model differs structurally from the Bookstore Model, but is representative of many financial models in that cash flows over multiple years need to be computed and summarized.
@@
 
-This deployment focused submodule ends with a short notebook discussing the different types of documentation needed in a typical project.
+The deployment focused submodule ends with a short notebook discussing the different types of documentation needed in a typical project.
 This includes code comments, \href{https://peps.python.org/pep-0257/}{docstrings}, \code{readme} files and generating documentation by creating \href{https://www.sphinx-doc.org/en/master/usage/restructuredtext/basics.html}{reStructuredText} files (similar to but more powerful than markdown) and using \href{https://www.sphinx-doc.org/en/master/index.html}{Sphinx}, the Python documentation generation tool.
@@
 Each of these examples are based on a real problems I encountered either in research or industrial projects.
-Several web based resources are shared including the \href{https://pbpython.com/}{Practical Business Python blog} written by Chris Moffitt, a data analytics professional (\cite{moffittPracticalBusinessPython2022}).
+Several web based resources are shared including the \href{https://pbpython.com/}{Practical Business Python blog} written by Chris Moffitt, a data analytics professional \citep{moffittPracticalBusinessPython2022}.
 This blog, which focuses on helping Excel-centric analytics professionals transition some of their work to Python, includes numerous posts on different aspects of Python and Excel integration.
@@
 
-Reshaping, or tidying, data \cite{wickhamTidyData2014} from wide to long format is a common data wrangling task especially in preparation for plotting.
+Reshaping, or tidying, data \citep{wickhamTidyData2014} from wide to long format is a common data wrangling task especially in preparation for plotting.
 In this example, not only do students have to grapple with reshaping the data in an Excel workbook, they also have to deal with the all too common problem of multi-row header lines in a spreadsheet.
@@
 Rumors have swirled for a number of years that Microsoft was considering adding Python as a first class language alternative to VBA within their MS Office suite of packages \cite{cimpanuMicrosoftConsidersAdding2017}.
-While this has not yet materialized, there continues to be significant interest in using Python to manipulate not only Excel workbooks but other office documents such as Powerpoint presentations \cite{cannyPythonpptx2013}.
+During the review process for this article, this development \href{https://techcommunity.microsoft.com/t5/excel-blog/announcing-python-in-excel-combining-the-power-of-python-and-the/ba-p/3893439}{finally materialized}, albeit in a limited roll out to beta testers.
+This has potentially enormous implications for the future of both Excel and Python.
+At this point, it is probably safe to say that this development further underscores the importance of business analytics students adding Python to their set of analytical toolbox.
+There continues to be significant interest in using Python to manipulate not only Excel workbooks but other office documents such as Powerpoint presentations \citep{cannyPythonpptx2013}.
 Given that the ubiquity of both Excel and Powerpoint in business education and practice, inclusion of this material in a business school Python based analytics course seems quite appropriate.
 Students can learn more advanced and extremely useful Python concepts and techniques within a context with which they are intimately familiar.
-Manipulation of MS Office documents is directly relevant to the job content of many of the students.
+Manipulation of MS Office documents is directly relevant to the job content of many of our students.
 This teaching module is a natural complement to the commonly taught spreadsheet based analytics and introductory data science courses found in many business schools and provides an interesting way to teach and learn Python for analytics.

### 2023-11-21 response_to_r1.md
finished revisions and response to reviewers

@@
 I'd like to thank the reviewer and the editor for their thoughtful comments on this paper. I have
-done by best to address them.
+done by best to address them and to improve the paper.
 
@@
 
-"One particularly relevant example are the courses in computation offered by the Naval Postgraduate School and described in Alderson (2022).
-In their courses, Jupyter notebooks (discussed in more detail in the next section) play a prominent role in providing an interactive development and learning environment for Python within an operations research context."
+> One particularly relevant example are the courses in computation offered by the Naval Postgraduate School and described in Alderson (2022).
+In their courses, Jupyter notebooks (discussed in more detail in the next section) play a prominent role in providing an interactive development and learning environment for Python within an operations research context.
 
-In the first version I avoid citing my own work in ITE even though it is quite relevant. It now seems appropriate now to add these and I have done
+In the first version I avoid citing my own work in ITE even though it is quite relevant. It now seems appropriate to add these and I have done
 that in the section describing my speadsheet based analytics course.
@@
 
-"At our institution, The EwP module is part of a business course entitled  [Advanced Analytics with
+> At our institution, The EwP module is part of a business course entitled  [Advanced Analytics with
 Python](http://www.sba.oakland.edu/faculty/isken/courses/aap) (AAP).
-The students taking this course have already taken the PCDA course within which they learned fundamental Python programming within the context of data analytics over seven weeks.
+The students taking this course have already taken the PCDA course within where learned fundamental Python programming within the context of data analytics over seven weeks.
 The PCDA course ends with intro to the [scikit-learn library](https://scikit-learn.org/stable/index.html) for doing predictive modeling in Python and the AAP course begins where the PCDA course ends.
@@
 website](http://www.sba.oakland.edu/faculty/isken/courses/pcda/)  and [AAP course
-website](http://www.sba.oakland.edu/faculty/isken/courses/aap/), you can see the topic outline for each course and how the EwP module is positioned."
+website](http://www.sba.oakland.edu/faculty/isken/courses/aap/), you can see the topic outline for each course and how the EwP module is positioned.
 
@@
 Anyone who is interested in adapting my materials for their course is free to do so.
-Everything other than all of the homework assignments are publicly accessible from the course websites.
+Everything other than the homework assignments and quizzes are publicly accessible from the course websites.
 That's why throughout this paper I provide hyperlinks to relevant parts of the EwP course website or course content.
@@
 The same tools are used in the AAP course.
-All of the technical and logistical details of both courses are described in excruciating detail in the public course websites linked to in the paper.
+All of the technical and logistical details of both courses are described in excruciating detail in the public course websites linked to in the paper - see [Software and pcda virtual machine](http://www.sba.oakland.edu/faculty/isken/courses/pcda/pcda_vm.html) for the PCDA course and [Setting up your analytics machine](http://www.sba.oakland.edu/faculty/isken/courses/aap/setup_analytics_machine.html) for the AAP course.
 
@@
 
-"Students are reminded of the importance of version control and we initialize a git repository for our budding package.
-The basics of version control were already covered in the first week of the course."
+> Students are reminded of the importance of version control and we initialize a git repository for our budding package.
+The basics of version control were already covered in the first week of the course.
 
 Again, the AAP course website main page at [http://www.sba.oakland.edu/faculty/isken/courses/aap/](http://www.sba.oakland.edu/faculty/isken/courses/aap/) contains a detailed topic outline and the reader can see what is covered when.
+
+The reason that we dive deeper into version control and project folder structure at this point is because we are now trying to create a deployable Python package containing the work we did in the previous three notebooks.
+These things matter in software development and we want to emphasize their importance to the students.
 
@@
 
-"Despite numerous calls for, and predictions of, the eradication of spreadsheets from the business world, Excel and other spreadsheets are \href{https://benn.substack.com/p/the-next-billion-programmers}{alive and well}.
+> Despite numerous calls for, and predictions of, the eradication of spreadsheets from the business world, Excel and other spreadsheets are \href{https://benn.substack.com/p/the-next-billion-programmers}{alive and well}.
 Their flexibility is unparalleled and it is unrealistic to expect them to disappear anytime soon.
 If anything, the [recent announcement by Microsoft](https://techcommunity.microsoft.com/t5/excel-blog/announcing-python-in-excel-combining-the-power-of-python-and-the/ba-p/3893439)
- that users will be able combine Python scripts with Excel formulas within the same workbook, will likely solidify Excel's place in the analytics space."
+ that users will be able combine Python scripts with Excel formulas within the same workbook, will likely solidify Excel's place in the analytics space.
 
@@
 
-3.3. Notebook 3: Monte-Carlo simulation (html version)
+> 3.3. Notebook 3: Monte-Carlo simulation ([html version](http://www.sba.oakland.edu/faculty/isken/excel_with_python/what_if_3_simulation.html))
 
@@
 
-"While all of the notebooks are publicly available (see Section 6), links to html versions of
+> While all of the notebooks are publicly available (see Section 6), links to html versions of
 each of the notebooks are provided throughout the document and readers are encouraged to use
-them while reading to see exactly how each topic is presented."
+them while reading to see exactly how each topic is presented.
 
-My intent is for the reader to open this html version in a browser tab and then they can see the entire notebook.
+My intent is for the reader to open the html versions of the notebooks in a browser tab and then they can see the entire notebook.
 Code snippets taken out of context are not nearly as useful as following the narrative flow of the notebook - this is the real strength of Jupyter notebooks for teaching.
 
-If the reviewers really want me to add more code screenshots, I can, but I'd really prefer readers look at the (html versions of) the actual notebooks.
+If the reviewer and editor want me to add more code screenshots, I can, but I'd really prefer readers look at the (html versions of) the actual notebooks.
 Of course, the actual Jupyter notebooks are also available from the course website.
@@
 
-I've added reference to this development in the opening section as I mentioned in Response 1 above. I've also added a bit in the last section:
+I've added reference to this development in the opening section as I mentioned in Response 1 above. I've also made the last section all about the future which opens with this expanded paragraph.
 
-"Rumors have swirled for a number of years that Microsoft was considering adding Python as a first class language alternative to VBA within their MS Office suite of packages \cite{cimpanuMicrosoftConsidersAdding2017}.
+> Rumors have swirled for a number of years that Microsoft was considering adding Python as a first class language alternative to VBA within their MS Office suite of packages \cite{cimpanuMicrosoftConsidersAdding2017}.
 During the review process for this article, this development [finally
@@
 This has potentially enormous implications for the future of both Excel and Python.
-At this point, it is probably safe to say that this development further underscores the importance of business analytics students adding Python to their set of analytical toolbox."
+At this point, it is probably safe to say that this development further underscores the importance of business analytics students adding Python to their analytical toolbox.
 

### 2023-11-21 python_excel_ite_r1.tex
finished revisions and response to reviewers

@@
 Spreadsheet based modeling has transformed the teaching of management science in business schools over the past twenty plus years.
-The late 1990's saw first editions of game changing textbooks \citep{winstonPracticalManagementScience2018,ragsdaleSpreadsheetModelingDecision2017} focused on the use of spreadsheets in teaching modeling and management science.
+The late 1990s saw first editions of game changing textbooks \citep{winstonPracticalManagementScience2018,ragsdaleSpreadsheetModelingDecision2017} focused on the use of spreadsheets in teaching modeling and management science.
 A series of conferences focused on teaching management science with spreadsheets started at Dartmouth in 1998 and several presentations from that conference made up the bulk of papers in the first two issues of the INFORMS Transactions on Education (ITE) journal \citep{baker2000:lp,bell2000:statistics,evans2000:simulation,carraway2000:mba,ragsdale2001:teaching,savage2001:blitzograms}.
@@
 The PCDA course, like many introductory data science courses, focuses mostly on data wrangling, exploratory data analysis (EDA), and statistical/machine learning based predictive models.
-In teaching EDA using tools like ggplot2, students are challenged to think about how they would go about trying to use Excel to create some of the plots that are easy to do with R and Python based plotting libraries.
+In teaching EDA using tools like [ggplot2](https://ggplot2.tidyverse.org/), students are challenged to think about how they would go about trying to use Excel to create some of the plots that are easy to do with R and Python based plotting libraries.
 There is some limited coverage of using Python to open Excel files and do worksheet manipulation.
@@
 At our institution, The EwP module is part of a business course entitled  \href{http://www.sba.oakland.edu/faculty/isken/courses/aap}{Advanced Analytics with Python} (AAP).
-The students taking this course have already taken the PCDA course within which they learned fundamental Python programming within the context of data analytics over seven weeks.
+The students taking this course have already taken the PCDA course where they learned fundamental Python programming within the context of data analytics over seven weeks.
 The PCDA course ends with intro to the \href{https://scikit-learn.org/stable/index.html}{scikit-learn library} for doing predictive modeling in Python and the AAP course begins where the PCDA course ends.
@@
 \begin{itemize}
-	\item familiarity with Jupyter notebooks and an integrated development environment such as \href{https://www.spyder-ide.org/}{Spyder}, \href{https://code.visualstudio.com/}{VSCode} or \href{https://www.jetbrains.com/pycharm/}{PyCharm},
 	\item variables, arithmetic and boolean operators, basic data types,
@@
 	\item basic use of NumPy and pandas for data wrangling and analysis,
-	\item basic plotting with matplotlib.
+	\item basic plotting with matplotlib,
+	\item familiarity with Jupyter notebooks and an integrated development environment such as \href{https://www.spyder-ide.org/}{Spyder}, \href{https://code.visualstudio.com/}{VSCode} or \href{https://www.jetbrains.com/pycharm/}{PyCharm}.
 \end{itemize}
@@
 All of the Python modules in the PCDA course as well as the entire AAP course are taught using a combination of Jupyter notebooks, Python scripts, and videos that walk readers through the notebooks.
-Jupyter notebooks, part of \href{https://jupyter.org/}{Project Jupyter}, are an example of \textit{literate computing} \citep{perez2015project} and allow one to create rich interactive documents that support iterative exploration and development.
+Jupyter notebooks, part of \href{https://jupyter.org/}{Project Jupyter}, are an example of \textit{literate computing} \citep{perez2015project} and allow the creation of rich interactive documents that support iterative exploration and development.
 
@@
 \includegraphics[scale=0.5]{cells_edit_mode}
-\caption{Notebook cells in edit model}
+\caption{Notebook cells in edit mode}
 \label{fig:cells_edit_mode}
@@
 
-There are two overarching learning objectives for this \href{http://www.sba.oakland.edu/faculty/isken/courses/aap/mod3a_whatif.html}{first EwP submodule} - see Figure \ref{fig:objectives_whatif} for how the objectives are presented to students via the course website.
+There are two overarching learning objectives for this \href{http://www.sba.oakland.edu/faculty/isken/courses/aap/mod3a_whatif.html}{first EwP submodule}.
 Students learn how Python might be used to do something they are quite familiar with doing in a spreadsheet -- model building and sensitivity analysis.
@@
 In this way, the computational problem to be solved drives the introduction of more advanced Python concepts and techniques instead of such things being presented in a vacuum.
+See Figure \ref{fig:objectives_whatif} for how the objectives are presented to students via the course website.
+
 I make it a point to show that there are often multiple possible approaches to a computational problem by being willing to abandon one approach and trying another.
@@
 This provides a great opportunity to remind students of one of the great strengths of the open source software ecosystem -- we can leverage work done in other packages and even look at actual source code to see how something was implemented.
-In this case, the well known \href{https://scikit-learn.org/stable/}{scikit-learn} package (\cite{scikit-learn}) has a \code{ParameterGrid} function that solves our scenario generation problem quite nicely.
+In this case, the well known \href{https://scikit-learn.org/stable/}{scikit-learn} package \citep{scikit-learn} has a \code{ParameterGrid} function that solves our scenario generation problem quite nicely.
 This example also reminds the students of the value of being able to make sense of API documentation and to understand object oriented code written by others.
@@
 Students are now ready to learn about the basics of turning our code into an installable package.
-This requires a basic understanding of two key Python files that are ubiquitous in Python pacakges - \code{\_\_init\_\_.py} and \code{setup.py}.
 Options are presented for installing our package locally so that we can use it for different modeling projects.
@@
 The deployment focused submodule ends with a short notebook discussing the different types of documentation needed in a typical project.
-This includes code comments, \href{https://peps.python.org/pep-0257/}{docstrings}, \code{readme} files and generating documentation by creating \href{https://www.sphinx-doc.org/en/master/usage/restructuredtext/basics.html}{reStructuredText} files (similar to but more powerful than markdown) and using \href{https://www.sphinx-doc.org/en/master/index.html}{Sphinx}, the Python documentation generation tool.
+This includes code comments, \href{https://peps.python.org/pep-0257/}{docstrings}, \code{readme} files and generating documentation by creating \href{https://www.sphinx-doc.org/en/master/usage/restructuredtext/basics.html}{reStructuredText} files (similar to, but more powerful than markdown) and using \href{https://www.sphinx-doc.org/en/master/index.html}{Sphinx}, the Python documentation generation tool.
 This is one of those tasks that few people like to do but can be very important for the long term success of a project.
@@
 While this first example is not necessarily Excel related, using Excel to consolidate csv files through a series of manual Copy-Paste operations is something that many Excel users will admit to having done.
-After getting familiar with pathlib, a library for working with file systems which has recently been added to Python's standard library, students are led through the development of a short procedure that creates a list of the csv files to concatenate and then uses pandas to read and combine the files in a \code{DataFrame} and writes out the consolidated data as a csv file.
+After getting familiar with \code{pathlib}, a library for working with file systems which has recently been added to Python's standard library, students are led through the development of a short procedure that creates a list of the csv files to concatenate and then uses pandas to read and combine the files in a \code{DataFrame} and writes out the consolidated data as a csv file.
 
@@
 
-\section{Software availability}
+\section{Software availability and classroom use}
 \label{sec:software}
@@
 It contains the original notebooks used in the blog posts that motivated this project.
-
-\section{Classroom experience and the future}
-\label{sec:classroom}
 
@@
 
+\section{The future of Python and Excel}
+\label{sec:future}
+
 Rumors have swirled for a number of years that Microsoft was considering adding Python as a first class language alternative to VBA within their MS Office suite of packages \cite{cimpanuMicrosoftConsidersAdding2017}.
@@
 This has potentially enormous implications for the future of both Excel and Python.
-At this point, it is probably safe to say that this development further underscores the importance of business analytics students adding Python to their set of analytical toolbox.
+At this point, it is probably safe to say that this development further underscores the importance of business analytics students adding Python to their analytical toolbox.
+
 There continues to be significant interest in using Python to manipulate not only Excel workbooks but other office documents such as Powerpoint presentations \citep{cannyPythonpptx2013}.

### 2023-11-21 response_to_r1.md
fixed links

@@
 
-As for assessment, in the final section of the paper, *Classroom experience and the future*, we described and made available for download an assignment that we have used for assessment of this material.
+As for assessment, in the section of the paper entitled, *Software availability and classroom use*, we describe and make available for download an assignment that we have used for assessment of this material.
 It covers all aspects of the EwP module.
@@
 
-> Despite numerous calls for, and predictions of, the eradication of spreadsheets from the business world, Excel and other spreadsheets are \href{https://benn.substack.com/p/the-next-billion-programmers}{alive and well}.
+> Despite numerous calls for, and predictions of, the eradication of spreadsheets from the business world, Excel and other spreadsheets are [alive and well](https://benn.substack.com/p/the-next-billion-programmers).
 Their flexibility is unparalleled and it is unrealistic to expect them to disappear anytime soon.

### 2023-11-21 python_excel_ite_r1.tex
revision 1 submitted

@@
 The PCDA course, like many introductory data science courses, focuses mostly on data wrangling, exploratory data analysis (EDA), and statistical/machine learning based predictive models.
-In teaching EDA using tools like [ggplot2](https://ggplot2.tidyverse.org/), students are challenged to think about how they would go about trying to use Excel to create some of the plots that are easy to do with R and Python based plotting libraries.
+In teaching EDA using tools like \href{https://ggplot2.tidyverse.org/}{ggplot2}, students are challenged to think about how they would go about trying to use Excel to create some of the plots that are easy to do with R and Python based plotting libraries.
 There is some limited coverage of using Python to open Excel files and do worksheet manipulation.

### 2023-11-21 response_to_r1.qmd
cleaning up

--- 2023-11-21 response_to_r1.md
@@
-# Response to reviewers of "Spreadsheet modeling and wrangling with Python"
+---
+title: Response to reviewers of "Spreadsheet modeling and wrangling with Python"
+date: 2023-11-21
+categories: [python, excel]
+citation: false
+output: html_document
+format:
+  html:
+    embed-resources: true
+---
 
@@
 My first request to the author is to include other related ITE publications either as part of the literature review or to the other relevant sections.  Here is a short list of links to start:
-•       https://pubsonline.informs.org/doi/pdf/10.1287/ited.2021.0261
-•       https://pubsonline.informs.org/doi/full/10.1287/ited.2023.0285
-•       https://pubsonline.informs.org/doi/pdf/10.1287/ited.2022.0275
+
+- https://pubsonline.informs.org/doi/pdf/10.1287/ited.2021.0261
+- https://pubsonline.informs.org/doi/full/10.1287/ited.2023.0285
+- https://pubsonline.informs.org/doi/pdf/10.1287/ited.2022.0275
 

### 2024-02-23 python_excel_ite_final.tex
final revision

--- 2023-11-21 python_excel_ite_r1.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\documentclass[ited]{informs3}                      % for a regular run
+\documentclass[ited]{informs3}                      % for a regular run
 %\documentclass[ited,nonblindrev]{informs3}          % for review, not blinded
 %\documentclass[ited,blindrev,copyedit]{informs3}    % spaced for copyediting
-\documentclass[ited,blindrev]{informs3}              % for review, blinded
+%\documentclass[ited,blindrev]{informs3}              % for review, blinded
 
@@
 \usepackage{graphicx}
-%\usepackage{minted}
+\usepackage{caption}
+\newenvironment{codelisting}{\captionsetup{type=listing}}{}
+
+\usepackage{minted}
 %\tcbuselibrary{minted} % tcolorbox minted library, required to use the "minted" tcb listing engine (this library is not loaded by the option [most])
+
+%\usepackage{listings}
+%\usepackage{color}
+%
+%\definecolor{dkgreen}{rgb}{0,0.6,0}
+%\definecolor{gray}{rgb}{0.5,0.5,0.5}
+%\definecolor{mauve}{rgb}{0.58,0,0.82}
+%
+%\lstset{frame=tb,
+%	language=Python,
+%	aboveskip=3mm,
+%	belowskip=3mm,
+%	showstringspaces=false,
+%	columns=flexible,
+%	basicstyle={\small\ttfamily},
+%	numbers=none,
+%	numberstyle=\tiny\color{gray},
+%	keywordstyle=\color{blue},
+%	commentstyle=\color{dkgreen},
+%	stringstyle=\color{mauve},
+%	breaklines=true,
+%	breakatwhitespace=true,
+%	tabsize=3
+%}
+
 \usepackage[colorlinks]{hyperref} % ALWAYS load this package LAST
@@
 In their courses, Jupyter notebooks (discussed in more detail in the next section) play a prominent role in providing an interactive development and learning environment for Python within an operations research context.
+In \citet{pythonprojects:ite}, an analytics project based course is described in which Python plays a significant role.
+Students use Python for a number of data wrangling, exploratory data analysis and modeling tasks.
+Through a course like this, students gain valuable experience with a tool that is becoming more of an expectation in the analytics world.
+Another effective and fun way that students can gain experience with Python is through using it solve puzzles such as Wordle \citep{wordle_2023} or the Advent of Code \citep{AOC}.
+
 In 2015, at Oakland University, we created a course called \href{http://www.sba.oakland.edu/faculty/isken/courses/pcda/}{Practical Computing for Data Analytics} (PCDA) that introduced both R and Python (along with a little Linux) for use in data analytics work.
 This new course was meant as a follow up to the \href{http://www.sba.oakland.edu/faculty/isken/courses/ba/}{spreadsheet based modeling course} we have had in place since 2001.
-
 The PCDA course, like many introductory data science courses, focuses mostly on data wrangling, exploratory data analysis (EDA), and statistical/machine learning based predictive models.
@@
 \centering
-\includegraphics[scale=0.5]{cells_edit_mode}
+\includegraphics[scale=0.5]{images/cells_edit_mode}
 \caption{Notebook cells in edit mode}
@@
 \centering
-\includegraphics[scale=0.5]{cells_after_running}
+\includegraphics[scale=0.5]{images/cells_after_running}
 \caption{Notebook cells after execution}
@@
 \begin{figure}[!htbp]
-	\includegraphics[width=1.0\textwidth]{objectives_whatif}
+	\includegraphics[width=1.0\textwidth]{images/objectives_whatif}
 	\caption{Learning objectives for Submodule 1}
@@
 \begin{figure}[!htbp]
-\includegraphics[width=1.0\textwidth]{output_2_0}
+\includegraphics[width=1.0\textwidth]{images/output_2_0}
 \caption{Typical spreadsheet model}
@@
 \centering
-\includegraphics[scale=0.5]{model_inputs}
+\includegraphics[scale=0.5]{images/model_inputs}
 \caption{Instantiating model inputs and creating output formulas}
@@
 \centering
-\includegraphics[scale=0.5]{function_datatable}
+\includegraphics[scale=0.5]{images/function_datatable}
 \caption{A 2-way data table}
@@
 \centering
-\includegraphics[scale=0.5]{datatable_scatter}
+\includegraphics[scale=0.5]{images/datatable_scatter}
 \caption{Scatter plot based on 2-way data table}
@@
 
-For our initial design of a \code{BookstoreModel} class we make all of the base inputs class attributes (properties) and then add method attributes to compute outputs such as costs, revenues and profits.
+For our initial design of a \code{BookstoreModel} class we make all of the base inputs class attributes (properties) and then add method attributes to compute outputs such as costs, revenues and profits - see Figure \ref{fig:oo_model}.
 Students can easily see the mapping between elements in the initial procedural model and this new object oriented version.
@@
 This section of the notebook ends with using the \code{BookstoreModel} class to create a new model object instance with all of its base inputs instantiated with values.
-Before moving on, students are presented with a challenges involving enhancements to the \code{BookstoreModel} class.
+Before moving on, students are presented with a challenge involving enhancements to the \code{BookstoreModel} class.
 Interspersing short coding challenges throughout the notebook gives students a chance to test themselves on their understanding and to get some coding practice.
+
+\begin{figure}[!htbp]
+	\centering
+	\includegraphics[scale=0.5]{images/oo_model}
+	\caption{Object oriented version of the model}
+	\label{fig:oo_model}
+\end{figure}
 
@@
 
-This first notebook draws to a close with the creation of a \code{data\_table} function that allows us to do sensitivity analysis with the objected oriented model.
-The payoff is the creation of the following faceted plot showing how profit varies for different order quantities and demand levels.
+This first notebook draws to a close with the creation of a \code{data\_table} function (Figure \ref{fig:oo_datatable_function}) that allows us to do sensitivity analysis with the objected oriented model.
+The payoff is the creation of the faceted plot in Figure \ref{fig:2way_dt} showing how profit varies for different order quantities and demand levels.
 We recap the main things learned in this notebook and prepare to add goal seeking capability to our model.
@@
 	\centering
-\includegraphics[width=1.0\textwidth]{two_way_dt}
+	\includegraphics[scale=0.5]{images/oo_datatable_function}
+	\caption{An n-way DataTable function}
+	\label{fig:oo_datatable_function}
+\end{figure}
+
+\begin{figure}[!htbp]
+\centering
+\includegraphics[width=1.0\textwidth]{images/two_way_dt}
 \caption{Faceted plot based on 2-way data table}
@@
 A \code{goal\_seek} function is created that implements a simple bisection search and we use it to find the break even demand point in our bookstore model.
+See Listing \ref{fig:goal_seek_function}.
 With \code{data\_table} and \code{goal\_seek} functions implemented, we are ready to move on to doing Monte-Carlo simulation.
+
+\begin{listing}
+\centering
+\begin{minted}[fontsize=\tiny]{python}
+def goal_seek(model, obj_fn, target, by_changing, a, b, N=100, verbose=False):
+    '''Approximate solution of f(x)=0 on interval [a,b] by bisection method.
+
+    Parameters
+    ----------
+    model : object
+        User defined object containing the appropriate methods and properties for doing the desired goal seek
+    obj_fn : function
+        The function for which we are trying to approximate a solution f(x)=target.
+    target : float
+        The goal
+    by_changing : string
+        Name of the input variable in model
+    a,b : numbers
+        The interval in which to search for a solution. The function returns
+        None if (f(a) - target) * (f(b) - target) >= 0 since a solution is not guaranteed.
+    N : (positive) integer
+        The number of iterations to implement.
+    verbose : boolean (default=False)
+        If True, root finding progress is reported
+
+    Returns
+    -------
+    x_N : number
+        The midpoint of the Nth interval computed by the bisection method. 
+        If all signs of values f(a_n), f(b_n) and f(m_n) are the same at any
+        iteration, the bisection method fails and return None.
+    '''  
+    # Clone the model
+    model_clone = copy.deepcopy(model)
+
+    # The following bisection search is a direct adaptation of
+    # https://www.math.ubc.ca/~pwalls/math-python/roots-optimization/bisection/  
+    setattr(model_clone, by_changing, a)
+    f_a_0 = getattr(model_clone, obj_fn)()
+    setattr(model_clone, by_changing, b)
+    f_b_0 = getattr(model_clone, obj_fn)()
+
+    if (f_a_0 - target) * (f_b_0 - target) >= 0:
+        # print("Bisection method fails.")
+        return None
+
+    # Initialize the end points
+    a_n = a
+    b_n = b
+    for n in range(1, N+1):
+        # Compute the midpoint
+        m_n = (a_n + b_n)/2
+
+        # Function value at midpoint
+        setattr(model_clone, by_changing, m_n)
+        f_m_n = getattr(model_clone, obj_fn)()
+
+        # Function value at a_n
+        setattr(model_clone, by_changing, a_n)
+        f_a_n = getattr(model_clone, obj_fn)()
+
+        # Function value at b_n
+        setattr(model_clone, by_changing, b_n)
+        f_b_n = getattr(model_clone, obj_fn)()
+
+        if verbose:
+            print(f"n = {n}, a_n = {a_n}, b_n = {b_n}, m_n = {m_n}, width = {b_n - a_n}")
+
+        # Figure out which half the root is in, or if we hit it exactly, or if the search failed
+        if (f_a_n - target) * (f_m_n - target) < 0:
+            a_n = a_n
+            b_n = m_n
+        elif (f_b_n - target) * (f_m_n - target) < 0:
+            a_n = m_n
+            b_n = b_n
+        elif f_m_n == target:
+            return m_n
+        else:
+            return None
+
+    # If we get here we hit iteration limit, return best solution found so far
+    return (a_n + b_n)/2
+\end{minted}
+\caption{\code{goal\_seek} function}
+\label{fig:goal_seek_function}
+\end{listing}
 
@@
 \centering
-\includegraphics[scale=0.5]{oo_sim_design_specs}
+\includegraphics[scale=0.5]{images/oo_sim_design_specs}
 \caption{Design specification for \code{simulate()} function}
@@
 By this point, students are well equipped to understand the code as it builds on basic Python concepts and more advanced concepts that we have already covered in the EwP module.
+See Listing \ref{fig:simulate_function}.
+
+\begin{listing}
+\centering
+\begin{minted}[fontsize=\tiny]{python}
+def simulate(model, random_inputs, outputs, scenario_inputs=None, keep_random_inputs=False):
+    '''Simulate model for one or more scenarios
+
+    Parameters
+    ----------
+    model : object
+        User defined object containing the appropriate methods and properties for computing outputs from inputs
+    random_intputs : dict of str to sequence of random variates
+        Keys are stochastic input variable names and values are sequence of $n$ random variates, where $n$ is the number of simulation replications
+    outputs : list of str
+        List of output variable names
+    scenario_inputs : optional (default is None), dict of str to sequence
+        Keys are deterministic input variable names and values are sequence of values for each scenario for this variable. Is consumed by
+        scikit-learn ParameterGrid() function. See https://scikit-learn.org/stable/modules/generated/sklearn.model_selection.ParameterGrid.html
+    keep_random_inputs : optional (default is False), boolean
+        If True, all the random input variates are included in the results dataframe
+
+    Returns
+    -------
+    results_df : list of dict
+        Values of all outputs for each simulation replication. If `scenario_inputs` is not None, then this is also for every combination of scenario inputs
+    '''
+
+    # Clone the model
+    model_clone = copy.deepcopy(model)
+
+    # Update clone with random_inputs
+    model_clone.update(random_inputs)
+
+    # Store raw simulation input values if desired
+    if keep_random_inputs:
+        scenario_base_vals = vars(model_clone)
+    else:
+        scenario_base_vals = vars(model)
+
+    # Initialize output counters and containers
+    scenario_num = 0
+    scenario_results = []
+
+    # Check if multiple scenarios
+    if scenario_inputs is not None:
+        # Create parameter grid for scenario inputs
+        sim_param_grid = list(ParameterGrid(scenario_inputs))
+
+        # Scenario loop
+        for params in sim_param_grid:
+            model_clone.update(params)
+            # Initialize scenario related outputs
+            result = {}
+            scenario_vals = copy.copy(params)
+            result['scenario_base_vals'] = scenario_base_vals
+            result['scenario_num'] = scenario_num
+            result['scenario_vals'] = scenario_vals
+            raw_output = {}
+
+            # Output measure loop
+            for output_name in outputs:
+                output_array = getattr(model_clone, output_name)()
+                raw_output[output_name] = output_array
+
+            # Gather results for this scenario
+            result['output'] = raw_output
+            scenario_results.append(result)
+            scenario_num += 1
+
+        return scenario_results
+
+    else:
+        # Similar logic to above, but only a single scenario
+        results = []
+        result = {}
+
+        result['scenario_base_vals'] = scenario_base_vals
+        result['scenario_num'] = scenario_num
+        result['scenario_vals'] = {}
+
+        raw_output = {}
+        for output_name in outputs:
+            output_array = getattr(model_clone, output_name)()
+            raw_output[output_name] = output_array
+
+        result['output'] = raw_output  
+        results.append(result)
+
+        return results
+\end{minted}
+\caption{\code{simulate} function}
+\label{fig:simulate_function}
+\end{listing}
+
 The rest of the notebook is spent trying out the \code{simulate()} function and ends with creating a grouped boxplot, see Figure \ref{fig:boxplot}, along with a faceted plot of profit histograms created with the popular \href{https://seaborn.pydata.org/}{Seaborn} package.
@@
 \centering
-\includegraphics[width=1.0\textwidth]{boxplots}
+\includegraphics[width=1.0\textwidth]{images/boxplots}
 \caption{Profit boxplots by order quantity}
@@
 \centering
-\includegraphics[scale=0.5]{sim_wrapup}
+\includegraphics[scale=0.5]{images/sim_wrapup}
 \caption{Closing thoughts in simulation notebook}
@@
 	\centering
-	\includegraphics[scale=0.5]{objectives_packaging}
+	\includegraphics[scale=0.5]{images/objectives_packaging}
 	\caption{Learning objectives for Submodule 2}
@@
 % Acknowledgments here
-\ACKNOWLEDGMENT{%
+%\ACKNOWLEDGMENT{%
 % Enter the text of acknowledgments here
-}% Leave this (end of acknowledgment)
+%}% Leave this (end of acknowledgment)
 

### 2024-02-23 what_if_1_model_datatable.tex
final revision

\documentclass[11pt]{article}

    \usepackage[breakable]{tcolorbox}
    \usepackage{parskip} % Stop auto-indenting (to mimic markdown behaviour)

    % Basic figure setup, for now with no caption control since it's done
    % automatically by Pandoc (which extracts ![](path) syntax from Markdown).
    \usepackage{graphicx}
    % Maintain compatibility with old templates. Remove in nbconvert 6.0
    \let\Oldincludegraphics\includegraphics
    % Ensure that by default, figures have no caption (until we provide a
    % proper Figure object with a Caption API and a way to capture that
    % in the conversion process - todo).
    \usepackage{caption}
    \DeclareCaptionFormat{nocaption}{}
    \captionsetup{format=nocaption,aboveskip=0pt,belowskip=0pt}

    \usepackage{float}
    \floatplacement{figure}{H} % forces figures to be placed at the correct location
    \usepackage{xcolor} % Allow colors to be defined
    \usepackage{enumerate} % Needed for markdown enumerations to work
    \usepackage{geometry} % Used to adjust the document margins
    \usepackage{amsmath} % Equations
    \usepackage{amssymb} % Equations
    \usepackage{textcomp} % defines textquotesingle
    % Hack from http://tex.stackexchange.com/a/47451/13684:
    \AtBeginDocument{%
        \def\PYZsq{\textquotesingle}% Upright quotes in Pygmentized code
    }
    \usepackage{upquote} % Upright quotes for verbatim code
    \usepackage{eurosym} % defines \euro

    \usepackage{iftex}
    \ifPDFTeX
        \usepackage[T1]{fontenc}
        \IfFileExists{alphabeta.sty}{
              \usepackage{alphabeta}
          }{
              \usepackage[mathletters]{ucs}
              \usepackage[utf8x]{inputenc}
          }
    \else
        \usepackage{fontspec}
        \usepackage{unicode-math}
    \fi

    \usepackage{fancyvrb} % verbatim replacement that allows latex
    \usepackage{grffile} % extends the file name processing of package graphics
                         % to support a larger range
    \makeatletter % fix for old versions of grffile with XeLaTeX
    \@ifpackagelater{grffile}{2019/11/01}
    {
      % Do nothing on new versions
    }
    {
      \def\Gread@@xetex#1{%
        \IfFileExists{"\Gin@base".bb}%
        {\Gread@eps{\Gin@base.bb}}%
        {\Gread@@xetex@aux#1}%
      }
    }
    \makeatother
    \usepackage[Export]{adjustbox} % Used to constrain images to a maximum size
    \adjustboxset{max size={0.9\linewidth}{0.9\paperheight}}

    % The hyperref package gives us a pdf with properly built
    % internal navigation ('pdf bookmarks' for the table of contents,
    % internal cross-reference links, web links for URLs, etc.)
    \usepackage{hyperref}
    % The default LaTeX title has an obnoxious amount of whitespace. By default,
    % titling removes some of it. It also provides customization options.
    \usepackage{titling}
    \usepackage{longtable} % longtable support required by pandoc >1.10
    \usepackage{booktabs}  % table support for pandoc > 1.12.2
    \usepackage{array}     % table support for pandoc >= 2.11.3
    \usepackage{calc}      % table minipage width calculation for pandoc >= 2.11.1
    \usepackage[inline]{enumitem} % IRkernel/repr support (it uses the enumerate* environment)
    \usepackage[normalem]{ulem} % ulem is needed to support strikethroughs (\sout)
                                % normalem makes italics be italics, not underlines
    \usepackage{mathrsfs}

    % Colors for the hyperref package
    \definecolor{urlcolor}{rgb}{0,.145,.698}
    \definecolor{linkcolor}{rgb}{.71,0.21,0.01}
    \definecolor{citecolor}{rgb}{.12,.54,.11}

    % ANSI colors
    \definecolor{ansi-black}{HTML}{3E424D}
    \definecolor{ansi-black-intense}{HTML}{282C36}
    \definecolor{ansi-red}{HTML}{E75C58}
    \definecolor{ansi-red-intense}{HTML}{B22B31}
    \definecolor{ansi-green}{HTML}{00A250}
    \definecolor{ansi-green-intense}{HTML}{007427}
    \definecolor{ansi-yellow}{HTML}{DDB62B}
    \definecolor{ansi-yellow-intense}{HTML}{B27D12}
    \definecolor{ansi-blue}{HTML}{208FFB}
    \definecolor{ansi-blue-intense}{HTML}{0065CA}
    \definecolor{ansi-magenta}{HTML}{D160C4}
    \definecolor{ansi-magenta-intense}{HTML}{A03196}
    \definecolor{ansi-cyan}{HTML}{60C6C8}
    \definecolor{ansi-cyan-intense}{HTML}{258F8F}
    \definecolor{ansi-white}{HTML}{C5C1B4}
    \definecolor{ansi-white-intense}{HTML}{A1A6B2}
    \definecolor{ansi-default-inverse-fg}{HTML}{FFFFFF}
    \definecolor{ansi-default-inverse-bg}{HTML}{000000}

    % common color for the border for error outputs.
    \definecolor{outerrorbackground}{HTML}{FFDFDF}

    % commands and environments needed by pandoc snippets
    % extracted from the output of `pandoc -s`
    \providecommand{\tightlist}{%
      \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
    \DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
    % Add ',fontsize=\small' for more characters per line
    \newenvironment{Shaded}{}{}
    \newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{{#1}}}}
    \newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{{#1}}}
    \newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
    \newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
    \newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
    \newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{{#1}}}}
    \newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{{#1}}}
    \newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
    \newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{{#1}}}
    \newcommand{\RegionMarkerTok}[1]{{#1}}
    \newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
    \newcommand{\NormalTok}[1]{{#1}}

    % Additional commands for more recent versions of Pandoc
    \newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.53,0.00,0.00}{{#1}}}
    \newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.73,0.40,0.53}{{#1}}}
    \newcommand{\ImportTok}[1]{{#1}}
    \newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.73,0.13,0.13}{\textit{{#1}}}}
    \newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}
    \newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}
    \newcommand{\VariableTok}[1]{\textcolor[rgb]{0.10,0.09,0.49}{{#1}}}
    \newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{{#1}}}}
    \newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.40,0.40,0.40}{{#1}}}
    \newcommand{\BuiltInTok}[1]{{#1}}
    \newcommand{\ExtensionTok}[1]{{#1}}
    \newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.74,0.48,0.00}{{#1}}}
    \newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.49,0.56,0.16}{{#1}}}
    \newcommand{\InformationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}
    \newcommand{\WarningTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}

    % Define a nice break command that doesn't care if a line doesn't already
    % exist.
    \def\br{\hspace*{\fill} \\* }
    % Math Jax compatibility definitions
    \def\gt{>}
    \def\lt{<}
    \let\Oldtex\TeX
    \let\Oldlatex\LaTeX
    \renewcommand{\TeX}{\textrm{\Oldtex}}
    \renewcommand{\LaTeX}{\textrm{\Oldlatex}}
    % Document parameters
    % Document title
    \title{what\_if\_1\_model\_datatable}

% Pygments definitions
\makeatletter
\def\PY@reset{\let\PY@it=\relax \let\PY@bf=\relax%
    \let\PY@ul=\relax \let\PY@tc=\relax%
    \let\PY@bc=\relax \let\PY@ff=\relax}
\def\PY@tok#1{\csname PY@tok@#1\endcsname}
\def\PY@toks#1+{\ifx\relax#1\empty\else%
    \PY@tok{#1}\expandafter\PY@toks\fi}
\def\PY@do#1{\PY@bc{\PY@tc{\PY@ul{%
    \PY@it{\PY@bf{\PY@ff{#1}}}}}}}
\def\PY#1#2{\PY@reset\PY@toks#1+\relax+\PY@do{#2}}

\@namedef{PY@tok@w}{\def\PY@tc##1{\textcolor[rgb]{0.73,0.73,0.73}{##1}}}
\@namedef{PY@tok@c}{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.24,0.48,0.48}{##1}}}
\@namedef{PY@tok@cp}{\def\PY@tc##1{\textcolor[rgb]{0.61,0.40,0.00}{##1}}}
\@namedef{PY@tok@k}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\@namedef{PY@tok@kp}{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\@namedef{PY@tok@kt}{\def\PY@tc##1{\textcolor[rgb]{0.69,0.00,0.25}{##1}}}
\@namedef{PY@tok@o}{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\@namedef{PY@tok@ow}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
\@namedef{PY@tok@nb}{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\@namedef{PY@tok@nf}{\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\@namedef{PY@tok@nc}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\@namedef{PY@tok@nn}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\@namedef{PY@tok@ne}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.80,0.25,0.22}{##1}}}
\@namedef{PY@tok@nv}{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\@namedef{PY@tok@no}{\def\PY@tc##1{\textcolor[rgb]{0.53,0.00,0.00}{##1}}}
\@namedef{PY@tok@nl}{\def\PY@tc##1{\textcolor[rgb]{0.46,0.46,0.00}{##1}}}
\@namedef{PY@tok@ni}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.44,0.44,0.44}{##1}}}
\@namedef{PY@tok@na}{\def\PY@tc##1{\textcolor[rgb]{0.41,0.47,0.13}{##1}}}
\@namedef{PY@tok@nt}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\@namedef{PY@tok@nd}{\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
\@namedef{PY@tok@s}{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\@namedef{PY@tok@sd}{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\@namedef{PY@tok@si}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.64,0.35,0.47}{##1}}}
\@namedef{PY@tok@se}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.67,0.36,0.12}{##1}}}
\@namedef{PY@tok@sr}{\def\PY@tc##1{\textcolor[rgb]{0.64,0.35,0.47}{##1}}}
\@namedef{PY@tok@ss}{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\@namedef{PY@tok@sx}{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\@namedef{PY@tok@m}{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\@namedef{PY@tok@gh}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
\@namedef{PY@tok@gu}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.50,0.00,0.50}{##1}}}
\@namedef{PY@tok@gd}{\def\PY@tc##1{\textcolor[rgb]{0.63,0.00,0.00}{##1}}}
\@namedef{PY@tok@gi}{\def\PY@tc##1{\textcolor[rgb]{0.00,0.52,0.00}{##1}}}
\@namedef{PY@tok@gr}{\def\PY@tc##1{\textcolor[rgb]{0.89,0.00,0.00}{##1}}}
\@namedef{PY@tok@ge}{\let\PY@it=\textit}
\@namedef{PY@tok@gs}{\let\PY@bf=\textbf}
\@namedef{PY@tok@gp}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
\@namedef{PY@tok@go}{\def\PY@tc##1{\textcolor[rgb]{0.44,0.44,0.44}{##1}}}
\@namedef{PY@tok@gt}{\def\PY@tc##1{\textcolor[rgb]{0.00,0.27,0.87}{##1}}}
\@namedef{PY@tok@err}{\def\PY@bc##1{{\setlength{\fboxsep}{\string -\fboxrule}\fcolorbox[rgb]{1.00,0.00,0.00}{1,1,1}{\strut ##1}}}}
\@namedef{PY@tok@kc}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\@namedef{PY@tok@kd}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\@namedef{PY@tok@kn}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\@namedef{PY@tok@kr}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\@namedef{PY@tok@bp}{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\@namedef{PY@tok@fm}{\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\@namedef{PY@tok@vc}{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\@namedef{PY@tok@vg}{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\@namedef{PY@tok@vi}{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\@namedef{PY@tok@vm}{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\@namedef{PY@tok@sa}{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\@namedef{PY@tok@sb}{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\@namedef{PY@tok@sc}{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\@namedef{PY@tok@dl}{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\@namedef{PY@tok@s2}{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\@namedef{PY@tok@sh}{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\@namedef{PY@tok@s1}{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\@namedef{PY@tok@mb}{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\@namedef{PY@tok@mf}{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\@namedef{PY@tok@mh}{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\@namedef{PY@tok@mi}{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\@namedef{PY@tok@il}{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\@namedef{PY@tok@mo}{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\@namedef{PY@tok@ch}{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.24,0.48,0.48}{##1}}}
\@namedef{PY@tok@cm}{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.24,0.48,0.48}{##1}}}
\@namedef{PY@tok@cpf}{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.24,0.48,0.48}{##1}}}
\@namedef{PY@tok@c1}{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.24,0.48,0.48}{##1}}}
\@namedef{PY@tok@cs}{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.24,0.48,0.48}{##1}}}

\def\PYZbs{\char`\\}
\def\PYZus{\char`\_}
\def\PYZob{\char`\{}
\def\PYZcb{\char`\}}
\def\PYZca{\char`\^}
\def\PYZam{\char`\&}
\def\PYZlt{\char`\<}
\def\PYZgt{\char`\>}
\def\PYZsh{\char`\#}
\def\PYZpc{\char`\%}
\def\PYZdl{\char`\$}
\def\PYZhy{\char`\-}
\def\PYZsq{\char`\'}
\def\PYZdq{\char`\"}
\def\PYZti{\char`\~}
% for compatibility with earlier versions
\def\PYZat{@}
\def\PYZlb{[}
\def\PYZrb{]}
\makeatother

    % For linebreaks inside Verbatim environment from package fancyvrb.
    \makeatletter
        \newbox\Wrappedcontinuationbox
        \newbox\Wrappedvisiblespacebox
        \newcommand*\Wrappedvisiblespace {\textcolor{red}{\textvisiblespace}}
        \newcommand*\Wrappedcontinuationsymbol {\textcolor{red}{\llap{\tiny$\m@th\hookrightarrow$}}}
        \newcommand*\Wrappedcontinuationindent {3ex }
        \newcommand*\Wrappedafterbreak {\kern\Wrappedcontinuationindent\copy\Wrappedcontinuationbox}
        % Take advantage of the already applied Pygments mark-up to insert
        % potential linebreaks for TeX processing.
        %        {, <, #, %, $, ' and ": go to next line.
        %        _, }, ^, &, >, - and ~: stay at end of broken line.
        % Use of \textquotesingle for straight quote.
        \newcommand*\Wrappedbreaksatspecials {%
            \def\PYGZus{\discretionary{\char`\_}{\Wrappedafterbreak}{\char`\_}}%
            \def\PYGZob{\discretionary{}{\Wrappedafterbreak\char`\{}{\char`\{}}%
            \def\PYGZcb{\discretionary{\char`\}}{\Wrappedafterbreak}{\char`\}}}%
            \def\PYGZca{\discretionary{\char`\^}{\Wrappedafterbreak}{\char`\^}}%
            \def\PYGZam{\discretionary{\char`\&}{\Wrappedafterbreak}{\char`\&}}%
            \def\PYGZlt{\discretionary{}{\Wrappedafterbreak\char`\<}{\char`\<}}%
            \def\PYGZgt{\discretionary{\char`\>}{\Wrappedafterbreak}{\char`\>}}%
            \def\PYGZsh{\discretionary{}{\Wrappedafterbreak\char`\#}{\char`\#}}%
            \def\PYGZpc{\discretionary{}{\Wrappedafterbreak\char`\%}{\char`\%}}%
            \def\PYGZdl{\discretionary{}{\Wrappedafterbreak\char`\$}{\char`\$}}%
            \def\PYGZhy{\discretionary{\char`\-}{\Wrappedafterbreak}{\char`\-}}%
            \def\PYGZsq{\discretionary{}{\Wrappedafterbreak\textquotesingle}{\textquotesingle}}%
            \def\PYGZdq{\discretionary{}{\Wrappedafterbreak\char`\"}{\char`\"}}%
            \def\PYGZti{\discretionary{\char`\~}{\Wrappedafterbreak}{\char`\~}}%
        }
        % Some characters . , ; ? ! / are not pygmentized.
        % This macro makes them "active" and they will insert potential linebreaks
        \newcommand*\Wrappedbreaksatpunct {%
            \lccode`\~`\.\lowercase{\def~}{\discretionary{\hbox{\char`\.}}{\Wrappedafterbreak}{\hbox{\char`\.}}}%
            \lccode`\~`\,\lowercase{\def~}{\discretionary{\hbox{\char`\,}}{\Wrappedafterbreak}{\hbox{\char`\,}}}%
            \lccode`\~`\;\lowercase{\def~}{\discretionary{\hbox{\char`\;}}{\Wrappedafterbreak}{\hbox{\char`\;}}}%
            \lccode`\~`\:\lowercase{\def~}{\discretionary{\hbox{\char`\:}}{\Wrappedafterbreak}{\hbox{\char`\:}}}%
            \lccode`\~`\?\lowercase{\def~}{\discretionary{\hbox{\char`\?}}{\Wrappedafterbreak}{\hbox{\char`\?}}}%
            \lccode`\~`\!\lowercase{\def~}{\discretionary{\hbox{\char`\!}}{\Wrappedafterbreak}{\hbox{\char`\!}}}%
            \lccode`\~`\/\lowercase{\def~}{\discretionary{\hbox{\char`\/}}{\Wrappedafterbreak}{\hbox{\char`\/}}}%
            \catcode`\.\active
            \catcode`\,\active
            \catcode`\;\active
            \catcode`\:\active
            \catcode`\?\active
            \catcode`\!\active
            \catcode`\/\active
            \lccode`\~`\~
        }
    \makeatother

    \let\OriginalVerbatim=\Verbatim
    \makeatletter
    \renewcommand{\Verbatim}[1][1]{%
        %\parskip\z@skip
        \sbox\Wrappedcontinuationbox {\Wrappedcontinuationsymbol}%
        \sbox\Wrappedvisiblespacebox {\FV@SetupFont\Wrappedvisiblespace}%
        \def\FancyVerbFormatLine ##1{\hsize\linewidth
            \vtop{\raggedright\hyphenpenalty\z@\exhyphenpenalty\z@
                \doublehyphendemerits\z@\finalhyphendemerits\z@
                \strut ##1\strut}%
        }%
        % If the linebreak is at a space, the latter will be displayed as visible
        % space at end of first line, and a continuation symbol starts next line.
        % Stretch/shrink are however usually zero for typewriter font.
        \def\FV@Space {%
            \nobreak\hskip\z@ plus\fontdimen3\font minus\fontdimen4\font
            \discretionary{\copy\Wrappedvisiblespacebox}{\Wrappedafterbreak}
            {\kern\fontdimen2\font}%
        }%

        % Allow breaks at special characters using \PYG... macros.
        \Wrappedbreaksatspecials
        % Breaks at punctuation characters . , ; ? ! and / need catcode=\active
        \OriginalVerbatim[#1,codes*=\Wrappedbreaksatpunct]%
    }
    \makeatother

    % Exact colors from NB
    \definecolor{incolor}{HTML}{303F9F}
    \definecolor{outcolor}{HTML}{D84315}
    \definecolor{cellborder}{HTML}{CFCFCF}
    \definecolor{cellbackground}{HTML}{F7F7F7}

    % prompt
    \makeatletter
    \newcommand{\boxspacing}{\kern\kvtcb@left@rule\kern\kvtcb@boxsep}
    \makeatother
    \newcommand{\prompt}[4]{
        {\ttfamily\llap{{\color{#2}[#3]:\hspace{3pt}#4}}\vspace{-\baselineskip}}
    }

    % Prevent overflowing lines due to hard-to-break entities
    \sloppy
    % Setup hyperref package
    \hypersetup{
      breaklinks=true,  % so long urls are correctly broken across lines
      colorlinks=true,
      urlcolor=urlcolor,
      linkcolor=linkcolor,
      citecolor=citecolor,
      }
    % Slightly bigger margins than the latex defaults

    \geometry{verbose,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}

\begin{document}

    \maketitle

    \hypertarget{excel-what-if-analysis-with-python---part-1-models-and-data-tables}{%
\subsection{Excel ``What if?'' analysis with Python - Part 1: Models and
Data
Tables}\label{excel-what-if-analysis-with-python---part-1-models-and-data-tables}}

    Excel is widely used for building and using models of business problems
to explore the impact of various model inputs on key outputs. Built in
``what if?'' tools such as Excel
\href{https://support.microsoft.com/en-us/office/calculate-multiple-results-by-using-a-data-table-e95e2487-6ca6-4413-ad12-77542a5ea50b}{Data
Tables} and
\href{https://support.microsoft.com/en-us/office/use-goal-seek-to-find-the-result-you-want-by-adjusting-an-input-value-320cb99e-f4a4-417f-b1c3-4f369d6e66c7}{Goal
Seek} are well known to power spreadsheet modelers. How might we do
similar modeling and analysis using Python?

While Python has been gaining momentum in the business analytics world,
it is often used for data wrangling, analysis and visualization of
tablular data using tools like pandas and matplotlib or Seaborn. You can
find some great examples at \href{https://pbpython.com/}{Chris Moffit's
Practical Business Python blog}. I use Python all the time for such
tasks and teach a course called
\href{http://www.sba.oakland.edu/faculty/isken/courses/mis5470/}{Practical
Computing for Data Analytics} that is Python (and R) based. But, it got
me to thinking. What about those things for which Excel is well suited
such as building formula based models and doing sensitivity analysis on
these models? What would those look like in Python?

For example, here's a high level screenshot of a model that I assign for
homework in my
\href{http://www.sba.oakland.edu/faculty/isken/courses/mis5460/}{MIS
4460/5460 Business Analytics class} (a spreadsheet based modeling
class). It's a really simple model in which we are selling a single
product that we produce. There is a fixed cost to producing the product
as well as a variable production cost per unit. We can sell the product
for some price and we believe that demand for the product is related to
the selling price through a power function. Let's assume for now that we
have sufficient capacity to produce to demand and that all inputs are
deterministic (we'll deal with simulating uncertainty later in this
document).

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{ }{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{k+kn}{from} \PY{n+nn}{IPython}\PY{n+nn}{.}\PY{n+nn}{display} \PY{k+kn}{import} \PY{n}{Image}
\PY{n}{Image}\PY{p}{(}\PY{n}{filename}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{images/tech\PYZus{}sales\PYZus{}model.png}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    The details aren't so important right now as is the overall structure of
the model. There's a few key inputs and some pretty straightforward
formulas for computing cost, revenue and profit. Notice the 1-way Data
Table being used to explore how profit varies for different selling
prices. There's a graph driven by the Data Table and some Text Boxes
used for annotation and summary interpretative comments. There's a
button that launches Goal Seek to find the break even selling price and
a 2-way Data Table (not shown) to explore the joint effect of selling
price and variable cost. Classic Excel modeling stuff. How might we go
about building a similar model using Python?

    What if we wanted to push it a little further and model some of the key
inputs with probability distributions to reflect our uncertainty about
their values? In the Excel world, we might use add-ins such as @Risk
which allow uncertain quantities to be directly modeled with probability
distributions. For example, we might have a key input such as the
exponent in the power function that relates selling price to demand that
is highly uncertain. By modeling it with a probability distribution and
then sampling from that distribution many times (essentially by
recalcing the spreadsheet) we can generate a bunch of possible values
for key outputs (e.g.~profit) and use statistics to summarize these
outputs using things like histograms and summary stats. Often this type
of simulation model is referred to as a \emph{Monte-Carlo} model to
suggest repeated sampling from one or more probability distributions
within an otherwise pretty static model. If you want to see such models
in action, check out my
\href{http://www.sba.oakland.edu/faculty/isken/courses/mis5460/simulation.html}{Simulation
Modeling with Excel page} from my Business Analytics course. Again, how
might we do this with Python?

    In the remainder of this notebook (and subsequent follow on notebooks),
we'll explore these questions and along the way introduce some basic
object-oriented (OO) programming concepts and other slightly more
advanced Python techniques. This is aimed at those who might have a
strong Excel based background but only a basic familiarity with Python
programming. We are going to use a slightly different problem example
than the one I've discussed above (we'll save that one for an ``exercise
for the motivated reader''). I've structured things as follows:

\textbf{Part 1 - Models and Data Tables: this notebook} * The bookstore
problem - Non-OO model for the bookstore problem - 1-, 2-, n-way Data
Tables with the non-OO model * Everything is an object in Python * An OO
model for the bookstore problem - 1-, 2-, n-way data tables with the OO
model

\textbf{Part 2 - Goal Seek: second notebook} * Goal Seek - a non-OO
hornet's nest - the OO approach

\textbf{Part 3 - Simulation: third notebook} * Monte-Carlo simulation

\textbf{Part 4 - Create package: fourth notebook} * share the results

\textbf{Part 5 - Simple GUI w/widgets: fifth notebook}

    \hypertarget{import-some-libraries}{%
\subsection{Import some libraries}\label{import-some-libraries}}

We'll need several Python libraries for this post.

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{1}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{k+kn}{import} \PY{n+nn}{copy}

\PY{k+kn}{import} \PY{n+nn}{numpy} \PY{k}{as} \PY{n+nn}{np}
\PY{k+kn}{import} \PY{n+nn}{pandas} \PY{k}{as} \PY{n+nn}{pd}
\PY{k+kn}{import} \PY{n+nn}{matplotlib}\PY{n+nn}{.}\PY{n+nn}{pyplot} \PY{k}{as} \PY{n+nn}{plt}
\PY{k+kn}{from} \PY{n+nn}{mpl\PYZus{}toolkits}\PY{n+nn}{.}\PY{n+nn}{mplot3d} \PY{k+kn}{import} \PY{n}{Axes3D}
\PY{k+kn}{import} \PY{n+nn}{seaborn} \PY{k}{as} \PY{n+nn}{sns}
\end{Verbatim}
\end{tcolorbox}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{2}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{o}{\PYZpc{}}\PY{k}{matplotlib} inline
\end{Verbatim}
\end{tcolorbox}

    \hypertarget{bookstore-model}{%
\subsection{Bookstore model}\label{bookstore-model}}

This example is based on one in the
\href{https://host.kelley.iu.edu/albrightbooks/}{spreadsheet modeling
textbook(s) I've used in my classes since 2001}. I started out using
Practical Management Science by Winston and Albright and switched to
their Business Analytics: Data Analysis and Decision Making (Albright
and Winston) around 2013ish. In both books, they introduce the ``Walton
Bookstore'' problem in the chapter on Monte-Carlo simulation. Here's the
basic problem (with a few modifications):

\begin{itemize}
\tightlist
\item
  we have to place an order for a perishable product (e.g.~a calendar),
\item
  there's a known unit cost for each one ordered,
\item
  we have a known selling price,
\item
  demand is uncertain but we can model it with some simple probability
  distribution,
\item
  for each unsold item, we can get a partial refund of our unit cost,
\item
  we need to select the order quantity for our one order for the year;
  orders can only be in multiples of 25.
\end{itemize}

    \hypertarget{base-model---non-oo-approach}{%
\subsubsection{Base model - non-OO
approach}\label{base-model---non-oo-approach}}

Starting simple, let's create some initialized variables for the base
inputs.

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{3}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{c+c1}{\PYZsh{} Base inputs}
\PY{n}{unit\PYZus{}cost} \PY{o}{=} \PY{l+m+mf}{7.50}
\PY{n}{selling\PYZus{}price} \PY{o}{=} \PY{l+m+mf}{10.00}
\PY{n}{unit\PYZus{}refund} \PY{o}{=} \PY{l+m+mf}{2.50}
\end{Verbatim}
\end{tcolorbox}

    Assume we've used historical data to estimate the mean and standard
deviation of demand. Furthermore, let's pretend that a histogram
revealed a relatively normal looking distribution. However, we are going
to start by:

\begin{itemize}
\tightlist
\item
  assuming away the uncertainty,
\item
  treat demand as deterministic using the mean,
\item
  we'll do some sensitivity analysis to the demand.
\end{itemize}

Later, we'll use Monte-Carlo simulation with normally distribution
demand.

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{4}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{c+c1}{\PYZsh{} Demand parameters}
\PY{n}{demand\PYZus{}mean} \PY{o}{=} \PY{l+m+mi}{193}
\PY{n}{demand\PYZus{}sd} \PY{o}{=} \PY{l+m+mi}{40}

\PY{c+c1}{\PYZsh{} Deterministic model}
\PY{n}{demand} \PY{o}{=} \PY{n}{demand\PYZus{}mean}
\end{Verbatim}
\end{tcolorbox}

    Finally, let's set the initial order quantity. This will be the variable
we'll focus on in the sensitivity analysis (along with demand). Of
course, if we pretend that we know demand is really equal to 193, then
we'd only consider ordering 175 or 200 and would pick the one leading to
higher profit. Let's set it to 200.

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{5}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{order\PYZus{}quantity} \PY{o}{=} \PY{l+m+mi}{200}
\end{Verbatim}
\end{tcolorbox}

    Now we can compute the various cost and revenue components and get to
the bottom line profit.

\textbf{QUESTION 1:} Complete the lines of code below (answer at bottom
of notebook)

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{6}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{order\PYZus{}cost} \PY{o}{=} \PY{n}{unit\PYZus{}cost} \PY{o}{*} \PY{n}{order\PYZus{}quantity}
\PY{n}{sales\PYZus{}revenue} \PY{o}{=} \PY{err}{?}\PY{err}{?}\PY{err}{?} \PY{o}{*} \PY{n}{selling\PYZus{}price}
\PY{n}{refund\PYZus{}revenue} \PY{o}{=} \PY{err}{?}\PY{err}{?}\PY{err}{?} \PY{o}{*} \PY{n}{unit\PYZus{}refund}
\PY{n}{profit} \PY{o}{=} \PY{n}{sales\PYZus{}revenue} \PY{o}{+} \PY{n}{refund\PYZus{}revenue} \PY{o}{\PYZhy{}} \PY{n}{order\PYZus{}cost}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}, frame=single, framerule=2mm, rulecolor=\color{outerrorbackground}]
\textcolor{ansi-cyan}{  Cell }\textcolor{ansi-green}{In[6], line 2}
\textcolor{ansi-red}{    sales\_revenue = ??? * selling\_price}
                    \^{}
\textcolor{ansi-red}{SyntaxError}\textcolor{ansi-red}{:} invalid syntax

    \end{Verbatim}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{7}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n+nb}{print}\PY{p}{(}\PY{n}{profit}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}, frame=single, framerule=2mm, rulecolor=\color{outerrorbackground}]
\textcolor{ansi-red}{---------------------------------------------------------------------------}
\textcolor{ansi-red}{NameError}                                 Traceback (most recent call last)
Cell \textcolor{ansi-green}{In[7], line 1}
\textcolor{ansi-green}{----> 1} \def\tcRGB{\textcolor[RGB]}\expandafter\tcRGB\expandafter{\detokenize{0,135,0}}{print}(\setlength{\fboxsep}{0pt}\colorbox{ansi-yellow}{profit\strut})

\textcolor{ansi-red}{NameError}: name 'profit' is not defined
    \end{Verbatim}

    Of course, working in a Jupyter notebook is different than working in an
Excel workbook. If we modify one of the base input values, the value of
\texttt{profit} is \textbf{not} going to magically update.

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{ }{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{unit\PYZus{}cost} \PY{o}{=} \PY{l+m+mf}{8.00}
\PY{n+nb}{print}\PY{p}{(}\PY{n}{profit}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    We either need to:

\begin{itemize}
\tightlist
\item
  rerun the code cell up above that computes the cost, revenue and
  profit,
\item
  or make the unit cost change up above when we first initialized and
  rerun all the cells,
\item
  or take a different approach\ldots{}
\end{itemize}

Since it feels like we are going to want to compute profit for different
combinations of base input values, it makes sense to create a function
to compute it. We'll do that eventually, but let's keep going with this
simplistic approach for a bit longer and learn some things about working
with vectors.

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{8}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{c+c1}{\PYZsh{} Reset entire model}

\PY{c+c1}{\PYZsh{} Base inputs}
\PY{n}{unit\PYZus{}cost} \PY{o}{=} \PY{l+m+mf}{7.50}
\PY{n}{selling\PYZus{}price} \PY{o}{=} \PY{l+m+mf}{10.00}
\PY{n}{unit\PYZus{}refund} \PY{o}{=} \PY{l+m+mf}{2.50}
\PY{n}{demand} \PY{o}{=} \PY{l+m+mi}{193}
\PY{n}{order\PYZus{}quantity} \PY{o}{=} \PY{l+m+mi}{200}

\PY{c+c1}{\PYZsh{} Intermediate variables}
\PY{n}{order\PYZus{}cost} \PY{o}{=} \PY{n}{unit\PYZus{}cost} \PY{o}{*} \PY{n}{order\PYZus{}quantity}
\PY{n}{sales\PYZus{}revenue} \PY{o}{=} \PY{n+nb}{min}\PY{p}{(}\PY{n}{order\PYZus{}quantity}\PY{p}{,} \PY{n}{demand}\PY{p}{)} \PY{o}{*} \PY{n}{selling\PYZus{}price}
\PY{n}{refund\PYZus{}revenue} \PY{o}{=} \PY{n+nb}{max}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n}{order\PYZus{}quantity} \PY{o}{\PYZhy{}} \PY{n}{demand}\PY{p}{)} \PY{o}{*} \PY{n}{unit\PYZus{}refund}

\PY{c+c1}{\PYZsh{} Key output variable}
\PY{n}{profit} \PY{o}{=} \PY{n}{sales\PYZus{}revenue} \PY{o}{+} \PY{n}{refund\PYZus{}revenue} \PY{o}{\PYZhy{}} \PY{n}{order\PYZus{}cost}

\PY{n}{message} \PY{o}{=} \PY{l+s+sa}{f}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+si}{\PYZob{}}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{Order Cost}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+si}{:}\PY{l+s+s2}{15}\PY{l+s+si}{\PYZcb{}}\PY{l+s+si}{\PYZob{}}\PY{n}{order\PYZus{}cost}\PY{l+s+si}{:}\PY{l+s+s2}{10.2f}\PY{l+s+si}{\PYZcb{}}\PY{l+s+s2}{ }\PY{l+s+se}{\PYZbs{}n}\PY{l+s+s2}{\PYZdq{}} \PYZbs{}
          \PY{l+s+sa}{f}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+si}{\PYZob{}}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{Sales Revenue}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+si}{:}\PY{l+s+s2}{15}\PY{l+s+si}{\PYZcb{}}\PY{l+s+si}{\PYZob{}}\PY{n}{sales\PYZus{}revenue}\PY{l+s+si}{:}\PY{l+s+s2}{10.2f}\PY{l+s+si}{\PYZcb{}}\PY{l+s+s2}{ }\PY{l+s+se}{\PYZbs{}n}\PY{l+s+s2}{\PYZdq{}} \PYZbs{}
          \PY{l+s+sa}{f}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+si}{\PYZob{}}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{Refund}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+si}{:}\PY{l+s+s2}{15}\PY{l+s+si}{\PYZcb{}}\PY{l+s+si}{\PYZob{}}\PY{n}{refund\PYZus{}revenue}\PY{l+s+si}{:}\PY{l+s+s2}{10.2f}\PY{l+s+si}{\PYZcb{}}\PY{l+s+s2}{ }\PY{l+s+se}{\PYZbs{}n}\PY{l+s+s2}{\PYZdq{}} \PYZbs{}
          \PY{l+s+sa}{f}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+si}{\PYZob{}}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{Profit}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+si}{:}\PY{l+s+s2}{15}\PY{l+s+si}{\PYZcb{}}\PY{l+s+si}{\PYZob{}}\PY{n}{profit}\PY{l+s+si}{:}\PY{l+s+s2}{10.2f}\PY{l+s+si}{\PYZcb{}}\PY{l+s+s2}{\PYZdq{}}

\PY{n+nb}{print}\PY{p}{(}\PY{n}{message}\PY{p}{)} 
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]
Order Cost        1500.00
Sales Revenue     1930.00
Refund              17.50
Profit             447.50
    \end{Verbatim}

    Notice we've used Python's newish approach to formatting strings called
\href{https://realpython.com/python-f-strings/}{f-strings}. See
https://pyformat.info/ for the basics on the format specification (using
examples comparing the old \% style formatting versus the newer
\texttt{str.format()} approach) or the
\href{https://docs.python.org/3/library/string.html\#format-specification-mini-language}{official
docs for all the gory details}.

    \hypertarget{sensitivity-analysis-analogous-to-a-1-way-data-table}{%
\subsubsection{Sensitivity analysis analogous to a 1-way Data
Table}\label{sensitivity-analysis-analogous-to-a-1-way-data-table}}

To compute \texttt{profit} for several different values of
\texttt{order\_quantity} in Excel, we could do a 1-way Data Table.

In Python, we could set \texttt{order\_quantity} equal to a vector and
then recompute profit. Our model should spit out a vector of profit
values corresponding to the order quantities in our
\texttt{order\_quantity} vector. The numpy function
\href{https://numpy.org/doc/stable/reference/generated/numpy.arange.html}{arange}
can be used to create the vector.

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{9}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{order\PYZus{}quantity} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{arange}\PY{p}{(}\PY{n}{start}\PY{o}{=}\PY{l+m+mi}{50}\PY{p}{,} \PY{n}{stop}\PY{o}{=}\PY{l+m+mi}{301}\PY{p}{,} \PY{n}{step}\PY{o}{=}\PY{l+m+mi}{25}\PY{p}{)}
\PY{n+nb}{print}\PY{p}{(}\PY{n}{order\PYZus{}quantity}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]
[ 50  75 100 125 150 175 200 225 250 275 300]
    \end{Verbatim}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{10}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n+nb}{type}\PY{p}{(}\PY{n}{order\PYZus{}quantity}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

            \begin{tcolorbox}[breakable, size=fbox, boxrule=.5pt, pad at break*=1mm, opacityfill=0]
\prompt{Out}{outcolor}{10}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
numpy.ndarray
\end{Verbatim}
\end{tcolorbox}

    \textbf{QUESTION} Why do we set \texttt{stop=301} instead of 300? Also,
what's the data type of \texttt{order\_quantity}?

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{11}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{order\PYZus{}cost} \PY{o}{=} \PY{n}{unit\PYZus{}cost} \PY{o}{*} \PY{n}{order\PYZus{}quantity}
\PY{n}{sales\PYZus{}revenue} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{minimum}\PY{p}{(}\PY{n}{order\PYZus{}quantity}\PY{p}{,} \PY{n}{demand}\PY{p}{)} \PY{o}{*} \PY{n}{selling\PYZus{}price}
\PY{n}{refund\PYZus{}revenue} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{maximum}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n}{order\PYZus{}quantity} \PY{o}{\PYZhy{}} \PY{n}{demand}\PY{p}{)} \PY{o}{*} \PY{n}{unit\PYZus{}refund}
\PY{n}{profit} \PY{o}{=} \PY{n}{sales\PYZus{}revenue} \PY{o}{+} \PY{n}{refund\PYZus{}revenue} \PY{o}{\PYZhy{}} \PY{n}{order\PYZus{}cost}
\PY{n+nb}{print}\PY{p}{(}\PY{n}{profit}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]
[125.  187.5 250.  312.5 375.  437.5 447.5 322.5 197.5  72.5 -52.5]
    \end{Verbatim}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{12}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{sales\PYZus{}revenue}
\end{Verbatim}
\end{tcolorbox}

            \begin{tcolorbox}[breakable, size=fbox, boxrule=.5pt, pad at break*=1mm, opacityfill=0]
\prompt{Out}{outcolor}{12}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
array([ 500.,  750., 1000., 1250., 1500., 1750., 1930., 1930., 1930.,
       1930., 1930.])
\end{Verbatim}
\end{tcolorbox}

    We can't just use \texttt{min()}( with a scalar (\texttt{demand})
compared to a vector (\texttt{order\_quantity}). Need to use numpy
functions since they will do the necessary
\href{https://numpy.org/doc/stable/user/basics.broadcasting.html}{broadcasting}
of demand values to fill out a vector the same size as the
\texttt{order\_quantity} vector. In R, broadcasting is usually referred
to as \emph{recycling}.

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{13}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{np}\PY{o}{.}\PY{n}{minimum}\PY{p}{(}\PY{n}{order\PYZus{}quantity}\PY{p}{,} \PY{n}{demand}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

            \begin{tcolorbox}[breakable, size=fbox, boxrule=.5pt, pad at break*=1mm, opacityfill=0]
\prompt{Out}{outcolor}{13}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
array([ 50,  75, 100, 125, 150, 175, 193, 193, 193, 193, 193])
\end{Verbatim}
\end{tcolorbox}

    This won't work:

\begin{verbatim}
min(order_quantity, demand)

---------------------------------------------------------------------------
ValueError                                Traceback (most recent call last)
<ipython-input-42-675af9b521e5> in <module>
----> 1 min(order_quantity, demand)

ValueError: The truth value of an array with more than one element is ambiguous. Use a.any() or a.all()
\end{verbatim}

    Here's a simple plot to show how profit varies by order quantity.

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{14}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{plt}\PY{o}{.}\PY{n}{plot}\PY{p}{(}\PY{n}{order\PYZus{}quantity}\PY{p}{,} \PY{n}{profit}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{Profit}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
\PY{n}{plt}\PY{o}{.}\PY{n}{xlabel}\PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{Order Quantity}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
\PY{n}{plt}\PY{o}{.}\PY{n}{ylabel}\PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{Profit}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
\PY{n}{plt}\PY{o}{.}\PY{n}{legend}\PY{p}{(}\PY{p}{)}
\PY{n}{plt}\PY{o}{.}\PY{n}{show}\PY{p}{(}\PY{p}{)}\PY{p}{;}
\end{Verbatim}
\end{tcolorbox}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{output_34_0.png}
    \end{center}
    { \hspace*{\fill} \\}

    \textbf{QUESTION} What about a 2-way Data Table? Why can't we just set
one of the other inputs (e.g.~demand) to a vector and do the same thing
we did above?

    One approach to creating a 2-way (actually, n-way) is to create a profit
function that we can call from within a list comprehension.

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{15}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{k}{def} \PY{n+nf}{bookstore\PYZus{}profit}\PY{p}{(}\PY{n}{unit\PYZus{}cost}\PY{p}{,} \PY{n}{selling\PYZus{}price}\PY{p}{,} \PY{n}{unit\PYZus{}refund}\PY{p}{,} \PY{n}{order\PYZus{}quantity}\PY{p}{,} \PY{n}{demand}\PY{p}{)}\PY{p}{:}
    \PY{l+s+sd}{\PYZsq{}\PYZsq{}\PYZsq{}}
\PY{l+s+sd}{    Compute profit in bookstore model}
\PY{l+s+sd}{    \PYZsq{}\PYZsq{}\PYZsq{}}
    \PY{n}{order\PYZus{}cost} \PY{o}{=} \PY{n}{unit\PYZus{}cost} \PY{o}{*} \PY{n}{order\PYZus{}quantity}
    \PY{n}{sales\PYZus{}revenue} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{minimum}\PY{p}{(}\PY{n}{order\PYZus{}quantity}\PY{p}{,} \PY{n}{demand}\PY{p}{)} \PY{o}{*} \PY{n}{selling\PYZus{}price}
    \PY{n}{refund\PYZus{}revenue} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{maximum}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n}{order\PYZus{}quantity} \PY{o}{\PYZhy{}} \PY{n}{demand}\PY{p}{)}
    \PY{n}{profit} \PY{o}{=} \PY{n}{sales\PYZus{}revenue} \PY{o}{+} \PY{n}{refund\PYZus{}revenue} \PY{o}{\PYZhy{}} \PY{n}{order\PYZus{}cost}
    \PY{k}{return} \PY{n}{profit}
\end{Verbatim}
\end{tcolorbox}

    A
\href{https://www.pythontutorial.net/python-basics/python-list-comprehensions/}{list
comprehension} can be used to create an n-way Data Table. List
comprehensions provide a nice syntactical shortcut to creating a list
based one looping over one or more iterables and computing associated
values to store in the list. As a simple first example that's related to
the model we are working on, let's compute the difference between demand
and order quantity for all the combinations of those variables in the
following ranges:

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{16}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{demand\PYZus{}range} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{arange}\PY{p}{(}\PY{l+m+mi}{50}\PY{p}{,} \PY{l+m+mi}{101}\PY{p}{,} \PY{l+m+mi}{10}\PY{p}{)}
\PY{n}{order\PYZus{}quantity\PYZus{}range} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{arange}\PY{p}{(}\PY{l+m+mi}{50}\PY{p}{,} \PY{l+m+mi}{101}\PY{p}{,} \PY{l+m+mi}{25}\PY{p}{)}
\PY{n+nb}{print}\PY{p}{(}\PY{n}{demand\PYZus{}range}\PY{p}{)}
\PY{n+nb}{print}\PY{p}{(}\PY{n}{order\PYZus{}quantity\PYZus{}range}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]
[ 50  60  70  80  90 100]
[ 50  75 100]
    \end{Verbatim}

    Let's first use a list comprehension to list the combinations (we'll
store each combo in a tuple). Each combination is a \emph{scenario} for
which we'd like to compute profit.

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{17}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n+nb}{print}\PY{p}{(}\PY{p}{[}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{o}\PY{p}{)} \PY{k}{for} \PY{n}{d} \PY{o+ow}{in} \PY{n}{demand\PYZus{}range} \PY{k}{for} \PY{n}{o} \PY{o+ow}{in} \PY{n}{order\PYZus{}quantity\PYZus{}range}\PY{p}{]}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]
[(50, 50), (50, 75), (50, 100), (60, 50), (60, 75), (60, 100), (70, 50), (70,
75), (70, 100), (80, 50), (80, 75), (80, 100), (90, 50), (90, 75), (90, 100),
(100, 50), (100, 75), (100, 100)]
    \end{Verbatim}

    This time I'll compute a list of differences between demand and order
quantity.

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{18}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n+nb}{print}\PY{p}{(}\PY{p}{[}\PY{n}{d} \PY{o}{\PYZhy{}} \PY{n}{o} \PY{k}{for} \PY{n}{d} \PY{o+ow}{in} \PY{n}{demand\PYZus{}range} \PY{k}{for} \PY{n}{o} \PY{o+ow}{in} \PY{n}{order\PYZus{}quantity\PYZus{}range}\PY{p}{]}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]
[0, -25, -50, 10, -15, -40, 20, -5, -30, 30, 5, -20, 40, 15, -10, 50, 25, 0]
    \end{Verbatim}

    Here's a 2-way with demand and order quantity - the other inputs are
held constant at their current values and the \texttt{bookstore\_profit}
function is used to compute profit. The list comprehension will generate
a list of tuples, \texttt{(demand,\ order\ quantity,\ profit)}. To
facilitate further analysis, we'll create a pandas DataFrame from this
list.

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{19}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{demand\PYZus{}range} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{arange}\PY{p}{(}\PY{l+m+mi}{50}\PY{p}{,} \PY{l+m+mi}{301}\PY{p}{,} \PY{l+m+mi}{5}\PY{p}{)}
\PY{n}{order\PYZus{}quantity\PYZus{}range} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{arange}\PY{p}{(}\PY{l+m+mi}{50}\PY{p}{,} \PY{l+m+mi}{301}\PY{p}{,} \PY{l+m+mi}{25}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{20}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{c+c1}{\PYZsh{} Create data table (as a list of tuples)}
\PY{n}{data\PYZus{}table\PYZus{}1} \PY{o}{=} \PY{p}{[}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{oq}\PY{p}{,} \PY{n}{bookstore\PYZus{}profit}\PY{p}{(}\PY{n}{unit\PYZus{}cost}\PY{p}{,} \PY{n}{selling\PYZus{}price}\PY{p}{,} \PY{n}{unit\PYZus{}refund}\PY{p}{,} \PY{n}{oq}\PY{p}{,} \PY{n}{d}\PY{p}{)}\PY{p}{)} 
                 \PY{k}{for} \PY{n}{d} \PY{o+ow}{in} \PY{n}{demand\PYZus{}range} \PY{k}{for} \PY{n}{oq} \PY{o+ow}{in} \PY{n}{order\PYZus{}quantity\PYZus{}range}\PY{p}{]}

\PY{c+c1}{\PYZsh{} Convert to dataframe}
\PY{n}{dtbl\PYZus{}1\PYZus{}df} \PY{o}{=} \PY{n}{pd}\PY{o}{.}\PY{n}{DataFrame}\PY{p}{(}\PY{n}{data\PYZus{}table\PYZus{}1}\PY{p}{,} \PY{n}{columns}\PY{o}{=}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{Demand}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{OrderQuantity}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{Profit}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{p}{)}
\PY{n}{dtbl\PYZus{}1\PYZus{}df}\PY{o}{.}\PY{n}{head}\PY{p}{(}\PY{l+m+mi}{25}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

            \begin{tcolorbox}[breakable, size=fbox, boxrule=.5pt, pad at break*=1mm, opacityfill=0]
\prompt{Out}{outcolor}{20}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
    Demand  OrderQuantity  Profit
0       50             50   125.0
1       50             75   -37.5
2       50            100  -200.0
3       50            125  -362.5
4       50            150  -525.0
5       50            175  -687.5
6       50            200  -850.0
7       50            225 -1012.5
8       50            250 -1175.0
9       50            275 -1337.5
10      50            300 -1500.0
11      55             50   125.0
12      55             75     7.5
13      55            100  -155.0
14      55            125  -317.5
15      55            150  -480.0
16      55            175  -642.5
17      55            200  -805.0
18      55            225  -967.5
19      55            250 -1130.0
20      55            275 -1292.5
21      55            300 -1455.0
22      60             50   125.0
23      60             75    52.5
24      60            100  -110.0
\end{Verbatim}
\end{tcolorbox}

    Here's a quick scatter plot showing how profit (mapped to color hue)
varies jointly by order quantity and demand. No surprises.

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{21}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{sns}\PY{o}{.}\PY{n}{set\PYZus{}style}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{darkgrid}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
\PY{n}{sns}\PY{o}{.}\PY{n}{scatterplot}\PY{p}{(}\PY{n}{x}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{Demand}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{y}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{OrderQuantity}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{hue}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{Profit}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{data}\PY{o}{=}\PY{n}{dtbl\PYZus{}1\PYZus{}df}\PY{p}{,} \PY{n}{palette}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{viridis}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{p}{;}
\end{Verbatim}
\end{tcolorbox}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{output_48_0.png}
    \end{center}
    { \hspace*{\fill} \\}

    For fun, let's try a 3d plot.

https://matplotlib.org/mpl\_toolkits/mplot3d/tutorial.html\#mplot3d-tutorial

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{23}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{c+c1}{\PYZsh{} Make the plot }
\PY{n}{fig} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{figure}\PY{p}{(}\PY{p}{)}
\PY{n}{ax} \PY{o}{=} \PY{n}{fig}\PY{o}{.}\PY{n}{add\PYZus{}subplot}\PY{p}{(}\PY{l+m+mi}{111}\PY{p}{,} \PY{n}{projection}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{3d}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
\PY{n}{ax}\PY{o}{.}\PY{n}{scatter}\PY{p}{(}\PY{n}{dtbl\PYZus{}1\PYZus{}df}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{OrderQuantity}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{p}{,} \PY{n}{dtbl\PYZus{}1\PYZus{}df}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{Demand}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{p}{,} \PY{n}{dtbl\PYZus{}1\PYZus{}df}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{Profit}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{p}{,} 
                \PY{n}{cmap}\PY{o}{=}\PY{n}{plt}\PY{o}{.}\PY{n}{cm}\PY{o}{.}\PY{n}{viridis}\PY{p}{,} \PY{n}{c}\PY{o}{=}\PY{n}{dtbl\PYZus{}1\PYZus{}df}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{Profit}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{p}{)}

\PY{n}{ax}\PY{o}{.}\PY{n}{set\PYZus{}xlabel}\PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{OrderQuantity}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
\PY{n}{ax}\PY{o}{.}\PY{n}{set\PYZus{}ylabel}\PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{Demand}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
\PY{n}{ax}\PY{o}{.}\PY{n}{set\PYZus{}zlabel}\PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{Profit}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}

\PY{n}{plt}\PY{o}{.}\PY{n}{show}\PY{p}{(}\PY{p}{)}\PY{p}{;}
\end{Verbatim}
\end{tcolorbox}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{output_50_0.png}
    \end{center}
    { \hspace*{\fill} \\}

    \hypertarget{everything-is-an-object-in-python}{%
\subsection{Everything is an object in
Python}\label{everything-is-an-object-in-python}}

So far we have used standard procedural Python for creating and using
our model. However, we want to explore using an object-oriented approach
to doing the same things we just did. Before doing that, we need to
review some basics regarding object-oriented programming.

Objects are the ``things'' we create, use, change and share in our
Python programs. Objects contain both data about the object and
functions for using and changing the object. If you've programmed in
Excel VBA, you've already worked with objects and their various
properties (data) and methods (functions). For example, consider a
\texttt{Worksheet} object. One of its properties is \texttt{Name} and we
can reference it with \texttt{Worksheet.Name}. To copy a worksheet, we
use \texttt{Worksheet.Copy} - \texttt{Copy} is a method.

\hypertarget{list-objects}{%
\subsubsection{List objects}\label{list-objects}}

In Python, everything is an object and we've already been using objects
in this notebook. It's probably not surprising that something like an
array or a list are objects. Let's see what kind of objects they are.

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{ }{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n+nb}{type}\PY{p}{(}\PY{n}{order\PYZus{}quantity\PYZus{}range}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{ }{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n+nb}{type}\PY{p}{(}\PY{n}{data\PYZus{}table\PYZus{}1}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    Let's explore some list methods. To start, we will create a list based
on the \texttt{order\_quantity\_range} and then do a few things with the
new list.

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{ }{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{oq\PYZus{}list} \PY{o}{=} \PY{n+nb}{list}\PY{p}{(}\PY{n}{order\PYZus{}quantity\PYZus{}range}\PY{p}{)}
\PY{n+nb}{print}\PY{p}{(}\PY{n}{oq\PYZus{}list}\PY{p}{)}
\PY{n+nb}{print}\PY{p}{(}\PY{n+nb}{type}\PY{p}{(}\PY{n}{oq\PYZus{}list}\PY{p}{)}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    A Python \texttt{class} is the template from which we create individual
objects, called \emph{instances} in our programs. The \texttt{list}
class lets us create list objects. Python has many built in classes with
numerous predefined methods and properties.

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{ }{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{c+c1}{\PYZsh{} Append a new item to end of list}
\PY{n}{oq\PYZus{}list}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{l+m+mi}{350}\PY{p}{)}
\PY{n+nb}{print}\PY{p}{(}\PY{n}{oq\PYZus{}list}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{ }{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{c+c1}{\PYZsh{} Reverse the order of our list}
\PY{n}{oq\PYZus{}list}\PY{o}{.}\PY{n}{reverse}\PY{p}{(}\PY{p}{)}
\PY{n+nb}{print}\PY{p}{(}\PY{n}{oq\PYZus{}list}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{ }{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{c+c1}{\PYZsh{} Reverse it back}
\PY{n}{oq\PYZus{}list}\PY{o}{.}\PY{n}{reverse}\PY{p}{(}\PY{p}{)}
\PY{n+nb}{print}\PY{p}{(}\PY{n}{oq\PYZus{}list}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    Notice that when we use the \texttt{append} and \texttt{reverse}
functions we need to include parentheses just as we do when calling any
function. In this case, we aren't passing any arguments into the
function, but often we will.

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{ }{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{c+c1}{\PYZsh{} Use sort() method to do a reverse sort of our list}
\PY{n}{oq\PYZus{}list}\PY{o}{.}\PY{n}{sort}\PY{p}{(}\PY{n}{reverse}\PY{o}{=}\PY{k+kc}{True}\PY{p}{)}
\PY{n+nb}{print}\PY{p}{(}\PY{n}{oq\PYZus{}list}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    What if we forget the parentheses?

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{ }{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n+nb}{print}\PY{p}{(}\PY{n}{oq\PYZus{}list}\PY{o}{.}\PY{n}{reverse}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    Python just tells us that \texttt{reverse} is a function. It's also
telling us that \texttt{reverse}, a function, is also an object. Yes,
functions are objects, too. That means we can do things like pass them
to other functions or modify some of their properties. Why would we ever
want to do this? We'll get to that.

    I'm sure you already know how to get items from a list by their index
number.

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{ }{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{c+c1}{\PYZsh{} Get the first element from the list}
\PY{n+nb}{print}\PY{p}{(}\PY{n}{oq\PYZus{}list}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    The bracket notation is really just a convenient shorthand for calling
one of the built in methods of list objects. We can use Python
\texttt{dir} function to see all of the attributes of any object.

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{ }{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n+nb}{print}\PY{p}{(}\PY{n+nb}{dir}\PY{p}{(}\PY{n}{oq\PYZus{}list}\PY{p}{)}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    Many of the attributes have both leading and trailing double
underscores, known as ``dunder'' attributes. These are just special
names reserved by Python for particular uses. In Python, you'll see use
of both single and double underscores, as both leading and trailing
characters. To dig deeper, check out
https://dbader.org/blog/meaning-of-underscores-in-python.

    In our example, using brackets to access a list item is really just a
different way of calling the \texttt{\_\_getitem\_\_} dunder method.

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{ }{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n+nb}{print}\PY{p}{(}\PY{n}{oq\PYZus{}list}\PY{o}{.}\PY{n+nf+fm}{\PYZus{}\PYZus{}getitem\PYZus{}\PYZus{}}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{)}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    Also notice that all of these functions worked \emph{in place} - they
operated directly on the object \texttt{oq\_list} and even modified it
for the sort related methods. Some methods will create a new instance of
the object and if you want to ``save'' it, you'll have to assign the
result back to your original object or create a new variable. If you've
used R, you might recall that the dplyr package \textbf{never} modifies
dataframes in place. The dplyr functions (e.g.~\texttt{select},
\texttt{mutate}, \texttt{filter}, \ldots{}) always return new objects
and we have to ``catch the results'' by assigning to a variable if we
wanted to actually modify our dataframe.

Returning to our list example, we'll make a copy and explore a bit.

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{ }{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{new\PYZus{}list} \PY{o}{=} \PY{n}{oq\PYZus{}list}\PY{o}{.}\PY{n}{copy}\PY{p}{(}\PY{p}{)}    \PY{c+c1}{\PYZsh{} Make a copy and save to new variable}
\PY{n+nb}{print}\PY{p}{(}\PY{n}{new\PYZus{}list}\PY{p}{)}              \PY{c+c1}{\PYZsh{} Check out the new variable}
\PY{n+nb}{print}\PY{p}{(}\PY{n}{new\PYZus{}list} \PY{o}{==} \PY{n}{oq\PYZus{}list}\PY{p}{)}   \PY{c+c1}{\PYZsh{} Does the new variable have the same value as the old variable?}
\PY{n+nb}{print}\PY{p}{(}\PY{n+nb}{type}\PY{p}{(}\PY{n}{new\PYZus{}list}\PY{p}{)} \PY{o}{==} \PY{n+nb}{type}\PY{p}{(}\PY{n}{oq\PYZus{}list}\PY{p}{)}\PY{p}{)}  \PY{c+c1}{\PYZsh{} Are the two variables of the same type?}
\PY{n+nb}{print}\PY{p}{(}\PY{n}{new\PYZus{}list} \PY{o+ow}{is} \PY{n}{oq\PYZus{}list}\PY{p}{)}   \PY{c+c1}{\PYZsh{} Is the new variable actually the same object as the old variable?}
\PY{n+nb}{print}\PY{p}{(}\PY{n+nb}{id}\PY{p}{(}\PY{n}{oq\PYZus{}list}\PY{p}{)}\PY{p}{,} \PY{n+nb}{id}\PY{p}{(}\PY{n}{new\PYZus{}list}\PY{p}{)}\PY{p}{)}  \PY{c+c1}{\PYZsh{} Every object has its own unique id number. Use id() to see it.}
\end{Verbatim}
\end{tcolorbox}

    The two lists have the same value and are of the same datatype, but are
\textbf{not} the same object.

    \textbf{QUESTION} If we reversed one of the two lists, would they still
be considered to have the same value? In other words, does order matter
with lists?

    Care must be taken when copying lists.

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{ }{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n+nb}{print}\PY{p}{(}\PY{n}{oq\PYZus{}list}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{ }{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{c+c1}{\PYZsh{} Instead of using copy method, just setting new variable value equal to existing list}
\PY{n}{new\PYZus{}oq\PYZus{}list} \PY{o}{=} \PY{n}{oq\PYZus{}list} 
\PY{n+nb}{print}\PY{p}{(}\PY{n}{new\PYZus{}oq\PYZus{}list}\PY{p}{)}
\PY{n}{oq\PYZus{}list}\PY{o}{.}\PY{n}{reverse}\PY{p}{(}\PY{p}{)} \PY{c+c1}{\PYZsh{} I\PYZsq{}m reversing the ORIGINAL list}
\PY{n+nb}{print}\PY{p}{(}\PY{n}{new\PYZus{}oq\PYZus{}list}\PY{p}{)}
\PY{n+nb}{print}\PY{p}{(}\PY{n+nb}{id}\PY{p}{(}\PY{n}{oq\PYZus{}list}\PY{p}{)}\PY{p}{,} \PY{n+nb}{id}\PY{p}{(}\PY{n}{new\PYZus{}oq\PYZus{}list}\PY{p}{)}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    Yikes! As you can see from the \texttt{id} numbers, both list variables
point to the same object in memory. Changing one list, changes the other
- they're the same object. So, how do you safely copy lists? You use the
\texttt{copy} method like we did above.

    Let's look at the attributes of a \texttt{numpy.array} object and see
how that compares to a list. They are both data structures that allow us
to store a bunch of values and get at them via an index number. What
differences do you think these two data structures might have?

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{ }{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n+nb}{dir}\PY{p}{(}\PY{n}{order\PYZus{}quantity\PYZus{}range}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    Wow,
\href{https://jakevdp.github.io/PythonDataScienceHandbook/02.02-the-basics-of-numpy-arrays.html}{numpy
arrays} have a whole lot more going on under the hood than do lists.
They play a huge role in doing data science work in Python with many
packages using numpy to store and work with data. Notice that there are
various mathematical methods such as \texttt{sum} and \texttt{mean} and
methods to convert arrays to lists and strings using the obviously named
methods. We see that there is a \texttt{dtype} property - it tells us
the data type of the array. Our array stores 64bit integers.

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{ }{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n+nb}{print}\PY{p}{(}\PY{n}{order\PYZus{}quantity\PYZus{}range}\PY{o}{.}\PY{n}{dtype}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \textbf{QUESTIONS} Why don't we use parens with \texttt{dtype}? Why
don't list objects have a \texttt{dtype} property?

    \hypertarget{what-about-numbers}{%
\subsubsection{What about numbers?}\label{what-about-numbers}}

Are things like integers and floating point numbers actually objects?

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{ }{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n+nb}{print}\PY{p}{(}\PY{n}{unit\PYZus{}cost}\PY{p}{)}
\PY{n+nb}{print}\PY{p}{(}\PY{n+nb}{type}\PY{p}{(}\PY{n}{unit\PYZus{}cost}\PY{p}{)}\PY{p}{)}
\PY{n+nb}{print}\PY{p}{(}\PY{n+nb}{dir}\PY{p}{(}\PY{n}{unit\PYZus{}cost}\PY{p}{)}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{ }{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{unit\PYZus{}cost}\PY{o}{.}\PY{n}{is\PYZus{}integer}\PY{p}{(}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{ }{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{unit\PYZus{}cost}\PY{o}{.}\PY{n}{hex}\PY{p}{(}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    Yep. Try out some of the attributes.

    \hypertarget{writing-our-own-classes}{%
\subsubsection{Writing our own classes}\label{writing-our-own-classes}}

While Python has many useful built in classes, true
\href{https://en.wikipedia.org/wiki/Object-oriented_programming}{object-oriented
programming} (OOP) involves creating our own classes from which we can
create object instances that are useful for the problem at hand. Python
is an OOP language but it does \textbf{not} force you to create classes
if you don't want to. You can write any blend you'd like of
\href{https://en.wikipedia.org/wiki/Procedural_programming}{procedural},
object oriented and
\href{https://en.wikipedia.org/wiki/Functional_programming}{functional
programming}. We aren't going to get into debates about which
programming paradigm is the ``best''. There are times when OOP makes a
lot of sense, and for those times, we'll use it. Other times, simply
creating useful functions (such as the \texttt{bookstore\_profit}
function we used earlier) is the appropriate thing to do.

Many business analysts who learn Excel VBA, don't realize that
\textbf{you can actually create classes in VBA}. Most stick with using
the built in Excel objects.

    \hypertarget{an-oo-version-of-the-bookstore-model}{%
\subsection{An OO version of the Bookstore
Model}\label{an-oo-version-of-the-bookstore-model}}

Let's create an OO version of the Bookstore Model and figure out how to
do n-way data tables just as we did with the non-OO version of the
model. Then we'll go on and figure out how to duplicate the
functionality of Excel's Goal Seek tool in Python. It's not clear
whether the non-OO or the OO version of our model will ultimately prove
best for this use case. Regardless, we'll learn a few things along the
way.

As an aside, there's a very useful site aimed at using Python for
quantitative economics and finance - https://quantecon.org/. Within
there are some terrific notebooks on various Python topics. Here are two
related to OOP:

\begin{itemize}
\tightlist
\item
  https://python-programming.quantecon.org/oop\_intro.html
\item
  https://python-programming.quantecon.org/python\_oop.html
\end{itemize}

    \hypertarget{initial-design-of-the-bookstoremodel-class}{%
\subsubsection{Initial design of the BookstoreModel
class}\label{initial-design-of-the-bookstoremodel-class}}

It seems like we would want all of the base inputs such as unit cost,
selling price, unit refund, order quantity and demand to be attributes
(properties) of the class. Here's the start of our class definition
code:

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{24}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{k}{class} \PY{n+nc}{BookstoreModel}\PY{p}{(}\PY{p}{)}\PY{p}{:}
    \PY{k}{def} \PY{n+nf+fm}{\PYZus{}\PYZus{}init\PYZus{}\PYZus{}}\PY{p}{(}\PY{n+nb+bp}{self}\PY{p}{,} \PY{n}{unit\PYZus{}cost}\PY{p}{,} \PY{n}{selling\PYZus{}price}\PY{p}{,} \PY{n}{unit\PYZus{}refund}\PY{p}{,} \PY{n}{order\PYZus{}quantity}\PY{p}{,} \PY{n}{demand}\PY{p}{)}\PY{p}{:}
        \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{unit\PYZus{}cost} \PY{o}{=} \PY{n}{unit\PYZus{}cost}
        \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{selling\PYZus{}price} \PY{o}{=} \PY{n}{selling\PYZus{}price}
        \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{unit\PYZus{}refund} \PY{o}{=} \PY{n}{unit\PYZus{}refund}
        \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{order\PYZus{}quantity} \PY{o}{=} \PY{n}{order\PYZus{}quantity}
        \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{demand} \PY{o}{=} \PY{n}{demand}
\end{Verbatim}
\end{tcolorbox}

    Some important things of note:

\begin{itemize}
\tightlist
\item
  we use the keyword \texttt{class} followed by the name of our class.
\item
  by convention, class names are capitalized (each word if multi-word).
\item
  the class declaration looks a little like defining a function but we
  use \texttt{class} instead of \texttt{def}.
\item
  there's nothing in the parentheses following the class name. We'll
  defer talking about this for now.
\item
  there's a single method (function) defined within the class and it is
  called \texttt{\_\_init\_\_}. From earlier in this document we know
  that this must be a built in dunder method that has some special
  purpose. From its name you've probably already guessed that it's the
  function called when a new object instance is first created. If you've
  programmed in Java or C++, this is kind of like a constructor.
\item
  the first argument to the \texttt{\_\_init\_\_} function is
  \texttt{self} and represents the object instance being created.
\item
  the remaining arguments to the \texttt{\_\_init\_\_} function are
  values that we want to use in the function. In this case, we simply
  want to create object properties corresponding to the base inputs for
  the bookstore model.
\item
  the rest of the lines within the \texttt{\_\_init\_\_} function are
  initializing our object properties to the values passed in.
\item
  in order to reference (i.e.~to get the value or set the value) an
  object property, we always preface the property with the \texttt{self}
  object, e.g., \texttt{self.unit\_cost}.
\item
  if you've programmed in Java you might be used to using ``setters and
  getters''. In general
  \href{https://www.python-course.eu/python3_properties.php}{we don't do
  that in Python} and instead just reference the property directly using
  dot notation.
\end{itemize}

    Well, our class isn't super useful yet, but let's try it out. Make sure
you run the code chunk above that defines the BookstoreModel class.

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{25}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{c+c1}{\PYZsh{} Reset the base inputs}
\PY{n}{unit\PYZus{}cost} \PY{o}{=} \PY{l+m+mf}{7.50}
\PY{n}{selling\PYZus{}price} \PY{o}{=} \PY{l+m+mf}{10.00}
\PY{n}{unit\PYZus{}refund} \PY{o}{=} \PY{l+m+mf}{2.50}
\PY{n}{demand} \PY{o}{=} \PY{l+m+mi}{193}
\PY{n}{order\PYZus{}quantity} \PY{o}{=} \PY{l+m+mi}{200}

\PY{c+c1}{\PYZsh{} Create a new BookstoreModel object}
\PY{n}{model\PYZus{}1} \PY{o}{=} \PY{n}{BookstoreModel}\PY{p}{(}\PY{n}{unit\PYZus{}cost}\PY{p}{,} \PY{n}{selling\PYZus{}price}\PY{p}{,} \PY{n}{unit\PYZus{}refund}\PY{p}{,} \PY{n}{order\PYZus{}quantity}\PY{p}{,} \PY{n}{demand}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    Let's print out one of the properties.

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{26}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n+nb}{print}\PY{p}{(}\PY{n}{model\PYZus{}1}\PY{o}{.}\PY{n}{unit\PYZus{}cost}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]
7.5
    \end{Verbatim}

    What happens if we try to print the model object itself?

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{27}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n+nb}{print}\PY{p}{(}\PY{n}{model\PYZus{}1}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]
<\_\_main\_\_.BookstoreModel object at 0x7fed59a46290>
    \end{Verbatim}

    Ok, not the most useful but we do see that we've got an object of the
right type and that the object lives in the \texttt{\_\_main\_\_}
\emph{namespace}. Namespaces are important in Python. They're even
mentioned in the Zen of Python. You can
\href{https://realpython.com/python-namespaces-scope/}{learn more about
namespaces here}.

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{28}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{k+kn}{import} \PY{n+nn}{this}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]
The Zen of Python, by Tim Peters

Beautiful is better than ugly.
Explicit is better than implicit.
Simple is better than complex.
Complex is better than complicated.
Flat is better than nested.
Sparse is better than dense.
Readability counts.
Special cases aren't special enough to break the rules.
Although practicality beats purity.
Errors should never pass silently.
Unless explicitly silenced.
In the face of ambiguity, refuse the temptation to guess.
There should be one-- and preferably only one --obvious way to do it.
Although that way may not be obvious at first unless you're Dutch.
Now is better than never.
Although never is often better than *right* now.
If the implementation is hard to explain, it's a bad idea.
If the implementation is easy to explain, it may be a good idea.
Namespaces are one honking great idea -- let's do more of those!
    \end{Verbatim}

    If you look back at our \texttt{bookstore\_profit} function, you'll see
that it first computes the various cost and revenue components and then
computes profit from that. So, it would seem useful to add some methods
to our BookstoreModel class that computed these things.

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{29}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{k}{class} \PY{n+nc}{BookstoreModel}\PY{p}{(}\PY{p}{)}\PY{p}{:}
    \PY{k}{def} \PY{n+nf+fm}{\PYZus{}\PYZus{}init\PYZus{}\PYZus{}}\PY{p}{(}\PY{n+nb+bp}{self}\PY{p}{,} \PY{n}{unit\PYZus{}cost}\PY{p}{,} \PY{n}{selling\PYZus{}price}\PY{p}{,} \PY{n}{unit\PYZus{}refund}\PY{p}{,} \PY{n}{order\PYZus{}quantity}\PY{p}{,} \PY{n}{demand}\PY{p}{)}\PY{p}{:}
        \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{unit\PYZus{}cost} \PY{o}{=} \PY{n}{unit\PYZus{}cost}
        \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{selling\PYZus{}price} \PY{o}{=} \PY{n}{selling\PYZus{}price}
        \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{unit\PYZus{}refund} \PY{o}{=} \PY{n}{unit\PYZus{}refund}
        \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{order\PYZus{}quantity} \PY{o}{=} \PY{n}{order\PYZus{}quantity}
        \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{demand} \PY{o}{=} \PY{n}{demand}

    \PY{k}{def} \PY{n+nf}{order\PYZus{}cost}\PY{p}{(}\PY{n+nb+bp}{self}\PY{p}{)}\PY{p}{:}
        \PY{l+s+sd}{\PYZdq{}\PYZdq{}\PYZdq{}Compute total order cost\PYZdq{}\PYZdq{}\PYZdq{}}
        \PY{k}{return} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{unit\PYZus{}cost} \PY{o}{*} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{order\PYZus{}quantity}

    \PY{k}{def} \PY{n+nf}{sales\PYZus{}revenue}\PY{p}{(}\PY{n+nb+bp}{self}\PY{p}{)}\PY{p}{:}
        \PY{l+s+sd}{\PYZdq{}\PYZdq{}\PYZdq{}Compute sales revenue\PYZdq{}\PYZdq{}\PYZdq{}}
        \PY{k}{return} \PY{n}{np}\PY{o}{.}\PY{n}{minimum}\PY{p}{(}\PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{order\PYZus{}quantity}\PY{p}{,} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{demand}\PY{p}{)} \PY{o}{*} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{selling\PYZus{}price}

    \PY{k}{def} \PY{n+nf}{refund\PYZus{}revenue}\PY{p}{(}\PY{n+nb+bp}{self}\PY{p}{)}\PY{p}{:}
        \PY{l+s+sd}{\PYZdq{}\PYZdq{}\PYZdq{}Compute revenue from refunds for unsold items\PYZdq{}\PYZdq{}\PYZdq{}}
        \PY{k}{return} \PY{n}{np}\PY{o}{.}\PY{n}{maximum}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{order\PYZus{}quantity} \PY{o}{\PYZhy{}} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{demand}\PY{p}{)} \PY{o}{*} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{unit\PYZus{}refund}

    \PY{k}{def} \PY{n+nf}{profit}\PY{p}{(}\PY{n+nb+bp}{self}\PY{p}{)}\PY{p}{:}
        \PY{l+s+sd}{\PYZsq{}\PYZsq{}\PYZsq{}}
\PY{l+s+sd}{        Compute profit}
\PY{l+s+sd}{        \PYZsq{}\PYZsq{}\PYZsq{}}
        \PY{n}{profit} \PY{o}{=} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{sales\PYZus{}revenue}\PY{p}{(}\PY{p}{)} \PY{o}{+} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{refund\PYZus{}revenue}\PY{p}{(}\PY{p}{)} \PY{o}{\PYZhy{}} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{order\PYZus{}cost}\PY{p}{(}\PY{p}{)}
        \PY{k}{return} \PY{n}{profit}
\end{Verbatim}
\end{tcolorbox}

    A few important things to note: * we added four new object functions
(methods) * each method only takes the \texttt{self} object as an input
argument. We don't pass in the properties such as \texttt{unit\_cost} or
\texttt{order\_quantity}. We don't need to since \texttt{self} knows
about these things via its properties. They are \emph{encapsulated}
within the object. * within each method we have to use
\texttt{self.someproperty} in our expressions * the \texttt{profit}
method calls some of the other methods and when it does, it prefaces the
method with \texttt{self.} and follows it with empty parens. The
\texttt{self} object is implicitly passed. * each method returns a value
* each method has a simple docstring

Before we can test out our enhanced class, we need to run the code cell
above and we need to create a new object instance based on the updated
class definition.

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{30}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{c+c1}{\PYZsh{} Create a new BookstoreModel object}
\PY{n}{model\PYZus{}2} \PY{o}{=} \PY{n}{BookstoreModel}\PY{p}{(}\PY{n}{unit\PYZus{}cost}\PY{p}{,} \PY{n}{selling\PYZus{}price}\PY{p}{,} \PY{n}{unit\PYZus{}refund}\PY{p}{,} \PY{n}{order\PYZus{}quantity}\PY{p}{,} \PY{n}{demand}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{31}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n+nb}{print}\PY{p}{(}\PY{n}{model\PYZus{}2}\PY{o}{.}\PY{n}{sales\PYZus{}revenue}\PY{p}{(}\PY{p}{)}\PY{p}{)}
\PY{n+nb}{print}\PY{p}{(}\PY{n}{model\PYZus{}2}\PY{o}{.}\PY{n}{profit}\PY{p}{(}\PY{p}{)}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]
1930.0
447.5
    \end{Verbatim}

    It would nice to have a way to print out a succinct summary of a
BookstoreModel object. Let's start by just listing out the properties.
We can use the Python \texttt{vars} function to help. This StackOverflow
post has some good info on getting a dictionary of object properties:

https://stackoverflow.com/questions/61517/python-dictionary-from-an-objects-fields

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{32}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n+nb}{print}\PY{p}{(}\PY{n+nb}{vars}\PY{p}{(}\PY{n}{model\PYZus{}2}\PY{p}{)}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]
\{'unit\_cost': 7.5, 'selling\_price': 10.0, 'unit\_refund': 2.5, 'order\_quantity':
200, 'demand': 193\}
    \end{Verbatim}

    The Pythonic way for creating a string representation of an object is to
implement a \texttt{\_\_str\_\_} dunder function for our object. This
function needs to return a string. Notice that the \texttt{vars}
function returns a dictionary. You can learn more about
\texttt{\_\_str\_\_} and the related \texttt{\_\_repr\_\_} dunder
functions at https://dbader.org/blog/python-repr-vs-str. In a nutshell,
we use \texttt{\_\_str\_\_} to create ``pretty'' string representations
of an object for our user and \texttt{\_\_repr\_\_} for an unambiguous
string representation that includes the object type.

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{33}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{k}{class} \PY{n+nc}{BookstoreModel}\PY{p}{(}\PY{p}{)}\PY{p}{:}
    \PY{k}{def} \PY{n+nf+fm}{\PYZus{}\PYZus{}init\PYZus{}\PYZus{}}\PY{p}{(}\PY{n+nb+bp}{self}\PY{p}{,} \PY{n}{unit\PYZus{}cost}\PY{p}{,} \PY{n}{selling\PYZus{}price}\PY{p}{,} \PY{n}{unit\PYZus{}refund}\PY{p}{,} \PY{n}{order\PYZus{}quantity}\PY{p}{,} \PY{n}{demand}\PY{p}{)}\PY{p}{:}
        \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{unit\PYZus{}cost} \PY{o}{=} \PY{n}{unit\PYZus{}cost}
        \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{selling\PYZus{}price} \PY{o}{=} \PY{n}{selling\PYZus{}price}
        \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{unit\PYZus{}refund} \PY{o}{=} \PY{n}{unit\PYZus{}refund}
        \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{order\PYZus{}quantity} \PY{o}{=} \PY{n}{order\PYZus{}quantity}
        \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{demand} \PY{o}{=} \PY{n}{demand}

    \PY{k}{def} \PY{n+nf}{order\PYZus{}cost}\PY{p}{(}\PY{n+nb+bp}{self}\PY{p}{)}\PY{p}{:}
        \PY{l+s+sd}{\PYZdq{}\PYZdq{}\PYZdq{}Compute total order cost\PYZdq{}\PYZdq{}\PYZdq{}}
        \PY{k}{return} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{unit\PYZus{}cost} \PY{o}{*} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{order\PYZus{}quantity}

    \PY{k}{def} \PY{n+nf}{sales\PYZus{}revenue}\PY{p}{(}\PY{n+nb+bp}{self}\PY{p}{)}\PY{p}{:}
        \PY{l+s+sd}{\PYZdq{}\PYZdq{}\PYZdq{}Compute sales revenue\PYZdq{}\PYZdq{}\PYZdq{}}
        \PY{k}{return} \PY{n}{np}\PY{o}{.}\PY{n}{minimum}\PY{p}{(}\PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{order\PYZus{}quantity}\PY{p}{,} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{demand}\PY{p}{)} \PY{o}{*} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{selling\PYZus{}price}

    \PY{k}{def} \PY{n+nf}{refund\PYZus{}revenue}\PY{p}{(}\PY{n+nb+bp}{self}\PY{p}{)}\PY{p}{:}
        \PY{l+s+sd}{\PYZdq{}\PYZdq{}\PYZdq{}Compute revenue from refunds for unsold items\PYZdq{}\PYZdq{}\PYZdq{}}
        \PY{k}{return} \PY{n}{np}\PY{o}{.}\PY{n}{maximum}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{order\PYZus{}quantity} \PY{o}{\PYZhy{}} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{demand}\PY{p}{)} \PY{o}{*} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{unit\PYZus{}refund}

    \PY{k}{def} \PY{n+nf}{profit}\PY{p}{(}\PY{n+nb+bp}{self}\PY{p}{)}\PY{p}{:}
        \PY{l+s+sd}{\PYZsq{}\PYZsq{}\PYZsq{}}
\PY{l+s+sd}{        Compute profit}
\PY{l+s+sd}{        \PYZsq{}\PYZsq{}\PYZsq{}}
        \PY{n}{profit} \PY{o}{=} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{sales\PYZus{}revenue}\PY{p}{(}\PY{p}{)} \PY{o}{+} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{refund\PYZus{}revenue}\PY{p}{(}\PY{p}{)} \PY{o}{\PYZhy{}} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{order\PYZus{}cost}\PY{p}{(}\PY{p}{)}
        \PY{k}{return} \PY{n}{profit}

    \PY{k}{def} \PY{n+nf+fm}{\PYZus{}\PYZus{}str\PYZus{}\PYZus{}}\PY{p}{(}\PY{n+nb+bp}{self}\PY{p}{)}\PY{p}{:}
        \PY{l+s+sd}{\PYZdq{}\PYZdq{}\PYZdq{}}
\PY{l+s+sd}{        String representation of bookstore inputs}
\PY{l+s+sd}{        \PYZdq{}\PYZdq{}\PYZdq{}}
        \PY{k}{return} \PY{n+nb}{str}\PY{p}{(}\PY{n+nb}{vars}\PY{p}{(}\PY{n+nb+bp}{self}\PY{p}{)}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    Notice we wrapped \texttt{vars} with \texttt{str} to convert the
dictionary to a string.

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{34}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{c+c1}{\PYZsh{} Reset the base inputs}
\PY{n}{unit\PYZus{}cost} \PY{o}{=} \PY{l+m+mf}{7.50}
\PY{n}{selling\PYZus{}price} \PY{o}{=} \PY{l+m+mf}{10.00}
\PY{n}{unit\PYZus{}refund} \PY{o}{=} \PY{l+m+mf}{2.50}
\PY{n}{demand} \PY{o}{=} \PY{l+m+mi}{193}
\PY{n}{order\PYZus{}quantity} \PY{o}{=} \PY{l+m+mi}{200}

\PY{c+c1}{\PYZsh{} Create a new BookstoreModel object}
\PY{n}{model\PYZus{}3} \PY{o}{=} \PY{n}{BookstoreModel}\PY{p}{(}\PY{n}{unit\PYZus{}cost}\PY{p}{,} \PY{n}{selling\PYZus{}price}\PY{p}{,} \PY{n}{unit\PYZus{}refund}\PY{p}{,} \PY{n}{order\PYZus{}quantity}\PY{p}{,} \PY{n}{demand}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{35}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n+nb}{print}\PY{p}{(}\PY{n}{model\PYZus{}3}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]
\{'unit\_cost': 7.5, 'selling\_price': 10.0, 'unit\_refund': 2.5, 'order\_quantity':
200, 'demand': 193\}
    \end{Verbatim}

    \textbf{CHALLENGE 1} Modify the \texttt{\_\_str\_\_} function so that it
also shows the computed value of \texttt{order\_cost},
\texttt{sales\_revenue}, \texttt{refund\_revenue} and \texttt{profit}.

\textbf{CHALLENGE 2} Add a \texttt{lost\_sales} method to the
\texttt{BookstoreModel} class that returns the number of units that we
could not sell because we didn't order enough.

    \hypertarget{way-data-table-with-oo-model}{%
\subsubsection{1-way Data Table with OO
model}\label{way-data-table-with-oo-model}}

It seems entirely plausible that we could simply set one of the inputs
to an array of values and then call the \texttt{profit} method.

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{36}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{c+c1}{\PYZsh{} Set property equal to an array}
\PY{n}{model\PYZus{}3}\PY{o}{.}\PY{n}{order\PYZus{}quantity} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{arange}\PY{p}{(}\PY{l+m+mi}{50}\PY{p}{,} \PY{l+m+mi}{301}\PY{p}{,} \PY{l+m+mi}{25}\PY{p}{)}
\PY{n}{model\PYZus{}3}\PY{o}{.}\PY{n}{order\PYZus{}quantity}
\end{Verbatim}
\end{tcolorbox}

            \begin{tcolorbox}[breakable, size=fbox, boxrule=.5pt, pad at break*=1mm, opacityfill=0]
\prompt{Out}{outcolor}{36}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
array([ 50,  75, 100, 125, 150, 175, 200, 225, 250, 275, 300])
\end{Verbatim}
\end{tcolorbox}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{37}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{c+c1}{\PYZsh{} Can we compute an array of profits? Try it.}
\end{Verbatim}
\end{tcolorbox}

    What about a 2-way data table using the OO model? We can't use the same
direct array passing approach because of \ldots{}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{38}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{c+c1}{\PYZsh{} This won\PYZsq{}t work unless the two arrays are the same size \PYZhy{} can\PYZsq{}t broadcast.}
\PY{c+c1}{\PYZsh{} Also, even if same size, doesn\PYZsq{}t do all combinations, just aligned elements.}
\PY{n}{demand\PYZus{}range} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{arange}\PY{p}{(}\PY{l+m+mi}{70}\PY{p}{,} \PY{l+m+mi}{321}\PY{p}{,} \PY{l+m+mi}{5}\PY{p}{)}
\PY{n}{order\PYZus{}quantity\PYZus{}range} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{arange}\PY{p}{(}\PY{l+m+mi}{70}\PY{p}{,} \PY{l+m+mi}{321}\PY{p}{,} \PY{l+m+mi}{10}\PY{p}{)}    \PY{c+c1}{\PYZsh{} Broadcast error}
\PY{c+c1}{\PYZsh{} order\PYZus{}quantity\PYZus{}range = np.arange(70, 321, 5)   \PYZsh{} Doesn\PYZsq{}t do all combinations}

\PY{c+c1}{\PYZsh{} Method 1: Set property equal to an array}
\PY{n}{model\PYZus{}3}\PY{o}{.}\PY{n}{demand} \PY{o}{=} \PY{n}{demand\PYZus{}range}
\PY{n}{model\PYZus{}3}\PY{o}{.}\PY{n}{order\PYZus{}quantity} \PY{o}{=} \PY{n}{order\PYZus{}quantity\PYZus{}range}

\PY{c+c1}{\PYZsh{} model\PYZus{}3.profit()  \PYZsh{} This line will trigger the broadcast error}
\end{Verbatim}
\end{tcolorbox}

    Recall that for the non-OO model, we could easily do an n-way data table
using a list comprehension that made use of the
\texttt{bookstore\_profit} function because we could explicitly loop
over ranges of any subset of the input variables and call the
\texttt{bookstore\_profit} function with all of the input arguments
specified. We can't do that with the OO model because the
\texttt{profit} function is a method of the object and we don't pass it
any input arguments other than \texttt{self}. So, we need a way to loop
over all the input ranges and update the bookstore model object's input
attributes and then call the \texttt{profit} method.

What to do?

    \hypertarget{a-decision-point-in-our-software-design-process}{%
\subsubsection{A decision point in our software design
process}\label{a-decision-point-in-our-software-design-process}}

An end user modeler analagous to your basic Excel power user isn't going
to write OO code to do a Data Table. The list comprehension approach
based on a global function (i.e.~the non-OO model) is much more likely.
However, a relatively generic Data Table function could be implemented
in an OO way and be quite reusable. Also, we still want to implement
Goal Seek and it's not clear whether the OO or non-OO approach makes the
most sense.

Well, let's forge ahead and design and create a Python based Data Table
function that accepts an OO BookstoreModel object as one of its inputs.
Surely, we'll learn some things along the way.

    \hypertarget{adding-an-update-method}{%
\paragraph{\texorpdfstring{Adding an \texttt{update}
method}{Adding an update method}}\label{adding-an-update-method}}

Since were are going to need to repeatedly update some subset of the
input parameters, let's add an \texttt{update} method to the
BookstoreModel class that takes a dictionary of property, value pairs.
The question is, \textbf{how do we update object attributes from a
dictionary specifying the attribute and the new value?}

This StackoverFlow post was quite helpful:

https://stackoverflow.com/questions/2466191/set-attributes-from-dictionary-in-python

The idea is that we use Python's \texttt{setattr} function. Using
\href{https://www.w3schools.com/python/ref_func_setattr.asp}{\texttt{setattr}}
is really easy:

\begin{verbatim}
setattr(object, attribute name, attribute value)
\end{verbatim}

As a specific example, it can be used like this.

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{39}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{model\PYZus{}4} \PY{o}{=} \PY{n}{BookstoreModel}\PY{p}{(}\PY{n}{unit\PYZus{}cost}\PY{p}{,} \PY{n}{selling\PYZus{}price}\PY{p}{,} \PY{n}{unit\PYZus{}refund}\PY{p}{,} \PY{n}{order\PYZus{}quantity}\PY{p}{,} \PY{n}{demand}\PY{p}{)}
\PY{n+nb}{print}\PY{p}{(}\PY{n}{model\PYZus{}4}\PY{p}{)}

\PY{c+c1}{\PYZsh{} We want to update a existing model object with these parameter values}
\PY{n}{new\PYZus{}params} \PY{o}{=} \PY{p}{\PYZob{}}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{unit\PYZus{}cost}\PY{l+s+s1}{\PYZsq{}}\PY{p}{:} \PY{l+m+mf}{8.5}\PY{p}{,} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{order\PYZus{}quantity}\PY{l+s+s1}{\PYZsq{}}\PY{p}{:} \PY{l+m+mi}{250}\PY{p}{\PYZcb{}}

\PY{c+c1}{\PYZsh{} Iterate over the keys in new\PYZus{}params}
\PY{k}{for} \PY{n}{key} \PY{o+ow}{in} \PY{n}{new\PYZus{}params}\PY{p}{:}
            \PY{n+nb}{setattr}\PY{p}{(}\PY{n}{model\PYZus{}4}\PY{p}{,} \PY{n}{key}\PY{p}{,} \PY{n}{new\PYZus{}params}\PY{p}{[}\PY{n}{key}\PY{p}{]}\PY{p}{)}

\PY{n+nb}{print}\PY{p}{(}\PY{n}{model\PYZus{}4}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]
\{'unit\_cost': 7.5, 'selling\_price': 10.0, 'unit\_refund': 2.5, 'order\_quantity':
200, 'demand': 193\}
\{'unit\_cost': 8.5, 'selling\_price': 10.0, 'unit\_refund': 2.5, 'order\_quantity':
250, 'demand': 193\}
    \end{Verbatim}

    Okay, here's our updated BookstoreModel class. Note that I've changed to
keyword arguments with default values.

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{40}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{k}{class} \PY{n+nc}{BookstoreModel}\PY{p}{(}\PY{p}{)}\PY{p}{:}
    \PY{k}{def} \PY{n+nf+fm}{\PYZus{}\PYZus{}init\PYZus{}\PYZus{}}\PY{p}{(}\PY{n+nb+bp}{self}\PY{p}{,} \PY{n}{unit\PYZus{}cost}\PY{o}{=}\PY{l+m+mi}{0}\PY{p}{,} \PY{n}{selling\PYZus{}price}\PY{o}{=}\PY{l+m+mi}{0}\PY{p}{,} \PY{n}{unit\PYZus{}refund}\PY{o}{=}\PY{l+m+mi}{0}\PY{p}{,} 
                 \PY{n}{order\PYZus{}quantity}\PY{o}{=}\PY{l+m+mi}{0}\PY{p}{,} \PY{n}{demand}\PY{o}{=}\PY{l+m+mi}{0}\PY{p}{)}\PY{p}{:}
        \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{unit\PYZus{}cost} \PY{o}{=} \PY{n}{unit\PYZus{}cost}
        \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{selling\PYZus{}price} \PY{o}{=} \PY{n}{selling\PYZus{}price}
        \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{unit\PYZus{}refund} \PY{o}{=} \PY{n}{unit\PYZus{}refund}
        \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{order\PYZus{}quantity} \PY{o}{=} \PY{n}{order\PYZus{}quantity}
        \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{demand} \PY{o}{=} \PY{n}{demand}

    \PY{k}{def} \PY{n+nf}{update}\PY{p}{(}\PY{n+nb+bp}{self}\PY{p}{,} \PY{n}{param\PYZus{}dict}\PY{p}{)}\PY{p}{:}
        \PY{l+s+sd}{\PYZdq{}\PYZdq{}\PYZdq{}}
\PY{l+s+sd}{        Update parameter values}
\PY{l+s+sd}{        \PYZdq{}\PYZdq{}\PYZdq{}}
        \PY{k}{for} \PY{n}{key} \PY{o+ow}{in} \PY{n}{param\PYZus{}dict}\PY{p}{:}
            \PY{n+nb}{setattr}\PY{p}{(}\PY{n+nb+bp}{self}\PY{p}{,} \PY{n}{key}\PY{p}{,} \PY{n}{param\PYZus{}dict}\PY{p}{[}\PY{n}{key}\PY{p}{]}\PY{p}{)}

    \PY{k}{def} \PY{n+nf}{order\PYZus{}cost}\PY{p}{(}\PY{n+nb+bp}{self}\PY{p}{)}\PY{p}{:}
        \PY{l+s+sd}{\PYZdq{}\PYZdq{}\PYZdq{}Compute total order cost\PYZdq{}\PYZdq{}\PYZdq{}}
        \PY{k}{return} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{unit\PYZus{}cost} \PY{o}{*} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{order\PYZus{}quantity}

    \PY{k}{def} \PY{n+nf}{sales\PYZus{}revenue}\PY{p}{(}\PY{n+nb+bp}{self}\PY{p}{)}\PY{p}{:}
        \PY{l+s+sd}{\PYZdq{}\PYZdq{}\PYZdq{}Compute sales revenue\PYZdq{}\PYZdq{}\PYZdq{}}
        \PY{k}{return} \PY{n}{np}\PY{o}{.}\PY{n}{minimum}\PY{p}{(}\PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{order\PYZus{}quantity}\PY{p}{,} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{demand}\PY{p}{)} \PY{o}{*} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{selling\PYZus{}price}

    \PY{k}{def} \PY{n+nf}{refund\PYZus{}revenue}\PY{p}{(}\PY{n+nb+bp}{self}\PY{p}{)}\PY{p}{:}
        \PY{l+s+sd}{\PYZdq{}\PYZdq{}\PYZdq{}Compute revenue from refunds for unsold items\PYZdq{}\PYZdq{}\PYZdq{}}
        \PY{k}{return} \PY{n}{np}\PY{o}{.}\PY{n}{maximum}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{order\PYZus{}quantity} \PY{o}{\PYZhy{}} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{demand}\PY{p}{)} \PY{o}{*} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{unit\PYZus{}refund}

    \PY{k}{def} \PY{n+nf}{profit}\PY{p}{(}\PY{n+nb+bp}{self}\PY{p}{)}\PY{p}{:}
        \PY{l+s+sd}{\PYZsq{}\PYZsq{}\PYZsq{}}
\PY{l+s+sd}{        Compute profit in bookstore model}
\PY{l+s+sd}{        \PYZsq{}\PYZsq{}\PYZsq{}}
        \PY{n}{profit} \PY{o}{=} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{sales\PYZus{}revenue}\PY{p}{(}\PY{p}{)} \PY{o}{+} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{refund\PYZus{}revenue}\PY{p}{(}\PY{p}{)} \PY{o}{\PYZhy{}} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{order\PYZus{}cost}\PY{p}{(}\PY{p}{)}
        \PY{k}{return} \PY{n}{profit}

    \PY{k}{def} \PY{n+nf+fm}{\PYZus{}\PYZus{}str\PYZus{}\PYZus{}}\PY{p}{(}\PY{n+nb+bp}{self}\PY{p}{)}\PY{p}{:}
        \PY{l+s+sd}{\PYZdq{}\PYZdq{}\PYZdq{}}
\PY{l+s+sd}{        Print dictionary of object attributes but don\PYZsq{}t include an underscore as first char}
\PY{l+s+sd}{        \PYZdq{}\PYZdq{}\PYZdq{}}
        \PY{k}{return} \PY{n+nb}{str}\PY{p}{(}\PY{n+nb}{vars}\PY{p}{(}\PY{n+nb+bp}{self}\PY{p}{)}\PY{p}{)}
        \PY{c+c1}{\PYZsh{}return str(\PYZob{}key: val for (key, val) in vars(self).items() if key[0] != \PYZsq{}\PYZus{}\PYZsq{}\PYZcb{})}
\end{Verbatim}
\end{tcolorbox}

    Let's try out our \texttt{update} method. Also, since now the
BookstoreModel class uses keyword arguments with defaults, I'm going to
create a new model object in a slightly different way:

\begin{itemize}
\tightlist
\item
  create a new model object and just take the zero default values for
  the input attributes
\item
  use the \texttt{update} method to set the input attributes via a
  dictionary
\end{itemize}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{41}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{base\PYZus{}inputs} \PY{o}{=} \PY{p}{\PYZob{}}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{unit\PYZus{}cost}\PY{l+s+s1}{\PYZsq{}}\PY{p}{:} \PY{l+m+mf}{7.50}\PY{p}{,}
              \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{selling\PYZus{}price}\PY{l+s+s1}{\PYZsq{}}\PY{p}{:} \PY{l+m+mf}{10.00}\PY{p}{,}
              \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{unit\PYZus{}refund}\PY{l+s+s1}{\PYZsq{}}\PY{p}{:} \PY{l+m+mf}{2.50}\PY{p}{,}
              \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{order\PYZus{}quantity}\PY{l+s+s1}{\PYZsq{}}\PY{p}{:} \PY{l+m+mi}{200}\PY{p}{,}
              \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{demand}\PY{l+s+s1}{\PYZsq{}}\PY{p}{:} \PY{l+m+mi}{193}\PY{p}{\PYZcb{}}

\PY{n}{model\PYZus{}5} \PY{o}{=} \PY{n}{BookstoreModel}\PY{p}{(}\PY{p}{)}
\PY{n}{model\PYZus{}5}\PY{o}{.}\PY{n}{update}\PY{p}{(}\PY{n}{base\PYZus{}inputs}\PY{p}{)}
\PY{n+nb}{print}\PY{p}{(}\PY{n}{model\PYZus{}5}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]
\{'unit\_cost': 7.5, 'selling\_price': 10.0, 'unit\_refund': 2.5, 'order\_quantity':
200, 'demand': 193\}
    \end{Verbatim}

    \hypertarget{how-to-generate-and-iterate-over-scenarios-to-create-a-2-way-data-table}{%
\paragraph{How to generate and iterate over scenarios to create a 2-way
Data
Table?}\label{how-to-generate-and-iterate-over-scenarios-to-create-a-2-way-data-table}}

Let's regroup and remind ourselves where we are in the development
process for our 2-way data table function. We've got two input
variables, \texttt{demand} and \texttt{order\_quantity} and we'd like to
specify a range over which each variable should vary. Then, we want to
consider all combinations of these variable values and compute the
associated, say, profit (really, any ouput of our model). It would be
convenient to be be able to specify our input ranges with a dictionary
whose keys are the input variable names and the values are any valid
Python iterable (such as a list or the output of some function that
returns a range of values). For example:

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{42}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{dt\PYZus{}param\PYZus{}ranges} \PY{o}{=} \PY{p}{\PYZob{}}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{demand}\PY{l+s+s1}{\PYZsq{}}\PY{p}{:} \PY{n}{np}\PY{o}{.}\PY{n}{arange}\PY{p}{(}\PY{l+m+mi}{70}\PY{p}{,} \PY{l+m+mi}{321}\PY{p}{,} \PY{l+m+mi}{25}\PY{p}{)}\PY{p}{,}
                   \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{order\PYZus{}quantity}\PY{l+s+s1}{\PYZsq{}}\PY{p}{:} \PY{n}{np}\PY{o}{.}\PY{n}{arange}\PY{p}{(}\PY{l+m+mi}{70}\PY{p}{,} \PY{l+m+mi}{321}\PY{p}{,} \PY{l+m+mi}{50}\PY{p}{)}\PY{p}{\PYZcb{}}
\end{Verbatim}
\end{tcolorbox}

    Our immediate goal is to convert this dictionary into a list of
dictionaries - one dictionary per scenario. Like this:

\begin{verbatim}
[{'demand': 70, 'order_quantity': 70},
 {'demand': 70, 'order_quantity': 120},
 {'demand': 70, 'order_quantity': 170},
 {'demand': 70, 'order_quantity': 220},
 {'demand': 70, 'order_quantity': 270},
 {'demand': 70, 'order_quantity': 320},
 {'demand': 95, 'order_quantity': 70},
 {'demand': 95, 'order_quantity': 120},
 ...
 ...
 ...
 {'demand': 320, 'order_quantity': 220},
 {'demand': 320, 'order_quantity': 270},
 {'demand': 320, 'order_quantity': 320}]

\end{verbatim}

Each of these component dictionaries is something we can use with our
new model object \texttt{update} method to set the corresponding values
for each scenario. Hmmm, feels like some clever iterating and zipping
and more iterating. Before doing that ourselves, can we do this using
the functionality of some existing Python library? Spoiler alert, yes,
we can.

    \hypertarget{learning-from-the-ideas-and-code-from-scikit-learn}{%
\paragraph{Learning from the ideas and code from
scikit-learn}\label{learning-from-the-ideas-and-code-from-scikit-learn}}

\href{https://scikit-learn.org/stable/index.html}{Scikit-learn} is an
extremely popular Python package for building, testing and using machine
learning models. In thinking about the input values for an n-way data
table, it feels similar to a
\href{https://scikit-learn.org/stable/modules/grid_search.html}{parameter
grid for a hyperparameter grid search in scikit learn}. Can we borrow
some of their patterns or code ideas?

https://scikit-learn.org/stable/modules/generated/sklearn.model\_selection.GridSearchCV.html\#sklearn.model\_selection.GridSearchCV

https://scikit-learn.org/stable/modules/grid\_search.html\#grid-search

Example from User Guide shows passing in two different grids to explore
as a list of dicts. Notice that each dict is like the setup for an Excel
Data Table - the key is just the name of some variable and the value is
some iterable. Don't worry about the details, just look at the structure
of \texttt{param\_grid}.

\begin{verbatim}
param_grid = [
  {'C': [1, 10, 100, 1000], 'kernel': ['linear']},
  {'C': [1, 10, 100, 1000], 'gamma': [0.001, 0.0001], 'kernel': ['rbf']},
 ]

\end{verbatim}

For example, the \texttt{GridSearchCV} takes an estimator object and a
parameter grid object. The parameter grid is a dict or list of dicts in
which the keys are the parameter names. SKLearn implements a
\texttt{ParameterGrid} class. Looking at its source, the docstring
contains the following example. Notice that the result of
\texttt{ParameterGrid(param\_grid)} is a list of dicts in which each
dict contains a specific combination of the parameter values, i.e., it
generates all the scenarios to explore.

\begin{verbatim}
Examples
--------
>>> from sklearn.model_selection import ParameterGrid
>>> param_grid = {'a': [1, 2], 'b': [True, False]}
>>> list(ParameterGrid(param_grid)) == (
...    [{'a': 1, 'b': True}, {'a': 1, 'b': False},
...     {'a': 2, 'b': True}, {'a': 2, 'b': False}])
True
\end{verbatim}

Hmm, we could use this class directly.

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{43}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{k+kn}{from} \PY{n+nn}{sklearn}\PY{n+nn}{.}\PY{n+nn}{model\PYZus{}selection}\PY{n+nn}{.}\PY{n+nn}{\PYZus{}search} \PY{k+kn}{import} \PY{n}{ParameterGrid}
\end{Verbatim}
\end{tcolorbox}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{44}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{dt\PYZus{}param\PYZus{}ranges} \PY{o}{=} \PY{p}{\PYZob{}}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{demand}\PY{l+s+s1}{\PYZsq{}}\PY{p}{:} \PY{n}{np}\PY{o}{.}\PY{n}{arange}\PY{p}{(}\PY{l+m+mi}{70}\PY{p}{,} \PY{l+m+mi}{321}\PY{p}{,} \PY{l+m+mi}{25}\PY{p}{)}\PY{p}{,}
                   \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{order\PYZus{}quantity}\PY{l+s+s1}{\PYZsq{}}\PY{p}{:} \PY{n}{np}\PY{o}{.}\PY{n}{arange}\PY{p}{(}\PY{l+m+mi}{70}\PY{p}{,} \PY{l+m+mi}{321}\PY{p}{,} \PY{l+m+mi}{50}\PY{p}{)}\PY{p}{\PYZcb{}}
\end{Verbatim}
\end{tcolorbox}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{45}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{dt\PYZus{}param\PYZus{}grid} \PY{o}{=} \PY{n+nb}{list}\PY{p}{(}\PY{n}{ParameterGrid}\PY{p}{(}\PY{n}{dt\PYZus{}param\PYZus{}ranges}\PY{p}{)}\PY{p}{)}
\PY{n+nb}{print}\PY{p}{(}\PY{n}{dt\PYZus{}param\PYZus{}grid}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]
[\{'demand': 70, 'order\_quantity': 70\}, \{'demand': 70, 'order\_quantity': 120\},
\{'demand': 70, 'order\_quantity': 170\}, \{'demand': 70, 'order\_quantity': 220\},
\{'demand': 70, 'order\_quantity': 270\}, \{'demand': 70, 'order\_quantity': 320\},
\{'demand': 95, 'order\_quantity': 70\}, \{'demand': 95, 'order\_quantity': 120\},
\{'demand': 95, 'order\_quantity': 170\}, \{'demand': 95, 'order\_quantity': 220\},
\{'demand': 95, 'order\_quantity': 270\}, \{'demand': 95, 'order\_quantity': 320\},
\{'demand': 120, 'order\_quantity': 70\}, \{'demand': 120, 'order\_quantity': 120\},
\{'demand': 120, 'order\_quantity': 170\}, \{'demand': 120, 'order\_quantity': 220\},
\{'demand': 120, 'order\_quantity': 270\}, \{'demand': 120, 'order\_quantity': 320\},
\{'demand': 145, 'order\_quantity': 70\}, \{'demand': 145, 'order\_quantity': 120\},
\{'demand': 145, 'order\_quantity': 170\}, \{'demand': 145, 'order\_quantity': 220\},
\{'demand': 145, 'order\_quantity': 270\}, \{'demand': 145, 'order\_quantity': 320\},
\{'demand': 170, 'order\_quantity': 70\}, \{'demand': 170, 'order\_quantity': 120\},
\{'demand': 170, 'order\_quantity': 170\}, \{'demand': 170, 'order\_quantity': 220\},
\{'demand': 170, 'order\_quantity': 270\}, \{'demand': 170, 'order\_quantity': 320\},
\{'demand': 195, 'order\_quantity': 70\}, \{'demand': 195, 'order\_quantity': 120\},
\{'demand': 195, 'order\_quantity': 170\}, \{'demand': 195, 'order\_quantity': 220\},
\{'demand': 195, 'order\_quantity': 270\}, \{'demand': 195, 'order\_quantity': 320\},
\{'demand': 220, 'order\_quantity': 70\}, \{'demand': 220, 'order\_quantity': 120\},
\{'demand': 220, 'order\_quantity': 170\}, \{'demand': 220, 'order\_quantity': 220\},
\{'demand': 220, 'order\_quantity': 270\}, \{'demand': 220, 'order\_quantity': 320\},
\{'demand': 245, 'order\_quantity': 70\}, \{'demand': 245, 'order\_quantity': 120\},
\{'demand': 245, 'order\_quantity': 170\}, \{'demand': 245, 'order\_quantity': 220\},
\{'demand': 245, 'order\_quantity': 270\}, \{'demand': 245, 'order\_quantity': 320\},
\{'demand': 270, 'order\_quantity': 70\}, \{'demand': 270, 'order\_quantity': 120\},
\{'demand': 270, 'order\_quantity': 170\}, \{'demand': 270, 'order\_quantity': 220\},
\{'demand': 270, 'order\_quantity': 270\}, \{'demand': 270, 'order\_quantity': 320\},
\{'demand': 295, 'order\_quantity': 70\}, \{'demand': 295, 'order\_quantity': 120\},
\{'demand': 295, 'order\_quantity': 170\}, \{'demand': 295, 'order\_quantity': 220\},
\{'demand': 295, 'order\_quantity': 270\}, \{'demand': 295, 'order\_quantity': 320\},
\{'demand': 320, 'order\_quantity': 70\}, \{'demand': 320, 'order\_quantity': 120\},
\{'demand': 320, 'order\_quantity': 170\}, \{'demand': 320, 'order\_quantity': 220\},
\{'demand': 320, 'order\_quantity': 270\}, \{'demand': 320, 'order\_quantity': 320\}]
    \end{Verbatim}

    Perfect! We have a list of dictionaries, each of which contains one of
the combinations of the \texttt{demand} and \texttt{order\_quantity}
attributes that we want in our data table. Isn't it nice that we just
implemented an \texttt{update} method that takes just such a dictionary
as its input. Almost like we planned it. :)

While there's nothing wrong with leveraging the skikit-learn library for
this task (after all, reusing objects is one of the whole reasons for
OOP), we really don't need all the generality provided by its
\texttt{ParameterGrid} class. We just need to do some careful iteration
through our dictionary of input ranges to create the same output. It
will give us a chance to use the very useful
\href{https://docs.python.org/3/library/itertools.html}{itertools
library} and see Python's \texttt{zip} function in action.

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{46}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{k+kn}{from} \PY{n+nn}{itertools} \PY{k+kn}{import} \PY{n}{product}
\end{Verbatim}
\end{tcolorbox}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{47}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{c+c1}{\PYZsh{} Look at the key, value pairs (tuples) in dt\PYZus{}param\PYZus{}ranges dictionary}
\PY{n}{dt\PYZus{}param\PYZus{}ranges}\PY{o}{.}\PY{n}{items}\PY{p}{(}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

            \begin{tcolorbox}[breakable, size=fbox, boxrule=.5pt, pad at break*=1mm, opacityfill=0]
\prompt{Out}{outcolor}{47}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
dict\_items([('demand', array([ 70,  95, 120, 145, 170, 195, 220, 245, 270, 295,
320])), ('order\_quantity', array([ 70, 120, 170, 220, 270, 320]))])
\end{Verbatim}
\end{tcolorbox}

    Now we can use
\href{https://docs.python.org/3.6/tutorial/controlflow.html\#unpacking-argument-lists}{unpacking}
along with the
\href{https://www.w3schools.com/python/ref_func_zip.asp}{zip function}
to get two tuples. One tuple will get stored in the variable
\texttt{keys} and the other in \texttt{values}.

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{48}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n+nb}{print}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{The original tuples}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
\PY{n+nb}{print}\PY{p}{(}\PY{o}{*}\PY{n}{dt\PYZus{}param\PYZus{}ranges}\PY{o}{.}\PY{n}{items}\PY{p}{(}\PY{p}{)}\PY{p}{)}
\PY{n+nb}{print}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+se}{\PYZbs{}n}\PY{l+s+s2}{The zipped tuples}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
\PY{n}{keys}\PY{p}{,} \PY{n}{values} \PY{o}{=} \PY{n+nb}{zip}\PY{p}{(}\PY{o}{*}\PY{n}{dt\PYZus{}param\PYZus{}ranges}\PY{o}{.}\PY{n}{items}\PY{p}{(}\PY{p}{)}\PY{p}{)}
\PY{n+nb}{print}\PY{p}{(}\PY{n}{keys}\PY{p}{,} \PY{n}{values}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]
The original tuples
('demand', array([ 70,  95, 120, 145, 170, 195, 220, 245, 270, 295, 320]))
('order\_quantity', array([ 70, 120, 170, 220, 270, 320]))

The zipped tuples
('demand', 'order\_quantity') (array([ 70,  95, 120, 145, 170, 195, 220, 245,
270, 295, 320]), array([ 70, 120, 170, 220, 270, 320]))
    \end{Verbatim}

    Now we have a tuple of the keys,
\texttt{(\textquotesingle{}demand\textquotesingle{},\ \textquotesingle{}order\_quantity\textquotesingle{})}
and a tuple containing the two arrays of associated values.

What we need next is a way to get all the combinations of the values in
the two arrays. This is a perfect job for the \texttt{itertools.product}
function. Here's a simple example of what \texttt{product} does.

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{49}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{colors} \PY{o}{=} \PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{red}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{blue}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{green}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}
\PY{n}{intensities} \PY{o}{=} \PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{pale}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{bright}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}
\PY{k}{for} \PY{n}{pair} \PY{o+ow}{in} \PY{n}{product}\PY{p}{(}\PY{n}{intensities}\PY{p}{,} \PY{n}{colors}\PY{p}{)}\PY{p}{:}
    \PY{n+nb}{print}\PY{p}{(}\PY{n}{pair}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]
('pale', 'red')
('pale', 'blue')
('pale', 'green')
('bright', 'red')
('bright', 'blue')
('bright', 'green')
    \end{Verbatim}

    The \texttt{values} variable is a tuple (that needs unpacking)
containing the two arrays for which we want the \texttt{product}. I'll
turn the result into a list so we can see it. Try it without the list to
see what you get.

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{50}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n+nb}{print}\PY{p}{(}\PY{n}{values}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]
(array([ 70,  95, 120, 145, 170, 195, 220, 245, 270, 295, 320]), array([ 70,
120, 170, 220, 270, 320]))
    \end{Verbatim}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{51}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n+nb}{print}\PY{p}{(}\PY{n+nb}{list}\PY{p}{(}\PY{n}{product}\PY{p}{(}\PY{o}{*}\PY{n}{values}\PY{p}{)}\PY{p}{)}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]
[(70, 70), (70, 120), (70, 170), (70, 220), (70, 270), (70, 320), (95, 70), (95,
120), (95, 170), (95, 220), (95, 270), (95, 320), (120, 70), (120, 120), (120,
170), (120, 220), (120, 270), (120, 320), (145, 70), (145, 120), (145, 170),
(145, 220), (145, 270), (145, 320), (170, 70), (170, 120), (170, 170), (170,
220), (170, 270), (170, 320), (195, 70), (195, 120), (195, 170), (195, 220),
(195, 270), (195, 320), (220, 70), (220, 120), (220, 170), (220, 220), (220,
270), (220, 320), (245, 70), (245, 120), (245, 170), (245, 220), (245, 270),
(245, 320), (270, 70), (270, 120), (270, 170), (270, 220), (270, 270), (270,
320), (295, 70), (295, 120), (295, 170), (295, 220), (295, 270), (295, 320),
(320, 70), (320, 120), (320, 170), (320, 220), (320, 270), (320, 320)]
    \end{Verbatim}

    Finally, we need to zip each of these tuples with the \texttt{keys}
variable and turn the resulting tuple into a dictionary. Let's recall
what it contains:

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{52}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{keys}
\end{Verbatim}
\end{tcolorbox}

            \begin{tcolorbox}[breakable, size=fbox, boxrule=.5pt, pad at break*=1mm, opacityfill=0]
\prompt{Out}{outcolor}{52}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
('demand', 'order\_quantity')
\end{Verbatim}
\end{tcolorbox}

    The \texttt{zip} function works like a zipper and pairs elements in the
same position in our two tuples.

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{53}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n+nb}{list}\PY{p}{(}\PY{n+nb}{zip}\PY{p}{(}\PY{n}{keys}\PY{p}{,} \PY{p}{(}\PY{l+m+mi}{70}\PY{p}{,} \PY{l+m+mi}{120}\PY{p}{)}\PY{p}{)}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

            \begin{tcolorbox}[breakable, size=fbox, boxrule=.5pt, pad at break*=1mm, opacityfill=0]
\prompt{Out}{outcolor}{53}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
[('demand', 70), ('order\_quantity', 120)]
\end{Verbatim}
\end{tcolorbox}

    \ldots{} and it's easy to turn this into a dictionary (which is what we
want).

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{54}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n+nb}{dict}\PY{p}{(}\PY{n+nb}{zip}\PY{p}{(}\PY{n}{keys}\PY{p}{,} \PY{p}{(}\PY{l+m+mi}{70}\PY{p}{,} \PY{l+m+mi}{120}\PY{p}{)}\PY{p}{)}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

            \begin{tcolorbox}[breakable, size=fbox, boxrule=.5pt, pad at break*=1mm, opacityfill=0]
\prompt{Out}{outcolor}{54}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\{'demand': 70, 'order\_quantity': 120\}
\end{Verbatim}
\end{tcolorbox}

    Ok, just put all these ideas together.

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{55}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{param\PYZus{}grid} \PY{o}{=} \PY{p}{[}\PY{p}{]}

\PY{n}{keys}\PY{p}{,} \PY{n}{values} \PY{o}{=} \PY{n+nb}{zip}\PY{p}{(}\PY{o}{*}\PY{n}{dt\PYZus{}param\PYZus{}ranges}\PY{o}{.}\PY{n}{items}\PY{p}{(}\PY{p}{)}\PY{p}{)}

\PY{k}{for} \PY{n}{scenario} \PY{o+ow}{in} \PY{n}{product}\PY{p}{(}\PY{o}{*}\PY{n}{values}\PY{p}{)}\PY{p}{:}
    \PY{n}{params} \PY{o}{=} \PY{n+nb}{dict}\PY{p}{(}\PY{n+nb}{zip}\PY{p}{(}\PY{n}{keys}\PY{p}{,} \PY{n}{scenario}\PY{p}{)}\PY{p}{)}
    \PY{n}{param\PYZus{}grid}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{params}\PY{p}{)}

\PY{n+nb}{print}\PY{p}{(}\PY{n}{param\PYZus{}grid}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]
[\{'demand': 70, 'order\_quantity': 70\}, \{'demand': 70, 'order\_quantity': 120\},
\{'demand': 70, 'order\_quantity': 170\}, \{'demand': 70, 'order\_quantity': 220\},
\{'demand': 70, 'order\_quantity': 270\}, \{'demand': 70, 'order\_quantity': 320\},
\{'demand': 95, 'order\_quantity': 70\}, \{'demand': 95, 'order\_quantity': 120\},
\{'demand': 95, 'order\_quantity': 170\}, \{'demand': 95, 'order\_quantity': 220\},
\{'demand': 95, 'order\_quantity': 270\}, \{'demand': 95, 'order\_quantity': 320\},
\{'demand': 120, 'order\_quantity': 70\}, \{'demand': 120, 'order\_quantity': 120\},
\{'demand': 120, 'order\_quantity': 170\}, \{'demand': 120, 'order\_quantity': 220\},
\{'demand': 120, 'order\_quantity': 270\}, \{'demand': 120, 'order\_quantity': 320\},
\{'demand': 145, 'order\_quantity': 70\}, \{'demand': 145, 'order\_quantity': 120\},
\{'demand': 145, 'order\_quantity': 170\}, \{'demand': 145, 'order\_quantity': 220\},
\{'demand': 145, 'order\_quantity': 270\}, \{'demand': 145, 'order\_quantity': 320\},
\{'demand': 170, 'order\_quantity': 70\}, \{'demand': 170, 'order\_quantity': 120\},
\{'demand': 170, 'order\_quantity': 170\}, \{'demand': 170, 'order\_quantity': 220\},
\{'demand': 170, 'order\_quantity': 270\}, \{'demand': 170, 'order\_quantity': 320\},
\{'demand': 195, 'order\_quantity': 70\}, \{'demand': 195, 'order\_quantity': 120\},
\{'demand': 195, 'order\_quantity': 170\}, \{'demand': 195, 'order\_quantity': 220\},
\{'demand': 195, 'order\_quantity': 270\}, \{'demand': 195, 'order\_quantity': 320\},
\{'demand': 220, 'order\_quantity': 70\}, \{'demand': 220, 'order\_quantity': 120\},
\{'demand': 220, 'order\_quantity': 170\}, \{'demand': 220, 'order\_quantity': 220\},
\{'demand': 220, 'order\_quantity': 270\}, \{'demand': 220, 'order\_quantity': 320\},
\{'demand': 245, 'order\_quantity': 70\}, \{'demand': 245, 'order\_quantity': 120\},
\{'demand': 245, 'order\_quantity': 170\}, \{'demand': 245, 'order\_quantity': 220\},
\{'demand': 245, 'order\_quantity': 270\}, \{'demand': 245, 'order\_quantity': 320\},
\{'demand': 270, 'order\_quantity': 70\}, \{'demand': 270, 'order\_quantity': 120\},
\{'demand': 270, 'order\_quantity': 170\}, \{'demand': 270, 'order\_quantity': 220\},
\{'demand': 270, 'order\_quantity': 270\}, \{'demand': 270, 'order\_quantity': 320\},
\{'demand': 295, 'order\_quantity': 70\}, \{'demand': 295, 'order\_quantity': 120\},
\{'demand': 295, 'order\_quantity': 170\}, \{'demand': 295, 'order\_quantity': 220\},
\{'demand': 295, 'order\_quantity': 270\}, \{'demand': 295, 'order\_quantity': 320\},
\{'demand': 320, 'order\_quantity': 70\}, \{'demand': 320, 'order\_quantity': 120\},
\{'demand': 320, 'order\_quantity': 170\}, \{'demand': 320, 'order\_quantity': 220\},
\{'demand': 320, 'order\_quantity': 270\}, \{'demand': 320, 'order\_quantity': 320\}]
    \end{Verbatim}

    Voila! We've recreated the same output as \texttt{ParameterGrid} gave
us.

\begin{quote}
This little exercise underscores the value of knowing the basics of OOP
if you are going to do analytics work in Python. Almost all of the main
libraries that you'll encounter (e.g.~pandas, matplotlib, scikit-learn)
are written in OO fashion. So, in order for you to be able to peek into
their source code, make sense of things, and maybe even borrow some code
as we did, you need to understand the basics of OOP in Python.
\end{quote}

    \hypertarget{creating-the-data_table-function}{%
\subsubsection{Creating the data\_table
function}\label{creating-the-data_table-function}}

Our plan is to create a function that takes three inputs:

\begin{itemize}
\tightlist
\item
  a model object such as \texttt{model\_5},
\item
  a dictionary containing the input attributes and associated ranges for
  the data table such as \texttt{dt\_param\_ranges},
\item
  a list of outputs which are methods of the the model object; for
  example \texttt{{[}\textquotesingle{}profit\textquotesingle{}{]}}
\end{itemize}

Since we don't want the \texttt{data\_table} function to modify our
passed in model object, we'll create a copy of the model within the
function. As we saw when copying list objects, we need to be careful
when copying user defined objects. We'll create what is known as a
\emph{deep copy} using the \texttt{copy} library. Learn more about
\href{https://docs.python.org/3/library/copy.html}{shallow and deep
copying here}.

Then we'll generate our parameter grid. I like to think of each element
in the parameter grid,
e.g.~\texttt{\{\textquotesingle{}demand\textquotesingle{}:\ 295,\ \textquotesingle{}order\_quantity\textquotesingle{}:\ 270\}},
as a \emph{scenario}. We loop over all the scenarios in the grid,
updating the model copy and computing the requested outputs. For each
scenario, we'll end up with a dictionary containing both the inputs and
outputs. For example,
\texttt{\{\textquotesingle{}demand\textquotesingle{}:\ 295,\ \textquotesingle{}order\_quantity\textquotesingle{}:\ 270,\ \textquotesingle{}profit\textquotesingle{}:405.0\}}.
Each such dictionary will get stored in a list and then at the very end
our function will convert this list to a pandas DataFrame and return it.

Here's our first version. Take note of the use of \texttt{getattr} to
indirectly call methods (e.g.~\texttt{profit}) of the model.

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{56}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{c+c1}{\PYZsh{} Demo of getattr for indirect method calling}
\PY{n+nb}{getattr}\PY{p}{(}\PY{n}{model\PYZus{}5}\PY{p}{,} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{profit}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}\PY{p}{(}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

            \begin{tcolorbox}[breakable, size=fbox, boxrule=.5pt, pad at break*=1mm, opacityfill=0]
\prompt{Out}{outcolor}{56}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
447.5
\end{Verbatim}
\end{tcolorbox}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{57}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{k}{def} \PY{n+nf}{data\PYZus{}table}\PY{p}{(}\PY{n}{model}\PY{p}{,} \PY{n}{scenario\PYZus{}inputs}\PY{p}{,} \PY{n}{outputs}\PY{p}{)}\PY{p}{:}
    \PY{l+s+sd}{\PYZsq{}\PYZsq{}\PYZsq{}Create n\PYZhy{}inputs by m\PYZhy{}outputs data table. }

\PY{l+s+sd}{    Parameters}
\PY{l+s+sd}{    \PYZhy{}\PYZhy{}\PYZhy{}\PYZhy{}\PYZhy{}\PYZhy{}\PYZhy{}\PYZhy{}\PYZhy{}\PYZhy{}}
\PY{l+s+sd}{    model : object}
\PY{l+s+sd}{        User defined object containing the appropriate methods and properties for computing outputs from inputs}
\PY{l+s+sd}{    scenario\PYZus{}inputs : dict of str to sequence}
\PY{l+s+sd}{        Keys are input variable names and values are sequence of values for each scenario for this variable.}
\PY{l+s+sd}{    outputs : list of str}
\PY{l+s+sd}{        List of output variable names}

\PY{l+s+sd}{    Returns}
\PY{l+s+sd}{    \PYZhy{}\PYZhy{}\PYZhy{}\PYZhy{}\PYZhy{}\PYZhy{}\PYZhy{}}
\PY{l+s+sd}{    results\PYZus{}df : pandas DataFrame}
\PY{l+s+sd}{        Contains values of all outputs for every combination of scenario inputs}
\PY{l+s+sd}{    \PYZsq{}\PYZsq{}\PYZsq{}}

    \PY{c+c1}{\PYZsh{} Clone the model using deepcopy}
    \PY{n}{model\PYZus{}clone} \PY{o}{=} \PY{n}{copy}\PY{o}{.}\PY{n}{deepcopy}\PY{p}{(}\PY{n}{model}\PY{p}{)}

    \PY{c+c1}{\PYZsh{} Create parameter grid}
    \PY{n}{dt\PYZus{}param\PYZus{}grid} \PY{o}{=} \PY{n+nb}{list}\PY{p}{(}\PY{n}{ParameterGrid}\PY{p}{(}\PY{n}{scenario\PYZus{}inputs}\PY{p}{)}\PY{p}{)}

    \PY{c+c1}{\PYZsh{} Create the table as a list of dictionaries}
    \PY{n}{results} \PY{o}{=} \PY{p}{[}\PY{p}{]}

    \PY{c+c1}{\PYZsh{} Loop over the scenarios}
    \PY{k}{for} \PY{n}{params} \PY{o+ow}{in} \PY{n}{dt\PYZus{}param\PYZus{}grid}\PY{p}{:}
        \PY{c+c1}{\PYZsh{} Update the model clone with scenario specific values}
        \PY{n}{model\PYZus{}clone}\PY{o}{.}\PY{n}{update}\PY{p}{(}\PY{n}{params}\PY{p}{)}
        \PY{c+c1}{\PYZsh{} Create a result dictionary based on a copy of the scenario inputs}
        \PY{n}{result} \PY{o}{=} \PY{n}{copy}\PY{o}{.}\PY{n}{copy}\PY{p}{(}\PY{n}{params}\PY{p}{)}
        \PY{c+c1}{\PYZsh{} Loop over the list of requested outputs}
        \PY{k}{for} \PY{n}{output} \PY{o+ow}{in} \PY{n}{outputs}\PY{p}{:}
            \PY{c+c1}{\PYZsh{} Compute the output.}
            \PY{n}{out\PYZus{}val} \PY{o}{=} \PY{n+nb}{getattr}\PY{p}{(}\PY{n}{model\PYZus{}clone}\PY{p}{,} \PY{n}{output}\PY{p}{)}\PY{p}{(}\PY{p}{)}
            \PY{c+c1}{\PYZsh{} Add the output to the result dictionary}
            \PY{n}{result}\PY{p}{[}\PY{n}{output}\PY{p}{]} \PY{o}{=} \PY{n}{out\PYZus{}val}

        \PY{c+c1}{\PYZsh{} Append the result dictionary to the results list}
        \PY{n}{results}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{result}\PY{p}{)}

    \PY{c+c1}{\PYZsh{} Convert the results list (of dictionaries) to a pandas DataFrame and return it}
    \PY{n}{results\PYZus{}df} \PY{o}{=} \PY{n}{pd}\PY{o}{.}\PY{n}{DataFrame}\PY{p}{(}\PY{n}{results}\PY{p}{)}
    \PY{k}{return} \PY{n}{results\PYZus{}df}
\end{Verbatim}
\end{tcolorbox}

    Okay, let's try it out. This time, I'll use \emph{dictionary unpacking}
to pass in the parameter values to our class.

\begin{itemize}
\tightlist
\item
  Create a dict called \texttt{base\_inputs} with values to use in the
  model
\item
  Then \textbf{base\_inputs} will \emph{unpack} the dictionary to key,
  value pairs that are used to specificy the keyword input arguments for
  out \texttt{BookstoreModel}.
\end{itemize}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{58}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{c+c1}{\PYZsh{} Create a dictionary of base input values}

\PY{n}{base\PYZus{}inputs} \PY{o}{=} \PY{p}{\PYZob{}}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{unit\PYZus{}cost}\PY{l+s+s1}{\PYZsq{}}\PY{p}{:} \PY{l+m+mf}{7.5}\PY{p}{,} 
               \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{selling\PYZus{}price}\PY{l+s+s1}{\PYZsq{}}\PY{p}{:} \PY{l+m+mf}{10.0}\PY{p}{,} 
               \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{unit\PYZus{}refund}\PY{l+s+s1}{\PYZsq{}}\PY{p}{:} \PY{l+m+mf}{2.5}\PY{p}{,} 
               \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{order\PYZus{}quantity}\PY{l+s+s1}{\PYZsq{}}\PY{p}{:} \PY{l+m+mi}{200}\PY{p}{,} 
               \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{demand}\PY{l+s+s1}{\PYZsq{}}\PY{p}{:} \PY{l+m+mi}{193}\PY{p}{\PYZcb{}}
\end{Verbatim}
\end{tcolorbox}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{59}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{c+c1}{\PYZsh{} Create a new model with inputs specified by base\PYZus{}inputs dict}
\PY{n}{model\PYZus{}6} \PY{o}{=} \PY{n}{BookstoreModel}\PY{p}{(}\PY{o}{*}\PY{o}{*}\PY{n}{base\PYZus{}inputs}\PY{p}{)}
\PY{n+nb}{print}\PY{p}{(}\PY{n}{model\PYZus{}6}\PY{p}{)}
\PY{n}{model\PYZus{}6}\PY{o}{.}\PY{n}{profit}\PY{p}{(}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]
\{'unit\_cost': 7.5, 'selling\_price': 10.0, 'unit\_refund': 2.5, 'order\_quantity':
200, 'demand': 193\}
    \end{Verbatim}

            \begin{tcolorbox}[breakable, size=fbox, boxrule=.5pt, pad at break*=1mm, opacityfill=0]
\prompt{Out}{outcolor}{59}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
447.5
\end{Verbatim}
\end{tcolorbox}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{60}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{c+c1}{\PYZsh{} Specify input ranges for scenarios (dictionary)}
\PY{c+c1}{\PYZsh{} 1\PYZhy{}way table}
\PY{n}{dt\PYZus{}param\PYZus{}ranges\PYZus{}1} \PY{o}{=} \PY{p}{\PYZob{}}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{demand}\PY{l+s+s1}{\PYZsq{}}\PY{p}{:} \PY{n}{np}\PY{o}{.}\PY{n}{arange}\PY{p}{(}\PY{l+m+mi}{70}\PY{p}{,} \PY{l+m+mi}{321}\PY{p}{,} \PY{l+m+mi}{25}\PY{p}{)}\PY{p}{\PYZcb{}}

\PY{c+c1}{\PYZsh{} 2\PYZhy{}way table}
\PY{n}{dt\PYZus{}param\PYZus{}ranges\PYZus{}2} \PY{o}{=} \PY{p}{\PYZob{}}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{demand}\PY{l+s+s1}{\PYZsq{}}\PY{p}{:} \PY{n}{np}\PY{o}{.}\PY{n}{arange}\PY{p}{(}\PY{l+m+mi}{70}\PY{p}{,} \PY{l+m+mi}{321}\PY{p}{,} \PY{l+m+mi}{25}\PY{p}{)}\PY{p}{,}
                     \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{order\PYZus{}quantity}\PY{l+s+s1}{\PYZsq{}}\PY{p}{:} \PY{n}{np}\PY{o}{.}\PY{n}{arange}\PY{p}{(}\PY{l+m+mi}{70}\PY{p}{,} \PY{l+m+mi}{321}\PY{p}{,} \PY{l+m+mi}{50}\PY{p}{)}\PY{p}{\PYZcb{}}

\PY{c+c1}{\PYZsh{} Specify desired outputs (list)}
\PY{n}{outputs} \PY{o}{=} \PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{profit}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{order\PYZus{}cost}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}

\PY{c+c1}{\PYZsh{} Use data\PYZus{}table function to create 1\PYZhy{}way data table}
\PY{n}{m6\PYZus{}dt1\PYZus{}df} \PY{o}{=} \PY{n}{data\PYZus{}table}\PY{p}{(}\PY{n}{model\PYZus{}6}\PY{p}{,} \PY{n}{dt\PYZus{}param\PYZus{}ranges\PYZus{}1}\PY{p}{,} \PY{n}{outputs}\PY{p}{)}
\PY{n}{m6\PYZus{}dt1\PYZus{}df}
\end{Verbatim}
\end{tcolorbox}

            \begin{tcolorbox}[breakable, size=fbox, boxrule=.5pt, pad at break*=1mm, opacityfill=0]
\prompt{Out}{outcolor}{60}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
    demand  profit  order\_cost
0       70  -475.0      1500.0
1       95  -287.5      1500.0
2      120  -100.0      1500.0
3      145    87.5      1500.0
4      170   275.0      1500.0
5      195   462.5      1500.0
6      220   500.0      1500.0
7      245   500.0      1500.0
8      270   500.0      1500.0
9      295   500.0      1500.0
10     320   500.0      1500.0
\end{Verbatim}
\end{tcolorbox}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{61}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{c+c1}{\PYZsh{} Use data\PYZus{}table function to create 2\PYZhy{}way data table}
\PY{n}{m6\PYZus{}dt2\PYZus{}df} \PY{o}{=} \PY{n}{data\PYZus{}table}\PY{p}{(}\PY{n}{model\PYZus{}6}\PY{p}{,} \PY{n}{dt\PYZus{}param\PYZus{}ranges\PYZus{}2}\PY{p}{,} \PY{n}{outputs}\PY{p}{)}
\PY{n}{m6\PYZus{}dt2\PYZus{}df}
\end{Verbatim}
\end{tcolorbox}

            \begin{tcolorbox}[breakable, size=fbox, boxrule=.5pt, pad at break*=1mm, opacityfill=0]
\prompt{Out}{outcolor}{61}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
    demand  order\_quantity  profit  order\_cost
0       70              70   175.0       525.0
1       70             120   -75.0       900.0
2       70             170  -325.0      1275.0
3       70             220  -575.0      1650.0
4       70             270  -825.0      2025.0
..     {\ldots}             {\ldots}     {\ldots}         {\ldots}
61     320             120   300.0       900.0
62     320             170   425.0      1275.0
63     320             220   550.0      1650.0
64     320             270   675.0      2025.0
65     320             320   800.0      2400.0

[66 rows x 4 columns]
\end{Verbatim}
\end{tcolorbox}

    Let's plot the 2-way results using Seaborn.

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{62}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{profit\PYZus{}dt\PYZus{}g} \PY{o}{=} \PY{n}{sns}\PY{o}{.}\PY{n}{FacetGrid}\PY{p}{(}\PY{n}{m6\PYZus{}dt2\PYZus{}df}\PY{p}{,} \PY{n}{col}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{order\PYZus{}quantity}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{sharey}\PY{o}{=}\PY{k+kc}{True}\PY{p}{,} \PY{n}{col\PYZus{}wrap}\PY{o}{=}\PY{l+m+mi}{3}\PY{p}{)}
\PY{n}{profit\PYZus{}dt\PYZus{}g} \PY{o}{=} \PY{n}{profit\PYZus{}dt\PYZus{}g}\PY{o}{.}\PY{n}{map}\PY{p}{(}\PY{n}{plt}\PY{o}{.}\PY{n}{plot}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{demand}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{profit}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{output_164_0.png}
    \end{center}
    { \hspace*{\fill} \\}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{63}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{profit\PYZus{}dt\PYZus{}g}\PY{o}{.}\PY{n}{savefig}\PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{images/two\PYZus{}way\PYZus{}dt.png}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    Perfect! The plots make sense in that the order quantity acts as an
upper limit on profit since we end up with unmet demand whenever demand
exceeds our order quantity.

    \hypertarget{concluding-thoughts-and-next-steps}{%
\subsection{Concluding thoughts and next
steps}\label{concluding-thoughts-and-next-steps}}

We've created an OO version of a simple spreadsheet type model that
lends itself pretty well to the type of sensitivity analysis we would do
with something like Excel's Data Table tool. Our Python
\texttt{data\_table} function actually allows \emph{n} input variables
and \emph{m} output variables. In Excel, you can either do a \(1xn\) (a
one-way Data Table) or \(2x1\) (a two-way Data Table).

Along the way we learned the basics of doing OOP in Python and a bunch
of more advanced Python functions and techniques. While that might have
seemed like a bunch of work, now we've got a reusable model object and
reusable data table function that we can build on.

In the next post, we'll build on this work and take on implementing an
Excel-style Goal Seek function in Python. We'll explore this using both
the OO and the non-OO models. Each leads to some interesting challenges.
Which approach do you think will end up being easier to use within a
Python \texttt{goal\_seek} function?

Then we'll use these same models for Monte-Carlo simulation. We'll be
creating our own \texttt{simulate} function to add to the
\texttt{data\_table} and \texttt{goal\_seek} functions that we created
in the first and second parts of this series.

In the fourth post in this series, we'll learn to create a Python
package containing our three ``what-if?'' functions so that we can share
our work and easily use these functions in different notebooks or Python
programs.

    \hypertarget{answers}{%
\subsection{ANSWERS}\label{answers}}

    \textbf{QUESTION 1:} Complete the lines of code below (answer at bottom
of notebook)

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{ }{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{order\PYZus{}cost} \PY{o}{=} \PY{n}{unit\PYZus{}cost} \PY{o}{*} \PY{n}{order\PYZus{}quantity}
\PY{n}{sales\PYZus{}revenue} \PY{o}{=} \PY{n+nb}{min}\PY{p}{(}\PY{n}{order\PYZus{}quantity}\PY{p}{,} \PY{n}{demand}\PY{p}{)} \PY{o}{*} \PY{n}{selling\PYZus{}price}
\PY{n}{refund\PYZus{}revenue} \PY{o}{=} \PY{n+nb}{max}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n}{order\PYZus{}quantity} \PY{o}{\PYZhy{}} \PY{n}{demand}\PY{p}{)} \PY{o}{*} \PY{n}{unit\PYZus{}refund}
\PY{n}{profit} \PY{o}{=} \PY{n}{sales\PYZus{}revenue} \PY{o}{+} \PY{n}{refund\PYZus{}revenue} \PY{o}{\PYZhy{}} \PY{n}{order\PYZus{}cost}
\end{Verbatim}
\end{tcolorbox}

    \textbf{CHALLENGE 2} Lost sales method

We could add the following method to our \texttt{BookstoreModel} class.

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{ }{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{k}{def} \PY{n+nf}{lost\PYZus{}sales}\PY{p}{(}\PY{n+nb+bp}{self}\PY{p}{)}\PY{p}{:}
    \PY{l+s+sd}{\PYZdq{}\PYZdq{}\PYZdq{}Compute number of items demanded but no inventory to sell\PYZdq{}\PYZdq{}\PYZdq{}}
    \PY{k}{return} \PY{n}{np}\PY{o}{.}\PY{n}{maximum}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{demand} \PY{o}{\PYZhy{}} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{order\PYZus{}quantity}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{ }{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]

\end{Verbatim}
\end{tcolorbox}

    % Add a bibliography block to the postdoc

\end{document}
